### 2022-01-12 manuscript/main.tex
backing up copy of manuscript

\documentclass[a4paper,man,natbib]{apa6}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}

% specify link style
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    citecolor=blue
}

\title{An item response theory analysis of the matrix reasoning item bank (MaRs-IB)}
\shorttitle{IRT analysis of MaRs-IB}
\author{Samuel Zorowitz$^1$, Nathaniel D. Daw$^{1,2}$}
\affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$Department of Psychology, Princeton University, USA}

\abstract{Your abstract here.}

\begin{document}
\maketitle

\section{Introduction}

Matrix reasoning tasks are one of the most well-studied and useful tasks in the behavioral sciences. In a typical administration, participants are presented with a 3x3 matrix in which 8 of 9 cells contains abstract figures or shapes. Participants are instructed to identify the missing figure from among several alternatives which are presented below the matrix. To do so, participants must use abstract reasoning to identify the piece that would complete the pattern. Matrix reasoning indexes fluid intelligence and is correlated with working memory capacity \citep{conway2002latent, kane2004generality, unsworth2005working, salthouse2014relations}. The progressive matrices has excellent predictive validity and is associated of academic performance \citep{roth2015intelligence} and college admission exams \citep{frey2004scholastic, koenig2008act}. In low-stakes testing environments, performance on the progressive matrices is associated with participant motivation and willingness to expend effort \citep{gignac2018moderate, gignac2019maximum}. For all of these reasons, the progressive matrices are increasingly used as a control measure in individual difference behavioral studies \citep{gillan2016characterizing, rouault2018psychiatric, moutoussis2021decision}.

Despite their utility, there are notable challenges in using matrix reasoning measures. The first is that of copyright. The most prominent measures, including the Raven's progressive matrices \citep{raven2003raven} and the Wechsler Abbreviated Scale of Intelligence matrix reasoning subtest \citep{wechsler1999wechsler}, are not free to use and copyright may prevent these pen-and-paper tasks from being adapted into computerized tasks. A second challenge is that measures that are free-to-use and in the public domain typically have very few items available. For example, the Hagen Matrices Test \citep{heydasch2014hagen} and International Cognitive Ability Resource \citep{condon2014international} have only 20 and 16 items available, respectively. The limited number of items raises the issue of practice effects \citep{ng1974applicability, bors2003effect}, especially in the context of online experiments where many lab groups may be testing the same participants.

To address these gaps, \cite{chierchia2019matrix} developed the matrix reasoning item bank (MaRs-IB). The MaRs-IB is an open-source, free-to-use set of 80 matrix reasoning items, each with three unique shape variants and two sets of distractors. The items span a wide range of complexity -- in terms of both their number of features and number of changing relations. Example items here?  In a sample of over 600 adults, adolescents, and children, the authors found that a 8-minute MaRs-IB test measure had excellent psychometric properties (high internal consistency and test-retest reliability) and convergent validity (correlations with ICAR matrix reasoning and digit span performance). Most importantly, the authors released item-level performance data for each item, thereby raising the possibility for other researchers to yield age-appropriate tasks of custom difficulty and duration.

There are, however, two impediments in researchers using the MaRs-IB norms. Both concern the design of the original MaRs-IB task. The initial MaRs-IB task was a fixed-time task, wherein participants completed as many items as possible in 8 minutes. The first issue then is that, because participants completed 33 items on average, there exists relatively little data on the majority of the MaRs-IB items (this is before taking into consideration the data available for each item variant). The second issue is that, because of individual differences in speed-accuracy trade-offs, the estimates of item difficulties may be biased. That is, items completed by fewer participants are more likely to have been reached by participants who had been sacrificing accuracy for speed, thereby giving a distorted view of the difficulty of these items.

Summary paragraph: here we test a large sample of adults, making sure that all items and variants were tested many times. We fix the number of items to prevent bias. Importantly, we use item response models to calibrate the data. This means we can accurately measure item difficulty, taking into considering participant ability, as well as discirmination and guessing. This in turn can be used to construct new tests with known properties, by test information or test functioning. We use this to construct three new short forms, which we test. We make all data and parameters public.y

ALL LINKS TO REPOS SHOULD BE HERE

\section{Calibration Study}

\subsection{Data collection}

A total of N=1584 participants were recruited from the Prolific Academic platform (https://www.prolific.co) to participate in an online behavioral experiment in July - August, 2021. Participants were eligible if they currently resided in the United States. This study was approved by the Institutional Review Board of Princeton University (\#7392), and all participants provided informed consent. Total study duration was approximately 6.4 minutes (2.4 sd) per participant. Participants received monetary compensation for their time (rate USD \$10/hr), plus an incentive-compatible bonus up to \$0.50 based on task performance (average payment: mean = \$1.30 [\$0.10 sd]). The experiment

cite https://doi.org/10.1111/bjop.12288, https://doi.org/10.1073/pnas.1018601108

After providing consent, participants completed a series of items from the matrix reasoning item bank. The design of the MaRs-IB has been described elsewhere \citep{chierchia2019matrix} and was based on publicly available code (https://app.gorilla.sc/openmaterials/36164). Briefly, each MaRs-IB item consisted of a 3 × 3 matrix. Eight of the nine resulting cells contained an abstract shape, while one cell on the bottom right-hand side of the matrix was empty. The participants' task was to complete the matrix by finding the missing shape among four possible alternatives. Items began with a 1200 ms fixation cross. Participants were then given up to 30 s to complete an item. After 25 s a clock appeared, indicating that 5 s remained before the next item began. An item ended when participants responded, or after 30 s had elapsed without response. Participants reviewed instructions before the task and were made to correctly complete three practice items before they could begin the task.

In order to ensure adequate sampling of all items, and in contrast to the original task design, all participants completed a pseudo-randomly selected subset of 16 out of 64 possible MaRs-IB items \footnote{We did not collect new data on 16 of the original MaRs-IB items. We excluded all Dimension 1 items, which are the simplest items and tend to have performance levels at ceiling. We also excluded Item 1 which, although not Dimension 1, also has performance levels near ceiling.} Item selection proceeded as follows: The 64 items were sorted into 16 groups by dimension. Participants were then randomly assigned one item from each of the groups. Importantly, participants were randomly assigned a clone of each item that varied in distractor type (md, pd) and item set (1, 2, 3). In this way, all items were sampled roughly 375 times (12 sd) and each clone was sampled 62 times (7 sd).

The data from multiple participants who completed the experiment were excluded prior to analysis. The data from N=7 participants were excluded for having browser resolutions smaller than necessary to see the task. The data from N=8 participants were excluded for missing responses on 4 or more trials. The data from N=70 participants were excluded for rapid guessing (defined as response times of less than 3 s) on 4 or more trials. In total, 83 of 1584 (5.24\%) were excluded, leaving the data from N=1501 participants for analysis.

The majority of participants in the calibration sample identified as women (men: N=670; women: N=811; non-binary or other: N=13; rather not say: N=2). Participants were 28.7 years old (9.9 sd) on average. The sample was relatively well-educated with the majority having a Bachelor's degree (N=507) or Master's degree or higher (N=322); fewer participants endorsed only completing some college (N=471) or only a high school degree (N=199).

The experiment was programmed in jsPsych \citep{de2015jspsych} and distributed using custom web-application software. The experiment code is available at \url{https://github.com/szorowi1/mars-irt}, and the web-software is available at \url{https://github.com/nivlab/nivturk}. A playable demo of the task is available at \url{https://nivlab.github.io/jspsych-demos/tasks/mars/experiment.html}.

\subsection{Descriptive statistics}

On average, participants completed 9.6 of 16 items correctly (sd = 3.1, IQR = 8.0 - 12.0) (Figure \ref{fig:fig01}a). Timing out occurred in only 1.8\% of items. A total of 43 participants (2.9\%) performed below chance (i.e. fewer than four items correct), while only 11 participants (0.7\%) responded correctly to all items. Thus, over 95\% of participants' observed scores were within a reasonable range. These results corroborate \cite{chierchia2019matrix} in that there were no obvious ceiling or floor effects, and that participants were motivated to perform despite minimal supervision in the online environment.

Across all 384 items, the average proportion correct was 0.60 (sd = 0.19, IQR = 0.45 - 0.74) (Figure \ref{fig:fig01}b). A total of 12 items (3.1\%) had performance beneath chance levels (25\%), and only 22 items (5.7\%) had performance levels above 90\%. As such, the majority of items sampled exhibited adequate performance with no obvious ceiling or floor effects.

Though not the primary focus of the analyses, we briefly investigated the relationship between participant' total scores, item difficulty, and response times. We fit a mixed-effects linear model to regress the log response time for trial $i$ on accuracy, the participant's observed score (minus their accuracy on that trial), the difficulty of the time (one minus the proportion correct), and the interaction of observed score and difficulty. We specified an intercept per participant to control for each participants' overall speed.

On average, participants spent 15.9 seconds to solve an item (sd = 7.2 s, IQR = 10.1 - 21.5 s). The mixed-effects model identified a significant effect of accuracy ($\beta = 0.067, t = 9.84, p < 1e^{-6}$), such that correct responses were on average slower than incorrect responses. Notably, there were large, significant effects of observed score ($\beta = 0.125, t = 16.948, p < 1e^{-6}$) and item difficulty ($\beta = 0.137, t = 44.326, p < 1e^{-6}$) such that participants were slower overall for more difficult items and for better performing participants. Finally, the interaction between observed score and item difficulty was also significant ($\beta = 0.058, t = 20.650, p < 1e^{-6}$), such that better performing participants slowed down even more for more difficult items.

The full pattern of response time results is summarized in Figure \ref{fig:fig01}c. In short, we corroborate the previous finding that item difficulty is associated with slower response times \citep{chierchia2019matrix}. In addition, we find two interesting patterns: better performing participants are slower on average, and this slowing is enhanced as items become more difficult. What this suggests then is that participants' overall performance on the MaRs-IB in this sample is reflective of speed-accuracy trade-offs likely reflective of individual differences in persistence or willingness to expend mental effort \citep{ranger2021effects}. That is, better performing participants are willing to spend more time on an item, whereas worse performing participants are more likely to give up sooner and guess as items become more difficult.

\begin{figure}
\centering
\includegraphics[width=1.0\textwidth]{figures/fig01.png}
\caption{\label{fig:fig01}This is a figure caption.}
\end{figure}

\subsection{Item response models}

Reflecting the inherent hierarchical structure of the MaRs-IB -- 80 item families with six clones each -- we estimated the psychometric properties of each item using item cloning models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}. All models were extensions of the three-parameter logistic (3PL) item repsonse model, where the probability of a correct response ($y_{ijk} = 1$) for person $i$ on item clone $k$ belonging to item family $j$ is:

\begin{equation}
p(y_{ijk} = 1) = \gamma_{jk} + (1-\gamma_{jk}) \cdot \text{logit}^{-1} \left( \alpha_{jk} \cdot \theta_i - \beta_{jk} \right)
\end{equation}

\noindent where $\theta_i$ is the latent ability for person $i$, and $\beta_{jk}$, $\alpha_{jk}$, and $\gamma_{jk}$ are the difficulty, discrimination, and guessing parameters for item clone $k$ of item family $j$. Because guessing parameters are difficult to estimate without very large amounts of data \citep{han2012fixing}, we fixed the guessing rate for each item to the nominal rate ($\gamma_{jk} = 0.25$).

To account for the nested structure of the MaRs-IB, the item difficulty and discrimination parameters were estimated using an additive multilevel item structure model \citep{cho2014additive}. Specifically, the difficulty of item clone $k$ belonging to item family $j$ was expressed as:

\begin{equation}
\beta_{jk} = \mu_\beta + \sum_{n=1}^N Q_{jn} \delta_{\beta n} + \epsilon_{\beta j} + \sum_{m=1}^M R_{km} \delta_{\beta m} + \epsilon_{\beta k}
\end{equation}

\noindent The difficulty parameter is thus modeled as an intercept ($\mu_\beta$), reflecting the average difficulty across all items, and four components:

\begin{enumerate}

\item $\sum_{n=1}^N Q_{jn} \delta_{\beta n}$: the effect of the level-1 attributes, where $Q_{jn}$ is the value of attribute $n$ for item family $j$, and $\delta_{\beta n}$ is the effect of attribute $n$. This component is the fixed effects contribution of item family attributes to item difficulty.

\item $\epsilon_{\beta j}$: the level-1 residual, or variance unexplained by the item family attributes. This component is the random effects contribution of item family attributes to item difficulty.

\item $\sum_{m=1}^M R_{km} \delta_{\beta m}$: the effect of the level-2 attributes, where $R_{km}$ is the value of attribute $m$ for item clone $k$, and $\delta_{\beta m}$ is the effect of attribute $m$. This component is the fixed effects contribution of item clone attributes to item difficulty.

\item $\epsilon_{\beta k}$: the level-2 residual, or variance unexplained by the item clone attributes. This component is the random effects contribution of item clone attributes to item difficulty.

\end{enumerate}

To elaborate, level-1 attributes describe the properties of an item family. Here we model the effects of two properties: feature number, or the number of unique shapes in each cell of an item; and rule number, or the number of changing relationships among the shapes (e.g. color change, shape change, position change) one has to consider in order to correctly solve the item. Example items illustrating these attributes are included in the Supplement.  Note that this definition of item rules deviates from other established taxonomies of matrix reasoning items \citep{carpenter1990one}, a point we return to in the Discussion.

In turn, level-2 attributes describe properties of an item clone. Here we model the effects of two additional properties: distractor type, or whether the distractors were generated following the minimal difference (MD) strategy and a paired difference (PD) strategy (see \cite{chierchia2019matrix} for details); and mean response time, or the average time it took to complete that item. The latter attribute was orthogonalized with respect to the three other item attributes and thus reflects any residual structure after accounting for all other modeled effects.

Similarly, item discrimination parameters were also expressed as the sum of an equivalent set of components:

\begin{equation}
\alpha_{jk} = \mu_\alpha + \sum_{n=1}^N Q_{jn} \delta_{\alpha n} + \epsilon_{\alpha j} + \sum_{m=1}^M R_{km} \delta_{\alpha m} + \epsilon_{\alpha k}
\end{equation}

\noindent The fixed effect components of the item discrimination parameters were estimated using the same level-1 and level-2 attributes as for the item difficulty parameters (i.e. number of features, number of rules, distractor type, average response time). During estimation, the item discrimination parameters were restricted to be in the range $\alpha_{jk} \in [0, 5]$.

Though complex, the item structure model above provides a natural way of measuring the quality of the MaRs-IB. In the perfect situation, the clones of each item family would be psychometrically equivalent. So too should item families with identical attributes (e.g. number of features, number of rules). This is, of course, unlikely in actuality and the item cloning model allows us to quantify the extent to which equivalent families and clones vary in their difficulty and discrimination.

To quantify the success of the item cloning process, we estimated a series of four models \citep{cho2014additive, lathrop2017item}:

\begin{enumerate}

\item \textit{Model 1}: Item parameters are expressed only as a function of their level-1 attributes. This model treats all clones within a family as interchangeable and all item families with the same attributes as equivalent.

\item \textit{Model 2}: Item parameters are expressed as a function of their level-1 attributes and residuals. This model treats all clones within a family as interchangeable, but now item families with the same attributes may psychometrically vary.

\item \textit{Model 3}: Item parameters are expressed as a function of their level-1 attributes, level-2 attributes, and level-1 residuals. This model allows clones within a family to vary, but all variation must be explained solely through their attributes.

\item \textit{Model 4}: Item parameters are expressed as a function of their level-1/2 attributes residuals. The full model in which item families and clones may vary in their psychometric properties.

\end{enumerate}

The resulting models were then submitted to model comparison. This procedure allows us to measure the success of the item generation process. If the psychometric properties of item families reflect only their attributes, then Model 1 should be preferred to Model 2; in contrast, if item families vary in their properties above and beyond what their attributes can explain, then Model 2 should be preferred. Similarly, if the clones of item family are truly equivalent, then Models 1 or 2 should be preferred to Models 3 or 4; in contrast, if item clones vary substantially (systematically or otherwise), then Models 3 or 4 should be preferred to Models 1 or 2.

Finally, we estimated person-level abilities following an item response modeling approach \citep{wilson2008explanatory}:

\begin{equation}
\theta_i = \sum_{p=1}^P X_{ip} \rho_p + \epsilon_i    
\end{equation}

\noindent where $X_{ip}$ is the value of attribute $p$ for person $i$, $\rho_p$ is the effect of attribute $n$, and $\epsilon_i$ is variance in ability unexplained by the person-level attributes.

We modeled person-level abilities as a function of four attributes: age, gender (coded as: male = -0.5, female = 0.5; all else = 0), average response time, and $\Delta$ response time. The last term reflected a person's adjustment of their response time as a function of item difficulty (defined as its corresponding proportion of correct responses), and thus measured a participant's tendency to slow down (or speed up) in response to more challenging items.

To ensure the identifiability of all models, person abilities were assumed to be $\theta_i \sim \mathcal{N} \left(0,1\right)$.

\begin{equation}
blah 
\end{equation}

All models were estimated within a Bayesian framework using Hamiltonian Monte Carlo as implemented in Stan (v2.26) \citep{carpenter2017stan}. For all models, four separate chains with randomised start values each took 7500 samples from the posterior. The first 5000 samples from each chain were discarded. As a result, 10000 post-warmup samples from the joint posterior were retained. The $\hat{R}$ values for all parameters was less than 1.01, indicating acceptable convergence between chains, and there were no divergent transitions in any chain.

Models were compared using Bayesian leave-one out cross-validation \citep{vehtari2017practical}, which has been found to perform better than other information criteria for comparing item response models \citep{luo2017performances}. Importantly, we computed the marginalized leave-one-cluster out (LOCO) cross-validation \citep{merkle2019bayesian}, which measures models' ability to generalize to held out items (rather than held out responses).

The fit of the winning model to the data was evaluated through several approaches. First, the fit of the models to the data was assessed by means of posterior predictive model checking \citep{gelman1996posterior}. Item fit was assessed using two two discrepancy statistics: Orlando and Thissen index ($S-\chi^2$) \citep{toribio2011discrepancy} and the generalized dimensionality discrepancy measure (GDDM) \citep{levy2009posterior, levy2011generalized}. The $S-\chi^2$ index is a measure of item fit based on the comparison of observed and expected proportions of participants who answer an item correctly at each raw-score group. The GDDM is a measure to test the model's unidimensionality assumption, which must not be violated in order to compute valid total scores from tests assembled from MaRs-IB items.

\subsection{Model comparison}

First we estimated Models 1-4 holding constant the item discrimination parameters ($\alpha = 1$), which allowed us to investigate the difficulty of MaRs-IB items independently. (Note that these models still included guessing parameters fixed to the nominal guessing rate. We also estimated models with no guessing parameter, i.e. 1PL or Rasch models, but model comparison revealed these to be inferior.) The goodness of fit of the four models was compared based on leave-one-cluster out (LOCO) cross-validation as summarized in Table \ref{table:1}.

First, Model 2 was preferred to Model 1 suggesting that there is variability in the difficulty of item families above and beyond what can be explained by their attributes. Similarly, Model 3 was preferred to Model 2 suggesting that there is systematic variability in the clones of each item family. Finally, Model 4 was preferred to Model 3 suggesting that there is variability in the difficulty of item clones above and beyond what can be explained by their respective attributes. In summary, the full model -- including all fixed and random effect terms -- was preferred to the three simpler models, indicating that item families and clones are nonequivalent even after controlling for their attributes.

On average, items were difficult (95\% HDI), but ranged considerably in difficulty. Over half of the variability in item difficulty (65.6\%) was explained by the fixed effects of item family (level-1) and clone (level-2) attributes. The associations between level-1 attributes and item difficulty were as expected: feature number was positively associated with difficulty (beta), such that the inclusion of one additional feature was associated with a 13.1\% reduction in expected accuracy. Similarly, rule number was positively associated with item difficulty (beta), such that the inclusion of one additional rule a 10.1\% reduction in expected accuracy. Turning next to level-2 attributes, distractor type was unexpectedly associated with item difficulty (beta), such that clones with MD distractors clones were associated with a 13.6\% reduction in accuracy compared to clones with PD distractors. This pattern is clearly evident in the raw item-level scores (Figure XX). Finally, average response time was positively associated with difficulty, such that a one standard deviation change in response time was associated with an 11.8\% reduction in accuracy.

The inclusion of item family (level-1) random effects explained an additional 18.5\% of the variance in item difficulty. In total, the percentage of variance in item difficulty not explained by residual variance from item clones (level-2) was 84.1\%. The residual variance was harder to explain (Figure XXb); indeed, there was no clear patterns in difficulty by item set, suggesting that at least part of residual variance may be explained by idiosynchratic perceptual features of item clones.

Next we estimated three models as above, but with two key differences: First, we no longer held constant the item discrimination parameters but rather estimated them alongside the item difficulty parameters. Second, whereas the item difficulty parameters were always estimated using the full additive multilevel item structure model (i.e. level-1/2 fixed and random effects), the item discrimination parameters were estimated adding one additional component per model (Model 1: level-1/2 fixed effects only; Model 2: fixed effects and level-1 random effects; Model 3: full model). The goodness of fit of the three models was again compared based on leave-one-cluster out (LOCO) cross-validation as summarized in Table \ref{table:2}.

First, all three models including discrimination parameters were preferred to the best-fitting model with discrimination parameters held constant. Model 2 was preferred to Model 1 suggesting that there is variability in the discrimination of item families not accounted for by their attributes. Importantly, Model 2 was preferred to Model 3 such that there was insufficient variability in discrimination in item clones to necessitate additional parameters. We return to this point in the Discussion.

Returning to variance partitioning, just less than half of the variance in item discrimination (44.9\%) was explained by attributes of the explained by attributes of item families and clones. Interestingly, only the number of item rules was positively associated with item discrimination (HDI); no other attributes were credibly related (HDIs here).

Finally, we turn to explanatory item response modeling. The largest effects were from the response times. Average response time blah. Delta response time blah. Interestingly, we found a negative correlation between age and ability, such that older participants tended to perform worse on average. Finally, we observed a small but credible association between gender and ability. Though we used the colorblind stimulus set, it is possible there was residual perceptual difficulty for men compared to women. Granted, this effect is small.

This paragraph should be on fit. Should mention: simulation study, posterior predictive check, item fit indices, and unidimensionality measure. You should probably review a couple of other RPM validation papers to make sure our methods are straight (e.g. doi.org/10.1037/a0027830)

\subsection{Discussion}

In discussing the success of the MaRs-IB generation process, some things to point out:

1. A lot of variance in item difficulty and discrimination can be explained by inherent features of the items. This is great!

2. Though there is quite a bit of residual variance at the item family level, this is fine as we only modeled two features. A point we return to in the Discussion.

3. At the clone level, it seems clear that item clones are not equivalent w.r.t. difficulty. For one thing, MD vs. PD big. Should describe level-2 standard deviation somewhere.

4. For clone discrimination, it seems less clear. One possibility is that there isn't much variability (cite Embertson). Another possibility is that we are underpowered. Cite Lathrop here for discussion of potential bias.

Summarize as good to go!

\section{Test assembly}

With 64 items, each with six clones, there is a vast universe of possible test forms we could construct. Beginning with our motivating problem -- abbreviated measures of nonverbal reasoning with the potential for retest -- we opted to construct three parallel short form measures. To do so, we used mixed integer programming \citep{der2005wj} to maximize the test information function (TIF) of each short form subject to the following constraints:

\begin{enumerate}

    \item Each short form was required to contain 12 items. This was chosen to minimize administration time of a form (2-4 minutes on average) and to achieve sufficient  reliability (Cronbach's $\alpha \geq 0.8$ in piloting).

    \item All short forms were required to be comprised of clones from the same item families. This was chosen to maximize the similarity of the short forms.

    \item Assigned clones could be of the MD or PD distractor type, but not both. This was chosen to prevent item overlap, as sometimes the MD and PD distractor strategies generate identical distractors.

    \item The difference in TIFs across short forms was required to be minimized. This was chosen to ensure that each short form had similar psychometric properties.

\end{enumerate}

\noindent The TIF was maximized at five ability levels ($\theta = -1.0, -0.5, 0.0, 0.5, 1.0$). Previous simulation studies have shown that target values at three to five well-chosen $\theta$s generally suffice \citep{der2005wj}. Solutions to the mixed integer programming problem were found using the \textit{mip} python package (v1.13.0) \citep{santos2020mixed}.

The test information functions (TIFs) and test characteristic curves (TCCs) for the three short forms are shown in Figure XX. As can be readily observed, the TIFs for each short form are remarkably similar reflecting the constraints of the MIP problem. Thus, the short forms should have similar levels of precision across ability levels. Similarly, the TCCs for each short form are similar, which was by no means guaranteed \citep{ali2016evaluation}. Thus, the short forms are expected to have similar total score distributions.

Thus we proceeded to test each short form. blah blah blah.

\section{Validation Study}

\subsection{Data Collection}

A total of N=347 participants were recruited from the Prolific Academic platform (https://www.prolific.co) to participate in an online behavioral experiment in November, 2021. Participants were eligible if they currently resided in the United States and had not participated in the calibration study. This study was approved by the Institutional Review Board of Princeton University (\#7392), and all participants provided informed consent. Total study duration was approximately 10.8 minutes (4.6 sd) per participant. Participants received monetary compensation for their time (rate USD \$10/hr), plus an incentive-compatible bonus up to \$0.75 based on task performance (average payment: mean = \$2.10 [\$0.18 sd]).

After providing consent, participants completed the experiment which was divided into three section. In the first, participants completed three short surveys: the 10-item need for cognition survey \citep{chiesi2018applying}, the 8-item PROMIS Cognitive Function-Short Forms (8a) \citep{iverson2021normative}, and the 8-item subject numeracy scale \citep{fagerlin2007measuring}. These measures were included as part of exploratory analyses that were not the central focus of this study and will not be reported in the main text (reported in Supplement). Next, participants completed one of the three 12-item MaRs-IRB short-forms described above. Besides for the items, the administration procedure of the short-forms was identical to that in the Calibration study. Finally, participants completed the 9-item abbreviated Raven's progressive matrices (form A) \citep{bilker2012development}. For consistency's sake, the administration procedure was matched to the MaRs-IB.

The data from a total of N=47 participants were excluded prior to analysis. Of these, N=17 participants were excluded for having not completed the entire experiment. Separately, the data from N=21 participants were excluded for failing one or more attention checks (infrequency items) embedded in the self-report measures \citep{zorowitz2021inattentive}. The data from N=13 participants were excluded for image loading problems on either the MaRs-IB or RPM tasks. Finally, N=6 participants were excluded for missing or rapid guessing on 4 or more trials during the MaRs-IB or RPM tasks. Note these are not mutually exclusive.

The majority of participants in the validation sample identified as women (men: N=138; women: N=147; non-binary or other: N=13; rather not say: N=2). Participants were 35.0 years old (13.2 sd) on average. The sample was relatively well-educated with the majority having a Bachelor's degree (N=121) or Master's degree or higher (N=45); fewer participants endorsed only completing some college (N=100) or only a high school degree or lower (N=33).

\subsection{Results}

The distribution of scores on the three MaRs-IB short-forms are summarized in Table \ref{table:2}. Across the short-forms, participants responded correctly on 8.0 of 12 items (sd = 2.5, IQR = 6 - 10). The distribution of scores were skewed towards to higher end, as would be expected given their respective test characteristic curves. A 1-way ANOVA did not reject the null hypothesis that the observed scores across test forms were equal (F(2,97) = 0.253, p = 0.777). Corroborating the predictions from the test assembly, the reliability of each short form was satisfactory (Table \ref{table:2}).

The distribution of scores on the RPM short-form is similarly summarized in Table \ref{table:2}. On average, participants responded correctly on 4.5 of 9 items (sd = 2.0, IQR = 3 - 9). Scores on the RPM were significantly lower on the RPM than the MaRs-IB ($t = 25.598, p < 1e^{-6}$), indicating that the items on the RPM short-form are more difficult on average than those selected for the MaRs-IB short-forms. The reliability of the measure was similarly good as the MaRs-IB (Table \ref{table:2}).

To measure convergent validity, blah blah blah. Need to write model here. Or should we write above?

\subsection{Discussion}

\section{General Discussion}

Paragraph XX: item discrimination random effects. Power analysis? also variance explained being similar to literature.

-----
Notes on RT effects: this paragraph should come later, as a discussion point on what we're actually measuring with the MaRs-IB (at least online)

Heitz (2014): in general, an increase in speed comes at the expense of accuracy (speed-accuracy trade-off)

Goldhammer (2015):

> Individual differences in test performance depend on both between-person differences in speed and ability (i.e., the location of individual speed-ability curves) and the adopted speed-ability compromise within persons (i.e., the location within a speed-ability curve). T

Goldhammer et al. (2015): test takers that increase their work pace have a lower effective ability

> On previous reasoning tasks: "These findings indicate that in tasks that require understanding new problem situations and in which the use of complex cognitive processes is mandatory, higher effective ability is usually associated with lower speed."

Ranger et al. (2021):

> Such a relation between response time and response accuracy or between work pace and effective ability is of uttermost importance as it implies that the responses depend on more than just the test taker’s capability.

> the capability is apportioned among two quantities, the effective ability, and the work pace, which constitute the actual performance of a test taker in a test; see Pohl and von Davier (2018) for a discussion on how to describe the performance of test takers.

As work pace increases, effective ability decreases (irrespective of true capability). Work rate may reflect other processes: persistence, or amount of time before participant gives up, mental effort costs, or foraging costs.

Basically, summarize results. All points are consistent with dual process theory (automatic vs. controlled processing). suggestins higher ability participants more deliberate, lower ability participants less deliberate. frame from perspective of persistence (ranger 2021).

Need to cite: https://doi.org/10.1016/j.intell.2020.101490

-----

Paragraph XX: future directions modeling speed-accuracy trade-offs

Paragraph XX: model more features as future direction. Carpenter taxonomy. Model by rule type. By perceptual features. automatic item generation as future direction. See \cite{lathrop2017item} for inspiration.

\bibliography{example}

\begin{table}
\centering
\begin{tabular}{l|c|c|c}
Model & elpd$_{loco}$ & loco & $\Delta$ loco \\\hline
1PG-F &  & -14516 & \\
1PG-FX &  & -13804 & \\
1PG-R &  & -13481 & \\
1PG-RX &  & -13293 & - \\
\end{tabular}
\caption{Table to test captions and labels.}
\label{table:1}
\end{table}

\begin{table}
    \centering
    \begin{tabular}{c|c|c|c}
    Measure & Mean (sd) & IQR & Reliability \\
    \hline
    MaRs-IB SFa & 7.9 (2.4) & 7 - 10 & 0.761 \\
    MaRs-IB SFb & 8.2 (2.6) & 7 - 10 & 0.810 \\
    MaRs-IB SFc & 7.9 (2.7) & 6 - 10 & 0.829 \\
    RPM SFa & 4.5 (2.0) & 3 - 9 & 0.791 \\
    \end{tabular}
    \caption{Caption}
    \label{table:2}
\end{table}

\end{document}

### 2022-01-16 manuscript/main.tex
writing day writing day

@@
 
-Despite their utility, there are notable challenges in using matrix reasoning measures.
-The first is that of copyright.
-The most prominent measures, including the Raven's progressive matrices \citep{raven2003raven} and the Wechsler Abbreviated Scale of Intelligence matrix reasoning subtest \citep{wechsler1999wechsler}, are not free to use and copyright may prevent these pen-and-paper tasks from being adapted into computerized tasks.
-A second challenge is that measures that are free-to-use and in the public domain typically have very few items available.
+Despite their utility, there are a number of challenges in using matrix reasoning measures in behavioral research.
+The first is the issue of copyright.
+The most prominent versions, including the Raven's progressive matrices \citep{raven2003raven} and the matrix subtest of the Wechsler Abbreviated Scale of Intelligence \citep{wechsler1999wechsler}, are not free to use and copyright may prevent these pen-and-paper tasks from being adapted into computerized tasks.
+A second challenge is those measures that are in the public domain typically have few items available.
 For example, the Hagen Matrices Test \citep{heydasch2014hagen} and International Cognitive Ability Resource \citep{condon2014international} have only 20 and 16 items available, respectively.
-The limited number of items raises the issue of practice effects \citep{ng1974applicability, bors2003effect}, especially in the context of online experiments where many lab groups may be testing the same participants.
+The availability of only a limited number of items raises the issue of practice effects \citep{ng1974applicability, bors2003effect}, especially in the context of online experiments where multiple groups of researchers may be unknowingly administering the same items to the same participants.
 
-To address these gaps, \cite{chierchia2019matrix} developed the matrix reasoning item bank (MaRs-IB).
-The MaRs-IB is an open-source, free-to-use set of 80 matrix reasoning items, each with three unique shape variants and two sets of distractors.
-The items span a wide range of complexity -- in terms of both their number of features and number of changing relations.
-Example items here?
-In a sample of over 600 adults, adolescents, and children, the authors found that a 8-minute MaRs-IB test measure had excellent psychometric properties (high internal consistency and test-retest reliability) and convergent validity (correlations with ICAR matrix reasoning and digit span performance).
-Most importantly, the authors released item-level performance data for each item, thereby raising the possibility for other researchers to yield age-appropriate tasks of custom difficulty and duration.
+In response to these challenges, \cite{chierchia2019matrix} developed the matrix reasoning item bank (MaRs-IB).
+The MaRs-IB is an open-source, free-to-use bank of 80 matrix reasoning items, where each item has three unique shape variants and two sets of distractors.
+The items were designed to vary considerably in their complexity (see Figure 1 for examples).
+In an initial validation study of a MaRs-IB short-form test, \cite{chierchia2019matrix} that measures of performance in a sample of 659 adults, adolescents, and children had excellent split-half and test-retest reliability.
+Moreover, performance on the MaRs-IB items were found to correlate with performance on the digit span tasks and ICAR matrix reasoning, indicating good convergent validity.
+Most importantly, the authors made available the percent correct data for each item (and variant) so that others researchers would be able to construct new MaRs-IB forms of custom difficulty and duration.
 
-There are, however, two impediments in researchers using the MaRs-IB norms.
-Both concern the design of the original MaRs-IB task.
-The initial MaRs-IB task was a fixed-time task, wherein participants completed as many items as possible in 8 minutes.
-The first issue then is that, because participants completed 33 items on average, there exists relatively little data on the majority of the MaRs-IB items (this is before taking into consideration the data available for each item variant).
-The second issue is that, because of individual differences in speed-accuracy trade-offs, the estimates of item difficulties may be biased.
-That is, items completed by fewer participants are more likely to have been reached by participants who had been sacrificing accuracy for speed, thereby giving a distorted view of the difficulty of these items.
+There are unfortunately two key obstacles preventing researchers from being fully able to do so, both of which stem from the design of the validation study.
+The original MaRs-IB short-form is a fixed-time and  fixed-order measure, where participants complete as many items as possible within 8-minutes.
+Because participants only completed 33 items on average, the first issue is that most items were completed by too few participants to establish reliable norms (an issue that is compounded when considering each item has six variants).
+Just over half the items (42/80) were completed by 100 or more participants.
+A second issue is that, because participants were free to choose for themselves whether to prioritize accuracy (number of correct responses) or productivity (number of items reached), it is likely that the percent correct data for items reached by fewer numbers of participants are biased.
+That is, items completed by fewer numbers of participants are more likely to have been completed by participants sacrificing accuracy for speed.
+Indeed, looking only at the easiest MaRs-IB items (dimension = 1), there is a strong positive correlation between number of completing participants and mean accuracy ($r = 0.673$).
 
-Summary paragraph: here we test a large sample of adults, making sure that all items and variants were tested many times.
-We fix the number of items to prevent bias.
-Importantly, we use item response models to calibrate the data.
-This means we can accurately measure item difficulty, taking into considering participant ability, as well as discirmination and guessing.
-This in turn can be used to construct new tests with known properties, by test information or test functioning.
-We use this to construct three new short forms, which we test.
-We make all data and parameters public.y
-
-ALL LINKS TO REPOS SHOULD BE HERE
+Here we wish to overcome the limitations of the previous study to further validate and make useful the MaRs-IB.
+First, we collected a novel dataset of response to MaRs-IB items in a large sample of adults (N=1501).
+Using a fixed-length design with counterbalancing, we ensured that all items, and their variants, were completed by a suitable number of participants.
+Next we analyzed participants' response data hierarchical item response models, which allowed us to measure all items' difficulty and discrimination while controlling for participants' varying abilities.
+The item response models also allowed us to measure directly the extent to which that item variants are truly equivalent.
+We then demonstrate how these item response parameters can be used to construct new short-forms with desirable properties.
+Using these methods, we introduce three new MaRs-IB short-form tests with good internal consistency and convergent validity.
+All data and code, including items' estimated parameters and example code for test assembly, is available at: \url{https://github.com/szorowi1/mars-irt}.
 
@@
 A total of N=1584 participants were recruited from the Prolific Academic platform (https://www.prolific.co) to participate in an online behavioral experiment in July - August, 2021.
-Participants were eligible if they currently resided in the United States.
+Participants were eligible if they were 18 years or older and currently resided in the United States.
 This study was approved by the Institutional Review Board of Princeton University (\#7392), and all participants provided informed consent.
@@
 Participants received monetary compensation for their time (rate USD \$10/hr), plus an incentive-compatible bonus up to \$0.50 based on task performance (average payment: mean = \$1.30 [\$0.10 sd]).
-The experiment
+Monetary incentives have previously been shown to motivate good performance in low-stakes testing environments \citep{duckworth2011role, gignac2018moderate}.
 
-cite https://doi.org/10.1111/bjop.12288, https://doi.org/10.1073/pnas.1018601108
-
-After providing consent, participants completed a series of items from the matrix reasoning item bank.
-The design of the MaRs-IB has been described elsewhere \citep{chierchia2019matrix} and was based on publicly available code (https://app.gorilla.sc/openmaterials/36164).
-Briefly, each MaRs-IB item consisted of a 3 × 3 matrix.
-Eight of the nine resulting cells contained an abstract shape, while one cell on the bottom right-hand side of the matrix was empty.
+After providing consent, participants completed a series of items from the MaRs-IB.
+The design of the MaRs-IB items have been described previously \citep{chierchia2019matrix}.
+Briefly, each MaRs-IB item consists of a 3 x 3 matrix.
+Eight of the nine resulting cells contain an abstract shape, while one cell on the bottom right-hand side of the matrix is empty.
 The participants' task was to complete the matrix by finding the missing shape among four possible alternatives.
@@
 Participants reviewed instructions before the task and were made to correctly complete three practice items before they could begin the task.
+Crucially, participants were instructed to respond carefully but to guess before the timer ran out.
+Our version of the MaRs-IB was based on publicly available code (\url{https://app.gorilla.sc/openmaterials/36164}).
 
-In order to ensure adequate sampling of all items, and in contrast to the original task design, all participants completed a pseudo-randomly selected subset of 16 out of 64 possible MaRs-IB items \footnote{We did not collect new data on 16 of the original MaRs-IB items.
-We excluded all Dimension 1 items, which are the simplest items and tend to have performance levels at ceiling.
-We also excluded Item 1 which, although not Dimension 1, also has performance levels near ceiling.} Item selection proceeded as follows: The 64 items were sorted into 16 groups by dimension.
-Participants were then randomly assigned one item from each of the groups.
-Importantly, participants were randomly assigned a clone of each item that varied in distractor type (md, pd) and item set (1, 2, 3).
-In this way, all items were sampled roughly 375 times (12 sd) and each clone was sampled 62 times (7 sd).
+In order to ensure adequate sampling of all items, and in contrast to the original MaRs-IB test form, all participants completed a pseudorandomly-selected set of 16 out of 64 possible MaRs-IB items \footnote{We did not collect new data for the 16 easiest items from the MaRs-IB (items 1, 2, 3, 4, 5, 7, 8, 9, 32, 33, 38, 41, 43, 48, 57, 68) as the original study or pilot testing found performance on these items to be at or near ceiling.}.
+Prior to data collection, we sorted the 64 items into 16 groups of four based on the number of changing relationships between the shapes of the matrix.
+Participants were then randomly assigned one item from each of the 16 groups.
+In this way, all participants received test forms of approximately equal difficulty.
+Importantly we also took steps to counterbalance the sampling of item variants, so that we had approximately the same number of responses available by item set (1, 2, or 3) and distractor type (minimal difference or or paired difference; see \cite{chierchia2019matrix} for details).
 
-The data from multiple participants who completed the experiment were excluded prior to analysis.
-The data from N=7 participants were excluded for having browser resolutions smaller than necessary to see the task.
-The data from N=8 participants were excluded for missing responses on 4 or more trials.
-The data from N=70 participants were excluded for rapid guessing (defined as response times of less than 3 s) on 4 or more trials.
-In total, 83 of 1584 (5.24\%) were excluded, leaving the data from N=1501 participants for analysis.
-
-The majority of participants in the calibration sample identified as women (men: N=670; women: N=811; non-binary or other: N=13; rather not say: N=2).
-Participants were 28.7 years old (9.9 sd) on average.
-The sample was relatively well-educated with the majority having a Bachelor's degree (N=507) or Master's degree or higher (N=322); fewer participants endorsed only completing some college (N=471) or only a high school degree (N=199).
-
-The experiment was programmed in jsPsych \citep{de2015jspsych} and distributed using custom web-application software.
-The experiment code is available at \url{https://github.com/szorowi1/mars-irt}, and the web-software is available at \url{https://github.com/nivlab/nivturk}.
-A playable demo of the task is available at \url{https://nivlab.github.io/jspsych-demos/tasks/mars/experiment.html}.
+To ensure data quality, the data from multiple participants who completed the experiment were excluded prior to analysis for one or more of the following reasons: for rapid guessing -- defined here as responding in less than 3 s -- on 4 or more trials N=70); for failing to respond on four or more trials (N=8); for minimizing or otherwise shrinking the browser window beyond what was required to view the stimuli (N=7).
+In total, 83 of 1584 (5.24\%) participants were excluded, leaving the data from N=1501 participants for analysis.
+Among these participants, each item (64 in total) was sampled approximately 375 times (sd = 12) and each item variant (384 in total) was sampled approximately 62 times (sd = 7).
 
@@
 
+The majority of participants in the sample identified as women (men: N=670; women: N=811; non-binary or other: N=13; rather not say: N=2) and were 28.7 years old (sd = 9.9) on average.
+The sample was relatively well-educated, with the majority having completed a Bachelor's degree (N=507) or Master's degree or higher (N=322).
+By comparison, fewer participants endorsed having completed only some college (N=471), only a high school degree (N=199), or preferring not to say (N=2).
+
 On average, participants completed 9.6 of 16 items correctly (sd = 3.1, IQR = 8.0 - 12.0) (Figure \ref{fig:fig01}a).
-Timing out occurred in only 1.8\% of items.
+Timing out occurred on only 1.8\% of trials.
 A total of 43 participants (2.9\%) performed below chance (i.e.
-fewer than four items correct), while only 11 participants (0.7\%) responded correctly to all items.
+fewer than four items correct), while only 11 participants (0.7\%) correctly responded to all 16 items.
 Thus, over 95\% of participants' observed scores were within a reasonable range.
-These results corroborate \cite{chierchia2019matrix} in that there were no obvious ceiling or floor effects, and that participants were motivated to perform despite minimal supervision in the online environment.
+These results corroborate \cite{chierchia2019matrix} in that there were no obvious ceiling or floor effects in performance, and that the majority of participants were sufficiently motivated to participate despite the low-stakes testing environment.
 
-Across all 384 items, the average proportion correct was 0.60 (sd = 0.19, IQR = 0.45 - 0.74) (Figure \ref{fig:fig01}b).
-A total of 12 items (3.1\%) had performance beneath chance levels (25\%), and only 22 items (5.7\%) had performance levels above 90\%.
-As such, the majority of items sampled exhibited adequate performance with no obvious ceiling or floor effects.
+Across all 384 unique item variants, the average proportion correct was 0.60 (sd = 0.19, IQR = 0.45 - 0.74) (Figure \ref{fig:fig01}b).
+A total of 12 items (3.1\%) had accuracy rates beneath chance levels (25\%), and only 22 items (5.7\%) had performance levels above 90\%.
+Mean performance on an item was significantly and negatively correlated with the number of changing relationships amongst its features (Spearman's $\rho = -0.482, p < 1e^{-6})$).
+Moreover, item-wise accuracy rates in this sample were remarkably similar to those reported by \cite{chierchia2019matrix} (Spearman's $\rho = 0.870, p < 1e^{-6})$), when restricting the latter to items with more than 200 available responses.
+These results further corroborate \cite{chierchia2019matrix} in that the majority of items exhibited good performance, with few items at ceiling or floor and passing of the manipulation check.
 
-Though not the primary focus of the analyses, we briefly investigated the relationship between participant' total scores, item difficulty, and response times.
-We fit a mixed-effects linear model to regress the log response time for trial $i$ on accuracy, the participant's observed score (minus their accuracy on that trial), the difficulty of the time (one minus the proportion correct), and the interaction of observed score and difficulty.
-We specified an intercept per participant to control for each participants' overall speed.
+Though not the primary focus of the current study, we briefly investigated the relationship between response accuracy and times.
+We used a mixed-effects model to predict participants' trial-level log-transformed response times as a function their response on that trial (correct or incorrect), their rest score (their total score on all other items), that item's difficulty (one minus the accuracy rate), and all possible interactions.
+We also specified an intercept per participant to control for each participants' overall speed.
 
-On average, participants spent 15.9 seconds to solve an item (sd = 7.2 s, IQR = 10.1 - 21.5 s).
-The mixed-effects model identified a significant effect of accuracy ($\beta = 0.067, t = 9.84, p < 1e^{-6}$), such that correct responses were on average slower than incorrect responses.
-Notably, there were large, significant effects of observed score ($\beta = 0.125, t = 16.948, p < 1e^{-6}$) and item difficulty ($\beta = 0.137, t = 44.326, p < 1e^{-6}$) such that participants were slower overall for more difficult items and for better performing participants.
-Finally, the interaction between observed score and item difficulty was also significant ($\beta = 0.058, t = 20.650, p < 1e^{-6}$), such that better performing participants slowed down even more for more difficult items.
+The complete summary of results from the regression analysis are presented in Table XX.
+On average, participants spent 15.9 s per item (sd = 7.2 s, IQR = 10.1 - 21.5 s).
+The mixed-effects model identified significant main effects of accuracy ($\beta = 0.067, t = 9.84, p < 1e^{-6}$), rest score ($\beta = 0.125, t = 16.948, p < 1e^{-6}$), and item difficulty ($\beta = 0.137, t = 44.326, p < 1e^{-6}$).
+On average then, response times were slower for correct responses, more accurate participants, and more difficult items.
+There was also a significant positive accuracy by difficulty interaction (correct responses were even slower for more difficult items) and a significant negative accuracy by score interaction (correct responses were faster for better performing participants).
+Finally, there was a significant positive difficulty by score interaction, such that better performing participants exhibited slower responses for more difficult items.
+The three-way interaction term was not statistically significant.
 
-The full pattern of response time results is summarized in Figure \ref{fig:fig01}c.
 In short, we corroborate the previous finding that item difficulty is associated with slower response times \citep{chierchia2019matrix}.
-In addition, we find two interesting patterns: better performing participants are slower on average, and this slowing is enhanced as items become more difficult.
-What this suggests then is that participants' overall performance on the MaRs-IB in this sample is reflective of speed-accuracy trade-offs likely reflective of individual differences in persistence or willingness to expend mental effort \citep{ranger2021effects}.
-That is, better performing participants are willing to spend more time on an item, whereas worse performing participants are more likely to give up sooner and guess as items become more difficult.
+In addition, we identified two notable patterns: better performing participants exhibited slower response times on average, which slowered further for more difficult times.
+This pattern of response time results is summarized in Figure \ref{fig:fig01}c.
+This pattern of results is consistent with controlled processing accounts, a point we return to in the Discussion.
 
@@
 
+Summary paragraph?
+
 \subsection{Item response models}
 
-Reflecting the inherent hierarchical structure of the MaRs-IB -- 80 item families with six clones each -- we estimated the psychometric properties of each item using item cloning models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
-All models were extensions of the three-parameter logistic (3PL) item repsonse model, where the probability of a correct response ($y_{ijk} = 1$) for person $i$ on item clone $k$ belonging to item family $j$ is:
+To measure the psychometric properties of each item, we used item response models.
+Given the hierarchical nature of the MaRs-IB -- 64 items with six variants each -- we specifically modeled participants' response data using item using a series of item cloning models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
+In item cloning models, the variants of an item (or "clones") are explicitly model as having been generated from the same template item (or "family").
+The basis of all of the models used here is the three-parameter logistic (3PL) item response model, where the probability of a correct response ($y_{ijk} = 1$) for person $i$ on item clone $k$ belonging to item family $j$ is:
 
@@
 \noindent where $\theta_i$ is the latent ability for person $i$, and $\beta_{jk}$, $\alpha_{jk}$, and $\gamma_{jk}$ are the difficulty, discrimination, and guessing parameters for item clone $k$ of item family $j$.
-Because guessing parameters are difficult to estimate without very large amounts of data \citep{han2012fixing}, we fixed the guessing rate for each item to the nominal rate ($\gamma_{jk} = 0.25$).
+Since guessing parameters are challenging to estimate in the absence of very large amounts of data \citep{han2012fixing}, we fixed the guessing parameter for each item clone to the nominal guessing rate ($\gamma_{jk} = 0.25$).
 
@@
 
-\noindent The difficulty parameter is thus modeled as an intercept ($\mu_\beta$), reflecting the average difficulty across all items, and four components:
+\noindent Every difficulty parameter is thus modeled as an intercept ($\mu_\beta$), reflecting the average difficulty across all items, and four components:
 
@@
 \item $\sum_{n=1}^N Q_{jn} \delta_{\beta n}$: the effect of the level-1 attributes, where $Q_{jn}$ is the value of attribute $n$ for item family $j$, and $\delta_{\beta n}$ is the effect of attribute $n$.
-This component is the fixed effects contribution of item family attributes to item difficulty.
+This component is the fixed effects contribution of item family to item difficulty.
 
-\item $\epsilon_{\beta j}$: the level-1 residual, or variance unexplained by the item family attributes. This component is the random effects contribution of item family attributes to item difficulty.
+\item $\epsilon_{\beta j}$: the level-1 residual, or variance unexplained by the item family attributes. This component is the random effects contribution of item family to item difficulty.
 
 \item $\sum_{m=1}^M R_{km} \delta_{\beta m}$: the effect of the level-2 attributes, where $R_{km}$ is the value of attribute $m$ for item clone $k$, and $\delta_{\beta m}$ is the effect of attribute $m$.
-This component is the fixed effects contribution of item clone attributes to item difficulty.
+This component is the fixed effects contribution of item clone to item difficulty.
 
-\item $\epsilon_{\beta k}$: the level-2 residual, or variance unexplained by the item clone attributes. This component is the random effects contribution of item clone attributes to item difficulty.
+\item $\epsilon_{\beta k}$: the level-2 residual, or variance unexplained by the item clone attributes. This component is the random effects contribution of item clone to item difficulty.
 
@@
 Here we model the effects of two properties: feature number, or the number of unique shapes in each cell of an item; and rule number, or the number of changing relationships among the shapes (e.g.
-color change, shape change, position change) one has to consider in order to correctly solve the item.
-Example items illustrating these attributes are included in the Supplement.
-Note that this definition of item rules deviates from other established taxonomies of matrix reasoning items \citep{carpenter1990one}, a point we return to in the Discussion.
+color change, shape change, position change) a participant must consider in order to solve the item.
 
 In turn, level-2 attributes describe properties of an item clone.
-Here we model the effects of two additional properties: distractor type, or whether the distractors were generated following the minimal difference (MD) strategy and a paired difference (PD) strategy (see \cite{chierchia2019matrix} for details); and mean response time, or the average time it took to complete that item.
-The latter attribute was orthogonalized with respect to the three other item attributes and thus reflects any residual structure after accounting for all other modeled effects.
+Here we model the effects of two additional properties: distractor type, or whether the distractors were generated following the minimal difference (MD) strategy and a paired difference (PD) strategy; and mean response time, or the average time spent on that item.
+Mean response time was orthogonalized with respect to the three other item attributes, and therefore reflects any residual structure after accounting for all other modeled effects.
 
@@
 
-\noindent The fixed effect components of the item discrimination parameters were estimated using the same level-1 and level-2 attributes as for the item difficulty parameters (i.e.
+\noindent The fixed effect components of the item discrimination parameters included the same level-1 and level-2 attributes as for the item difficulty parameters (i.e.
 number of features, number of rules, distractor type, average response time).
@@
 
-Though complex, the item structure model above provides a natural way of measuring the quality of the MaRs-IB.
-In the perfect situation, the clones of each item family would be psychometrically equivalent.
-So too should item families with identical attributes (e.g.
-number of features, number of rules).
-This is, of course, unlikely in actuality and the item cloning model allows us to quantify the extent to which equivalent families and clones vary in their difficulty and discrimination.
+Though complex, item cloning models provide a natural means of evaluating the success of the MaRs-IB item generation process.
+For example, modeling an item's properties as a function of its attributes allows us to quantify the extent to which item design worked as intended (e.g.
+that an increase in the number of changing relationships between shapes corresponds to an increase in item difficulty).
+The inclusion of level-1 random effects allows us to quantify the extent to which item families with identical attributes (e.g.
+number of features, number of changing relations) are psychometrically equivalent.
+Similarly, the inclusion of level-2 random effects allows us to quantify the extent to which clones of an item family are psychometrically equivalent.
 
-To quantify the success of the item cloning process, we estimated a series of four models \citep{cho2014additive, lathrop2017item}:
+In order to test the hypotheses outlined above, we fit a series of nested item cloning models \citep{cho2014additive, lathrop2017item}:
 
@@
 
-\item \textit{Model 1}: Item parameters are expressed only as a function of their level-1 attributes.
+\item \textit{Model a}: Item parameters are expressed only as a function of their level-1 attributes.
 This model treats all clones within a family as interchangeable and all item families with the same attributes as equivalent.
 
-\item \textit{Model 2}: Item parameters are expressed as a function of their level-1 attributes and residuals.
-This model treats all clones within a family as interchangeable, but now item families with the same attributes may psychometrically vary.
+\item \textit{Model b}: Item parameters are expressed as a function of their level-1 attributes and residuals.
+This model treats all clones within a family as interchangeable, but now item families with the same attributes may vary in their properties.
 
-\item \textit{Model 3}: Item parameters are expressed as a function of their level-1 attributes, level-2 attributes, and level-1 residuals.
+\item \textit{Model c}: Item parameters are expressed as a function of their level-1 attributes, level-2 attributes, and level-1 residuals.
 This model allows clones within a family to vary, but all variation must be explained solely through their attributes.
 
-\item \textit{Model 4}: Item parameters are expressed as a function of their level-1/2 attributes residuals. The full model in which item families and clones may vary in their psychometric properties.
+\item \textit{Model d}: Item parameters are expressed as a function of their level-1/2 attributes residuals. The full model in which item families and clones may vary in their psychometric properties.
 
@@
 
-The resulting models were then submitted to model comparison.
-This procedure allows us to measure the success of the item generation process.
-If the psychometric properties of item families reflect only their attributes, then Model 1 should be preferred to Model 2; in contrast, if item families vary in their properties above and beyond what their attributes can explain, then Model 2 should be preferred.
-Similarly, if the clones of item family are truly equivalent, then Models 1 or 2 should be preferred to Models 3 or 4; in contrast, if item clones vary substantially (systematically or otherwise), then Models 3 or 4 should be preferred to Models 1 or 2.
+The resulting model fits are then compared as a test of the hypotheses above.
+For example, if the psychometric properties of items can be solely explained by their attributes, then Model A should be preferred to B; in contrast, if item families psychometrically vary beyond what their attributes can explain, then Model B should be preferred.
+Similarly, if the clones of an item family are truly equivalent, then Models A or B should be preferred; in contrast, if item clones vary (as a function of their attributes or otherwise) then model C or D should be preferred.
 
-Finally, we estimated person-level abilities following an item response modeling approach \citep{wilson2008explanatory}:
+Finally, we estimated person-level abilities following an explanatory item response modeling approach \citep{wilson2008explanatory}:
 
@@
 
-\noindent where $X_{ip}$ is the value of attribute $p$ for person $i$, $\rho_p$ is the effect of attribute $n$, and $\epsilon_i$ is variance in ability unexplained by the person-level attributes.
+\noindent where $X_{ip}$ is the value of attribute $p$ for person $i$, $\rho_p$ is the partial correlation of attribute $n$, and $\epsilon_i$ is variance in ability unexplained by the person-level attributes.
 
 We modeled person-level abilities as a function of four attributes: age, gender (coded as: male = -0.5, female = 0.5; all else = 0), average response time, and $\Delta$ response time.
-The last term reflected a person's adjustment of their response time as a function of item difficulty (defined as its corresponding proportion of correct responses), and thus measured a participant's tendency to slow down (or speed up) in response to more challenging items.
+The last term reflected the degree to which a participant adjusted their response time as a function of item difficulty, and thus measured a participant's tendency to slow down (or speed up) in response to more challenging items (Figure \ref{fig:fig01}c).
 
-To ensure the identifiability of all models, person abilities were assumed to be $\theta_i \sim \mathcal{N} \left(0,1\right)$.
-
-\begin{equation}
-blah 
-\end{equation}
+To ensure the identifiability of all models, person abilities were constrained to have a mean of zero and a variance of 1.0, such that the residual variance was defined as $V(\epsilon) = 1 - \sum \rho_p^2$.
+Note that all person-level attributes were standard-scored prior being entered into the model.
 
@@
 The first 5000 samples from each chain were discarded.
-As a result, 10000 post-warmup samples from the joint posterior were retained.
-The $\hat{R}$ values for all parameters was less than 1.01, indicating acceptable convergence between chains, and there were no divergent transitions in any chain.
+As such, 10,000 post-warmup samples from the joint posterior were retained.
+The $\hat{R}$ values for all parameters was equal to or less than 1.01, indicating acceptable convergence between chains, and there were no divergent transitions in any chain.
 
-Models were compared using Bayesian leave-one out cross-validation \citep{vehtari2017practical}, which has been found to perform better than other information criteria for comparing item response models \citep{luo2017performances}.
-Importantly, we computed the marginalized leave-one-cluster out (LOCO) cross-validation \citep{merkle2019bayesian}, which measures models' ability to generalize to held out items (rather than held out responses).
+Models were compared using Bayesian leave-one out cross-validation \citep{vehtari2017practical}, which has been found to perform better than more traditional information criteria for comparing item response models \citep{luo2017performances}.
+Importantly, we computed the conditional leave-one-cluster out (LOCO) cross- validation \citep{merkle2019bayesian}, which measures a model's ability to generalize to held out items (rather than held out responses).
 
-The fit of the winning model to the data was evaluated through several approaches.
-First, the fit of the models to the data was assessed by means of posterior predictive model checking \citep{gelman1996posterior}.
-Item fit was assessed using two two discrepancy statistics: Orlando and Thissen index ($S-\chi^2$) \citep{toribio2011discrepancy} and the generalized dimensionality discrepancy measure (GDDM) \citep{levy2009posterior, levy2011generalized}.
-The $S-\chi^2$ index is a measure of item fit based on the comparison of observed and expected proportions of participants who answer an item correctly at each raw-score group.
-The GDDM is a measure to test the model's unidimensionality assumption, which must not be violated in order to compute valid total scores from tests assembled from MaRs-IB items.
+The fit of the best-fitting model to the data was evaluated through posterior predictive model checks \citep{gelman1996posterior, levy2017bayesian}.
+Item fit was assessed using the Orlando and Thissen index ($S-\chi^2$) \citep{toribio2011discrepancy}, which compares the observed and expected proportions of participants who answer an item correctly at each raw-score group.
+The unidimensionality of the data was assessed using the generalized dimensionality discrepancy measure (GDDM) \citep{levy2011generalized}, which tests for violation of the assumption of local independence.
+This latter measure is crucial for ensuring there is not non-negligible levels of local item dependence in the data, which may otherwise lower the efficiency of future test forms due to the inclusion of correlated, redundant items.
 
@@
 We also estimated models with no guessing parameter, i.e.
-1PL or Rasch models, but model comparison revealed these to be inferior.) The goodness of fit of the four models was compared based on leave-one-cluster out (LOCO) cross-validation as summarized in Table \ref{table:1}.
+1PL or Rasch models, but model comparison revealed these to be inferior.)
+
+The goodness of fit of the four models was compared based on leave-one-cluster out (LOCO) cross-validation as summarized in Table \ref{table:1}.
+The winning model was Model 1d, in which item difficulties were modeled as a function of all level-1 and level-2 fixed and random effects.
+This has multiple interpretations, which we will unpack in turn.
+First, this implies that there was systematic variability in item difficulty at the family-level beyond what could explained by their attributes (this is reflected...
+
+[above to be combined with below when time comes]
 
@@
 
-On average, items were difficult (95\% HDI), but ranged considerably in difficulty.
+The average item difficulty was $\beta_\mu = 0.177$ (95\% HDI = 0.118 - 0.238) but this ranged considerably across items (sd = 1.431, 95\% HDI = 1.370 - 1.492).
+
 Over half of the variability in item difficulty (65.6\%) was explained by the fixed effects of item family (level-1) and clone (level-2) attributes.
@@
 See \cite{lathrop2017item} for inspiration.
+Note that this definition of item rules deviates from other established taxonomies of matrix reasoning items \citep{carpenter1990one}, a point we return to in the Discussion.
 
-\bibliography{example}
+\bibliography{references}
 
@@
     \centering
-\begin{tabular}{l|c|c|c}
-Model & elpd$_{loco}$ & loco & $\Delta$ loco \\\hline
-1PG-F &  & -14516 & \\
-1PG-FX &  & -13804 & \\
-1PG-R &  & -13481 & \\
-1PG-RX &  & -13293 & - \\
+    \begin{tabular}{c|c|c|c|c|c|c|c}
+    \hline
+    Model & FE-1 & RE-1 & FE-2 & RE-2 & psis-loco & $\Delta$ psis-loco (se) \\
+    \hline
+    1a & x &   &   &   & -14516.85 & -1223.69 (44.94)\\
+    1b & x & x &   &   & -13804.74 & -511.58 (28.77) \\
+    1c & x & x & x &   & -13481.12 & -187.96 (17.25) \\
+    1d & x & x & x & x & -13293.16 & - \\
+    \hline
+    2a & x &   & x &   & -13209.38 & -2.13 (2.04) \\
+    2b & x & x & x &   & -13207.25 & - \\
+    2c & x & x & x & x & -13208.69 & 1.44 (0.85) \\
+    \hline
     \end{tabular}
-\caption{Table to test captions and labels.}
+    \caption{\label{tab:1} Goodness of fit of item response models to MaRs-IB data. Pareto-smoothed importance sampling}
     \label{table:1}
@@
 
+\begin{table}
+\begin{center}
+\begin{tabular}{lrrrrrr}
+\hline
+&  Coef. & Std.Err. & z & P$> |$z$|$ & [0.025 & 0.975]  \\
+\hline
+Intercept &  2.648 &    0.009 & 287.757 &       0.000 &  2.630 &  2.666  \\
+accuracy &  0.039 &    0.007 &   5.367 &       0.000 &  0.025 &  0.053  \\
+difficulty &  0.073 &    0.005 &  13.292 &       0.000 &  0.062 &  0.083  \\
+accuracy:difficulty &  0.097 &    0.007 &  14.351 &       0.000 &  0.084 &  0.110  \\
+score &  0.167 &    0.009 &  19.204 &       0.000 &  0.150 &  0.184  \\
+accuracy:score & -0.064 &    0.007 &  -8.764 &       0.000 & -0.078 & -0.050  \\
+difficulty:score &  0.032 &    0.005 &   6.549 &       0.000 &  0.023 &  0.042  \\
+accuracy:difficulty:score &  0.012 &    0.006 &   1.955 &       0.051 & -0.000 &  0.025  \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
 \end{document}

### 2022-01-18 manuscript/main.tex
more edits

@@
 Across all 384 unique item variants, the average proportion correct was 0.60 (sd = 0.19, IQR = 0.45 - 0.74) (Figure \ref{fig:fig01}b).
-A total of 12 items (3.1\%) had accuracy rates beneath chance levels (25\%), and only 22 items (5.7\%) had performance levels above 90\%.
+A total of 12 items (3.1\%) had accuracy rates beneath chance levels, and only 22 items (5.7\%) had performance levels above 90\%.
 Mean performance on an item was significantly and negatively correlated with the number of changing relationships amongst its features (Spearman's $\rho = -0.482, p < 1e^{-6})$).
@@
 
-Models were compared using Bayesian leave-one out cross-validation \citep{vehtari2017practical}, which has been found to perform better than more traditional information criteria for comparing item response models \citep{luo2017performances}.
+The goodness of fit of the models was compared were compared using Bayesian leave-one out cross-validation \citep{vehtari2017practical}, which has been found to perform better than more traditional information criteria for comparing item response models \citep{luo2017performances}.
 Importantly, we computed the conditional leave-one-cluster out (LOCO) cross- validation \citep{merkle2019bayesian}, which measures a model's ability to generalize to held out items (rather than held out responses).
@@
 
-First we estimated Models 1-4 holding constant the item discrimination parameters ($\alpha = 1$), which allowed us to investigate the difficulty of MaRs-IB items independently.
-(Note that these models still included guessing parameters fixed to the nominal guessing rate.
-We also estimated models with no guessing parameter, i.e.
-1PL or Rasch models, but model comparison revealed these to be inferior.)
+First we estimated the four models described above holding the item discrimination parameters fixed ($\alpha = 1$).
+Note that these models still included a fixed guessing parameter ($\gamma = 0.25$).
+We also estimated models with no guessing parameters (i.e.
+one-parameter logistic or Rasch models), but model comparison indicated these models were worse fits to the data (Table XX).
 
-The goodness of fit of the four models was compared based on leave-one-cluster out (LOCO) cross-validation as summarized in Table \ref{table:1}.
-The winning model was Model 1d, in which item difficulties were modeled as a function of all level-1 and level-2 fixed and random effects.
-This has multiple interpretations, which we will unpack in turn.
-First, this implies that there was systematic variability in item difficulty at the family-level beyond what could explained by their attributes (this is reflected...
+The goodness of fit of the four models was compared using Bayesian leave-one-cluster out cross-validation as summarized in Table \ref{table:1}.
+The full model (level-1/2 fixed \& random effects) fit the response data better than the model with only level-1 fixed effects ($\Delta$ LOCO = 1223.69, se = 44.94), the model with level-1 fixed and random effects ($\Delta$ LOCO 511.58, se = 28.77), and the model with all but level-2 random effects ($\Delta$ LOCO = 187.96, se = 17.25).
+Therefore, the two-level model with fixed and random item difficulty effects is chosen as the estimated model to discuss further.
 
-[above to be combined with below when time comes]
+Overall, the average item difficulty was $\mu_\beta = 0.177$ (95\% HDI = 0.118 - 0.238).
+Difficulty also ranged considerably across items (sd = 1.431, 95\% HDI = 1.370 - 1.492).
+We next unpack the sources of variability in item difficulty, beginning first with the level-1 attributes.
+As expected, a one-unit change in the number of item features was credibly associated with an increase in item difficulty ($\delta_1$ = 0.132, 95\% HDI = 0.087 - 0.174).
+Similarly, a one-unit change in the number of changing relations was also associated with an increase in item difficulty ($\delta_2$ = 0.112, 95\% HDI = 0.080 - 0.144).
+Together, these two attributes explained approximately 65.5\% of the variance in difficulty across item families (35.0\% across all items).
+This result supports the success of the item generation process, insofar that item difficulty is associated with and well-explained by the core item attributes.
+It also highlights, however, that item families with identical attributes are not exchangeable as confirmed by the better fit of Model 1b compared to Model 1a ($\Delta$ LOCO = 712.11, se = 34.61).
 
-First, Model 2 was preferred to Model 1 suggesting that there is variability in the difficulty of item families above and beyond what can be explained by their attributes.
-Similarly, Model 3 was preferred to Model 2 suggesting that there is systematic variability in the clones of each item family.
-Finally, Model 4 was preferred to Model 3 suggesting that there is variability in the difficulty of item clones above and beyond what can be explained by their respective attributes.
-In summary, the full model -- including all fixed and random effect terms -- was preferred to the three simpler models, indicating that item families and clones are nonequivalent even after controlling for their attributes.
+Turning next to the level-2 attributes, there was an unexpected association between distractor type and item difficulty such that otherwise identical items were more difficult with MD than PD distractors ($\delta_3$ = 0.252 , 95\% HDI = 0.218 - 0.287).
+There was also an association between average response time and item difficulty ($\delta_4$ = 0.122, 95\% HDI = 0.093 - 0.151).
+In total, the controllable fixed effects (number of features, number of changing relations, distractor type), explained nearly half (48.2\%) of the variance in item difficulty.
+Compared to a model with only level-1 fixed and random effects, the improved fits of Models 1c ($\Delta$ = 323.62, se = 23.36) and 1d ($\Delta$ LOCO = 187.96, se = 17.25) indicates that item clones are not equivalent in their difficulty.
+Taking into account the controllable fixed effects and level-1 random effects, the total explained variance in item difficulty is approximately 72.0\%.
+To summarize, although item variants are not strictly exchangeable, only the minority of variance in item difficulty can be attributed to unmodeled effects (e.g.
+perceptual features).
 
-The average item difficulty was $\beta_\mu = 0.177$ (95\% HDI = 0.118 - 0.238) but this ranged considerably across items (sd = 1.431, 95\% HDI = 1.370 - 1.492).
+Next we estimated the item cloning models again, but no longer holding the item discrimination parameters constant.
+In all models, the item difficulty parameters were estimated with fixed and random effects at both levels.
+In contrast, for Model 2a the item discrimination parameters were modeled as a function of level-1 and level-2 fixed effects only.
+In Model 2b, the item discrimination parameters were modeled as a function of the fixed effects and level-1 random effects.
+Finally, in Model 2c, the item discrimination parameters were modeled with fixed and random effects at both levels.
 
-Over half of the variability in item difficulty (65.6\%) was explained by the fixed effects of item family (level-1) and clone (level-2) attributes.
-The associations between level-1 attributes and item difficulty were as expected: feature number was positively associated with difficulty (beta), such that the inclusion of one additional feature was associated with a 13.1\% reduction in expected accuracy.
-Similarly, rule number was positively associated with item difficulty (beta), such that the inclusion of one additional rule a 10.1\% reduction in expected accuracy.
-Turning next to level-2 attributes, distractor type was unexpectedly associated with item difficulty (beta), such that clones with MD distractors clones were associated with a 13.6\% reduction in accuracy compared to clones with PD distractors.
-This pattern is clearly evident in the raw item-level scores (Figure XX).
-Finally, average response time was positively associated with difficulty, such that a one standard deviation change in response time was associated with an 11.8\% reduction in accuracy.
+The goodness of fit of the three models is summarized in Table \ref{table:1}.
+Notably, the best-fitting model was not the full model but instead the model with only level-1 random effects (Model 2b).
+This was preferred to the model with only fixed effects ($\Delta$ LOCO = 2.13, se = 2.04) and the model with both level-1 and level-2 random effects ($\Delta$ LOCO = -1.44, se = 0.85).
+We note that the authors of the LOO metric recommend that standard errors be at least twice as large as the difference between models to be confident that one model truly provides a better fit.
+\footnote{\url{https://discourse.mc-stan.org/t/loo-comparison-in-reference-to-standard-error/4009}} As such, we proceed by interpreting Model 2b as the overall best-fitting model with caution.
+Further note that this model still fit substantially better than the winning model with fixed item discrimination parameters ($\Delta$ LOCO = 85.91, se = 6.33).
 
-The inclusion of item family (level-1) random effects explained an additional 18.5\% of the variance in item difficulty.
-In total, the percentage of variance in item difficulty not explained by residual variance from item clones (level-2) was 84.1\%.
-The residual variance was harder to explain (Figure XXb); indeed, there was no clear patterns in difficulty by item set, suggesting that at least part of residual variance may be explained by idiosynchratic perceptual features of item clones.
+Overall, the average item discrimination was $\mu_\alpha = 1.298$ (95\% HDI = 1.212 - 1.385), and exhibited modest variance across items (sd = 0.221, 95\% HDI = 0.121 - 0.322).
+Among the fixed-effects, there was only a credible association between the number of changing relations and item discrimination ($\delta_{\alpha_2}$ = 0.067, 95\% HDI = 0.011 - 0.123).
+There was no such association for the number of item features ($\delta_{\alpha_1}$ =-0.037 95\% HDI = -0.090 - 0.017), distractor type ($\delta_{\alpha_3}$ -0.036, 95\% HDI = -0.082, 0.006), or average response time ($\delta_{\alpha_4}$ = -0.025, 95\% HDI = -0.075 - 0.024).
+Regardless, the controllable fixed effects (number of features, number of changing relations, distractor type) explained approximately (39.2\%) of the variance in item discrimination.
+In summary then, there is evidence that items vary in their discrimination but not strong evidence to suggest that item variants vary substantially from one another.
+We return to this point in the discussion.
 
-Next we estimated three models as above, but with two key differences: First, we no longer held constant the item discrimination parameters but rather estimated them alongside the item difficulty parameters.
-Second, whereas the item difficulty parameters were always estimated using the full additive multilevel item structure model (i.e.
-level-1/2 fixed and random effects), the item discrimination parameters were estimated adding one additional component per model (Model 1: level-1/2 fixed effects only; Model 2: fixed effects and level-1 random effects; Model 3: full model).
-The goodness of fit of the three models was again compared based on leave-one-cluster out (LOCO) cross-validation as summarized in Table \ref{table:2}.
+Briefly, we turn to the person-level predictors of performance.
+Corroborating the mixed-effects regression model, average response time was positively associated with ability ($\rho_1$ = 0.427, 95\% HDI = 0.374 - 0.478), as was $\Delta$ response time ($\rho_2$ = 0.368, 95\% HDI = 0.313 - 0.424).
+There was a negative relationship between age and ability ($\rho_3$ = -0.299, 95\% HDI = -0.353 -- -0.246).
+Finally, there was a small and unexpected association between gender and ability ($\rho_4$ = 0.128, 95\% HDI = 0.078 - 0.182), such that women performed slightly better than men on average.
+Though we used the colorblind-friendly MaRs-IB stimuli for this study, it is possible that residual effects may persist.
 
-First, all three models including discrimination parameters were preferred to the best-fitting model with discrimination parameters held constant.
-Model 2 was preferred to Model 1 suggesting that there is variability in the discrimination of item families not accounted for by their attributes.
-Importantly, Model 2 was preferred to Model 3 such that there was insufficient variability in discrimination in item clones to necessitate additional parameters.
-We return to this point in the Discussion.
-
-Returning to variance partitioning, just less than half of the variance in item discrimination (44.9\%) was explained by attributes of the explained by attributes of item families and clones.
-Interestingly, only the number of item rules was positively associated with item discrimination (HDI); no other attributes were credibly related (HDIs here).
-
-Finally, we turn to explanatory item response modeling.
-The largest effects were from the response times.
-Average response time blah.
-Delta response time blah.
-Interestingly, we found a negative correlation between age and ability, such that older participants tended to perform worse on average.
-Finally, we observed a small but credible association between gender and ability.
-Though we used the colorblind stimulus set, it is possible there was residual perceptual difficulty for men compared to women.
-Granted, this effect is small.
-
-This paragraph should be on fit.
-Should mention: simulation study, posterior predictive check, item fit indices, and unidimensionality measure.
-You should probably review a couple of other RPM validation papers to make sure our methods are straight (e.g.
-doi.org/10.1037/a0027830)
+Finally, we validated the fit of the winning model using several posterior predictive model checking measures.
+Across the 384 items, the average posterior predictive p-value for the $S-\chi^2$ index of item fit was $p = 0.510$ (sd = 0.242).
+Only 14 items (3.64\%) indicated misfit with posterior predictive p-values less than 0.05 or greater than 0.95.
+Moreover, the posterior predictive p-value for the test of undimensionality was $p = 0.803$.
+Though this suggests some evidence for local dependence, it is within a tolerable range.
+Overall then, the best-fitting model provides an adequate fit to the data.
 
@@
 
-In discussing the success of the MaRs-IB generation process, some things to point out:
+In this first study, we recruited a large sample of adults to complete items from the MaRs-IB, making sure that each item (and its variants) were completed by a sizable number of participants.
+Using hierarchical item response models, we then measured the psychometric properties (i.e.
+difficulty and discrimination) of each item.
+Corroborating the initial study from \cite{chierchia2019matrix}, we found that the items of the MaRs-IB vary greatly in their difficulty -- a prerequisite for measuring nonverbal reasoning across the ability spectrum.
+We also found that the items of the MaRs-IB exhibited medium-to-large levels of discrimination \citep{baker2017basics}, similar to what has been reported in previous item response theory analyses of the Raven's progressive matrices \citep{chiesi2012using}.
 
-1. A lot of variance in item difficulty and discrimination can be explained by inherent features of the items. This is great!
+Moreover, using item cloning models we were able to test several important properties of the MaRs-IB items.
+First, we confirmed that item difficulty is a function of and to a moderate extent explained by its attributes (namely, the number of features and the number of changing relations).
+Relatedly, we identified that an item's discrimination is a function and partially explained by its number of changing relations, which is similar to what has been reported in other investigations of matrix reasoning tasks \citep{embretson1999generating}.
+Second, we found that item variants are not equivalent and exchangeable with respect to difficulty.
+Unexpectedly, and contrary to \cite{chierchia2019matrix}, distractor type emerged as a strong predictor of item difficulty.
+All else being equal, item variants using MD distractors were harder than the same items using PD distractors.
+Importantly, once taking this (and the residual variance in difficulty at the item family level) into account, a large majority of item difficulty (72.0\%) was accounted for, suggesting that a minority of variance in difficulty across item variants stems from unmodeled features such as their superficial perceptual properties.
 
-2. Though there is quite a bit of residual variance at the item family level, this is fine as we only modeled two features. A point we return to in the Discussion.
-
-3. At the clone level, it seems clear that item clones are not equivalent w.r.t. difficulty. For one thing, MD vs. PD big. Should describe level-2 standard deviation somewhere.
-
-4.
-For clone discrimination, it seems less clear.
-One possibility is that there isn't much variability (cite Embertson).
-Another possibility is that we are underpowered.
-Cite Lathrop here for discussion of potential bias.
-
-Summarize as good to go!
+summary sentence here saying that the MaRs-IB has many desirable properties. We next move onto test construction.
 
@@
 
-\begin{table}
-\begin{center}
-\begin{tabular}{lrrrrrr}
-\hline
-&  Coef. & Std.Err. & z & P$> |$z$|$ & [0.025 & 0.975]  \\
-\hline
-Intercept &  2.648 &    0.009 & 287.757 &       0.000 &  2.630 &  2.666  \\
-accuracy &  0.039 &    0.007 &   5.367 &       0.000 &  0.025 &  0.053  \\
-difficulty &  0.073 &    0.005 &  13.292 &       0.000 &  0.062 &  0.083  \\
-accuracy:difficulty &  0.097 &    0.007 &  14.351 &       0.000 &  0.084 &  0.110  \\
-score &  0.167 &    0.009 &  19.204 &       0.000 &  0.150 &  0.184  \\
-accuracy:score & -0.064 &    0.007 &  -8.764 &       0.000 & -0.078 & -0.050  \\
-difficulty:score &  0.032 &    0.005 &   6.549 &       0.000 &  0.023 &  0.042  \\
-accuracy:difficulty:score &  0.012 &    0.006 &   1.955 &       0.051 & -0.000 &  0.025  \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
 \end{document}

### 2022-01-18 manuscript/main.tex
test assembly updates

@@
 
-With 64 items, each with six clones, there is a vast universe of possible test forms we could construct.
-Beginning with our motivating problem -- abbreviated measures of nonverbal reasoning with the potential for retest -- we opted to construct three parallel short form measures.
+With 64 item templates, with 6 variants each, there is a vast number of possible test forms we could construct using the MaRs-IB.
+Returning to our motivating problem -- the availability of abbreviated matrix reasoning measures with the potential for retest -- we opted to construct three parallel short form measures.
 To do so, we used mixed integer programming \citep{der2005wj} to maximize the test information function (TIF) of each short form subject to the following constraints:
@@
     \item Each short form was required to contain 12 items.
-This was chosen to minimize administration time of a form (2-4 minutes on average) and to achieve sufficient  reliability (Cronbach's $\alpha \geq 0.8$ in piloting).
+This was chosen to minimize the administration time of a given form (2-4 minutes on average) and to achieve good score reliability (Cronbach's $\alpha \geq 0.8$).
 
-    \item All short forms were required to be comprised of clones from the same item families. This was chosen to maximize the similarity of the short forms.
+    \item All short forms were required to be comprised of variants from the same item families. This was chosen to maximize the similarity of the short forms.
 
-    \item Assigned clones could be of the MD or PD distractor type, but not both.
-This was chosen to prevent item overlap, as sometimes the MD and PD distractor strategies generate identical distractors.
+    \item Variants could be of the MD or PD distractor type, but not both. This chosen to prevent item redundancy (i.e. including the same puzzle twice in one short form).
 
-    \item The difference in TIFs across short forms was required to be minimized. This was chosen to ensure that each short form had similar psychometric properties.
+    \item The difference in TIFs between short forms was required to be minimized.
+This was chosen to ensure that each short form had similar psychometric properties (i.e.
+equivalent reliability across ability levels).
 
@@
 
-The test information functions (TIFs) and test characteristic curves (TCCs) for the three short forms are shown in Figure XX.
-As can be readily observed, the TIFs for each short form are remarkably similar reflecting the constraints of the MIP problem.
-Thus, the short forms should have similar levels of precision across ability levels.
-Similarly, the TCCs for each short form are similar, which was by no means guaranteed \citep{ali2016evaluation}.
-Thus, the short forms are expected to have similar total score distributions.
+The results of the test assembly are presented in Figure \ref{fig:fig02}.
+As expected, the MIP solver selected for items that were more discriminating on average (Figure \ref{fig:fig02}a).Though this was not explicit constraint, the test characteristic curves (TCCs) of the three short forms are markedly similar, each with an expected total score of 7.7 out of 12 items.
+Note that, because low ability participants are expected to guess an item correctly 25\% of time, the lower asymptote of the expected score distribution is 3 and the total score distribution is left-skewed.
+Finally, the TIFs across the three short forms are remarkably similar \ref{fig:fig02}b) as would be expected given the assembly constraints.
+As such, the short forms are predicted to have similar levels of precision across ability levels.
+Indeed, the expected marginal score reliability for each test form are similar (short form 1: $\alpha = 0.794$; short form 2: $\alpha = 0.795$; short form 3: $\alpha = 0.792$).
 
-Thus we proceeded to test each short form. blah blah blah.
+\begin{figure}
+\centering
+\includegraphics[width=1.0\textwidth]{figures/fig02.png}
+\caption{\label{fig:fig02}This is a figure caption.}
+\end{figure}
+
+Next, we proceeded to test each short form in a new sample of participants.
+The goals for this study were twofold.
+First, we sought to confirm that each short form would have the properties predicted above.
+Second, we sought to establish the convergent validity of the short forms by correlating total scores with performance on an abbreviated version of the Raven's progressive matrices.
 

### 2022-01-19 04_Validation/scripts/fit_3pl.py
checkpoint progress

@@ -1,0 +1,103 @@
+import os, sys
+import numpy as np
+from os.path import dirname
+from pandas import read_csv
+from cmdstanpy import CmdStanModel
+ROOT_DIR = dirname(dirname(os.path.realpath(__file__)))
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Define parameters.
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## I/O parameters.
+stan_model = '3pl'
+
+## Sampling parameters.
+iter_warmup   = 5000
+iter_sampling = 2500
+chains = 4
+thin = 1
+parallel_chains = 4
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Load and prepare data.
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## Load data.
+mars = read_csv(os.path.join('data', 'mars.csv'))
+rpm  = read_csv(os.path.join('data', 'rpm.csv'))
+metadata = read_csv(os.path.join('data', 'metadata.csv'))
+
+## Apply rejections.
+reject = read_csv(os.path.join(ROOT_DIR, 'data', 'reject.csv'))
+mars = mars[mars.subject.isin(reject.query('reject == 0').subject)].reset_index(drop=True)
+rpm  = rpm[rpm.subject.isin(reject.query('reject == 0').subject)].reset_index(drop=True)
+metadata  = metadata[metadata.subject.isin(reject.query('reject == 0').subject)].reset_index(drop=True)
+
+## Score missing data.
+mars['accuracy'] = mars['accuracy'].fillna(0)
+rpm['accuracy'] = rpm['accuracy'].fillna(0)
+
+## Format metadata.
+zscore = lambda x: (x - np.nanmean(x)) / np.nanstd(x)
+gender = {'Male':-0.5, 'Female':0.5, 'Other': 0, 'Rather not say': 0}
+metadata['age'] = zscore(metadata.age)
+metadata['gender'] = zscore(metadata.gender.replace(gender))
+metadata['rpm'] = zscore(rpm.groupby('subject').accuracy.sum().values)
+
+## Load item parameters.
[…]

### 2022-01-19 04_Validation/scripts/fit_summability.py

@@ -1,74 +1,0 @@
-import os, sys
-import numpy as np
-from os.path import dirname
-from pandas import read_csv
-from cmdstanpy import CmdStanModel
-ROOT_DIR = dirname(dirname(os.path.realpath(__file__)))
-
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-### Define parameters.
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-
-## I/O parameters.
-fdat = sys.argv[1]
-stan_model = 'tetrachoric'
-
-## Sampling parameters.
-iter_warmup   = 5000
-iter_sampling = 2500
-chains = 4
-thin = 1
-parallel_chains = 4
-
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-### Load and prepare data.
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-
-## Load data.
-data = read_csv(os.path.join(ROOT_DIR, 'data', f'{fdat}.csv'))
-
-## Apply rejections.
-reject = read_csv(os.path.join(ROOT_DIR, 'data', 'reject.csv'))
-data = data.loc[data.subject.isin(reject.query('reject==0').subject)]
-
-## Score missing data.
-data['accuracy'] = data['accuracy'].fillna(0)
-
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-### Assemble data for Stan.
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-
-## Define response data.
-Y = data.pivot_table('accuracy','subject','trial').values
-
-## Define metadata.
-J, K = Y.shape
-
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-### Fit Stan Model.
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-
-## Assemble data.
-dd = dict(J=J, K=K, Y=Y)
-
-## Load StanModel
-StanModel = CmdStanModel(stan_file=os.path.join('stan_models',f'{stan_model}.stan'))
-
-## Fit Stan model.
[…]

### 2022-01-19 manuscript/main.tex
push edits

@@
 
-To ensure data quality, the data from multiple participants who completed the experiment were excluded prior to analysis for one or more of the following reasons: for rapid guessing -- defined here as responding in less than 3 s -- on 4 or more trials N=70); for failing to respond on four or more trials (N=8); for minimizing or otherwise shrinking the browser window beyond what was required to view the stimuli (N=7).
+To ensure data quality, the data from multiple participants who completed the experiment were excluded prior to analysis for one or more of the following reasons: for rapid guessing -- defined here as responding in less than 3 s -- on 4 or more trials (N=70); for failing to respond on four or more trials (N=8); for minimizing or otherwise shrinking the browser window beyond what was required to view the stimuli (N=7).
 In total, 83 of 1584 (5.24\%) participants were excluded, leaving the data from N=1501 participants for analysis.
@@
 In the first, participants completed three short surveys: the 10-item need for cognition survey \citep{chiesi2018applying}, the 8-item PROMIS Cognitive Function-Short Forms (8a) \citep{iverson2021normative}, and the 8-item subject numeracy scale \citep{fagerlin2007measuring}.
-These measures were included as part of exploratory analyses that were not the central focus of this study and will not be reported in the main text (reported in Supplement).
+These measures were included as part of planned exploratory analyses and will not be reported in the main text (all correlations between these self-report measures and performance on the matrix reasoning tasks are reported in Supplement).
 Next, participants completed one of the three 12-item MaRs-IRB short-forms described above.
-Besides for the items, the administration procedure of the short-forms was identical to that in the Calibration study.
+The administration procedure of the short-forms was identical to that in the first study.
 Finally, participants completed the 9-item abbreviated Raven's progressive matrices (form A) \citep{bilker2012development}.
-For consistency's sake, the administration procedure was matched to the MaRs-IB.
+To minimize bias, the administration procedure for the progressive matrices was matched to that for the MaRs-IB items.
 
-The data from a total of N=47 participants were excluded prior to analysis.
-Of these, N=17 participants were excluded for having not completed the entire experiment.
-Separately, the data from N=21 participants were excluded for failing one or more attention checks (infrequency items) embedded in the self-report measures \citep{zorowitz2021inattentive}.
-The data from N=13 participants were excluded for image loading problems on either the MaRs-IB or RPM tasks.
-Finally, N=6 participants were excluded for missing or rapid guessing on 4 or more trials during the MaRs-IB or RPM tasks.
-Note these are not mutually exclusive.
+To ensure data quality, the data from multiple participants were excluded prior to analysis for one or more of the following reasons: for failing to complete all three sections of the experiment (N=17); for failing one or more attention checks \citep{zorowitz2021inattentive} in the self-report measures (N=21); for experiencing technical difficulties during the loading of the stimuli (N=11); for rapid guessing on 4 or more trials (N=3); or for failing to respond on four or more trials (N=1).
+In total, 47 of 347 (13.54\%) participants were excluded, leaving the data from N=300 participants for analysis.
 
-The majority of participants in the validation sample identified as women (men: N=138; women: N=147; non-binary or other: N=13; rather not say: N=2).
+\subsection{Descriptive statistics}
+
+The majority of participants identified as women (men: N=138; women: N=147; non-binary or other: N=13; rather not say: N=2).
 Participants were 35.0 years old (13.2 sd) on average.
-The sample was relatively well-educated with the majority having a Bachelor's degree (N=121) or Master's degree or higher (N=45); fewer participants endorsed only completing some college (N=100) or only a high school degree or lower (N=33).
-
-\subsection{Results}
+The sample was relatively well-educated, with the majority having completed a Bachelor's degree (N=121) or Master's degree or higher (N=45).
+By comparison, fewer participants endorsed having completed only some college (N=100), only a high school degree (N=33), or preferring not to say (N=1).
 
 The distribution of scores on the three MaRs-IB short-forms are summarized in Table \ref{table:2}.
-Across the short-forms, participants responded correctly on 8.0 of 12 items (sd = 2.5, IQR = 6 - 10).
-The distribution of scores were skewed towards to higher end, as would be expected given their respective test characteristic curves.
-A 1-way ANOVA did not reject the null hypothesis that the observed scores across test forms were equal (F(2,97) = 0.253, p = 0.777).
-Corroborating the predictions from the test assembly, the reliability of each short form was satisfactory (Table \ref{table:2}).
+Consistent with model predictions, participants on average responded correctly on 8.0 of 12 items (sd = 2.5, IQR = 6 - 10).
+Similarly, the distribution of total scores were right-shifted \ref{fig:fig03} as would be expected given the nominal guessing rate of 25\%.
+A one-way ANOVA comparing the total scores across the three short-forms did not reject the null hypothesis (F(2,297) = 0.253, p = 0.777).
+Finally, the marginal score reliability for each short form was satisfactory as predicted (Table \ref{table:2}).
+In summary, the distribution of scores across the MaRs-IB short forms were consistent with the predictions based on the test assembly, indicating that the psychometric properties of these items were well calibrated in the previous study.
 
-The distribution of scores on the RPM short-form is similarly summarized in Table \ref{table:2}.
+The distribution of scores on the abbreviated Raven's progressive matrices is also summarized in Table \ref{table:2}.
 On average, participants responded correctly on 4.5 of 9 items (sd = 2.0, IQR = 3 - 9).
-Scores on the RPM were significantly lower on the RPM than the MaRs-IB ($t = 25.598, p < 1e^{-6}$), indicating that the items on the RPM short-form are more difficult on average than those selected for the MaRs-IB short-forms.
-The reliability of the measure was similarly good as the MaRs-IB (Table \ref{table:2}).
+Scores on the RPM were significantly lower on the RPM than the MaRs-IB ($t = 25.598, p < 1e^{-6}$), indicating either that the RPM items are more difficult or less susceptible to guessing (consistent with these items having 6 or 8 distractors).
+The marginal score reliability for the RPM was similar to that of the MaRs-IB forms (Table \ref{table:2}).
+
+\begin{figure}
+\centering
+\includegraphics[width=1.0\textwidth]{figures/fig03.png}
+\caption{\label{fig:fig03}This is a figure caption.}
+\end{figure}
+
+\subsection{Convergent validity}
 
@@
     \hline
-    MaRs-IB SFa & 7.9 (2.4) & 7 - 10 & 0.761 \\
-    MaRs-IB SFb & 8.2 (2.6) & 7 - 10 & 0.810 \\
-    MaRs-IB SFc & 7.9 (2.7) & 6 - 10 & 0.829 \\
-    RPM SFa & 4.5 (2.0) & 3 - 9 & 0.791 \\
+    MaRs-IB SF-1 & 7.9 (2.4) & 7 - 10 & 0.761 \\
+    MaRs-IB SF-2 & 8.2 (2.6) & 7 - 10 & 0.810 \\
+    MaRs-IB SF-3 & 7.9 (2.7) & 6 - 10 & 0.829 \\
+    RPM SF-A & 4.5 (2.0) & 3 - 9 & 0.791 \\
     \end{tabular}

### 2022-01-21 manuscript/main.tex
updates to manuscript

@@
 
-\begin{equation}
+\begin{equation} \label{eq:1}
 p(y_{ijk} = 1) = \gamma_{jk} + (1-\gamma_{jk}) \cdot \text{logit}^{-1} \left( \alpha_{jk} \cdot \theta_i - \beta_{jk} \right)
@@
 
-\begin{equation}
+\begin{equation} \label{eq:4}
 \theta_i = \sum_{p=1}^P X_{ip} \rho_p + \epsilon_i    
@@
 
-\section{Test assembly}
+\section{Validation Study}
+
+\subsection{Test assembly}
 
@@
 
-\section{Validation Study}
-
 \subsection{Data Collection}
@@
 
+\begin{table}
+    \centering
+    \begin{tabular}{c|c|c|c}
+    Measure & Mean (sd) & IQR & Reliability \\
+    \hline
+    MaRs-IB SF-1 & 7.9 (2.4) & 7 - 10 & 0.761 \\
+    MaRs-IB SF-2 & 8.2 (2.6) & 7 - 10 & 0.810 \\
+    MaRs-IB SF-3 & 7.9 (2.7) & 6 - 10 & 0.829 \\
+    RPM SF-A & 4.5 (2.0) & 3 - 9 & 0.791 \\
+    \end{tabular}
+    \caption{Caption}
+    \label{table:2}
+\end{table}
+
 \begin{figure}
@@
 
-\subsection{Convergent validity}
+To measure the convergent validity between the MaRs-IB and RPM measures, we again utilized an explanatory item response model framework.
+Specifically we modeled participants' ability on the MaRs-IB short forms as in Eq \ref{eq:4}, this time entering participants gender, age, and total score on the RPM as covariates.
+Participants' individual responses were again modeled using the 3PL model (eq \ref{eq:1}), where the difficulty and discrimination parameters for each item were fixed to the posterior average as estimated in the item cloning models.
 
-To measure convergent validity, blah blah blah. Need to write model here. Or should we write above?
+The partial correlation between total scores on the RPM and ability on the MaRs-IB was $\rho = 0.527$ (95\% HDI = 0.457 - 0.577).
+Given the reliability of both measures, we should not expect correlations much larger than $\rho = 0.624$ (or $0.79^2$).
+As such, this result indicates good convergent validity between the two matrix reasoning measures.
+Finally, in this smaller sample, neither age nor gender was credibly associated with performance on the MaRs-IB short forms (age: $\rho$ = -0.123, 95\% HDI = -0.253 - 0.013; gender: $\rho$ = -0.025, 95\% HDI = -0.160 - 0.108).
+
+Would a paragraph on MaRs-IB administration time be helpful? To briefly say that participants spent, on average, 3 minutes on the MaRs-IB items making it a short and valid measure.
 
 \subsection{Discussion}
+
+Do we need a short discussion here or is it sufficient to simply move to general discussion?
 
@@
 
-\begin{table}
-    \centering
-    \begin{tabular}{c|c|c|c}
-    Measure & Mean (sd) & IQR & Reliability \\
-    \hline
-    MaRs-IB SF-1 & 7.9 (2.4) & 7 - 10 & 0.761 \\
-    MaRs-IB SF-2 & 8.2 (2.6) & 7 - 10 & 0.810 \\
-    MaRs-IB SF-3 & 7.9 (2.7) & 6 - 10 & 0.829 \\
-    RPM SF-A & 4.5 (2.0) & 3 - 9 & 0.791 \\
-    \end{tabular}
-    \caption{Caption}
-    \label{table:2}
-\end{table}
-
 \end{document}

### 2022-01-23 04_Validation/scripts/reliability.R
update reliability scripts, add correlation data

@@ -1,0 +1,67 @@
+library('tidyr')
+library('lavaan')
+library('semTools')
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Reliability (MaRs-IB)
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## Load data.
+mars = read.csv('../data/mars.csv')
+
+## Apply rejections.
+reject = read.csv('../data/reject.csv')
+mars = mars[is.element(mars$subject, reject[reject$reject==0,'subject']),];
+
+## Score missing data.
+mars$accuracy = ifelse(is.na(mars$accuracy), 0, mars$accuracy);
+
+## Define formula.
+mod1f <- 'score =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9 + x10 + x11 + x12'
+
+for (i in 1:3) {
+  
+  ## Restrict data to current short form. Pivot DataFrame.
+  sf = pivot_wider(mars[mars$short_form==i,], id_cols=subject, names_from=trial, 
+                   names_prefix = 'x', values_from=accuracy, values_fill=0);
+  
+  ## Fit model.
+  fit1f <- cfa(mod1f, data=sf, std.lv=T, ordered=T, estimator='WLSMV');
+  
+  ## Compute reliability.
+  alpha = reliability(fit1f);
+  
+  ## Print to console.
+  print(paste('mars-sf', i, ':', alpha[2]));
+  
+}
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Reliability (RPM)
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## Load data.
+rpm = read.csv('../data/rpm.csv')
+
+## Apply rejections.
+reject = read.csv('../data/reject.csv')
+rpm = rpm[is.element(rpm$subject, reject[reject$reject==0,'subject']),];
+
+## Score missing data.
+rpm$accuracy = ifelse(is.na(rpm$accuracy), 0, rpm$accuracy);
+
+## Define formula.
+mod1f <- 'score =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9'
+
+## Restrict data to current short form. Pivot DataFrame.
[…]

### 2022-04-07 manuscript/main.tex
rewrites

--- 2022-01-19 manuscript/main.tex
@@
 
-\title{An item response theory analysis of the matrix reasoning item bank (MaRs-IB)}
+% override apa6 section headers
+% https://tex.stackexchange.com/questions/125537/how-to-modify-subsubsection-header-apa6-cls
+\makeatletter
+\renewcommand{\subsubsection}{\@startsection{subsubsection}{3}
+  {\z@}%
+  {\b@level@two@skip}{\e@level@two@skip}%
+  {\normalfont\normalsize\bfseries}}
+\makeatother
+
+% override apa6 first paragraph index
+% https://tex.stackexchange.com/questions/155028/indentation-apa6-class-for-2nd-paragraphs-after-title
+% \makeatletter
+%   \b@level@one@skip=-2.5ex plus -1ex minus -.2ex
+%   \b@level@two@skip=-2.5ex plus -1ex minus -.2ex
+% \makeatother
+
+\title{An item response theory analysis of the Matrix Reasoning Item Bank (MaRs-IB)}
 \shorttitle{IRT analysis of MaRs-IB}
@@
 
-Matrix reasoning tasks are one of the most well-studied and useful tasks in the behavioral sciences.
-In a typical administration, participants are presented with a 3x3 matrix in which 8 of 9 cells contains abstract figures or shapes.
-Participants are instructed to identify the missing figure from among several alternatives which are presented below the matrix.
-To do so, participants must use abstract reasoning to identify the piece that would complete the pattern.
+Matrix reasoning tasks are among the most well-studied and useful measures in the behavioral sciences.
+Matrix reasoning requires participants to use abstract reasoning to solve puzzles.
 Matrix reasoning indexes fluid intelligence and is correlated with working memory capacity \citep{conway2002latent, kane2004generality, unsworth2005working, salthouse2014relations}.
-The progressive matrices has excellent predictive validity and is associated of academic performance \citep{roth2015intelligence} and college admission exams \citep{frey2004scholastic, koenig2008act}.
 In low-stakes testing environments, performance on the progressive matrices is associated with participant motivation and willingness to expend effort \citep{gignac2018moderate, gignac2019maximum}.
-For all of these reasons, the progressive matrices are increasingly used as a control measure in individual difference behavioral studies \citep{gillan2016characterizing, rouault2018psychiatric, moutoussis2021decision}.
+The progressive matrices has excellent external validity, and is predictive of academic performance \citep{roth2015intelligence} and scores on college admission exams \citep{frey2004scholastic, koenig2008act}.
+% Thus, matrix reasoning tasks are increasingly utilized as control measure in individual difference behavioral studies \citep{gillan2016characterizing, rouault2018psychiatric, moutoussis2021decision}.
 
-Despite their utility, there are a number of challenges in using matrix reasoning measures in behavioral research.
-The first is the issue of copyright.
-The most prominent versions, including the Raven's progressive matrices \citep{raven2003raven} and the matrix subtest of the Wechsler Abbreviated Scale of Intelligence \citep{wechsler1999wechsler}, are not free to use and copyright may prevent these pen-and-paper tasks from being adapted into computerized tasks.
-A second challenge is those measures that are in the public domain typically have few items available.
-For example, the Hagen Matrices Test \citep{heydasch2014hagen} and International Cognitive Ability Resource \citep{condon2014international} have only 20 and 16 items available, respectively.
-The availability of only a limited number of items raises the issue of practice effects \citep{ng1974applicability, bors2003effect}, especially in the context of online experiments where multiple groups of researchers may be unknowingly administering the same items to the same participants.
+Unfortunately there are a number of challenges in using matrix reasoning tasks as part of behavioral research.
+One challenge is the problem of copyright.
+The most prominent matrix reasoning tasks, such as the Raven's progressive matrices \citep{raven2003raven} and the WAIS matrix subtest \citep{wechsler1999wechsler}, are typically not free to use and have restrictions against being computerized.
+A second challenge is that, across the measures in the public domain, there are relatively few items available.
+For example, the Hagen Matrices Test \citep{heydasch2014hagen} and ICAR matrix reasoning test \citep{condon2014international} have only 20 and 16 items, respectively.
+The availability of only a limited number of items raises the possibility of practice effects \citep{ng1974applicability, bors2003effect}, especially in the age of online experiments where multiple groups of researchers may be unwittingly administering the same test to many of the same participants.
 
-In response to these challenges, \cite{chierchia2019matrix} developed the matrix reasoning item bank (MaRs-IB).
-The MaRs-IB is an open-source, free-to-use bank of 80 matrix reasoning items, where each item has three unique shape variants and two sets of distractors.
-The items were designed to vary considerably in their complexity (see Figure 1 for examples).
-In an initial validation study of a MaRs-IB short-form test, \cite{chierchia2019matrix} that measures of performance in a sample of 659 adults, adolescents, and children had excellent split-half and test-retest reliability.
-Moreover, performance on the MaRs-IB items were found to correlate with performance on the digit span tasks and ICAR matrix reasoning, indicating good convergent validity.
-Most importantly, the authors made available the percent correct data for each item (and variant) so that others researchers would be able to construct new MaRs-IB forms of custom difficulty and duration.
+In response to these challenges, \cite{chierchia2019matrix} developed, and released into the public domain, the matrix reasoning item bank (MaRs-IB).
+The MaRs-IB includes 80 matrix reasoning puzzles.
+Each puzzle consists of a 3x3 matrix containing abstract shapes in eight out of nine cells.
+To solve a puzzle, participants must deduce which of four response options completes the matrix based on the pattern of changing relations among the shapes across cells (Figure 1).
+The puzzles in the MaRs-IB vary considerably in their complexity, both in the number of shapes per cell and in the number of changing relations across shapes.
+Importantly each template puzzle has six clones, which are matched in terms of complexity but differ in their perceptual properties (one of three shape sets) and their response options (one of two distractor sets).
+Thus, the MaRs-IB offer considerable utility insofar that they are free to use, have many items, and many unique items to prevent practice effects.
 
-There are unfortunately two key obstacles preventing researchers from being fully able to do so, both of which stem from the design of the validation study.
-The original MaRs-IB short-form is a fixed-time and  fixed-order measure, where participants complete as many items as possible within 8-minutes.
-Because participants only completed 33 items on average, the first issue is that most items were completed by too few participants to establish reliable norms (an issue that is compounded when considering each item has six variants).
+A preliminary investigation of the psychometric properties of the MaRs-IB items proved promising.
+Indeed, \cite{chierchia2019matrix} found that scores on a short-form matrix reasoning test using MaRs-IB items had good split-half reliability and test-retest reliability.
+The authors also found that performance observed scores on this test were moderately correlated with a digit span task and the ICAR matrix reasoning test, indicating good convergent validity.
+% note the release of summary statistics here?
+
+Unfortunately, there are two notable limitations in this initial investigation.
+The first is that, due to the design of the matrix reasoning test, there is little data available for the majority of the items in the MaRs-IB.
 Just over half the items (42/80) were completed by 100 or more participants.
-A second issue is that, because participants were free to choose for themselves whether to prioritize accuracy (number of correct responses) or productivity (number of items reached), it is likely that the percent correct data for items reached by fewer numbers of participants are biased.
-That is, items completed by fewer numbers of participants are more likely to have been completed by participants sacrificing accuracy for speed.
-Indeed, looking only at the easiest MaRs-IB items (dimension = 1), there is a strong positive correlation between number of completing participants and mean accuracy ($r = 0.673$).
+As such, there is too little data to characterize the psychometric properties of each item with reasonable confidence (an issue that is exacerbated when considering each template item has six variants).
+A second issue is that, also due to the design of the matrix reasoning test, participants were allowed to choose for themselves whether to prioritize accuracy (number of correct responses) or productivity (number of items reached).
+As a consequence, items later in the sequence are likely to be contaminated by participants sacrificing accuracy for speed, thereby distorting the summary statistics for these items.\footnote{appendix: Indeed, looking only at the easiest MaRs-IB items (dimension = 1), there is a strong positive correlation between number of completing participants and mean accuracy ($r = 0.673$).} Thus, caution is warranted in interpreting the results of the original study and additional validation is warranted.
 
-Here we wish to overcome the limitations of the previous study to further validate and make useful the MaRs-IB.
-First, we collected a novel dataset of response to MaRs-IB items in a large sample of adults (N=1501).
-Using a fixed-length design with counterbalancing, we ensured that all items, and their variants, were completed by a suitable number of participants.
-Next we analyzed participants' response data hierarchical item response models, which allowed us to measure all items' difficulty and discrimination while controlling for participants' varying abilities.
+%Better intro sentence tk: mention that this is a more extensive psychometric validation of the MaRs-IB in a much larger dataset, with three major improvements.
+The most important contribution of the current work is the evaluation of the psychometric properties of (almost) every item in the MaRs-IB using item response theory (IRT) \citep{embretson2013item, de2013theory}.
+IRT models confer several important benefits for the purposes of psychometric validation.
+First they provide an estimate of the difficulty and discrimination of each item, where the latter is an index of how informative or reliable an item is.
+Through these parameters, IRT models allows for the computation of the test information function (TIF), which measures how sensitive a test is at measuring performance at varying levels of ability.
+Finally, IRT make possible optimal test assembly \citep{van1998optimal}, or design of new tests with maximal sensitivity to measuring performance in a specified range of abilities under varying constraints (e.g.
+test length, time limits, parallel forms with equal TIF).
+
+A second contribution offered here is an analysis of the relationships between attributes of MaRs-IB items and their psychometric properties using explanatory item response models \citep{de2004explanatory, wilson2008explanatory}.
+Explanatory IRT models provide a framework for measuring how item parameters (e.g.
+difficulty) are shaped by their features.
+As such, these models provide means to evaluate how successful was the generation of the items in the MaRs-IB.
+For example, we can quantify the extent to which features of item complexity (e.g.
+number of elements, number of changing relations) predict item difficulty.
+Moreover, we can determine whether two items that vary only in their distractor set are psychometrically equivalent.
+Explanatory IRT models have the additional potential benefit of producing more accurate estimates of item parameters \citep{neuhaus2006separating}.
+
+%item clones would be isomorphic (psychometrically equivalent).
+In any realistic situation, the 2P-F is misspecified as it ignores the psychometric deviations introduced in the item cloning process (Glas & van der Linden, 2003).
+
+A third contribution is the decomposition of variance in item functioning from modeled and unmodeled sources using additive multilevel item structure (AMIS) models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
+This framework provides another means by which to evaluate the success of the design of the MaRs-IB by quantifying the residual variance in item parameters after accounting for item attributes.
+For example, two items of equal complexity may turn out not be psychometrically equivalent if unmodeled attributes (e.g.
+the types of changing relations) implicitly impact item functioning.
+Similarly, two item clones may not be psychometrically equivalent due to differences in their superficial properties.
+Quantifying the degree of error in the item cloning process is critical, as the degree of nonequivalent dictates whether clones are exchangeable.
+That is, if clones are not psychometrically equivalent, then they cannot be used interchangeably in parallel short forms.
+
+The paper proceeds as follows.
+First we analyze participants' response data hierarchical item response models, which allowed us to measure all items' difficulty and discrimination while controlling for participants' varying abilities.
 The item response models also allowed us to measure directly the extent to which that item variants are truly equivalent.
@@
 Using these methods, we introduce three new MaRs-IB short-form tests with good internal consistency and convergent validity.
+
 All data and code, including items' estimated parameters and example code for test assembly, is available at: \url{https://github.com/szorowi1/mars-irt}.
@@
 
-\subsection{Data collection}
+\subsection{Objectives}
 
-A total of N=1584 participants were recruited from the Prolific Academic platform (https://www.prolific.co) to participate in an online behavioral experiment in July - August, 2021.
-Participants were eligible if they were 18 years or older and currently resided in the United States.
+The purpose of the first study was to calibrate the items of the MaRs-IB, using response data collected from a large number of adult participants from the general population.
+In particular, we sought to accomplish the following three objectives: (1) to characterize the psychometric properties of the items in the MaRs-IB using item response models; (2) to investigate the associations between item attributes and their psychometric properties, and (3) to determine the equivalence of items with similar complexities, and clones of the same item templates.
+
+\subsection{Methods}
+
+\subsubsection{Participants}
+
+A total of N=1584 participants were recruited from the Prolific Academic platform (\url{https://www.prolific.co}) to participate in an online experiment between July - August, 2021.
+Participants were eligible if they were 18 years or older and lived in the United States.
+Total study duration was approximately 6.4 minutes (sd = 2.4) per participant.
+Participants received monetary compensation for their time (rate: \$10 USD/hr), plus a performance-based bonus up to \$0.50 USD.
+On average, participants earned a total of \$1.30 USD (sd = \$0.10).
+We offered performance bonuses as they have been found to motivate performance in low-stakes testing environments \citep{duckworth2011role, gignac2018moderate}.
 This study was approved by the Institutional Review Board of Princeton University (\#7392), and all participants provided informed consent.
-Total study duration was approximately 6.4 minutes (2.4 sd) per participant.
-Participants received monetary compensation for their time (rate USD \$10/hr), plus an incentive-compatible bonus up to \$0.50 based on task performance (average payment: mean = \$1.30 [\$0.10 sd]).
-Monetary incentives have previously been shown to motivate good performance in low-stakes testing environments \citep{duckworth2011role, gignac2018moderate}.
 
-After providing consent, participants completed a series of items from the MaRs-IB.
+To ensure data quality, the data from multiple participants were excluded prior to analysis (see Exclusion Criteria below) leaving the data from N=1501 participants for analysis.
+The majority of participants in this sample identified as women (men: N=670; women: N=811; non-binary or other: N=13; rather not say: N=2) and were 28.7 years old (sd = 9.9) on average.
+The sample was relatively well-educated, with the majority having completed a Bachelor's degree (N=507) or Master's degree or higher (N=322).
+By comparison, fewer participants endorsed having completed only some college (N=471), only a high school degree (N=199), or preferred not to say (N=2).
+
+\subsubsection{Procedure}
+
+After providing consent, participants completed 16 items from the MaRs-IB.
 The design of the MaRs-IB items have been described previously \citep{chierchia2019matrix}.
 Briefly, each MaRs-IB item consists of a 3 x 3 matrix.
-Eight of the nine resulting cells contain an abstract shape, while one cell on the bottom right-hand side of the matrix is empty.
-The participants' task was to complete the matrix by finding the missing shape among four possible alternatives.
-Items began with a 1200 ms fixation cross.
-Participants were then given up to 30 s to complete an item.
-After 25 s a clock appeared, indicating that 5 s remained before the next item began.
-An item ended when participants responded, or after 30 s had elapsed without response.
-Participants reviewed instructions before the task and were made to correctly complete three practice items before they could begin the task.
-Crucially, participants were instructed to respond carefully but to guess before the timer ran out.
-Our version of the MaRs-IB was based on publicly available code (\url{https://app.gorilla.sc/openmaterials/36164}).
+Eight of the nine resulting cells contain abstract shapes, while one cell on the bottom right-hand side of the matrix is empty.
+Participants are instructed to ``complete the matrix'' by finding the missing shape from among four possible alternatives.
 
-In order to ensure adequate sampling of all items, and in contrast to the original MaRs-IB test form, all participants completed a pseudorandomly-selected set of 16 out of 64 possible MaRs-IB items \footnote{We did not collect new data for the 16 easiest items from the MaRs-IB (items 1, 2, 3, 4, 5, 7, 8, 9, 32, 33, 38, 41, 43, 48, 57, 68) as the original study or pilot testing found performance on these items to be at or near ceiling.}.
-Prior to data collection, we sorted the 64 items into 16 groups of four based on the number of changing relationships between the shapes of the matrix.
-Participants were then randomly assigned one item from each of the 16 groups.
+Each trial of the experiment began with a fixation cross, lasting  1200 ms, after which a MaRs-IB item was presented.
+Participants were then given up to 30 s to solve the puzzle.
+After 25 s a clock appeared, indicating that 5 s remained before presentation of the next item.
+A trial ended when participants responded, or after 30 s had elapsed without response.
+Before the trials began, participants reviewed instructions and were made to correctly complete three practice items.
+Participants were told to respond carefully but to guess before the timer ran out.
+The format of the experiment and instructions were adapted from code made publicly available by the original authors (\url{https://app.gorilla.sc/openmaterials/36164}).
+
+In order to ensure sufficient and uniform sampling of the MaRs-IB items, all participants completed 16 items pseudorandomly selected from the 64 available MaRs-IB items.\footnote{We did not collect new data for the 16 easiest items from the MaRs-IB (items 1, 2, 3, 4, 5, 7, 8, 9, 32, 33, 38, 41, 43, 48, 57, 68) as the original study or pilot testing found performance on these items to be at or near ceiling.} The item sets were constructed as follows: First, we subdivided the 64 items into 16 sets of four based on their complexity.
+Participants were then randomly assigned one item from each of the 16 sets.
 In this way, all participants received test forms of approximately equal difficulty.
-Importantly we also took steps to counterbalance the sampling of item variants, so that we had approximately the same number of responses available by item set (1, 2, or 3) and distractor type (minimal difference or or paired difference; see \cite{chierchia2019matrix} for details).
+We also counterbalanced which variant of each item was presented across participants, such that we had approximately the same number of responses available by shape set (1, 2, or 3) and distractor type (minimal difference or or paired difference; see \cite{chierchia2019matrix} for details).
 
-To ensure data quality, the data from multiple participants who completed the experiment were excluded prior to analysis for one or more of the following reasons: for rapid guessing -- defined here as responding in less than 3 s -- on 4 or more trials (N=70); for failing to respond on four or more trials (N=8); for minimizing or otherwise shrinking the browser window beyond what was required to view the stimuli (N=7).
-In total, 83 of 1584 (5.24\%) participants were excluded, leaving the data from N=1501 participants for analysis.
-Among these participants, each item (64 in total) was sampled approximately 375 times (sd = 12) and each item variant (384 in total) was sampled approximately 62 times (sd = 7).
+\subsubsection{Exclusion criteria}
 
-\subsection{Descriptive statistics}
+The data from multiple participants who completed the experiment were excluded prior to analysis for one or more of the following reasons: rapid guessing (responding in less than 3 s) on four or more trials (N=70); failing to respond on four or more trials (N=8); or minimizing the browser window to a size smaller than the minimally required dimensions (N=7).
+In total 83 of 1584 (5.2\%) participants were excluded, leaving the data from N=1501 participants available for analysis.
+Among these participants, each item template (64 in total) was sampled approximately 375 times (sd = 12) and each item clone (384 in total) was sampled approximately 62 times (sd = 7).
 
-The majority of participants in the sample identified as women (men: N=670; women: N=811; non-binary or other: N=13; rather not say: N=2) and were 28.7 years old (sd = 9.9) on average.
-The sample was relatively well-educated, with the majority having completed a Bachelor's degree (N=507) or Master's degree or higher (N=322).
-By comparison, fewer participants endorsed having completed only some college (N=471), only a high school degree (N=199), or preferring not to say (N=2).
+\subsubsection{Item response models}
+
+We employed item response models to characterize the psychometric properties of each item.
+Specifically, we used additive multilevel item structure (AMIS) models with random residuals \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
+In item structure models, item clones are modeled as descending from a template item.
+The basis of all of these models was the three-parameter logistic (3PL) item response model, where the probability of a correct response ($y_{ijk} = 1$) for person $i$ on item clone $k$ belonging to item family $j$ is:
+
+\begin{equation} \label{eq:1}
+p(y_{ijk} = 1) = \gamma_{jk} + (1-\gamma_{jk}) \cdot \text{logit}^{-1} \left( \alpha_{jk} \cdot \theta_i - \beta_{jk} \right)
+\end{equation}
+
+\noindent where $\theta_i$ is the latent ability for person $i$, and $\beta_{jk}$, $\alpha_{jk}$, and $\gamma_{jk}$ are the difficulty, discrimination, and guessing parameters for item clone $k$ of item family $j$.
+Because guessing parameters are challenging to reliably estimate in the absence of very large amounts of data \citep{han2012fixing}, we fixed the guessing parameter for each item clone to the nominal guessing rate ($\gamma_{jk} = 0.25$).
+
+The difficulty and discrimination parameters were estimated following an additive multilevel item structure.
+Concretely, the difficulty of item clone $k$ belonging to item template $j$ was expressed as:
+
+\begin{equation}
+\beta_{jk} = \mu_\beta + \sum_{n=1}^N Q_{jn} \delta_{\beta n} + \epsilon_{\beta j} + \sum_{m=1}^M R_{km} \delta_{\beta m} + \epsilon_{\beta k}
+\end{equation}
+
+\noindent To elaborate, each item difficulty parameter was modeled as an intercept ($\mu_\beta$), reflecting the average difficulty across all items, and four additional components:
+
+\begin{enumerate}
+
+\item $\sum_{n=1}^N Q_{jn} \delta_{\beta n}$: the effect of the template (level-1) attributes on item difficulty, where $Q_{jn}$ is the value of attribute $n$ for item template $j$, and $\delta_{\beta n}$ is the effect of attribute $n$.
+
+\item $\epsilon_{\beta j}$: the template (level-1) residual, or residual variability in item difficulty of the item template unexplained by its attributes.
+
+\item $\sum_{m=1}^M R_{km} \delta_{\beta m}$: the effect of the clone (level-2) attributes on item difficulty, where $R_{km}$ is the value of attribute $m$ for item clone $k$, and $\delta_{\beta m}$ is the effect of attribute $m$.
+
+\item $\epsilon_{\beta k}$: the clone (level-2) residual, or residual variability in item difficulty of the item clone unexplained by its attributes.
+
+\end{enumerate}
+
+Similarly, item discrimination parameters were also modeled as the sum of an equivalent set of components:
+
+\begin{equation}
+\alpha_{jk} = \mu_\alpha + \sum_{n=1}^N Q_{jn} \delta_{\alpha n} + \epsilon_{\alpha j} + \sum_{m=1}^M R_{km} \delta_{\alpha m} + \epsilon_{\alpha k}
+\end{equation}
+
+\noindent where the interpretation of each component is the same as for item difficulty.
+
+The AMIS model framework provides a straightforward means of achieving Aims 2 and 3.
+Via the attribute effect components, we can test for any associations between item attributes and parameters (Aim 2).
+So too, via the error terms, we can quantify the magnitude of residual variance in item parameters from unmodeled sources both at the level of item templates and item clones (Aim 3).
+If the residual variance in item parameters across item clones is sufficiently large, then we can conclude item clones are neither psychometrically equivalent nor exchangeable.
+
+In the following analyses, we consider two template (level-1) attributes: feature number, or the number of unique shapes in each cell of an item; and rule number, or the number of changing relationships among the shapes (e.g.
+color change, shape change, position change) a participant must consider in order to solve the item.
+We predicted that both attributes would be positively related to item difficulty.
+In contrast, we had no hypotheses about the relationship between these attributes and item discrimination.
+
+In turn, we consider two clone (level-2) attributes: distractor type, or MD-type distractors versus PD-type distractors; and mean response time, or the average time spent deliberating on that item.
+Based on the preliminary findings of \cite{chierchia2019matrix}, we predicted there would be no association between distractor type and either type of item parameter.
+We note that, in contrast to all of the other modeled attributes, average response time is not an intrinsic property of an item.
+We opted to include it in our models, however, because of its previously established relationship with item difficulty \citep{chierchia2019matrix}, and because including response time as a covariate has previously been shown to improve parameter estimation \citep{bertling2018using}.
+To prevent any collinearity between the attributes, mean response time was orthogonalized with respect to the three other item attributes, and therefore reflects any residual structure after accounting for all other modeled effects.\footnote{Note that we intentionally do not include shape set, or the superficial properties of an item, as covariates in our model.
+This is due to the would-be complexity of doing so.
+Though each item has three clones, the clones do not all have the same perceptual patterns.
+There are XX unique patterns.
+Future studies might consider even larger datasets to study perceptual influence.}
+
+Finally, we estimated person-level abilities following an explanatory item response modeling approach \citep{wilson2008explanatory}:
+
+\begin{equation} \label{eq:4}
+\theta_i = \sum_{p=1}^P X_{ip} \rho_p + \epsilon_i    
+\end{equation}
+
+\noindent where $X_{ip}$ is the value of attribute $p$ for person $i$, $\rho_p$ is the partial correlation of attribute $n$, and $\epsilon_i$ is variance in ability unexplained by the person-level attributes.
+
+\subsubsection{Model Fitting}
+
+All models were estimated within a Bayesian framework using Hamiltonian Monte Carlo as implemented in Stan (v2.22) \citep{carpenter2017stan}.
+For all models, four separate chains with randomised start values each took 7500 samples from the posterior.
+The first 5,000 samples from each chain were discarded.
+As such, 10,000 post-warmup samples from the joint posterior were retained.
+The $\hat{R}$ values for all parameters were equal to or less than 1.01, indicating acceptable convergence between chains, and there were no divergent transitions in any chain.
+
+During estimation, the item discrimination parameters were restricted to be in the range $\alpha_{jk} \in [0, 5]$.
+
+To ensure the identifiability of all models, person abilities were constrained to have a mean of zero and a variance of 1.0, such that the residual variance was defined as $V(\epsilon) = 1 - \sum \rho_p^2$.
+Note that all person-level attributes were standard-scored prior being entered into the model.
+
+\subsubsection{Model comparison}
+
+The goodness of fit of the models was compared using Bayesian leave-one out cross-validation \citep{vehtari2017practical}, which has been found to perform better than more traditional information criteria for comparing item response models \citep{luo2017performances}.
+Importantly, we computed the conditional leave-one-cluster out (LOCO) cross- validation \citep{merkle2019bayesian}, which measures a model's ability to generalize to held out items (rather than held out responses or held out participants) --- i.e., for inferring generalization to additional items constructed using the same features rather than to additional participants sampled from the same population.
+
+\subsubsection{Goodness of fit}
+
+The fit of the best-fitting model to the data was evaluated through posterior predictive model checks \citep{gelman1996posterior, levy2017bayesian}.
+Item fit was assessed using the Orlando and Thissen index ($S-\chi^2$) \citep{toribio2011discrepancy}, which compares the observed and expected proportions of participants who answer an item correctly at each raw-score group.
+The unidimensionality of the data was assessed using the generalized dimensionality discrepancy measure (GDDM) \citep{levy2011generalized}, which tests for violation of the assumption of local independence.
+This latter measure is crucial for ensuring there are negligible levels of local item dependence in the data, which may otherwise lower the efficiency of future test forms due to the inclusion of correlated, redundant items.
+
+\subsection{Results}
+
+\subsubsection{Descriptive statistics}
 
@@
 
-Though not the primary focus of the current study, we briefly investigated the relationship between response accuracy and times.
-We used a mixed-effects model to predict participants' trial-level log-transformed response times as a function their response on that trial (correct or incorrect), their rest score (their total score on all other items), that item's difficulty (one minus the accuracy rate), and all possible interactions.
-We also specified an intercept per participant to control for each participants' overall speed.
+We investigated the relationship between response accuracy and times using a random-intercepts linear regression model.
+Participants' log-transformed, item-level response times were predicted as a function of item accuracy, rest score (observed score for all other items), and item difficulty (one minus the proportion correct on that item).
+The results are summarized in Table XX.
+On average, participants spent 15.9 s (sd = 7.2 s) on each item.
+There were significant main effects of accuracy ($\beta = 0.067$, $t = 9.84$, $p < 0.001$), rest score ($\beta = 0.125$, $t = 16.95$, $p < 0.001$), and item difficulty ($\beta = 0.137, t = 44.33, p < 0.001$) such that response times were slower for correct responses, better-performing participants, and more difficult items.
 
-The complete summary of results from the regression analysis are presented in Table XX.
-On average, participants spent 15.9 s per item (sd = 7.2 s, IQR = 10.1 - 21.5 s).
-The mixed-effects model identified significant main effects of accuracy ($\beta = 0.067, t = 9.84, p < 1e^{-6}$), rest score ($\beta = 0.125, t = 16.948, p < 1e^{-6}$), and item difficulty ($\beta = 0.137, t = 44.326, p < 1e^{-6}$).
-On average then, response times were slower for correct responses, more accurate participants, and more difficult items.
-There was also a significant positive accuracy by difficulty interaction (correct responses were even slower for more difficult items) and a significant negative accuracy by score interaction (correct responses were faster for better performing participants).
-Finally, there was a significant positive difficulty by score interaction, such that better performing participants exhibited slower responses for more difficult items.
+There was also a significant accuracy by difficulty interaction (correct responses were even slower for more difficult items), and a significant accuracy by score interaction (correct responses were faster for better performing participants).
+Finally, there was a significant difficulty by score interaction, such that better performing participants exhibited slower responses for more difficult items.
+This pattern of response time results is visible in Figure \ref{fig:fig01}c.
 The three-way interaction term was not statistically significant.
-
-In short, we corroborate the previous finding that item difficulty is associated with slower response times \citep{chierchia2019matrix}.
-In addition, we identified two notable patterns: better performing participants exhibited slower response times on average, which slowered further for more difficult times.
-This pattern of response time results is summarized in Figure \ref{fig:fig01}c.
-This pattern of results is consistent with controlled processing accounts, a point we return to in the Discussion.
 
@@
 \includegraphics[width=1.0\textwidth]{figures/fig01.png}
-\caption{\label{fig:fig01}This is a figure caption.}
+\caption{\label{fig:fig01}Summary of performance on the MaRs-IB items.
+(A) The distribution of total scores across all 1501 participants.
+(B) The distribution of proportion correct responses across all 384 items.
+(C) The distribution of participants' median response times across items broken down by participants' total scores and item difficulty (one minus proportion correct responses).
+Error bars indicate bootstrapped 95\% confidence intervals around the mean.}
 \end{figure}
 
-Summary paragraph?
+\subsubsection{Alternative item structure models}
 
-\subsection{Item response models}
+We fit a series of nested IRT models in order to identify the model that best explained participants' response data.
+The models varied in how item parameters were estimated, with successive models allowing for greater variability in parameters across item templates and clones.
+Identifying the model that most accurately captures the structure of item parameters is important in the context of item cloning, as improperly modeled variance can result in the biased estimation of item parameters \citep{lathrop2017item}.
 
-To measure the psychometric properties of each item, we used item response models.
-Given the hierarchical nature of the MaRs-IB -- 64 items with six variants each -- we specifically modeled participants' response data using item using a series of item cloning models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
-In item cloning models, the variants of an item (or "clones") are explicitly model as having been generated from the same template item (or "family").
-The basis of all of the models used here is the three-parameter logistic (3PL) item response model, where the probability of a correct response ($y_{ijk} = 1$) for person $i$ on item clone $k$ belonging to item family $j$ is:
+We fit the models in two waves.
+The first wave was comprised of four models in which the item discrimination parameters were fixed ($\alpha = 1$).\footnote{Note that these models still included a fixed guessing parameter ($\gamma = 0.25$).
+We also estimated models with no guessing parameters (i.e.
+one-parameter logistic or Rasch models), but model comparison indicated these models were worse fits to the data (Table SXX).} We did this in order to minimize the number of total models that we would otherwise need to fit (i.e.
+all permutations of the four models below, applied to item difficulty and discrimination parameters independently).
+The first four models were as follows:
 
-\begin{equation}
-p(y_{ijk} = 1) = \gamma_{jk} + (1-\gamma_{jk}) \cdot \text{logit}^{-1} \left( \alpha_{jk} \cdot \theta_i - \beta_{jk} \right)
-\end{equation}
+\begin{itemize}
 
-\noindent where $\theta_i$ is the latent ability for person $i$, and $\beta_{jk}$, $\alpha_{jk}$, and $\gamma_{jk}$ are the difficulty, discrimination, and guessing parameters for item clone $k$ of item family $j$.
-Since guessing parameters are challenging to estimate in the absence of very large amounts of data \citep{han2012fixing}, we fixed the guessing parameter for each item clone to the nominal guessing rate ($\gamma_{jk} = 0.25$).
+\item Model 1a: Item difficulty is a function only of template attributes (number of shapes, number of changing relations).
+In this model, items of equal complexity are psychometrically equivalent as are all item clones.
 
-To account for the nested structure of the MaRs-IB, the item difficulty and discrimination parameters were estimated using an additive multilevel item structure model \citep{cho2014additive}.
-Specifically, the difficulty of item clone $k$ belonging to item family $j$ was expressed as:
+\item Model 1b: Item difficulty is a function of template attributes and residuals.
+In this model, items of equal complexity may not be equally difficult, but all clones of an item are still psychometrically equivalent.
 
-\begin{equation}
-\beta_{jk} = \mu_\beta + \sum_{n=1}^N Q_{jn} \delta_{\beta n} + \epsilon_{\beta j} + \sum_{m=1}^M R_{km} \delta_{\beta m} + \epsilon_{\beta k}
-\end{equation}
+\item Model 1c: Item difficulty is a function of both template and clone attributes (including distractor type and average response time), and template-level residuals.
+In this model, item clones may systematically vary in their difficulty as a function of their attributes.
 
-\noindent Every difficulty parameter is thus modeled as an intercept ($\mu_\beta$), reflecting the average difficulty across all items, and four components:
+\item Model 1d: Item difficulty is function of both template/clone attributes and residuals.
+In this model, neither item templates nor clones are psychometrically equivalent.
+The degree of nonequivalence is dependent on the magnitude of the residual terms.
 
-\begin{enumerate}
+\end{itemize}
 
-\item $\sum_{n=1}^N Q_{jn} \delta_{\beta n}$: the effect of the level-1 attributes, where $Q_{jn}$ is the value of attribute $n$ for item family $j$, and $\delta_{\beta n}$ is the effect of attribute $n$.
-This component is the fixed effects contribution of item family to item difficulty.
+The goodness-of-fit of the four models is summarized in Table \ref{table:1}.
+Model 1d demonstrated a better fit of the data than Model 1a ($\Delta$ LOCO = 1223.69, se = 44.94), Model 1b ($\Delta$ LOCO 511.58, se = 28.77), and Model 1c ($\Delta$ LOCO = 187.96, se = 17.25).
+Thus, Model 1d was chosen as the starting point for the second wave of model fitting.
 
-\item $\epsilon_{\beta j}$: the level-1 residual, or variance unexplained by the item family attributes. This component is the random effects contribution of item family to item difficulty.
+Next we estimated a second set of three models in which the item discrimination parameter was a free parameter.
+For all models, the item difficulty parameters were specified according to Model 1d.
+The structure of the three models were as follows:
 
-\item $\sum_{m=1}^M R_{km} \delta_{\beta m}$: the effect of the level-2 attributes, where $R_{km}$ is the value of attribute $m$ for item clone $k$, and $\delta_{\beta m}$ is the effect of attribute $m$.
-This component is the fixed effects contribution of item clone to item difficulty.
+\begin{itemize}
 
-\item $\epsilon_{\beta k}$: the level-2 residual, or variance unexplained by the item clone attributes. This component is the random effects contribution of item clone to item difficulty.
+    \item Model 2a: Item discrimination is a function only of both template and clone attributes.
+Here we include all attributes at once having found that they were all predictive of item difficulty (see below).
 
-\end{enumerate}
+    \item Model 2b: Item discrimination is a function of both item attributes and template residuals. In this model, items of equal complexity may not be equally discriminating.
 
-To elaborate, level-1 attributes describe the properties of an item family.
-Here we model the effects of two properties: feature number, or the number of unique shapes in each cell of an item; and rule number, or the number of changing relationships among the shapes (e.g.
-color change, shape change, position change) a participant must consider in order to solve the item.
+    \item Model 2c: the item discrimination parameters were modeled with fixed and random effects at both levels.
+In this model, neither item templates nor clones are equivalently discriminating.
+The degree of nonequivalence is dependent on the magnitude of the residual terms.
 
-In turn, level-2 attributes describe properties of an item clone.
-Here we model the effects of two additional properties: distractor type, or whether the distractors were generated following the minimal difference (MD) strategy and a paired difference (PD) strategy; and mean response time, or the average time spent on that item.
-Mean response time was orthogonalized with respect to the three other item attributes, and therefore reflects any residual structure after accounting for all other modeled effects.
+\end{itemize}
 
-Similarly, item discrimination parameters were also expressed as the sum of an equivalent set of components:
+The goodness-of-fit of these models is summarized in Table \ref{table:1}.
+Notably, we did not find the most complex model (Model 2c) to be the best-fitting model.
+Although Model 2c was preferred to Model 2a ($\Delta$ LOCO = 2.13, se = 2.04), it was not preferred to Model 2b ($\Delta$ LOCO = -1.44, se = 0.85).
+%We note that the authors of the LOO metric recommend that standard errors be at least twice as large as the difference between models to be confident that one model truly provides a better fit.
+However, these differences in LOCO values from each model to the next were each within two standard errors of the mean, indicating only weak predictive improvement of Model 2b over the others \citep{vehtari2022cv}.
+As such we proceed by select Model 2b as the best-fitting model overall, but proceed in discussing it with caution.
+We also note that Model 2b fit clearly better than Model 1d ($\Delta$ LOCO = 85.91, se = 6.33, supporting the inclusion of varying item discrimination parameters).
 
-\begin{equation}
-\alpha_{jk} = \mu_\alpha + \sum_{n=1}^N Q_{jn} \delta_{\alpha n} + \epsilon_{\alpha j} + \sum_{m=1}^M R_{km} \delta_{\alpha m} + \epsilon_{\alpha k}
-\end{equation}
+\subsubsection{Objective 1}
 
-\noindent The fixed effect components of the item discrimination parameters included the same level-1 and level-2 attributes as for the item difficulty parameters (i.e.
-number of features, number of rules, distractor type, average response time).
-During estimation, the item discrimination parameters were restricted to be in the range $\alpha_{jk} \in [0, 5]$.
+Across all items, the average item difficulty was $\mu_\beta = 0.177$ (95\% HDI = 0.118 -- 0.238).
+There was also considerable variance in item difficulty (sd = 1.431, 95\% HDI = 1.370 -- 1.492), the smallest ($\beta$ = -3.704, 95\% HDI = -4.687 -- -2.775) and largest ($\beta$ = 3.497, 95\% HDI = 2.481 -- 4.635) item difficulty parameters spanning a wide range.
+For an average ability participant, these estimates translate to an expected average accuracy of 58.4\% across all items, and essentially ceiling (floor) performance for the least (most) difficult items.
+Thus, items in MaRs-IB span a substantial range of difficulty.
 
-Though complex, item cloning models provide a natural means of evaluating the success of the MaRs-IB item generation process.
-For example, modeling an item's properties as a function of its attributes allows us to quantify the extent to which item design worked as intended (e.g.
-that an increase in the number of changing relationships between shapes corresponds to an increase in item difficulty).
-The inclusion of level-1 random effects allows us to quantify the extent to which item families with identical attributes (e.g.
-number of features, number of changing relations) are psychometrically equivalent.
-Similarly, the inclusion of level-2 random effects allows us to quantify the extent to which clones of an item family are psychometrically equivalent.
+In turn, the average item discrimination was $\mu_\alpha = 1.298$ (95\% HDI = 1.212 - 1.385).
+Variability in item discrimination was modest in comparison to item difficulty (sd = 0.221, 95\% HDI = 0.121 - 0.322).
+Item difficulty and discrimination were uncorrelated across items ($\rho$ = -0.060, 95\% HDI = -0.379 - 0.250).
+Following convention \citep{baker2017basics}, the items in the MaRs-IB exhibited medium ($\alpha \in$ 0.65 - 1.34) to high ($\alpha \in$ 1.35 – 1.69) levels of discrimination and are thus suitable for measuring matrix reasoning ability.
 
-In order to test the hypotheses outlined above, we fit a series of nested item cloning models \citep{cho2014additive, lathrop2017item}:
+\subsubsection{Objective 2}
 
-\begin{enumerate}
+Next we inspected the associations between item attributes and parameters.
+As predicted, a one-unit change in the number of item features was associated with an increase in item difficulty ($\delta_{\beta_1}$ = 0.579, 95\% HDI = 0.389 -- 0.770), or an approximate 10.6\% reduction in accuracy for a participant of average ability.
+Similarly, a one-unit change in the number of changing relations was also associated with an increase in item difficulty ($\delta_{\beta_2}$ = 0.514, 95\% HDI = 0.378 -- 0.663), or a 9.5\% reduction in accuracy.
 
-\item \textit{Model a}: Item parameters are expressed only as a function of their level-1 attributes.
-This model treats all clones within a family as interchangeable and all item families with the same attributes as equivalent.
+Looking next at the clone-level attributes, there was an unexpected association between distractor type and item difficulty ($\delta_{\beta_3}$ = 1.105, 95\% HDI = 0.940 -- 1.269).
+That is, all else equal, an item with MD distractors was associated with an expected 19.9\% reduction in accuracy compared to items with PD distractors.
+Finally, as expected, there was a positive association between mean response time and item difficulty ($\delta_{\beta_4}$ = 0.541, 95\% HDI = 0.414 -- 0.671).
 
-\item \textit{Model b}: Item parameters are expressed as a function of their level-1 attributes and residuals.
-This model treats all clones within a family as interchangeable, but now item families with the same attributes may vary in their properties.
+In contrast, there was only one credible association between an item attribute and item discrimination parameters.
+A one-unit change in the number of changing relations was associated with a small but credible increase in item discrimination  $\delta_{\alpha_2}$ = 0.020, 95\% HDI = 0.003 -- 0.037).
+There was no reliable association between item discrimination and the number of item features ($\delta_{\alpha_1}$ = -0.015, 95\% HDI = -0.036 -- 0.008), distractor type  ($\delta_{\alpha_3}$ = -0.029, 95\% HDI = -0.065 -- 0.005), or average response time ($\delta_{\alpha_4}$ = -0.010, 95\% HDI = -0.030 -- 0.009).
 
-\item \textit{Model c}: Item parameters are expressed as a function of their level-1 attributes, level-2 attributes, and level-1 residuals.
-This model allows clones within a family to vary, but all variation must be explained solely through their attributes.
+\subsubsection{Objective 3}
 
-\item \textit{Model d}: Item parameters are expressed as a function of their level-1/2 attributes residuals. The full model in which item families and clones may vary in their psychometric properties.
+Finally, we inspected the relative contributions of modeled and unmodeled sources to variability in item functioning.
+The two modeled item-level attributes, number of features and changing relations, explained approximately 65.5\% of the variance in difficulty across the 64 item templates (or 35.0\% of the variance across all 384 items).
+In conjunction with distractor type, these three item attributes explained nearly half (48.2\%) of the variance in item difficulty across all 384 items.
+Nevertheless, there was non-negligible residual variability in item difficulty at the template- and clone-level as evidenced by the comparatively better fit of item structure models including random effects terms at both levels.
+As such, neither item templates nor clones can be considered exchangeable with respect to difficulty.
 
-\end{enumerate}
+By contrast, the three inherent item attributes explained  approximately 39.2\% of the variance in item discrimination across item templates.
+In contrast to item difficulty, model comparison suggested there was only non-negligible residual variance in item discrimination at the template-level.
+Indeed, a model that included random effects at the clone-level offered a comparably worse fit.
+As such, there was not credible evidence that item clones are not exchangeable with respect to discrimination.
 
-The resulting model fits are then compared as a test of the hypotheses above.
-For example, if the psychometric properties of items can be solely explained by their attributes, then Model A should be preferred to B; in contrast, if item families psychometrically vary beyond what their attributes can explain, then Model B should be preferred.
-Similarly, if the clones of an item family are truly equivalent, then Models A or B should be preferred; in contrast, if item clones vary (as a function of their attributes or otherwise) then model C or D should be preferred.
+\subsubsection{Person-level effects}
 
-Finally, we estimated person-level abilities following an explanatory item response modeling approach \citep{wilson2008explanatory}:
+We investigated person-level abilities as a function of four covariates: age, gender, average response time, and $\Delta$ response time.
+This last term reflects the degree to which a participant slowed down (or sped up) their response time as a function of item difficulty (Figure \ref{fig:fig01}c).
+All covariates were standard-scored prior to analysis with the exception of gender, which was binary coded (male = -0.5, female = 0.5).
 
-\begin{equation}
-\theta_i = \sum_{p=1}^P X_{ip} \rho_p + \epsilon_i    
-\end{equation}
+There was a negative association between age and ability ($\rho_1$ = -0.299, 95\% HDI = -0.353 -- -0.246).
+There was a small but credible association between gender and ability ($\rho_2$ = 0.128, 95\% HDI = 0.078 -- 0.182), such that women performed marginally better than men.
+Average response time was positively associated with ability ($\rho_3$ = 0.427, 95\% HDI = 0.374 -- 0.478), as was $\Delta$ response time ($\rho_4$ = 0.368, 95\% HDI = 0.313 -- 0.424), indicating a speed-accuracy tradeoff.
 
-\noindent where $X_{ip}$ is the value of attribute $p$ for person $i$, $\rho_p$ is the partial correlation of attribute $n$, and $\epsilon_i$ is variance in ability unexplained by the person-level attributes.
-
-We modeled person-level abilities as a function of four attributes: age, gender (coded as: male = -0.5, female = 0.5; all else = 0), average response time, and $\Delta$ response time.
-The last term reflected the degree to which a participant adjusted their response time as a function of item difficulty, and thus measured a participant's tendency to slow down (or speed up) in response to more challenging items (Figure \ref{fig:fig01}c).
-
-To ensure the identifiability of all models, person abilities were constrained to have a mean of zero and a variance of 1.0, such that the residual variance was defined as $V(\epsilon) = 1 - \sum \rho_p^2$.
-Note that all person-level attributes were standard-scored prior being entered into the model.
-
-All models were estimated within a Bayesian framework using Hamiltonian Monte Carlo as implemented in Stan (v2.26) \citep{carpenter2017stan}.
-For all models, four separate chains with randomised start values each took 7500 samples from the posterior.
-The first 5000 samples from each chain were discarded.
-As such, 10,000 post-warmup samples from the joint posterior were retained.
-The $\hat{R}$ values for all parameters was equal to or less than 1.01, indicating acceptable convergence between chains, and there were no divergent transitions in any chain.
-
-The goodness of fit of the models was compared were compared using Bayesian leave-one out cross-validation \citep{vehtari2017practical}, which has been found to perform better than more traditional information criteria for comparing item response models \citep{luo2017performances}.
-Importantly, we computed the conditional leave-one-cluster out (LOCO) cross- validation \citep{merkle2019bayesian}, which measures a model's ability to generalize to held out items (rather than held out responses).
-
-The fit of the best-fitting model to the data was evaluated through posterior predictive model checks \citep{gelman1996posterior, levy2017bayesian}.
-Item fit was assessed using the Orlando and Thissen index ($S-\chi^2$) \citep{toribio2011discrepancy}, which compares the observed and expected proportions of participants who answer an item correctly at each raw-score group.
-The unidimensionality of the data was assessed using the generalized dimensionality discrepancy measure (GDDM) \citep{levy2011generalized}, which tests for violation of the assumption of local independence.
-This latter measure is crucial for ensuring there is not non-negligible levels of local item dependence in the data, which may otherwise lower the efficiency of future test forms due to the inclusion of correlated, redundant items.
-
-\subsection{Model comparison}
-
-First we estimated the four models described above holding the item discrimination parameters fixed ($\alpha = 1$).
-Note that these models still included a fixed guessing parameter ($\gamma = 0.25$).
-We also estimated models with no guessing parameters (i.e.
-one-parameter logistic or Rasch models), but model comparison indicated these models were worse fits to the data (Table XX).
-
-The goodness of fit of the four models was compared using Bayesian leave-one-cluster out cross-validation as summarized in Table \ref{table:1}.
-The full model (level-1/2 fixed \& random effects) fit the response data better than the model with only level-1 fixed effects ($\Delta$ LOCO = 1223.69, se = 44.94), the model with level-1 fixed and random effects ($\Delta$ LOCO 511.58, se = 28.77), and the model with all but level-2 random effects ($\Delta$ LOCO = 187.96, se = 17.25).
-Therefore, the two-level model with fixed and random item difficulty effects is chosen as the estimated model to discuss further.
-
-Overall, the average item difficulty was $\mu_\beta = 0.177$ (95\% HDI = 0.118 - 0.238).
-Difficulty also ranged considerably across items (sd = 1.431, 95\% HDI = 1.370 - 1.492).
-We next unpack the sources of variability in item difficulty, beginning first with the level-1 attributes.
-As expected, a one-unit change in the number of item features was credibly associated with an increase in item difficulty ($\delta_1$ = 0.132, 95\% HDI = 0.087 - 0.174).
-Similarly, a one-unit change in the number of changing relations was also associated with an increase in item difficulty ($\delta_2$ = 0.112, 95\% HDI = 0.080 - 0.144).
-Together, these two attributes explained approximately 65.5\% of the variance in difficulty across item families (35.0\% across all items).
-This result supports the success of the item generation process, insofar that item difficulty is associated with and well-explained by the core item attributes.
-It also highlights, however, that item families with identical attributes are not exchangeable as confirmed by the better fit of Model 1b compared to Model 1a ($\Delta$ LOCO = 712.11, se = 34.61).
-
-Turning next to the level-2 attributes, there was an unexpected association between distractor type and item difficulty such that otherwise identical items were more difficult with MD than PD distractors ($\delta_3$ = 0.252 , 95\% HDI = 0.218 - 0.287).
-There was also an association between average response time and item difficulty ($\delta_4$ = 0.122, 95\% HDI = 0.093 - 0.151).
-In total, the controllable fixed effects (number of features, number of changing relations, distractor type), explained nearly half (48.2\%) of the variance in item difficulty.
-Compared to a model with only level-1 fixed and random effects, the improved fits of Models 1c ($\Delta$ = 323.62, se = 23.36) and 1d ($\Delta$ LOCO = 187.96, se = 17.25) indicates that item clones are not equivalent in their difficulty.
-Taking into account the controllable fixed effects and level-1 random effects, the total explained variance in item difficulty is approximately 72.0\%.
-To summarize, although item variants are not strictly exchangeable, only the minority of variance in item difficulty can be attributed to unmodeled effects (e.g.
-perceptual features).
-
-Next we estimated the item cloning models again, but no longer holding the item discrimination parameters constant.
-In all models, the item difficulty parameters were estimated with fixed and random effects at both levels.
-In contrast, for Model 2a the item discrimination parameters were modeled as a function of level-1 and level-2 fixed effects only.
-In Model 2b, the item discrimination parameters were modeled as a function of the fixed effects and level-1 random effects.
-Finally, in Model 2c, the item discrimination parameters were modeled with fixed and random effects at both levels.
-
-The goodness of fit of the three models is summarized in Table \ref{table:1}.
-Notably, the best-fitting model was not the full model but instead the model with only level-1 random effects (Model 2b).
-This was preferred to the model with only fixed effects ($\Delta$ LOCO = 2.13, se = 2.04) and the model with both level-1 and level-2 random effects ($\Delta$ LOCO = -1.44, se = 0.85).
-We note that the authors of the LOO metric recommend that standard errors be at least twice as large as the difference between models to be confident that one model truly provides a better fit.
-\footnote{\url{https://discourse.mc-stan.org/t/loo-comparison-in-reference-to-standard-error/4009}} As such, we proceed by interpreting Model 2b as the overall best-fitting model with caution.
-Further note that this model still fit substantially better than the winning model with fixed item discrimination parameters ($\Delta$ LOCO = 85.91, se = 6.33).
-
-Overall, the average item discrimination was $\mu_\alpha = 1.298$ (95\% HDI = 1.212 - 1.385), and exhibited modest variance across items (sd = 0.221, 95\% HDI = 0.121 - 0.322).
-Among the fixed-effects, there was only a credible association between the number of changing relations and item discrimination ($\delta_{\alpha_2}$ = 0.067, 95\% HDI = 0.011 - 0.123).
-There was no such association for the number of item features ($\delta_{\alpha_1}$ =-0.037 95\% HDI = -0.090 - 0.017), distractor type ($\delta_{\alpha_3}$ -0.036, 95\% HDI = -0.082, 0.006), or average response time ($\delta_{\alpha_4}$ = -0.025, 95\% HDI = -0.075 - 0.024).
-Regardless, the controllable fixed effects (number of features, number of changing relations, distractor type) explained approximately (39.2\%) of the variance in item discrimination.
-In summary then, there is evidence that items vary in their discrimination but not strong evidence to suggest that item variants vary substantially from one another.
-We return to this point in the discussion.
-
-Briefly, we turn to the person-level predictors of performance.
-Corroborating the mixed-effects regression model, average response time was positively associated with ability ($\rho_1$ = 0.427, 95\% HDI = 0.374 - 0.478), as was $\Delta$ response time ($\rho_2$ = 0.368, 95\% HDI = 0.313 - 0.424).
-There was a negative relationship between age and ability ($\rho_3$ = -0.299, 95\% HDI = -0.353 -- -0.246).
-Finally, there was a small and unexpected association between gender and ability ($\rho_4$ = 0.128, 95\% HDI = 0.078 - 0.182), such that women performed slightly better than men on average.
-Though we used the colorblind-friendly MaRs-IB stimuli for this study, it is possible that residual effects may persist.
+\subsubsection{Model Fit}
 
@@
 
-\section{Test assembly}
+\section{Validation Study}
+
+\subsection{Optimal test assembly}
 
 With 64 item templates, with 6 variants each, there is a vast number of possible test forms we could construct using the MaRs-IB.
-Returning to our motivating problem -- the availability of abbreviated matrix reasoning measures with the potential for retest -- we opted to construct three parallel short form measures.
+Returning to our motivating problem --- the availability of abbreviated matrix reasoning measures with the potential for retest --- we opted to construct three parallel short form measures.
 To do so, we used mixed integer programming \citep{der2005wj} to maximize the test information function (TIF) of each short form subject to the following constraints:
@@
 
-    \item Variants could be of the MD or PD distractor type, but not both. This chosen to prevent item redundancy (i.e. including the same puzzle twice in one short form).
+    \item Variants could be of the MD or PD distractor type, but not both. This was chosen to prevent item redundancy (i.e. including the same puzzle twice in one short form).
 
@@
 The results of the test assembly are presented in Figure \ref{fig:fig02}.
-As expected, the MIP solver selected for items that were more discriminating on average (Figure \ref{fig:fig02}a).Though this was not explicit constraint, the test characteristic curves (TCCs) of the three short forms are markedly similar, each with an expected total score of 7.7 out of 12 items.
+As expected, the MIP solver selected for items that were more discriminating on average (Figure \ref{fig:fig02}a).
+Though this was not an explicit constraint, the test characteristic curves (TCCs) of the three short forms are markedly similar, each with an expected total score of 7.7 out of 12 items.
 Note that, because low ability participants are expected to guess an item correctly 25\% of time, the lower asymptote of the expected score distribution is 3 and the total score distribution is left-skewed.
-Finally, the TIFs across the three short forms are remarkably similar \ref{fig:fig02}b) as would be expected given the assembly constraints.
+Finally, the TIFs across the three short forms are remarkably similar (Figure~\ref{fig:fig02}b) as would be expected given the assembly constraints.
 As such, the short forms are predicted to have similar levels of precision across ability levels.
@@
 \includegraphics[width=1.0\textwidth]{figures/fig02.png}
-\caption{\label{fig:fig02}This is a figure caption.}
+\caption{\label{fig:fig02}Results from the test assembly of three parallel MaRs-IB short form (SF) measures.
+(A) The psychometric properties of the items selected for each form, compared to all remaining items.
+Points represent the posterior means of the item difficulty and discrimination  parameters.
+(B) The test characteristic curves (TCCs) for each short form.
+The degree of overlap highlights the similarity of expected scores for each test form.
+(C) The test information functions (TIFs) for each short form.
+The degree of overlap highlights the similarity of reliability for each test form.}
 \end{figure}
@@
 
-\section{Validation Study}
+\subsection{Methods}
 
-\subsection{Data Collection}
+\subsubsection{Participants}
 
 A total of N=347 participants were recruited from the Prolific Academic platform (https://www.prolific.co) to participate in an online behavioral experiment in November, 2021.
-Participants were eligible if they currently resided in the United States and had not participated in the calibration study.
+Participants were eligible if they currently resided in the United States and had not participated in the previous study.
+Total study duration was approximately 10.8 minutes (sd = 4.6) per participant.
+Participants received monetary compensation for their time (rate USD \$10/hr), plus a performance-based bonus up to \$0.75 based on task performance.
+Participants earned on average \$2.10 USD (sd = \$0.18 sd).
 This study was approved by the Institutional Review Board of Princeton University (\#7392), and all participants provided informed consent.
-Total study duration was approximately 10.8 minutes (4.6 sd) per participant.
-Participants received monetary compensation for their time (rate USD \$10/hr), plus an incentive-compatible bonus up to \$0.75 based on task performance (average payment: mean = \$2.10 [\$0.18 sd]).
 
-After providing consent, participants completed the experiment which was divided into three section.
-In the first, participants completed three short surveys: the 10-item need for cognition survey \citep{chiesi2018applying}, the 8-item PROMIS Cognitive Function-Short Forms (8a) \citep{iverson2021normative}, and the 8-item subject numeracy scale \citep{fagerlin2007measuring}.
-These measures were included as part of planned exploratory analyses and will not be reported in the main text (all correlations between these self-report measures and performance on the matrix reasoning tasks are reported in Supplement).
-Next, participants completed one of the three 12-item MaRs-IRB short-forms described above.
-The administration procedure of the short-forms was identical to that in the first study.
-Finally, participants completed the 9-item abbreviated Raven's progressive matrices (form A) \citep{bilker2012development}.
-To minimize bias, the administration procedure for the progressive matrices was matched to that for the MaRs-IB items.
-
-To ensure data quality, the data from multiple participants were excluded prior to analysis for one or more of the following reasons: for failing to complete all three sections of the experiment (N=17); for failing one or more attention checks \citep{zorowitz2021inattentive} in the self-report measures (N=21); for experiencing technical difficulties during the loading of the stimuli (N=11); for rapid guessing on 4 or more trials (N=3); or for failing to respond on four or more trials (N=1).
-In total, 47 of 347 (13.54\%) participants were excluded, leaving the data from N=300 participants for analysis.
-
-\subsection{Descriptive statistics}
-
+To ensure data quality, the data from multiple participants were excluded prior to analysis (see Exclusion Criteria below) leaving the data from N=300 participants for analysis.
 The majority of participants identified as women (men: N=138; women: N=147; non-binary or other: N=13; rather not say: N=2).
@@
 The sample was relatively well-educated, with the majority having completed a Bachelor's degree (N=121) or Master's degree or higher (N=45).
-By comparison, fewer participants endorsed having completed only some college (N=100), only a high school degree (N=33), or preferring not to say (N=1).
+By comparison, fewer participants endorsed having completed only some college (N=100), only a high school degree (N=33), or preferred not to say (N=1).
 
-The distribution of scores on the three MaRs-IB short-forms are summarized in Table \ref{table:2}.
-Consistent with model predictions, participants on average responded correctly on 8.0 of 12 items (sd = 2.5, IQR = 6 - 10).
-Similarly, the distribution of total scores were right-shifted \ref{fig:fig03} as would be expected given the nominal guessing rate of 25\%.
-A one-way ANOVA comparing the total scores across the three short-forms did not reject the null hypothesis (F(2,297) = 0.253, p = 0.777).
-Finally, the marginal score reliability for each short form was satisfactory as predicted (Table \ref{table:2}).
-In summary, the distribution of scores across the MaRs-IB short forms were consistent with the predictions based on the test assembly, indicating that the psychometric properties of these items were well calibrated in the previous study.
+\subsubsection{Procedure}
 
-The distribution of scores on the abbreviated Raven's progressive matrices is also summarized in Table \ref{table:2}.
-On average, participants responded correctly on 4.5 of 9 items (sd = 2.0, IQR = 3 - 9).
-Scores on the RPM were significantly lower on the RPM than the MaRs-IB ($t = 25.598, p < 1e^{-6}$), indicating either that the RPM items are more difficult or less susceptible to guessing (consistent with these items having 6 or 8 distractors).
-The marginal score reliability for the RPM was similar to that of the MaRs-IB forms (Table \ref{table:2}).
+The study procedure was divided into three sections.
+After providing consent, participants first completed three short surveys: the 10-item need for cognition survey \citep{chiesi2018applying}, the 8-item PROMIS Cognitive Function-Short Form (8a) \citep{iverson2021normative}, and the 8-item subject numeracy scale \citep{fagerlin2007measuring}.
+These measures were included as part of exploratory analyses that are reported in the Supplement.
+Afterwards participants completed one of the three 12-item MaRs-IB short-form measures described above.
+The administration procedure of the short-forms was identical to that in the first study.
+Finally, participants completed the 9-item abbreviated Raven's progressive matrices (RPM; form A) \citep{bilker2012development}.
+The presentation format and instructions for the RPM items was the same as for the MaRs-IB items.
+
+\subsubsection{Exclusion criteria}
+
+To ensure data quality, the data from multiple participants were excluded prior to analysis for one or more of the following reasons: failing to complete all three sections of the experiment (N=17); failing one or more attention checks \citep{zorowitz2021inattentive} embedded in the self-report measures (N=21); experiencing technical difficulties during the experiment (N=11); rapid guessing on four or more items (N=3); or failing to respond on four or more items (N=1).
+In total, 47 of 347 (13.5\%) participants were excluded leaving the data from N=300 participants for analysis.
+
+\subsubsection{Analysis}
+
+To measure the convergent validity between the MaRs-IB and RPM measures, we fit an additional explanatory item response model.
+Here participants' ability on the MaRs-IB short forms were estimated as a function of their gender, age, and observed score on the RPM (Eq \ref{eq:4}).
+The probability of a participant correctly solving a MaRs-IB was defined according to the 3PL model (Eq \ref{eq:1}), where the difficulty and discrimination parameters for an item were fixed to the posterior mean of the estimates obtained in the previous study.
+Thus, the model estimates the partial correlation between RPM performance and MaRs-IB ability controlling for the age and gender of the participants.
+
+The IRT model was estimated within a Bayesian framework using Hamiltonian Monte Carlo as implemented in Stan (v2.22) \citep{carpenter2017stan}.
+Four separate chains with randomised start values each took 7500 samples from the posterior.
+The first 5,000 samples from each chain were discarded.
+As such, 10,000 post-warmup samples from the joint posterior were retained.
+The $\hat{R}$ values for all parameters were equal to or less than 1.01, indicating acceptable convergence between chains, and there were no divergent transitions in any chain.
+
+We calculated the ordinal reliability of the MaRs-IB and RPM short-form measures using the lavaan \citep{lavaan} and semTools \citep{semtools} packages available in R.
+
+\subsection{Results}
+
+Performance on the MaRs-IB short-forms is summarized in Table \ref{table:3}.
+On average, participants completed the measure in 174 s (sd = 54.0 s) and responded correctly on 8.0 of 12 items (sd = 2.5).
+The distribution of observed scores was right-shifted (Figure~\ref{fig:fig03}), as was expected given the nominal guessing rate.
+A one-way ANOVA comparing the average score across the three short-forms was not statistically significant (F(2,297) = 0.253, p = 0.777).
+Importantly, the ordinal reliability of each measure was satisfactory (all $\alpha > 0.75$).
+In sum, performance on the MaRs-IB short form were consistent with predicted made during test assembly, indicating that the psychometric properties of the items were adequately calibrated during in the previous study.
+
+Performance on the abbreviated RPM measure is also summarized in Table \ref{table:3}.
+On average, participants completed the measure in 125 s (sd = 40.1 s) and responded correctly on 4.5 of 9 items (sd = 2.0).
+The proportion of correct responses on on the RPM was significantly lower than that observed for the MaRs-IB ($t = 13.028$, $p < 0.001$), suggesting that the RPM is more difficult or less biased by guessing.\footnote{In contrast to the MaRs-IB, items in the RPM have either six or eight response options.
+As such, the nominal guessing rate is expected to be lower for the RPM.} The ordinal reliability for the RPM was also adequate.
+
+The partial correlation between observed scores on the RPM and ability on the MaRs-IB was $\rho = 0.527$ (95\% HDI = 0.457 - 0.577).
+Considering the attenuation of correlations between two imperfect measures \citep{spearman1961proof}, we should not expect correlations much larger than $\rho = 0.64$ (or $0.80^2$).
+As such, this result indicates good convergent validity between the two matrix reasoning measures.
+Briefly, in this smaller sample, neither age nor gender was credibly associated with performance on the MaRs-IB short forms (age: $\rho$ = -0.123, 95\% HDI = -0.253 -- 0.013; gender: $\rho$ = -0.025, 95\% HDI = -0.160 -- 0.108).
+
+To summarize this section, we used the item parameters estimated as part of the calibration study, and optimal assembly methods, to construct three new MaRs-IB short form measures.
+As predicted, these measures were quick to administer, had adequate reliability, and good convergent validity with a second matrix reasoning measure.
+
+\section{General Discussion}
+
+Here we have provided the most comprehensive interrogation of of the MaRs-IB to date.
+Using item structure models fit to data collected from a large sample of adult participants, we found that the MaRs-IB possesses desirable psychometric properties.
+The items in the MaRs-IB span a considerable range of difficulty and exhibit medium-to-high levels of discrimination.
+As such the MaRs-IB is suitable for measuring matrix reasoning across the ability spectrum in the adults.
+We also verified that the design of the MaRs-IB was in part successful: as intended, item attributes such as feature number and rule number were positively associated with, and explained much of the variability, in item difficulty.
+Contrary to our expectations, however, we also found that the design of the MaRs-IB had unintended consequences: distractor type was also associated with item difficulty.
+Moreover, we found non-negligible variability in the difficulty of item clones even after controlling for distractor type.
+Crucially, item clones in the MaRs-IB therefore cannot be treated as exchangeable.
+
+With the estimated item parameters, we then constructed three new MaRs-IB short-form measures using optimal test assembly methods.
+These abbreviated test measures were designed to measure matrix reasoning ability with maximal reliability in the general adult population under several desirable constraints (i.e.
+12 item limit, administration time of 2-4 minutes).
+In a second sample of participants, we found that each short form measure had adequate reliability (Cronbach $\alpha \approx 0.8$) and good convergent validity with an abbreviated Raven's progressive matrices measure.
+Moreover, the observed scores on the short-forms followed the expected distribution as predicted by their test characteristic curves.
+These results demonstrate the success of the item calibration study in producing accurate estimates of the item parameters.
+
+Of additional note are the associations we found between accuracy and response time.
+The best-performing participants spent more time on each item, consistent with a speed-accuracy trade-off account of performance \citep{heitz2014speed}.
+Interestingly, we also found an interaction between participant ability and item difficulty; that is, the best-performing participants increased their deliberation time as items become more difficult, whereas the worst-performing participants maintained approximately the same work rate.
+These results are consistent with a persistence account of participant ability, wherein performance reflects in part a willingness to invest time in the solution process \citep{ranger2021effects}.
+As such, the MaRs-IB may be suitable not only for measuring matrix reasoning ability but also mental effort costs [cite], opportunity costs [cite], or other  motivational factors \citep{duckworth2011role} related to the tendency to give up (at least in low-stakes test settings; see \cite{shaw2020reasoning} for a counter-example).
+Moreover, the use of cognitive models may help to disentangle the relative contributions of these factors to performance on the MaRs-IB \citep{ranger2014accumulator}.
+
+The current investigation of the MaRs-IB is not without its own limitations.
+One notable limitation is our sample.
+Here we analyzed response data collected from an online, adult sample that was relatively young and well-educated.
+We cannot guarantee that the psychometric properties of the MaRs-IB reported here will generalize to other populations or testing contexts.
+Future researchers may consider replicating the current study in other populations of interest (e.g.
+children, clinical samples).
+By using item response models here, however, we make possible the opportunity for future IRT ``linking'' studies.
+IRT linking describes a set of methods to establish the comparability of item and ability parameters estimated from data collected from two or more groups that differ in ability \citep{lee2018irt}.
+Future studies employing these methods could provide insight not only in how the functioning of the MaRs-IB may change in different populations, but also how matrix reasoning ability differs across samples.
+
+A second limitation is in our investigation of sources of variance.
+Here we only investigated the contributions of three item attributes.
+There are far more sources to have investigated (e.g.
+type of rule change such as color change versus shape change).
+Indeed, there are other item features taxonomies we could have explored \citep{carpenter1990one}.
+We also did not explicitly model perceptual features as has been done before.
+Although doing so would not change our results about exchangeability of item clones, it would allow for more precise estimates of item parameters and improved test assembly.
+Also allows for possibility of automatic item generation.
+Room for future research.
+
+In summary, we have provided an in-depth psychometric validation of the MaRs-IB, finding that it is suitable for measuring matrix reasoning ability in the general adult population.
+We release the item parameters so that researchers interested in measuring matrix reasoning can construct maximally reliable tests for their own research.
+
+\bibliography{references}
+
+\begin{table}[]
+\centering
+\begin{tabular*}{\textwidth}{lc@{\hskip 6mm}r@{\hskip 6mm}c}
+\toprule
+Predictor &  Coef. (SE) & Z-score (p-value) & 95\% CI  \\
+\midrule
+Accuracy & 0.039 (0.007) &   5.367 (<0.001) &  0.025 -- 0.053  \\
+Score &  0.167 (0.009) &  19.204 (<0.001) &  0.150 -- 0.184  \\
+Difficulty & 0.073 (0.005) &  13.292 (<0.001) &  0.062 -- 0.083  \\
+Accuracy x Score & -0.064 (0.007) &  -8.764 (<0.001) & -0.078 -- -0.050  \\
+Accuracy x Difficulty & 0.097 (0.007) &  14.351 (<0.001) &  0.084 -- 0.110  \\
+Score x Difficulty &  0.032 (0.005) &   6.549 (<0.001) &  0.023 -- 0.042  \\
+Accuracy x Score x Difficulty &  0.012 (0.006) & 1.955 (0.051) & -0.000 -- 0.025  \\
+\bottomrule
+\end{tabular*}
+\caption{\label{table:1}\normalfont Summary of the random-intercepts linear regression model predicting log-transformed response time as a function of accuracy, participant rest score, and item difficulty.}
+\end{table}
+
+\begin{table}
+    \centering
+    \begin{tabular*}{\textwidth}{cccccccc}
+    \toprule
+    Model & FE-1 & RE-1 & FE-2 & RE-2 & psis-loco & $\Delta$ psis-loco (se) \\
+    \midrule
+    1a & x &   &   &   & -14516.85 & -1223.69 (44.94)\\
+    1b & x & x &   &   & -13804.74 & -511.58 (28.77) \\
+    1c & x & x & x &   & -13481.12 & -187.96 (17.25) \\
+    1d & x & x & x & x & -13293.16 & - \\
+    \midrule
+    2a & x &   & x &   & -13209.38 & -2.13 (2.04) \\
+    2b & x & x & x &   & -13207.25 & - \\
+    2c & x & x & x & x & -13208.69 & 1.44 (0.85) \\
+    \bottomrule
+    \end{tabular*}
+    \caption{\label{tab:2} \normalfont Goodness of fit of item response models to MaRs-IB data. Pareto-smoothed importance sampling}
+    \label{table:2}
+\end{table}
+
+\begin{table}
+    \centering
+    \begin{tabular*}{\textwidth}{lccccc}
+    \toprule
+    Measure & Sample & Task Time (sd) & Mean Score (sd) & IQR & Reliability ($\alpha$) \\
+    \midrule
+    MaRs-IB SF-1 & 103 & 174.9 s (53.3) & 7.9 (2.4) & 7 - 10 & 0.761 \\
+    MaRs-IB SF-2 & 98  & 169.0 s (53.0) & 8.2 (2.6) & 7 - 10 & 0.810 \\
+    MaRs-IB SF-3 & 99  & 178.5 s (55.7) & 7.9 (2.7) & 6 - 10 & 0.829 \\
+    RPM SF-A     & 300 & 124.6 s (40.1) & 4.5 (2.0) & 3 - 9 & 0.791 \\
+    \bottomrule
+    \end{tabular*}
+    \caption{\label{table:3} \normalfont Summary of performance on the MaRs-IB and Raven's progressive matrices (RPM) short form (SF) measures. IQR = interquartile range.}
+\end{table}
 
@@
 \includegraphics[width=1.0\textwidth]{figures/fig03.png}
-\caption{\label{fig:fig03}This is a figure caption.}
+\caption{\label{fig:fig03}The joint distribution of total scores on the MaRs-IB short form and abbreviated Raven's progressive matrices (RPM) measures.
+The distribution of observed scores for each measure are shown in the margins.}
 \end{figure}
 
-\subsection{Convergent validity}
-
-To measure convergent validity, blah blah blah. Need to write model here. Or should we write above?
-
-\subsection{Discussion}
-
-\section{General Discussion}
-
-Paragraph XX: item discrimination random effects. Power analysis? also variance explained being similar to literature.
-
------
-Notes on RT effects: this paragraph should come later, as a discussion point on what we're actually measuring with the MaRs-IB (at least online)
-
-Heitz (2014): in general, an increase in speed comes at the expense of accuracy (speed-accuracy trade-off)
-
-Goldhammer (2015):
-
-> Individual differences in test performance depend on both between-person differences in speed and ability (i.e., the location of individual speed-ability curves) and the adopted speed-ability compromise within persons (i.e., the location within a speed-ability curve).
-T
-
-Goldhammer et al. (2015): test takers that increase their work pace have a lower effective ability
-
-> On previous reasoning tasks: "These findings indicate that in tasks that require understanding new problem situations and in which the use of complex cognitive processes is mandatory, higher effective ability is usually associated with lower speed."
-
-Ranger et al. (2021):
-
-> Such a relation between response time and response accuracy or between work pace and effective ability is of uttermost importance as it implies that the responses depend on more than just the test taker’s capability.
-
-> the capability is apportioned among two quantities, the effective ability, and the work pace, which constitute the actual performance of a test taker in a test; see Pohl and von Davier (2018) for a discussion on how to describe the performance of test takers.
-
-As work pace increases, effective ability decreases (irrespective of true capability).
-Work rate may reflect other processes: persistence, or amount of time before participant gives up, mental effort costs, or foraging costs.
-
-Basically, summarize results.
-All points are consistent with dual process theory (automatic vs.
-controlled processing).
-suggestins higher ability participants more deliberate, lower ability participants less deliberate.
-frame from perspective of persistence (ranger 2021).
-
-Need to cite: https://doi.org/10.1016/j.intell.2020.101490
-
------
-
-Paragraph XX: future directions modeling speed-accuracy trade-offs
-
-Paragraph XX: model more features as future direction.
-Carpenter taxonomy.
-Model by rule type.
-By perceptual features.
-automatic item generation as future direction.
-See \cite{lathrop2017item} for inspiration.
-Note that this definition of item rules deviates from other established taxonomies of matrix reasoning items \citep{carpenter1990one}, a point we return to in the Discussion.
-
-\bibliography{references}
-
-\begin{table}
-    \centering
-    \begin{tabular}{c|c|c|c|c|c|c|c}
-    \hline
-    Model & FE-1 & RE-1 & FE-2 & RE-2 & psis-loco & $\Delta$ psis-loco (se) \\
-    \hline
-    1a & x &   &   &   & -14516.85 & -1223.69 (44.94)\\
-    1b & x & x &   &   & -13804.74 & -511.58 (28.77) \\
-    1c & x & x & x &   & -13481.12 & -187.96 (17.25) \\
-    1d & x & x & x & x & -13293.16 & - \\
-    \hline
-    2a & x &   & x &   & -13209.38 & -2.13 (2.04) \\
-    2b & x & x & x &   & -13207.25 & - \\
-    2c & x & x & x & x & -13208.69 & 1.44 (0.85) \\
-    \hline
-    \end{tabular}
-    \caption{\label{tab:1} Goodness of fit of item response models to MaRs-IB data. Pareto-smoothed importance sampling}
-    \label{table:1}
-\end{table}
-
-\begin{table}
-    \centering
-    \begin{tabular}{c|c|c|c}
-    Measure & Mean (sd) & IQR & Reliability \\
-    \hline
-    MaRs-IB SF-1 & 7.9 (2.4) & 7 - 10 & 0.761 \\
-    MaRs-IB SF-2 & 8.2 (2.6) & 7 - 10 & 0.810 \\
-    MaRs-IB SF-3 & 7.9 (2.7) & 6 - 10 & 0.829 \\
-    RPM SF-A & 4.5 (2.0) & 3 - 9 & 0.791 \\
-    \end{tabular}
-    \caption{Caption}
-    \label{table:2}
-\end{table}
-
 \end{document}

### 2022-04-19 manuscript/main.tex
updated manuscript

@@
 
-Matrix reasoning tasks are among the most well-studied and useful measures in the behavioral sciences.
-Matrix reasoning requires participants to use abstract reasoning to solve puzzles.
-Matrix reasoning indexes fluid intelligence and is correlated with working memory capacity \citep{conway2002latent, kane2004generality, unsworth2005working, salthouse2014relations}.
-In low-stakes testing environments, performance on the progressive matrices is associated with participant motivation and willingness to expend effort \citep{gignac2018moderate, gignac2019maximum}.
-The progressive matrices has excellent external validity, and is predictive of academic performance \citep{roth2015intelligence} and scores on college admission exams \citep{frey2004scholastic, koenig2008act}.
-% Thus, matrix reasoning tasks are increasingly utilized as control measure in individual difference behavioral studies \citep{gillan2016characterizing, rouault2018psychiatric, moutoussis2021decision}.
+Matrix reasoning tasks are among the most widely used measures of cognitive ability in the behavioral sciences.
+Much of their popularity undoubtedly reflects their versatility in use.
+Matrix reasoning tasks are strong (albeit impure) indicators of general intelligence \citep{gignac2015raven} and working memory capacity \citep{kane2004generality, unsworth2005working}.
+They are predictive of important real-world outcomes such as childhood academic performance \citep{roth2015intelligence} and performance on college entrance exams \citep{frey2004scholastic, koenig2008act}.
+In low-stakes testing settings, matrix reasoning tasks can additionally function as measures of motivation, willingness to expend mental effort, and other aspects of personality \citep{duckworth2011role, gignac2019maximum}.
+Performance on matrix reasoning tasks can also serve as an instrumental variable in studies of psychiatric populations; that is, it can be used to control for general disruptions to cognitive ability where specific domains of cognition are primarily of interest \citep{gillan2016characterizing, rouault2018psychiatric, moutoussis2021decision}.
 
-Unfortunately there are a number of challenges in using matrix reasoning tasks as part of behavioral research.
-One challenge is the problem of copyright.
-The most prominent matrix reasoning tasks, such as the Raven's progressive matrices \citep{raven2003raven} and the WAIS matrix subtest \citep{wechsler1999wechsler}, are typically not free to use and have restrictions against being computerized.
-A second challenge is that, across the measures in the public domain, there are relatively few items available.
-For example, the Hagen Matrices Test \citep{heydasch2014hagen} and ICAR matrix reasoning test \citep{condon2014international} have only 20 and 16 items, respectively.
-The availability of only a limited number of items raises the possibility of practice effects \citep{ng1974applicability, bors2003effect}, especially in the age of online experiments where multiple groups of researchers may be unwittingly administering the same test to many of the same participants.
+Unfortunately there are a number of impediments to using matrix reasoning tasks as part of behavioral research.
+One such challenge is the problem of copyright.
+The most prominent examples of matrix reasoning tasks, such as Raven's progressive matrices \citep{raven2003raven} and the WAIS matrix subtest \citep{wechsler1999wechsler}, are not free to use and have legal restrictions against being computerized.
+A second challenge is that, across all of the matrix reasoning tests in the public domain, there are relatively few unique items available.
+The Hagen matrices test \citep{heydasch2014hagen} and ICAR matrix reasoning test \citep{condon2014international}, for example, have only 20 and 16 items respectively.
+The availability of only a limited number of items raises the possibility of repeat exposure effects, which threaten the validity of these measures  \citep{ng1974applicability, bors2003effect}.
+This problem is exacerbated in the current age of online experiments, where multiple groups of researchers may be unintentionally administering the same test to the same participants recruited from the same online labor platforms.
 
-In response to these challenges, \cite{chierchia2019matrix} developed, and released into the public domain, the matrix reasoning item bank (MaRs-IB).
-The MaRs-IB includes 80 matrix reasoning puzzles.
-Each puzzle consists of a 3x3 matrix containing abstract shapes in eight out of nine cells.
-To solve a puzzle, participants must deduce which of four response options completes the matrix based on the pattern of changing relations among the shapes across cells (Figure 1).
-The puzzles in the MaRs-IB vary considerably in their complexity, both in the number of shapes per cell and in the number of changing relations across shapes.
-Importantly each template puzzle has six clones, which are matched in terms of complexity but differ in their perceptual properties (one of three shape sets) and their response options (one of two distractor sets).
-Thus, the MaRs-IB offer considerable utility insofar that they are free to use, have many items, and many unique items to prevent practice effects.
+Noting these challenges, \cite{chierchia2019matrix} developed and made publicly available the matrix reasoning item bank (MaRs-IB), a collection of 80 unique matrix reasoning puzzles.
+Each puzzle in the MaRs-IB consists of a 3x3 matrix containing abstract shapes in eight out of nine cells (Figure \ref{fig:fig00}).
+To solve a puzzle, participants must deduce which of four response options correctly completes the matrix based on the pattern of changing relations among the shapes across cells.
+The puzzles in the MaRs-IB were designed to span a considerable range of complexity, both in the number of unique shapes and changing relations in and across cells.
+To further increase the reusability of the MaRs-IB and minimize practice effects, each puzzle (hereafter referred to as a template item) has six clones.
+An item's clones share identical structural properties (i.e.
+number of shapes, number of changing relations) but may vary in their distractors (two unique sets per item) or perceptual features (three unique shape sets per item).
+Thus, the MaRs-IB addresses several issues associated with other matrix reasoning tests.
 
-A preliminary investigation of the psychometric properties of the MaRs-IB items proved promising.
-Indeed, \cite{chierchia2019matrix} found that scores on a short-form matrix reasoning test using MaRs-IB items had good split-half reliability and test-retest reliability.
-The authors also found that performance observed scores on this test were moderately correlated with a digit span task and the ICAR matrix reasoning test, indicating good convergent validity.
-% note the release of summary statistics here?
+\cite{chierchia2019matrix} also conducted an initial study of the psychometric properties of the MaRs-IB.
+A sample of 659 participants completed a fixed-length and fixed-order test measure in which they had eight minutes to complete as many MaRs-IB items as possible.
+The authors found that performance on this test measure had good split-half and test-retest reliability.
+Moreover they found that participants' ability to solve puzzles in the MaRs-IB was moderately predictive of their performance on a digit span task and the ICAR matrix reasoning test, indicating good convergent validity.
+At the conclusion of this study, \cite{chierchia2019matrix} publicly released a summary of the difficulty (i.e.
+proportion correct responses) of each item in the MaRs-IB so that future researchers could construct new test measures of custom difficulty and duration.
 
-Unfortunately, there are two notable limitations in this initial investigation.
-The first is that, due to the design of the matrix reasoning test, there is little data available for the majority of the items in the MaRs-IB.
-Just over half the items (42/80) were completed by 100 or more participants.
-As such, there is too little data to characterize the psychometric properties of each item with reasonable confidence (an issue that is exacerbated when considering each template item has six variants).
-A second issue is that, also due to the design of the matrix reasoning test, participants were allowed to choose for themselves whether to prioritize accuracy (number of correct responses) or productivity (number of items reached).
-As a consequence, items later in the sequence are likely to be contaminated by participants sacrificing accuracy for speed, thereby distorting the summary statistics for these items.\footnote{appendix: Indeed, looking only at the easiest MaRs-IB items (dimension = 1), there is a strong positive correlation between number of completing participants and mean accuracy ($r = 0.673$).} Thus, caution is warranted in interpreting the results of the original study and additional validation is warranted.
+Unfortunately, there are two critical limitations in this study that undermine the utility of the summary statistics.
+The first is that, given design of the test, there was relatively little data collected for the majority of the items in the MaRs-IB.
+Indeed, only half (42 of 80 items) were completed by 100 or more participants.
+As such, one cannot be confident about the reported functioning of a majority of the items (an issue that is exacerbated when considering that each item has six clones).
+A second issue is participants were allowed to choose for themselves whether to prioritize accuracy (number of correct responses) or productivity (number of items reached).
+As a consequence, the response data for items that came later in the fixed-order test were increasingly likely to be produced by participants sacrificing accuracy for speed, thereby biasing the proportion correct measure for these items (see the Supplementary materials for evidence of this point).
+Therefore, additional investigation of the functioning and psychometric properties of the MaRs-IB is warranted.
 
-%Better intro sentence tk: mention that this is a more extensive psychometric validation of the MaRs-IB in a much larger dataset, with three major improvements.
-The most important contribution of the current work is the evaluation of the psychometric properties of (almost) every item in the MaRs-IB using item response theory (IRT) \citep{embretson2013item, de2013theory}.
-IRT models confer several important benefits for the purposes of psychometric validation.
-First they provide an estimate of the difficulty and discrimination of each item, where the latter is an index of how informative or reliable an item is.
-Through these parameters, IRT models allows for the computation of the test information function (TIF), which measures how sensitive a test is at measuring performance at varying levels of ability.
-Finally, IRT make possible optimal test assembly \citep{van1998optimal}, or design of new tests with maximal sensitivity to measuring performance in a specified range of abilities under varying constraints (e.g.
-test length, time limits, parallel forms with equal TIF).
+Here we provide a more extensive study and psychometric validation of the MaRs-IB.
+The most important contribution of the current work is the characterization of (almost) every item in the MaRs-IB using item response theory (IRT) \citep{embretson2013item, de2013theory}.
+IRT models confer several important advantages for the purposes of psychometric validation.
+First they provide an estimate of the difficulty and discrimination of each item, where the latter is an index of how much information an item provides given a level of ability.
+Using these quantities, IRT models in turn facilitate the computation of the test information function (TIF), which measures the reliability of a test measure across a range of ability.
+Finally, IRT models make possible optimal test assembly  \citep{van1998optimal}, or the design of new tests with maximal reliability given researcher-specified constraints (e.g.
+test duration, test difficulty).
 
-A second contribution offered here is an analysis of the relationships between attributes of MaRs-IB items and their psychometric properties using explanatory item response models \citep{de2004explanatory, wilson2008explanatory}.
-Explanatory IRT models provide a framework for measuring how item parameters (e.g.
-difficulty) are shaped by their features.
-As such, these models provide means to evaluate how successful was the generation of the items in the MaRs-IB.
-For example, we can quantify the extent to which features of item complexity (e.g.
-number of elements, number of changing relations) predict item difficulty.
-Moreover, we can determine whether two items that vary only in their distractor set are psychometrically equivalent.
-Explanatory IRT models have the additional potential benefit of producing more accurate estimates of item parameters \citep{neuhaus2006separating}.
+A second contribution provided here is an investigation of the relationship between the attributes of MaRs-IB items and their psychometric properties using explanatory item response models \citep{de2004explanatory, wilson2008explanatory}.
+The utility of the MaRs-IB depends in part on its items exhibiting certain expected characteristics.
+Across items we should expect complexity (e.g.
+more shapes per cell, more changing relations) to be positively associated with difficulty.
+Moreover if item clones are exchangeable (i.e.
+perfect replicas with identical psychometric properties), then we should expect no association between distractor type and difficulty.
+Thus explanatory IRT models offer an opportunity to evaluate the success of the MaRs-IB design.
+As a secondary benefit, explanatory IRT models also provide more accurate estimates of item parameters \citep{neuhaus2006separating}.
 
-%item clones would be isomorphic (psychometrically equivalent).
-In any realistic situation, the 2P-F is misspecified as it ignores the psychometric deviations introduced in the item cloning process (Glas & van der Linden, 2003).
+A third contribution is the decomposition of variance in item functioning, into modeled and unmodeled sources, using additive multilevel item structure (AMIS) models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
+First we use AMIS models to quantify the residual variance in difficulty and discrimination across items after accounting for their respective attributes.
+At the item-level, smaller residual variances indicate that items are functioning as intended.
+We also use AMIS models to quantify the variance in difficulty and discrimination across item clones.
+If item clones are to be treated as exchangeable (i.e.
+able to be used interchangeably), then the residual variance in their item parameters must be negligible.
+AMIS models therefore provide a second means by which to evaluate the success of the MaRs-IB design.
 
-A third contribution is the decomposition of variance in item functioning from modeled and unmodeled sources using additive multilevel item structure (AMIS) models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
-This framework provides another means by which to evaluate the success of the design of the MaRs-IB by quantifying the residual variance in item parameters after accounting for item attributes.
-For example, two items of equal complexity may turn out not be psychometrically equivalent if unmodeled attributes (e.g.
-the types of changing relations) implicitly impact item functioning.
-Similarly, two item clones may not be psychometrically equivalent due to differences in their superficial properties.
-Quantifying the degree of error in the item cloning process is critical, as the degree of nonequivalent dictates whether clones are exchangeable.
-That is, if clones are not psychometrically equivalent, then they cannot be used interchangeably in parallel short forms.
+The remainder of the paper proceeds as follows.
+First we report a calibration study, in which a large sample of participants each completed a subset of items from the MaRs-IB.
+Using their response data, we then fit a series of multilevel item structure models in order to estimate item parameters for each item (and their clones).
+Using these same models, we also interrogate the relationship between item attributes and functioning within and cross items in the MaRs-IB.
+Then we report a second validation study, in which a smaller sample of participants completed one of three new MaRs-IB short-form tests.
+These tests --- designed using optimal test assembly methods applied to the item parameters estimated during the calibration study --- were found to have good psychometric properties and convergent validity.
+This second study provides a blueprint for how to construct new MaRs-IB test measures using the item parameter estimates that have been released as part of this study.
 
-The paper proceeds as follows.
-First we analyze participants' response data hierarchical item response models, which allowed us to measure all items' difficulty and discrimination while controlling for participants' varying abilities.
-The item response models also allowed us to measure directly the extent to which that item variants are truly equivalent.
-We then demonstrate how these item response parameters can be used to construct new short-forms with desirable properties.
-Using these methods, we introduce three new MaRs-IB short-form tests with good internal consistency and convergent validity.
+All data, code, and model outputs (including the item parameter estimtes) are publicly available at: \url{https://github.com/dawlab/mars-irt}.
 
-All data and code, including items' estimated parameters and example code for test assembly, is available at: \url{https://github.com/szorowi1/mars-irt}.
-
+\clearpage
 \section{Calibration Study}
@@
 
-The purpose of the first study was to calibrate the items of the MaRs-IB, using response data collected from a large number of adult participants from the general population.
+The purpose of the first study was to calibrate the items of the MaRs-IB using response data collected from a large number of adult participants from the general population.
 In particular, we sought to accomplish the following three objectives: (1) to characterize the psychometric properties of the items in the MaRs-IB using item response models; (2) to investigate the associations between item attributes and their psychometric properties, and (3) to determine the equivalence of items with similar complexities, and clones of the same item templates.
@@
 To ensure data quality, the data from multiple participants were excluded prior to analysis (see Exclusion Criteria below) leaving the data from N=1501 participants for analysis.
-The majority of participants in this sample identified as women (men: N=670; women: N=811; non-binary or other: N=13; rather not say: N=2) and were 28.7 years old (sd = 9.9) on average.
-The sample was relatively well-educated, with the majority having completed a Bachelor's degree (N=507) or Master's degree or higher (N=322).
-By comparison, fewer participants endorsed having completed only some college (N=471), only a high school degree (N=199), or preferred not to say (N=2).
+In these participants, the majority identified as women (men: N=670; women: N=811; non-binary or other: N=13; rather not say: N=2) and the average age was 28.7 years old (sd = 9.9).
+The sample was relatively well-educated with the majority having completed a Bachelor's degree (N=507) or Master's degree or higher (N=322).
+Comparatively fewer participants completed only some college (N=471), only a high school degree (N=199), or preferred not to say (N=2).
 
@@
 Eight of the nine resulting cells contain abstract shapes, while one cell on the bottom right-hand side of the matrix is empty.
-Participants are instructed to ``complete the matrix'' by finding the missing shape from among four possible alternatives.
+Participants are instructed to ``complete the matrix'' by identifying the missing shape from among four possible alternatives.
 
-Each trial of the experiment began with a fixation cross, lasting  1200 ms, after which a MaRs-IB item was presented.
-Participants were then given up to 30 s to solve the puzzle.
-After 25 s a clock appeared, indicating that 5 s remained before presentation of the next item.
+The presentation of each item was preceded by a fixation cross, which lasted for 1200 ms.
+Upon presentation of the item, participants were given up to 30 s to solve the puzzle.
+After 25 s elapsed, a clock appeared to count down the remaining 5 s.
 A trial ended when participants responded, or after 30 s had elapsed without response.
 Before the trials began, participants reviewed instructions and were made to correctly complete three practice items.
-Participants were told to respond carefully but to guess before the timer ran out.
-The format of the experiment and instructions were adapted from code made publicly available by the original authors (\url{https://app.gorilla.sc/openmaterials/36164}).
+Participants were instructed to respond carefully, but to guess if they could not solve the puzzle before the timer ran out.
+The format of the experiment and instructions were adapted from code publicly released as part of the original study (\url{https://app.gorilla.sc/openmaterials/36164}).
 
-In order to ensure sufficient and uniform sampling of the MaRs-IB items, all participants completed 16 items pseudorandomly selected from the 64 available MaRs-IB items.\footnote{We did not collect new data for the 16 easiest items from the MaRs-IB (items 1, 2, 3, 4, 5, 7, 8, 9, 32, 33, 38, 41, 43, 48, 57, 68) as the original study or pilot testing found performance on these items to be at or near ceiling.} The item sets were constructed as follows: First, we subdivided the 64 items into 16 sets of four based on their complexity.
-Participants were then randomly assigned one item from each of the 16 sets.
-In this way, all participants received test forms of approximately equal difficulty.
-We also counterbalanced which variant of each item was presented across participants, such that we had approximately the same number of responses available by shape set (1, 2, or 3) and distractor type (minimal difference or or paired difference; see \cite{chierchia2019matrix} for details).
+In order to ensure sufficient sampling of every item (and their clones) in the MaRs-IB, participants were administered a pseudorandomly-selected set of 16 out of 64 total items.\footnote{We did not collect new data for the 16 easiest items in the MaRs-IB (items 1, 2, 3, 4, 5, 7, 8, 9, 32, 33, 38, 41, 43, 48, 57, 68) because performance on these items was at or near ceiling.} Item sets were constructed as follows: We subdivided the 64 items into 16 sets of four based on their dimensionality (a measure of item complexity defined in \cite{chierchia2019matrix}).
+Participants were randomly assigned one item from each of the 16 sets.
+As a consequence, all participants received test forms of roughly equal difficulty.
+Importantly we also counterbalanced the assignment of item clones across participants, such that we had an approximately uniform number of responses available for each item by shape set (1, 2, or 3) and distractor type (minimal difference, MD, or or paired difference, PD; see \cite{chierchia2019matrix} for details).
 
@@
 
-The data from multiple participants who completed the experiment were excluded prior to analysis for one or more of the following reasons: rapid guessing (responding in less than 3 s) on four or more trials (N=70); failing to respond on four or more trials (N=8); or minimizing the browser window to a size smaller than the minimally required dimensions (N=7).
-In total 83 of 1584 (5.2\%) participants were excluded, leaving the data from N=1501 participants available for analysis.
-Among these participants, each item template (64 in total) was sampled approximately 375 times (sd = 12) and each item clone (384 in total) was sampled approximately 62 times (sd = 7).
+The data from multiple participants who completed the experiment were excluded prior to analysis for one or more of the following reasons: rapid guessing (responding in less than 3 s) on four or more trials (N=70); failing to respond on four or more trials (N=8); or minimizing the browser window to a dimension smaller than the minimal requirements (N=7).
+In total, 83 of 1584 (5.2\%) participants were excluded leaving the data from N=1501 participants available for analysis.
+Across these participants, each item (64 total) was administered approximately 375 times (sd = 12) and each item clone (384 total) was administered approximately 62 times (sd = 7).
 
@@
 We employed item response models to characterize the psychometric properties of each item.
-Specifically, we used additive multilevel item structure (AMIS) models with random residuals \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
-In item structure models, item clones are modeled as descending from a template item.
-The basis of all of these models was the three-parameter logistic (3PL) item response model, where the probability of a correct response ($y_{ijk} = 1$) for person $i$ on item clone $k$ belonging to item family $j$ is:
+Specifically, we used additive multilevel item structure (AMIS) models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
+In AMIS models, item parameters are defined according to a hierarchical structure in which clones descend from item templates.
+At the core of of all of the AMIS models used here is the three parameter logistic (3PL) item response model, where the probability of a correct response ($y_{ijk} = 1$) for person $i$ on item clone $k$ belonging to item template $j$ is:
 
@@
 \noindent where $\theta_i$ is the latent ability for person $i$, and $\beta_{jk}$, $\alpha_{jk}$, and $\gamma_{jk}$ are the difficulty, discrimination, and guessing parameters for item clone $k$ of item family $j$.
-Because guessing parameters are challenging to reliably estimate in the absence of very large amounts of data \citep{han2012fixing}, we fixed the guessing parameter for each item clone to the nominal guessing rate ($\gamma_{jk} = 0.25$).
+Because the estimates of guessing parameters are often unreliable in the absence of very large amounts of data \citep{han2012fixing}, we fixed the guessing parameter for every item clone to the nominal guessing rate ($\gamma_{jk} = 0.25$).
 
@@
 
-\noindent To elaborate, each item difficulty parameter was modeled as an intercept ($\mu_\beta$), reflecting the average difficulty across all items, and four additional components:
+\noindent such that each item difficulty parameter was modeled as an intercept ($\mu_\beta$), reflecting the average difficulty across all items, and four additional components:
 
@@
 
-\item $\sum_{n=1}^N Q_{jn} \delta_{\beta n}$: the effect of the template (level-1) attributes on item difficulty, where $Q_{jn}$ is the value of attribute $n$ for item template $j$, and $\delta_{\beta n}$ is the effect of attribute $n$.
+\item $\sum_{n=1}^N Q_{jn} \delta_{\beta n}$: the effect of the template (level 1) attributes on item difficulty, where $Q_{jn}$ is the value of attribute $n$ for item template $j$, and $\delta_{\beta n}$ is the effect of attribute $n$.
 
-\item $\epsilon_{\beta j}$: the template (level-1) residual, or residual variability in item difficulty of the item template unexplained by its attributes.
+\item $\epsilon_{\beta j}$: the template (level 1) residual, or residual variability in item difficulty of the item template unexplained by its attributes.
 
-\item $\sum_{m=1}^M R_{km} \delta_{\beta m}$: the effect of the clone (level-2) attributes on item difficulty, where $R_{km}$ is the value of attribute $m$ for item clone $k$, and $\delta_{\beta m}$ is the effect of attribute $m$.
+\item $\sum_{m=1}^M R_{km} \delta_{\beta m}$: the effect of the clone (level 2) attributes on item difficulty, where $R_{km}$ is the value of attribute $m$ for item clone $k$, and $\delta_{\beta m}$ is the effect of attribute $m$.
 
-\item $\epsilon_{\beta k}$: the clone (level-2) residual, or residual variability in item difficulty of the item clone unexplained by its attributes.
+\item $\epsilon_{\beta k}$: the clone (level 2) residual, or residual variability in item difficulty of the item clone unexplained by its attributes.
 
@@
 
-Similarly, item discrimination parameters were also modeled as the sum of an equivalent set of components:
+\noindent In turn, item discrimination parameters were expressed as the sum of an equivalent set of components:
 
@@
 
-The AMIS model framework provides a straightforward means of achieving Aims 2 and 3.
-Via the attribute effect components, we can test for any associations between item attributes and parameters (Aim 2).
-So too, via the error terms, we can quantify the magnitude of residual variance in item parameters from unmodeled sources both at the level of item templates and item clones (Aim 3).
-If the residual variance in item parameters across item clones is sufficiently large, then we can conclude item clones are neither psychometrically equivalent nor exchangeable.
-
-In the following analyses, we consider two template (level-1) attributes: feature number, or the number of unique shapes in each cell of an item; and rule number, or the number of changing relationships among the shapes (e.g.
-color change, shape change, position change) a participant must consider in order to solve the item.
-We predicted that both attributes would be positively related to item difficulty.
+In our analyses, we considered two template level (level 1) attributes: feature number, or the number of elements in each cell of an item; and rule number, or the number of changing relationships across the cells of an item (e.g.
+color change, shape change, position change).
+We predicted that both attributes would be positively associated with item difficulty; that is, all else equal, items with a great number of features or rules would be more difficult to solve.
 In contrast, we had no hypotheses about the relationship between these attributes and item discrimination.
 
-In turn, we consider two clone (level-2) attributes: distractor type, or MD-type distractors versus PD-type distractors; and mean response time, or the average time spent deliberating on that item.
-Based on the preliminary findings of \cite{chierchia2019matrix}, we predicted there would be no association between distractor type and either type of item parameter.
-We note that, in contrast to all of the other modeled attributes, average response time is not an intrinsic property of an item.
-We opted to include it in our models, however, because of its previously established relationship with item difficulty \citep{chierchia2019matrix}, and because including response time as a covariate has previously been shown to improve parameter estimation \citep{bertling2018using}.
-To prevent any collinearity between the attributes, mean response time was orthogonalized with respect to the three other item attributes, and therefore reflects any residual structure after accounting for all other modeled effects.\footnote{Note that we intentionally do not include shape set, or the superficial properties of an item, as covariates in our model.
-This is due to the would-be complexity of doing so.
-Though each item has three clones, the clones do not all have the same perceptual patterns.
-There are XX unique patterns.
-Future studies might consider even larger datasets to study perceptual influence.}
+We also considered two clone level (level 2) attributes: distractor type, or the use of MD-type versus PD-type response alternatives; and mean response time, or the average time participants spent deliberating on that item clone.
+We note that, in contrast to all other modeled attributes, average response time is not an intrinsic property of an item.
+We opted to include it in our models, however, because including response time as a covariate has previously been shown to improve parameter estimation \citep{bertling2018using}.
+To prevent possible collinearity between the attributes, mean response time was orthogonalized with respect to the three other item attributes; therefore, any effects of response time reflects residual structure after accounting for all other item attributes.
+Based on \cite{chierchia2019matrix}, we predicted there would be no association between distractor type and item difficulty, and a positive association between mean response time and item difficulty.
+As with the template-level attributes, we had no hypotheses about the relationship between the clone-level attributes and item discrimination.
 
-Finally, we estimated person-level abilities following an explanatory item response modeling approach \citep{wilson2008explanatory}:
+The AMIS model framework provides a natural means of accomplishing the three goals of the calibration the study.
+By fitting an AMIS model, we estimate the item parameters and quantify the psychometric properties for every item, and their clones, in the MaRs-IB (Aim 1).
+By modeling the effects of item attributes, we can evaluate the success of the MaRs-IB design by investigating whether item attributes influence item functioning as expected (e.g.
+a positive association between rule number and item difficulty; Aim 2).
+Finally, by estimating the magnitude of residual error in item parameters from unmodeled sources of variance (e.g.
+perceptual features), we can test whether the MaRs-IB possesses certain desirable structural properties (e.g.
+psychometric equivalence and exchangeability among item clones; Aim 3).
+
+We also estimated person-level abilities following an explanatory item response modeling approach \citep{wilson2008explanatory}:
 
@@
 \noindent where $X_{ip}$ is the value of attribute $p$ for person $i$, $\rho_p$ is the partial correlation of attribute $n$, and $\epsilon_i$ is variance in ability unexplained by the person-level attributes.
+
+We investigated person-level abilities as a function of four covariates: age, gender, average response time, and $\Delta$ response time.
+This last term reflects the degree of change in a participant's deliberation time as a function of item difficulty (defined as one minus the proportion of correct responses for that item).
+All covariates were standard-scored prior to analysis with the exception of gender, which was binary coded (male = -0.5, female = 0.5).
 
@@
 During estimation, the item discrimination parameters were restricted to be in the range $\alpha_{jk} \in [0, 5]$.
-
-To ensure the identifiability of all models, person abilities were constrained to have a mean of zero and a variance of 1.0, such that the residual variance was defined as $V(\epsilon) = 1 - \sum \rho_p^2$.
-Note that all person-level attributes were standard-scored prior being entered into the model.
+To ensure the identifiability of all models, person abilities were constrained to have a mean of zero and a variance of 1.0.
+In other words, the residual variance in ability was specified as $V(\epsilon) = 1 - \sum \rho_p^2$.
 
@@
 The goodness of fit of the models was compared using Bayesian leave-one out cross-validation \citep{vehtari2017practical}, which has been found to perform better than more traditional information criteria for comparing item response models \citep{luo2017performances}.
-Importantly, we computed the conditional leave-one-cluster out (LOCO) cross- validation \citep{merkle2019bayesian}, which measures a model's ability to generalize to held out items (rather than held out responses or held out participants) --- i.e., for inferring generalization to additional items constructed using the same features rather than to additional participants sampled from the same population.
+We computed the conditional leave-one-cluster out (LOCO) cross-validation \citep{merkle2019bayesian}, which measures a model's ability to generalize to held out items (rather than held out responses or held out participants) --- i.e., for inferring generalization to additional items constructed using the same features rather than to additional participants sampled from the same population.
 
@@
 
-The fit of the best-fitting model to the data was evaluated through posterior predictive model checks \citep{gelman1996posterior, levy2017bayesian}.
-Item fit was assessed using the Orlando and Thissen index ($S-\chi^2$) \citep{toribio2011discrepancy}, which compares the observed and expected proportions of participants who answer an item correctly at each raw-score group.
-The unidimensionality of the data was assessed using the generalized dimensionality discrepancy measure (GDDM) \citep{levy2011generalized}, which tests for violation of the assumption of local independence.
-This latter measure is crucial for ensuring there are negligible levels of local item dependence in the data, which may otherwise lower the efficiency of future test forms due to the inclusion of correlated, redundant items.
+The fit of the best-fitting model to the data was evaluated using posterior predictive model checking \citep{gelman1996posterior, levy2017bayesian}.
+A sample of predicted responses was generated for each sample of item parameters and the posterior predictive $p$ (PPP) value was computed based on two discrepancy statistics:  the Orlando and Thissen index ($S-\chi^2$) \citep{toribio2011discrepancy} and the standardized generalized dimensionality discrepancy measure (SGDDM) \citep{levy2015standardized}.
+
+The $S-\chi^2$ discrepancy statistic is a measure of item fit based on the comparison of observed and expected proportions of examinees who answer an item correctly at each raw-score group.
+A global measure of fit at the test level is obtained by summing the discrepancy values over the items.
+SGDDM is a measure of model dimensionality.
+More specifically, SGDDM may be interpreted as the mean absolute conditional correlation between all non-redundant pairs of items, which allows for the identification of inter-item correlations beyond those explained by the model.
+This latter measure is important as it tests for the presence of non-negligible response dependence across items, which may otherwise lower the efficiency of future test forms due to the inclusion of nonindependent or  redundant items.
 
@@
 
-On average, participants completed 9.6 of 16 items correctly (sd = 3.1, IQR = 8.0 - 12.0) (Figure \ref{fig:fig01}a).
+On average, participants completed 9.6 of 16 items correctly (sd = 3.1, IQR = 8.0 -- 12.0) (Figure \ref{fig:fig01}a).
 Timing out occurred on only 1.8\% of trials.
 A total of 43 participants (2.9\%) performed below chance (i.e.
-fewer than four items correct), while only 11 participants (0.7\%) correctly responded to all 16 items.
-Thus, over 95\% of participants' observed scores were within a reasonable range.
-These results corroborate \cite{chierchia2019matrix} in that there were no obvious ceiling or floor effects in performance, and that the majority of participants were sufficiently motivated to participate despite the low-stakes testing environment.
+fewer than four items correct), while only 11 participants (0.7\%) solved all 16 of their items.
+Thus, over 95\% of observed scores were in a reasonable range.
+These results corroborate those reported by \cite{chierchia2019matrix} in that there were no obvious ceiling or floor effects in performance, and that the majority of participants were sufficiently motivated to participate in spite of the low-stakes testing environment.
 
-Across all 384 unique item variants, the average proportion correct was 0.60 (sd = 0.19, IQR = 0.45 - 0.74) (Figure \ref{fig:fig01}b).
-A total of 12 items (3.1\%) had accuracy rates beneath chance levels, and only 22 items (5.7\%) had performance levels above 90\%.
-Mean performance on an item was significantly and negatively correlated with the number of changing relationships amongst its features (Spearman's $\rho = -0.482, p < 1e^{-6})$).
-Moreover, item-wise accuracy rates in this sample were remarkably similar to those reported by \cite{chierchia2019matrix} (Spearman's $\rho = 0.870, p < 1e^{-6})$), when restricting the latter to items with more than 200 available responses.
-These results further corroborate \cite{chierchia2019matrix} in that the majority of items exhibited good performance, with few items at ceiling or floor and passing of the manipulation check.
+Across all 384 item clones, the average proportion correct was 0.60 (sd = 0.19, IQR = 0.45 - 0.74; Figure \ref{fig:fig01}b).
+A total of 12 items (3.1\%) exhibited performance levels beneath chance, and only 22 items (5.7\%) had performance levels near ceiling ($\geq$ 90\%).
+Across items, the proportion of correct responses was negatively correlated with rule number ($\rho$ = -0.482, p < 0.001).
+These results further corroborate those reported by \cite{chierchia2019matrix} in that the MaRs-IB exhibited good functioning with performance levels for only a few item clones at ceiling or floor.
 
-We investigated the relationship between response accuracy and times using a random-intercepts linear regression model.
-Participants' log-transformed, item-level response times were predicted as a function of item accuracy, rest score (observed score for all other items), and item difficulty (one minus the proportion correct on that item).
-The results are summarized in Table XX.
+We investigated the relationship between accuracy and response time using a mixed effects (random intercepts) linear regression model.
+Participants' log-transformed response times were predicted as a function of that trial accuracy, rest score (participants' observed scores on all other items), and item difficulty (one minus the proportion correct for that item).
+The mixed-effects model was estimated using the \textit{statsmodels} python package (v0.12.2) \citep{seabold2010statsmodels}).
+
+The results are summarized in Table \ref{table:1}.
 On average, participants spent 15.9 s (sd = 7.2 s) on each item.
-There were significant main effects of accuracy ($\beta = 0.067$, $t = 9.84$, $p < 0.001$), rest score ($\beta = 0.125$, $t = 16.95$, $p < 0.001$), and item difficulty ($\beta = 0.137, t = 44.33, p < 0.001$) such that response times were slower for correct responses, better-performing participants, and more difficult items.
+There were significant main effects of accuracy ($\beta = 0.067$, $t = 9.84$, $p < 0.001$), rest score ($\beta = 0.125$, $t = 16.95$, $p < 0.001$), and item difficulty ($\beta = 0.137, t = 44.33, p < 0.001$).
+That is, response times were slower for correct responses, better-performing participants, and more difficult items.
 
-There was also a significant accuracy by difficulty interaction (correct responses were even slower for more difficult items), and a significant accuracy by score interaction (correct responses were faster for better performing participants).
-Finally, there was a significant difficulty by score interaction, such that better performing participants exhibited slower responses for more difficult items.
+There were also significant accuracy by difficulty and  accuracy by rest score interactions.
+Correct responses were even slower for more difficult items, but faster for better performing participants.
+There was also a significant difficulty by score interaction, such that better performing participants exhibited slower responses for more difficult items.
 This pattern of response time results is visible in Figure \ref{fig:fig01}c.
-The three-way interaction term was not statistically significant.
-
-\begin{figure}
-\centering
-\includegraphics[width=1.0\textwidth]{figures/fig01.png}
-\caption{\label{fig:fig01}Summary of performance on the MaRs-IB items.
-(A) The distribution of total scores across all 1501 participants.
-(B) The distribution of proportion correct responses across all 384 items.
-(C) The distribution of participants' median response times across items broken down by participants' total scores and item difficulty (one minus proportion correct responses).
-Error bars indicate bootstrapped 95\% confidence intervals around the mean.}
-\end{figure}
+The three-way interaction term was not significant.
 
@@
 
-We fit a series of nested IRT models in order to identify the model that best explained participants' response data.
-The models varied in how item parameters were estimated, with successive models allowing for greater variability in parameters across item templates and clones.
-Identifying the model that most accurately captures the structure of item parameters is important in the context of item cloning, as improperly modeled variance can result in the biased estimation of item parameters \citep{lathrop2017item}.
+We fit a series of nested AMIS models in order to identify the model that best predicted participants' response data.
+The models varied in how item parameters were specified, with each successive model allowing for greater flexibility in item parameter estimation.
+For the goals of current study, it is important to identify the model that most accurately reflects variability in item functioning, as insufficiently flexible models can result in the biased estimation of item parameters \citep{lathrop2017item}.
 
-We fit the models in two waves.
-The first wave was comprised of four models in which the item discrimination parameters were fixed ($\alpha = 1$).\footnote{Note that these models still included a fixed guessing parameter ($\gamma = 0.25$).
-We also estimated models with no guessing parameters (i.e.
-one-parameter logistic or Rasch models), but model comparison indicated these models were worse fits to the data (Table SXX).} We did this in order to minimize the number of total models that we would otherwise need to fit (i.e.
-all permutations of the four models below, applied to item difficulty and discrimination parameters independently).
-The first four models were as follows:
+We fit AMIS models in two waves.
+In the first wave of models, the item discrimination parameters were fixed ($\alpha = 1$).\footnote{Note that these models still included a fixed guessing parameter ($\gamma = 0.25$).
+We also estimated models with no guessing parameter (i.e.
+one-parameter logistic or Rasch models), but model comparison indicated these models provided worse fits to the data (Table SXX).} We applied this restriction to the first set of models in order to identify the best structure for the item difficulty parameters independent of the discrimination parameters.
+In the first wave, we fit four models:
 
@@
 
-\item Model 1a: Item difficulty is a function only of template attributes (number of shapes, number of changing relations).
-In this model, items of equal complexity are psychometrically equivalent as are all item clones.
+\item Model 1: Item difficulty is a function only of item template attributes (i.e.
+feature number, rule number).
+In this model, item with identical attributes are psycho- metrically equivalent, as are all the clones belonging to an item.
 
-\item Model 1b: Item difficulty is a function of template attributes and residuals.
-In this model, items of equal complexity may not be equally difficult, but all clones of an item are still psychometrically equivalent.
+\item Model 2: Item difficulty is a function of item template attributes and residuals.
+In this model, item templates with identical attributes may not be equally difficult; however, all clones belonging to an item are still psychometrically equivalent.
 
-\item Model 1c: Item difficulty is a function of both template and clone attributes (including distractor type and average response time), and template-level residuals.
-In this model, item clones may systematically vary in their difficulty as a function of their attributes.
+\item Model 3: Item difficulty is a function of both item template and clone attributes (i.e.
+distractor type, mean response time), as well as template-level residuals.
+Item clones may now vary in their difficulty as a function of their attributes.
 
-\item Model 1d: Item difficulty is function of both template/clone attributes and residuals.
-In this model, neither item templates nor clones are psychometrically equivalent.
-The degree of nonequivalence is dependent on the magnitude of the residual terms.
+\item Model 4: Item difficulty is function of attributes and residuals at both levels.
+In this model, neither item templates nor clones with identical attributes are necessarily equally difficult, depending on the magnitude of the residual terms.
 
@@
 The goodness-of-fit of the four models is summarized in Table \ref{table:1}.
-Model 1d demonstrated a better fit of the data than Model 1a ($\Delta$ LOCO = 1223.69, se = 44.94), Model 1b ($\Delta$ LOCO 511.58, se = 28.77), and Model 1c ($\Delta$ LOCO = 187.96, se = 17.25).
-Thus, Model 1d was chosen as the starting point for the second wave of model fitting.
+Model 4 demonstrated a better fit of the data than Model 1 ($\Delta$ LOCO = 1223.69, se = 44.94), Model 2 ($\Delta$ LOCO 511.58, se = 28.77), and Model 3 ($\Delta$ LOCO = 187.96, se = 17.25).
+Thus, Model 4 was selected as the starting point for the second wave of model fitting.
 
-Next we estimated a second set of three models in which the item discrimination parameter was a free parameter.
-For all models, the item difficulty parameters were specified according to Model 1d.
-The structure of the three models were as follows:
+Next we estimated a second set of models in which the item discrimination parameter was a free parameter.
+For all models, the item difficulty parameters were specified according to Model 4.
+In the second wave, we fit three models:
 
@@
 
-    \item Model 2a: Item discrimination is a function only of both template and clone attributes.
-Here we include all attributes at once having found that they were all predictive of item difficulty (see below).
+    \item Model 5: Item discrimination is a function of both template and clone attributes.
+In this model, item templates with identical attributes will be equally discriminating; however, item clones may vary in their discrimination as a function of their attributes.
 
-    \item Model 2b: Item discrimination is a function of both item attributes and template residuals. In this model, items of equal complexity may not be equally discriminating.
+    \item Model 6: Item discrimination is a function of all item attributes and template-level residuals.
+In this model, item templates with identical attributes may not necessarily be equally discriminating.
 
-    \item Model 2c: the item discrimination parameters were modeled with fixed and random effects at both levels.
-In this model, neither item templates nor clones are equivalently discriminating.
-The degree of nonequivalence is dependent on the magnitude of the residual terms.
+    \item Model 7: Item discrimination is a function of attributes and residuals at both levels.
+In this model, neither item templates nor clones with identical attributes are necessarily equally discriminating, depending on the magnitude of the residual terms.
 
@@
 
-The goodness-of-fit of these models is summarized in Table \ref{table:1}.
-Notably, we did not find the most complex model (Model 2c) to be the best-fitting model.
-Although Model 2c was preferred to Model 2a ($\Delta$ LOCO = 2.13, se = 2.04), it was not preferred to Model 2b ($\Delta$ LOCO = -1.44, se = 0.85).
-%We note that the authors of the LOO metric recommend that standard errors be at least twice as large as the difference between models to be confident that one model truly provides a better fit.
-However, these differences in LOCO values from each model to the next were each within two standard errors of the mean, indicating only weak predictive improvement of Model 2b over the others \citep{vehtari2022cv}.
-As such we proceed by select Model 2b as the best-fitting model overall, but proceed in discussing it with caution.
-We also note that Model 2b fit clearly better than Model 1d ($\Delta$ LOCO = 85.91, se = 6.33, supporting the inclusion of varying item discrimination parameters).
+The goodness-of-fit of the three models is summarized in Table \ref{table:1}.
+Notably, we did not find the most complex model (Model 7) to be the best-fitting model.
+Although Model 7 was preferred to Model 5 ($\Delta$ LOCO = 2.13, se = 2.04), it was not preferred to Model 6 ($\Delta$ LOCO = -1.44, se = 0.85).
+However, these differences in LOCO values from each model to the next were each within two standard errors of the mean, indicating only weak predictive improvement of Model 6 over the others \citep{vehtari2022cv}.
+As such we will select Model 6 as the best-fitting model overall, but proceed in discussing it with caution.
+We also note that Model 6 fit clearly better than Model 4 ($\Delta$ LOCO = 85.91, se = 6.33, supporting the inclusion of varying item discrimination parameters).
 
-\subsubsection{Objective 1}
+\subsubsection{Aim 1: Estimate item parameters for all items in the MaRs-IB}
 
 Across all items, the average item difficulty was $\mu_\beta = 0.177$ (95\% HDI = 0.118 -- 0.238).
-There was also considerable variance in item difficulty (sd = 1.431, 95\% HDI = 1.370 -- 1.492), the smallest ($\beta$ = -3.704, 95\% HDI = -4.687 -- -2.775) and largest ($\beta$ = 3.497, 95\% HDI = 2.481 -- 4.635) item difficulty parameters spanning a wide range.
-For an average ability participant, these estimates translate to an expected average accuracy of 58.4\% across all items, and essentially ceiling (floor) performance for the least (most) difficult items.
-Thus, items in MaRs-IB span a substantial range of difficulty.
+There was considerable variability in item difficulty across items (sd = 1.431, 95\% HDI = 1.370 -- 1.492), wih the smallest ($\beta$ = -3.704, 95\% HDI = -4.687 -- -2.775) and largest ($\beta$ = 3.497, 95\% HDI = 2.481 -- 4.635) item difficulty parameters spanning a wide range.
+For an average ability participant ($\theta = 0$), these parameter estimates translate to an expected average accuracy of 58.4\% across all items, and ceiling (floor) performance for the least (most) difficult item.
 
 In turn, the average item discrimination was $\mu_\alpha = 1.298$ (95\% HDI = 1.212 - 1.385).
-Variability in item discrimination was modest in comparison to item difficulty (sd = 0.221, 95\% HDI = 0.121 - 0.322).
-Item difficulty and discrimination were uncorrelated across items ($\rho$ = -0.060, 95\% HDI = -0.379 - 0.250).
+Variability in item discrimination was modest by comparison (sd = 0.221, 95\% HDI = 0.121 - 0.322).
+Item difficulty and discrimination were uncorrelated across items ($\rho$ = -0.060, 95\% HDI = -0.379 -- 0.250).
 Following convention \citep{baker2017basics}, the items in the MaRs-IB exhibited medium ($\alpha \in$ 0.65 - 1.34) to high ($\alpha \in$ 1.35 – 1.69) levels of discrimination and are thus suitable for measuring matrix reasoning ability.
 
-\subsubsection{Objective 2}
+\subsubsection{Aim 2: Quantify the associations between item attributes and functioning}
 
 Next we inspected the associations between item attributes and parameters.
-As predicted, a one-unit change in the number of item features was associated with an increase in item difficulty ($\delta_{\beta_1}$ = 0.579, 95\% HDI = 0.389 -- 0.770), or an approximate 10.6\% reduction in accuracy for a participant of average ability.
-Similarly, a one-unit change in the number of changing relations was also associated with an increase in item difficulty ($\delta_{\beta_2}$ = 0.514, 95\% HDI = 0.378 -- 0.663), or a 9.5\% reduction in accuracy.
+As predicted, a one-unit change in feature number was associated with an increase in item difficulty ($\delta_{\beta_1}$ = 0.579, 95\% HDI = 0.389 -- 0.770), or an approximate 10.6\% reduction in accuracy for a participant of average ability.
+Similarly, a one-unit change in rule number was also associated with an increase in item difficulty ($\delta_{\beta_2}$ = 0.514, 95\% HDI = 0.378 -- 0.663), or a 9.5\% reduction in accuracy.
 
 Looking next at the clone-level attributes, there was an unexpected association between distractor type and item difficulty ($\delta_{\beta_3}$ = 1.105, 95\% HDI = 0.940 -- 1.269).
-That is, all else equal, an item with MD distractors was associated with an expected 19.9\% reduction in accuracy compared to items with PD distractors.
-Finally, as expected, there was a positive association between mean response time and item difficulty ($\delta_{\beta_4}$ = 0.541, 95\% HDI = 0.414 -- 0.671).
+All else being equal, an item with MD distractors was associated with an expected 19.9\% reduction in accuracy compared to items with PD distractors.
+As expected, there was a positive association between mean response time and item difficulty ($\delta_{\beta_4}$ = 0.541, 95\% HDI = 0.414 -- 0.671).
 
 In contrast, there was only one credible association between an item attribute and item discrimination parameters.
-A one-unit change in the number of changing relations was associated with a small but credible increase in item discrimination  $\delta_{\alpha_2}$ = 0.020, 95\% HDI = 0.003 -- 0.037).
-There was no reliable association between item discrimination and the number of item features ($\delta_{\alpha_1}$ = -0.015, 95\% HDI = -0.036 -- 0.008), distractor type  ($\delta_{\alpha_3}$ = -0.029, 95\% HDI = -0.065 -- 0.005), or average response time ($\delta_{\alpha_4}$ = -0.010, 95\% HDI = -0.030 -- 0.009).
+A one-unit change in rule number was associated with a small increase in item discrimination $\delta_{\alpha_2}$ = 0.020, 95\% HDI = 0.003 -- 0.037).
+There were no reliable associations between item discrimination and feature number ($\delta_{\alpha_1}$ = -0.015, 95\% HDI = -0.036 -- 0.008), distractor type ($\delta_{\alpha_3}$ = -0.029, 95\% HDI = -0.065 -- 0.005), or average response time ($\delta_{\alpha_4}$ = -0.010, 95\% HDI = -0.030 -- 0.009).
 
-\subsubsection{Objective 3}
+\subsubsection{Objective 3: Decompose the variability in item functioning}
 
 Finally, we inspected the relative contributions of modeled and unmodeled sources to variability in item functioning.
-The two modeled item-level attributes, number of features and changing relations, explained approximately 65.5\% of the variance in difficulty across the 64 item templates (or 35.0\% of the variance across all 384 items).
-In conjunction with distractor type, these three item attributes explained nearly half (48.2\%) of the variance in item difficulty across all 384 items.
-Nevertheless, there was non-negligible residual variability in item difficulty at the template- and clone-level as evidenced by the comparatively better fit of item structure models including random effects terms at both levels.
-As such, neither item templates nor clones can be considered exchangeable with respect to difficulty.
+The two modeled item-level attributes, feature number and rule number, explained approximately 65.5\% of the variance in difficulty across the 64 item templates (or 35.0\% of the variance across all 384 item clones).
+In conjunction with distractor type, the three item attributes explained nearly half (48.2\%) of the variance in item difficulty across all 384 item clones.
+Nevertheless, there was non-negligible residual variability in item difficulty at the template and clone levels as evidenced by the comparatively better fit of an item structure model that included random effects at both levels.
+That is, the results indicate that neither item templates nor clones with identical attributes can be assumed to be equal in difficulty.
 
-By contrast, the three inherent item attributes explained  approximately 39.2\% of the variance in item discrimination across item templates.
-In contrast to item difficulty, model comparison suggested there was only non-negligible residual variance in item discrimination at the template-level.
-Indeed, a model that included random effects at the clone-level offered a comparably worse fit.
-As such, there was not credible evidence that item clones are not exchangeable with respect to discrimination.
+Feature number, rule number, and distractor type cumulatievely explained approximately 39.2\% of the variance in item discrimination across the 64 item templates.
+In contrast to item difficulty, model comparison indicated there was non-negligible residual variability in item discrimination only at template level.
+Indeed, a model that included random effects at the clone level produced a comparatively worse fit to the data.
+As such, the results indicate that item templates with identical attributes cannot be assumed to be equally discriminating; however, there was not sufficient evidence to refute that item clones are equally discriminating.
+We note this result should be interpreted with caution given the relatively weak predictive improvement of the best-fitting model compared to other similar models.
 
@@
 
-We investigated person-level abilities as a function of four covariates: age, gender, average response time, and $\Delta$ response time.
-This last term reflects the degree to which a participant slowed down (or sped up) their response time as a function of item difficulty (Figure \ref{fig:fig01}c).
-All covariates were standard-scored prior to analysis with the exception of gender, which was binary coded (male = -0.5, female = 0.5).
-
+In the best-fitting model, there were several credible associations between person-level attributes and ability.
 There was a negative association between age and ability ($\rho_1$ = -0.299, 95\% HDI = -0.353 -- -0.246).
-There was a small but credible association between gender and ability ($\rho_2$ = 0.128, 95\% HDI = 0.078 -- 0.182), such that women performed marginally better than men.
-Average response time was positively associated with ability ($\rho_3$ = 0.427, 95\% HDI = 0.374 -- 0.478), as was $\Delta$ response time ($\rho_4$ = 0.368, 95\% HDI = 0.313 -- 0.424), indicating a speed-accuracy tradeoff.
+There was also a small association between gender and ability ($\rho_2$ = 0.128, 95\% HDI = 0.078 -- 0.182), such that women performed marginally better than men.
+Average response time was positively associated with ability ($\rho_3$ = 0.427, 95\% HDI = 0.374 -- 0.478), as was $\Delta$ response time ($\rho_4$ = 0.368, 95\% HDI = 0.313 -- 0.424), suggesting a speed-accuracy trade-off in performance.
 
@@
 
-Finally, we validated the fit of the winning model using several posterior predictive model checking measures.
-Across the 384 items, the average posterior predictive p-value for the $S-\chi^2$ index of item fit was $p = 0.510$ (sd = 0.242).
-Only 14 items (3.64\%) indicated misfit with posterior predictive p-values less than 0.05 or greater than 0.95.
-Moreover, the posterior predictive p-value for the test of undimensionality was $p = 0.803$.
-Though this suggests some evidence for local dependence, it is within a tolerable range.
+Finally, we validated the fit of the best-fitting model using several posterior predictive model checking measures.
+Across the 384 items, the average posterior predictive p-value for the $S-\chi^2$ index of item fit was $p = 0.510$.
+Only 14 of 384 items (3.64\%) indicated misfit with PPP-values less than 0.05 or greater than 0.95.
+Moreover, the PPP-value for the test of residual dependence was $p = 0.803$.
+Though this suggests some evidence for response dependence, it is within a tolerable range.
 Overall then, the best-fitting model provides an adequate fit to the data.
@@
 
-In this first study, we recruited a large sample of adults to complete items from the MaRs-IB, making sure that each item (and its variants) were completed by a sizable number of participants.
-Using hierarchical item response models, we then measured the psychometric properties (i.e.
+In this study, we recruited a large sample of adults to complete items from the MaRs-IB, making sure that each item (and its clones) were completed by a sizable number of participants.
+Using multilevel item structure models, we then estimated the psychometric properties (i.e.
 difficulty and discrimination) of each item.
-Corroborating the initial study from \cite{chierchia2019matrix}, we found that the items of the MaRs-IB vary greatly in their difficulty -- a prerequisite for measuring nonverbal reasoning across the ability spectrum.
-We also found that the items of the MaRs-IB exhibited medium-to-large levels of discrimination \citep{baker2017basics}, similar to what has been reported in previous item response theory analyses of the Raven's progressive matrices \citep{chiesi2012using}.
+Corroborating the findings of a preliminary investigation of the MaRs-IB \citep{chierchia2019matrix}, we found that the items of the MaRs-IB vary greatly in their difficulty -- a prerequisite for measuring nonverbal reasoning across the ability spectrum.
+We also found that the items of the MaRs-IB exhibited medium-to-large levels of discrimination \citep{baker2017basics}, similar to what has been found in investigations of other measures of matrix reasoning \citep{chiesi2012using}.
 
-Moreover, using item cloning models we were able to test several important properties of the MaRs-IB items.
-First, we confirmed that item difficulty is a function of and to a moderate extent explained by its attributes (namely, the number of features and the number of changing relations).
-Relatedly, we identified that an item's discrimination is a function and partially explained by its number of changing relations, which is similar to what has been reported in other investigations of matrix reasoning tasks \citep{embretson1999generating}.
-Second, we found that item variants are not equivalent and exchangeable with respect to difficulty.
-Unexpectedly, and contrary to \cite{chierchia2019matrix}, distractor type emerged as a strong predictor of item difficulty.
-All else being equal, item variants using MD distractors were harder than the same items using PD distractors.
-Importantly, once taking this (and the residual variance in difficulty at the item family level) into account, a large majority of item difficulty (72.0\%) was accounted for, suggesting that a minority of variance in difficulty across item variants stems from unmodeled features such as their superficial perceptual properties.
+We also investigated whether the MaRs-IB exhibited several desirable properties.
+We confirmed that the difficulty of MaRs-IB items is a function of, and to a moderate extent explained by, item attributes (namely, feature number and rule number).
+Notably, we also identified that item discrimination is associated with rule number, in line with previous investigations of matrix reasoning measures \citep{embretson1999generating}.
+Crucially, we found evidence that item clones are not equally difficult.
+Unexpectedly, and contrary to the results of \cite{chierchia2019matrix}, distractor type emerged as a robust predictor of item difficulty.
+All else being equal, items were much harder when using MD-type distractors compared to PD-type distactors.
+Moroever, there was non-negligible residual variance in item difficulty across clones even after taking into consideration all modeled sources of variance.
+Therefore item clones in the MaRs-IB are not psychometrically equivalent and should not be treated as exchangeable.
 
-summary sentence here saying that the MaRs-IB has many desirable properties. We next move onto test construction.
+In summary, the results of the calibration study confirm the MaRs-IB to be a useful and psychometrically-sound resource for measuring matrix reasoning ability.
+The items span a great range of difficulty with medium-to-large levels of discrimination, thus making the MaRs-IB suitable for measuring matrix reasoning across a wide range of ability levels.
+That item clones vary in their difficulty, and should not therefore be treated as exchangeable, does not undermine this point.
+Furthermore, the item parameter estimates derived here can be used to design new MaRs-IB measures with good reliability and potential for reuse, as we demonstrate in the next study.
 
@@
 
+\subsection{Objectives}
+
+The purpose of the second study was to provide a demonstration of how to use the item parameter estimates derived in the previous study to design new MaRs-IB measures.
+Specifically, we use optimal test assembly methods to construct three parallel short form measures with maximal reliability under time length constraints.
+
 \subsection{Optimal test assembly}
 
-With 64 item templates, with 6 variants each, there is a vast number of possible test forms we could construct using the MaRs-IB.
-Returning to our motivating problem --- the availability of abbreviated matrix reasoning measures with the potential for retest --- we opted to construct three parallel short form measures.
+With 64 item templates, each with six clones, there is a vast number of possible MaRs-IB test forms we could construct.
+Returning to our motivating problem --- the availability of abbreviated matrix reasoning measures with the potential for reuse --- we decided to construct three parallel short form measures.
 To do so, we used mixed integer programming \citep{der2005wj} to maximize the test information function (TIF) of each short form subject to the following constraints:
@@
     \item Each short form was required to contain 12 items.
-This was chosen to minimize the administration time of a given form (2-4 minutes on average) and to achieve good score reliability (Cronbach's $\alpha \geq 0.8$).
+This was chosen to minimize the administration time of a given form (2-4 minutes on average) and to achieve sufficient test reliability (Cronbach's $\alpha \geq 0.8$).
 
-    \item All short forms were required to be comprised of variants from the same item families. This was chosen to maximize the similarity of the short forms.
+    \item All short forms were required to be made up of clones from the same items. This constraint was adopted in order to maximize the similarity of the short forms.
 
-    \item Variants could be of the MD or PD distractor type, but not both. This was chosen to prevent item redundancy (i.e. including the same puzzle twice in one short form).
+    \item Clones selected for inclusion could either use MD-type or PD-type distractors, but not both.
+This was chosen to prevent item redundancy (i.e.
+including the same puzzle twice in one short form).
 
-    \item The difference in TIFs between short forms was required to be minimized.
-This was chosen to ensure that each short form had similar psychometric properties (i.e.
-equivalent reliability across ability levels).
+    \item The difference in TIFs between short forms should be minimized.
+This constraint was adopted to ensure that each short form had approximately equal psychometric properties (i.e.
+equivalent measurement reliability across ability levels).
 
@@
 \noindent The TIF was maximized at five ability levels ($\theta = -1.0, -0.5, 0.0, 0.5, 1.0$).
-Previous simulation studies have shown that target values at three to five well-chosen $\theta$s generally suffice \citep{der2005wj}.
+Previous simulation studies have shown that target values at three to five well-chosen $\theta$s are generally sufficient \citep{der2005wj}.
 Solutions to the mixed integer programming problem were found using the \textit{mip} python package (v1.13.0) \citep{santos2020mixed}.
@@
 As expected, the MIP solver selected for items that were more discriminating on average (Figure \ref{fig:fig02}a).
-Though this was not an explicit constraint, the test characteristic curves (TCCs) of the three short forms are markedly similar, each with an expected total score of 7.7 out of 12 items.
-Note that, because low ability participants are expected to guess an item correctly 25\% of time, the lower asymptote of the expected score distribution is 3 and the total score distribution is left-skewed.
-Finally, the TIFs across the three short forms are remarkably similar (Figure~\ref{fig:fig02}b) as would be expected given the assembly constraints.
-As such, the short forms are predicted to have similar levels of precision across ability levels.
-Indeed, the expected marginal score reliability for each test form are similar (short form 1: $\alpha = 0.794$; short form 2: $\alpha = 0.795$; short form 3: $\alpha = 0.792$).
+Though this was not an explicit constraint, the test characteristic curves (TCCs) of the three short forms are markedly similar (Figure \ref{fig:fig02}b).
+Based on the TCCs, the expected observed score for each measure is 7.7 out of 12 items.
+(Note that, because the lowest ability participants are expected to guess an item correctly 25\% of time, the lower asymptote of the TCCs is at a score of 3.) Finally, the TIFs across the three short forms are remarkably similar (Figure~\ref{fig:fig02}c), as is expected given the assembly constraints.
+As such, we should expect each short form to have approximately equal precision in measuring matrix reasoning ability across the spectrum of ability.
+Indeed, the simulated marginal score reliability for each test form are similar (short form 1: $\alpha = 0.794$; short form 2: $\alpha = 0.795$; short form 3: $\alpha = 0.792$).
 
-\begin{figure}
-\centering
-\includegraphics[width=1.0\textwidth]{figures/fig02.png}
-\caption{\label{fig:fig02}Results from the test assembly of three parallel MaRs-IB short form (SF) measures.
-(A) The psychometric properties of the items selected for each form, compared to all remaining items.
-Points represent the posterior means of the item difficulty and discrimination  parameters.
-(B) The test characteristic curves (TCCs) for each short form.
-The degree of overlap highlights the similarity of expected scores for each test form.
-(C) The test information functions (TIFs) for each short form.
-The degree of overlap highlights the similarity of reliability for each test form.}
-\end{figure}
-
-Next, we proceeded to test each short form in a new sample of participants.
-The goals for this study were twofold.
-First, we sought to confirm that each short form would have the properties predicted above.
-Second, we sought to establish the convergent validity of the short forms by correlating total scores with performance on an abbreviated version of the Raven's progressive matrices.
+Next, we proceeded to administer each short form to a second sample of participants.
+Our aim was to verify that observed scores on each short form conformed to our expectations based on the TCCs and TIFs.
+Successfully doing so would provide additional evidence that we were successful in calibrating the functioning of the items in the MaRs-IB.
+We also sought to establish the convergent validity by correlating observed scores on the short form measures with performance on an abbreviated version of the Raven's progressive matrices.
 
@@
 
-A total of N=347 participants were recruited from the Prolific Academic platform (https://www.prolific.co) to participate in an online behavioral experiment in November, 2021.
-Participants were eligible if they currently resided in the United States and had not participated in the previous study.
-Total study duration was approximately 10.8 minutes (sd = 4.6) per participant.
-Participants received monetary compensation for their time (rate USD \$10/hr), plus a performance-based bonus up to \$0.75 based on task performance.
-Participants earned on average \$2.10 USD (sd = \$0.18 sd).
+A total of N=347 participants were recruited from the Prolific Academic platform to participate in an online behavioral experiment in November, 2021.
+Participants were eligible if they currently resided in the United States and had not participated in the calibration study.
+Study duration was on average 10.8 minutes (sd = 4.6).
+Participants received monetary compensation for their time (rate USD \$10/hr) plus a bonus up to \$0.75 based on task performance.
+Participants earned on average \$2.10 USD (sd = \$0.18).
 This study was approved by the Institutional Review Board of Princeton University (\#7392), and all participants provided informed consent.
@@
 The majority of participants identified as women (men: N=138; women: N=147; non-binary or other: N=13; rather not say: N=2).
-Participants were 35.0 years old (13.2 sd) on average.
+Participants were 35.0 years old (sd = 13.2) on average.
 The sample was relatively well-educated, with the majority having completed a Bachelor's degree (N=121) or Master's degree or higher (N=45).
@@
 
-The study procedure was divided into three sections.
+The study was divided into three parts.
 After providing consent, participants first completed three short surveys: the 10-item need for cognition survey \citep{chiesi2018applying}, the 8-item PROMIS Cognitive Function-Short Form (8a) \citep{iverson2021normative}, and the 8-item subject numeracy scale \citep{fagerlin2007measuring}.
-These measures were included as part of exploratory analyses that are reported in the Supplement.
-Afterwards participants completed one of the three 12-item MaRs-IB short-form measures described above.
+These measures were included as part of exploratory analyses that are reported in the supplementary materials.
+Afterwards participants completed one of the three 12-item MaRs-IB short-form measures.
 The administration procedure of the short-forms was identical to that in the first study.
 Finally, participants completed the 9-item abbreviated Raven's progressive matrices (RPM; form A) \citep{bilker2012development}.
-The presentation format and instructions for the RPM items was the same as for the MaRs-IB items.
+To maximize consistency in administration, the presentation format and instructions for the RPM measure was the same as for the MaRs-IB test forms.
 
@@
 The probability of a participant correctly solving a MaRs-IB was defined according to the 3PL model (Eq \ref{eq:1}), where the difficulty and discrimination parameters for an item were fixed to the posterior mean of the estimates obtained in the previous study.
-Thus, the model estimates the partial correlation between RPM performance and MaRs-IB ability controlling for the age and gender of the participants.
+Thus, the model estimates the partial correlation between RPM performance and MaRs-IB ability controlling for the age and gender of the participants, and the differences in item functioning across test forms.
 
@@
 Performance on the MaRs-IB short-forms is summarized in Table \ref{table:3}.
-On average, participants completed the measure in 174 s (sd = 54.0 s) and responded correctly on 8.0 of 12 items (sd = 2.5).
+On average, participants completed the measure in 174 s (sd = 54.0 s) and correctly solved 8.0 of 12 items (sd = 2.5).
 The distribution of observed scores was right-shifted (Figure~\ref{fig:fig03}), as was expected given the nominal guessing rate.
 A one-way ANOVA comparing the average score across the three short-forms was not statistically significant (F(2,297) = 0.253, p = 0.777).
-Importantly, the ordinal reliability of each measure was satisfactory (all $\alpha > 0.75$).
-In sum, performance on the MaRs-IB short form were consistent with predicted made during test assembly, indicating that the psychometric properties of the items were adequately calibrated during in the previous study.
+Importantly, the marginal score reliability of each measure was satisfactory (all $\alpha > 0.75$).
+In sum, performance on each MaRs-IB test measure conformed to expectations, indicating that item functioning was adequately calibrated in the previous study.
 
@@
 The proportion of correct responses on on the RPM was significantly lower than that observed for the MaRs-IB ($t = 13.028$, $p < 0.001$), suggesting that the RPM is more difficult or less biased by guessing.\footnote{In contrast to the MaRs-IB, items in the RPM have either six or eight response options.
-As such, the nominal guessing rate is expected to be lower for the RPM.} The ordinal reliability for the RPM was also adequate.
+As such, the nominal guessing rate is expected to be lower for the RPM.} The marginal score reliability for the RPM was also adequate.
 
 The partial correlation between observed scores on the RPM and ability on the MaRs-IB was $\rho = 0.527$ (95\% HDI = 0.457 - 0.577).
-Considering the attenuation of correlations between two imperfect measures \citep{spearman1961proof}, we should not expect correlations much larger than $\rho = 0.64$ (or $0.80^2$).
-As such, this result indicates good convergent validity between the two matrix reasoning measures.
-Briefly, in this smaller sample, neither age nor gender was credibly associated with performance on the MaRs-IB short forms (age: $\rho$ = -0.123, 95\% HDI = -0.253 -- 0.013; gender: $\rho$ = -0.025, 95\% HDI = -0.160 -- 0.108).
+Considering the attenuation of correlations between two imperfect measures \citep{spearman1961proof}, we should not expect correlations much larger than $\rho \approx 0.64$ (or $0.80 \times 0.80$).
+Therefore, this result indicates good convergent validity between the two matrix reasoning measures.
+Briefly, in this second sample, neither age nor gender was credibly associated with performance on the MaRs-IB short forms (age: $\rho$ = -0.123, 95\% HDI = -0.253 -- 0.013; gender: $\rho$ = -0.025, 95\% HDI = -0.160 -- 0.108).
 
-To summarize this section, we used the item parameters estimated as part of the calibration study, and optimal assembly methods, to construct three new MaRs-IB short form measures.
-As predicted, these measures were quick to administer, had adequate reliability, and good convergent validity with a second matrix reasoning measure.
+To summarize, we used the estimated item parameters in conjunction with optimal assembly methods to construct three parallel MaRs-IB short forms.
+As expected, these new measures were quick to administer, exhibited score distributions in line with model predictions, and demonstrated good internal reliability and convergent validity with a second matrix reasoning measure.
+Together, these results highlight our success in accurately calibrating the functioning of MaRs-IB items, and consequently the feasibility of constructing new matrix reasoning measures.
 
@@
 
-Here we have provided the most comprehensive interrogation of of the MaRs-IB to date.
-Using item structure models fit to data collected from a large sample of adult participants, we found that the MaRs-IB possesses desirable psychometric properties.
+Here we have provided the most comprehensive investigation of the MaRs-IB to date.
+Using item structure models fit to data collected from a large sample of adult participants, we found that the MaRs-IB possesses many desirable psychometric properties.
 The items in the MaRs-IB span a considerable range of difficulty and exhibit medium-to-high levels of discrimination.
-As such the MaRs-IB is suitable for measuring matrix reasoning across the ability spectrum in the adults.
-We also verified that the design of the MaRs-IB was in part successful: as intended, item attributes such as feature number and rule number were positively associated with, and explained much of the variability, in item difficulty.
-Contrary to our expectations, however, we also found that the design of the MaRs-IB had unintended consequences: distractor type was also associated with item difficulty.
-Moreover, we found non-negligible variability in the difficulty of item clones even after controlling for distractor type.
-Crucially, item clones in the MaRs-IB therefore cannot be treated as exchangeable.
+The MaRs-IB is therefore suitable for measuring matrix reasoning across the ability spectrum in the adults.
+We also verified that the design of the MaRs-IB was in part successful.
+As intended, item attributes such as feature number and rule number were positively associated with, and explained much of the variability, in item difficulty.
+We also found some limitations in the design of the MaRs-IB.
+For example, distractor type was also associated with item difficulty.
+Moreover, we also found non-negligible residual variance in item difficulty across clones after controlling for item attributes.
+As a consequence, item clones in the MaRs-IB cannot therefore be treated as psychometrically equivalent and exchangeable.
 
@@
 Interestingly, we also found an interaction between participant ability and item difficulty; that is, the best-performing participants increased their deliberation time as items become more difficult, whereas the worst-performing participants maintained approximately the same work rate.
-These results are consistent with a persistence account of participant ability, wherein performance reflects in part a willingness to invest time in the solution process \citep{ranger2021effects}.
-As such, the MaRs-IB may be suitable not only for measuring matrix reasoning ability but also mental effort costs [cite], opportunity costs [cite], or other  motivational factors \citep{duckworth2011role} related to the tendency to give up (at least in low-stakes test settings; see \cite{shaw2020reasoning} for a counter-example).
-Moreover, the use of cognitive models may help to disentangle the relative contributions of these factors to performance on the MaRs-IB \citep{ranger2014accumulator}.
+These results are consistent with a persistence account of participant ability, wherein good performance reflects in part a willingness to invest time in the solution process \citep{ranger2021effects}.
+As such, the MaRs-IB may be suitable not only for measuring matrix reasoning ability but also mental effort costs \citep{kool2018mental}, opportunity costs \citep{otto2019opportunity}, or other motivational factors \citep{duckworth2011role} related to the tendency to exert effort or give up.
+The use of cognitive models may help to disentangle the relative contributions of these factors to performance on the MaRs-IB \citep{ranger2014accumulator}.
 
@@
 One notable limitation is our sample.
-Here we analyzed response data collected from an online, adult sample that was relatively young and well-educated.
-We cannot guarantee that the psychometric properties of the MaRs-IB reported here will generalize to other populations or testing contexts.
+Here we analyzed response data collected from an online adult sample which was relatively young and well-educated.
+We cannot guarantee that the psychometric properties of the MaRs-IB found here will generalize to other populations or testing contexts.
 Future researchers may consider replicating the current study in other populations of interest (e.g.
@@
 IRT linking describes a set of methods to establish the comparability of item and ability parameters estimated from data collected from two or more groups that differ in ability \citep{lee2018irt}.
-Future studies employing these methods could provide insight not only in how the functioning of the MaRs-IB may change in different populations, but also how matrix reasoning ability differs across samples.
+Future studies utilizing these methods could provide insight not only in how the functioning of the MaRs-IB might change in across populations, but also how matrix reasoning ability differs across populations.
 
-A second limitation is in our investigation of sources of variance.
-Here we only investigated the contributions of three item attributes.
-There are far more sources to have investigated (e.g.
-type of rule change such as color change versus shape change).
-Indeed, there are other item features taxonomies we could have explored \citep{carpenter1990one}.
-We also did not explicitly model perceptual features as has been done before.
-Although doing so would not change our results about exchangeability of item clones, it would allow for more precise estimates of item parameters and improved test assembly.
-Also allows for possibility of automatic item generation.
-Room for future research.
+A second limitation is in our relatively limited investigation of sources of variance in item functioning.
+Here we investigated the contributions of only three item attributes.
+There were many item properties that we elected not to model (e.g.
+type of changing relation such as color change or shape change).
+Previous investigations of matrix reasoning tasks have proposed other taxonomies of item attributes that affect item functioning \citep{carpenter1990one}.
+Incorporating additional attributes into our models may have further increased the precision of our parameter estimates and might  provide additional knowledge for guiding test assembly.
+Future studies should consider more deeply investigating the possible relationships between other attributes of items in the MaRs-IB and their functioning.
 
-In summary, we have provided an in-depth psychometric validation of the MaRs-IB, finding that it is suitable for measuring matrix reasoning ability in the general adult population.
-We release the item parameters so that researchers interested in measuring matrix reasoning can construct maximally reliable tests for their own research.
+In conclusion, we have provided an psychometric validation of the MaRs-IB, finding that it is suitable for measuring matrix reasoning ability in the general adult population.
+We hope that the materials and results presented here will encourage researchers interested in measuring matrix reasoning using the MaRs-IB in their studies.
 
@@
 \end{tabular*}
-\caption{\label{table:1}\normalfont Summary of the random-intercepts linear regression model predicting log-transformed response time as a function of accuracy, participant rest score, and item difficulty.}
+\caption{\label{table:1}\normalfont Summary of the mixed effects linear regression model predicting log-transformed response time as a function of trial accuracy, participant rest score (total correct on all other items), and item difficulty (one minus the proportion correct for an item).
+\newline lme4 syntax: log(RT) $\sim$ accuracy * score * difficulty + (1 | subject).}
 \end{table}
@@
     \centering
-    \begin{tabular*}{\textwidth}{cccccccc}
+    \begin{tabular}{cccccccc}
     \toprule
@@
     \midrule
-    1a & x &   &   &   & -14516.85 & -1223.69 (44.94)\\
-    1b & x & x &   &   & -13804.74 & -511.58 (28.77) \\
-    1c & x & x & x &   & -13481.12 & -187.96 (17.25) \\
-    1d & x & x & x & x & -13293.16 & - \\
+    1 & x &   &   &   & -14516.85 & -1223.69 (44.94)\\
+    2 & x & x &   &   & -13804.74 & -511.58 (28.77) \\
+    3 & x & x & x &   & -13481.12 & -187.96 (17.25) \\
+    4 & x & x & x & x & -13293.16 & - \\
     \midrule
-    2a & x &   & x &   & -13209.38 & -2.13 (2.04) \\
-    2b & x & x & x &   & -13207.25 & - \\
-    2c & x & x & x & x & -13208.69 & 1.44 (0.85) \\
+    5 & x &   & x &   & -13209.38 & -2.13 (2.04) \\
+    6 & x & x & x &   & -13207.25 & - \\
+    7 & x & x & x & x & -13208.69 & 1.44 (0.85) \\
     \bottomrule
-    \end{tabular*}
-    \caption{\label{tab:2} \normalfont Goodness of fit of item response models to MaRs-IB data. Pareto-smoothed importance sampling}
+    \end{tabular}
+    \caption{\label{tab:2} \normalfont Goodness of fit of item response models to MaRs-IB data.
+PSIS = Pareto-smoothed importance sampling.
+Note: LOCO values are presented in original, not deviance, scale so larger numbers indicate better fit.}
     \label{table:2}
@@
     \centering
-    \begin{tabular*}{\textwidth}{lccccc}
+    \begin{tabular*}{\textwidth}{llcccc}
     \toprule
@@
     \midrule
-    MaRs-IB SF-1 & 103 & 174.9 s (53.3) & 7.9 (2.4) & 7 - 10 & 0.761 \\
-    MaRs-IB SF-2 & 98  & 169.0 s (53.0) & 8.2 (2.6) & 7 - 10 & 0.810 \\
-    MaRs-IB SF-3 & 99  & 178.5 s (55.7) & 7.9 (2.7) & 6 - 10 & 0.829 \\
-    RPM SF-A     & 300 & 124.6 s (40.1) & 4.5 (2.0) & 3 - 9 & 0.791 \\
+    MaRs-IB SF-1 & N=103 & 174.9 s (53.3) & 7.9 (2.4) & 7 - 10 & 0.761 \\
+    MaRs-IB SF-2 & N=98  & 169.0 s (53.0) & 8.2 (2.6) & 7 - 10 & 0.810 \\
+    MaRs-IB SF-3 & N=99  & 178.5 s (55.7) & 7.9 (2.7) & 6 - 10 & 0.829 \\
+    RPM SF-A     & N=300 & 124.6 s (40.1) & 4.5 (2.0) & 3 - 9 & 0.791 \\
     \bottomrule
@@
 \end{table}
+
+\begin{figure}
+\centering
+\includegraphics[width=1.0\textwidth]{figures/fig00.png}
+\caption{\label{fig:fig00} Example items from the MaRs-IB.
+(A) A simple item containing a single feature and one relation change (i.e.
+only the colour changes).
+(B) A harder item containing three features and six relation changes.
+In both panels, the first option is the correct solution.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=1.0\textwidth]{figures/fig01.png}
+\caption{\label{fig:fig01}Summary of performance on the MaRs-IB items.
+(A) The distribution of total scores across all 1501 participants.
+(B) The distribution of proportion correct responses across all 384 items.
+(C) The distribution of participants' median response times across items broken down by participants' total scores and item difficulty (one minus proportion correct responses).
+Error bars indicate bootstrapped 95\% confidence intervals around the mean.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=1.0\textwidth]{figures/fig02.png}
+\caption{\label{fig:fig02}Results from the test assembly of three parallel MaRs-IB short form (SF) measures.
+(A) The psychometric properties of the items selected for each form, compared to all remaining items.
+Points represent the posterior means of the item difficulty and discrimination  parameters.
+(B) The test characteristic curves (TCCs) for each short form.
+The degree of overlap highlights the similarity of expected scores for each test form.
+(C) The test information functions (TIFs) for each short form.
+The degree of overlap highlights the similarity of reliability for each test form.}
+\end{figure}
 

### 2022-04-19 manuscript/supplement.tex
updated manuscript

\documentclass[a4paper,man,natbib]{apa6}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}

% override apa6 first paragraph index
% https://tex.stackexchange.com/questions/155028/indentation-apa6-class-for-2nd-paragraphs-after-title
% \makeatletter
%   \b@level@one@skip=-2.5ex plus -1ex minus -.2ex
%   \b@level@two@skip=-2.5ex plus -1ex minus -.2ex
% \makeatother

\title{Supplemental materials for ``An item response theory analysis of the Matrix Reasoning Item Bank (MaRs-IB)''}
\shorttitle{Supplemental materials for ``IRT analysis of MaRs-IB''}
\author{Samuel Zorowitz$^1$, Nathaniel D. Daw$^{1,2}$}
\affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$Department of Psychology, Princeton University, USA}

\setcounter{figure}{0}
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}

\begin{document}
\maketitle

\section*{Speed-accuracy trade-offs in Chierchia et al. (2019)}

To investigate the possibility of speed-accuracy trade-offs in the original MaRs-IB data, we looked at the proportion of correct responses to the easiest items (dimension 1/2 items) as a function of the number of participants having reached that item. In all partitions of these data, we found robust positive correlations between proportion correct and number reached (dimension 1 items: $\rho$ = 0.514, p = 0.050; dimension 2 items: $\rho$ = 0.767, p < 0.001; combined: $\rho$ = 0.695, p < 0.001). This result suggests that the participants that did reach items later in the sequence were able to do by sacrificing accuracy for speed. As such, the item-level performance summary released as part of Chierchia et al. (2019) are likely biased indicators of item functioning.

\begin{table}
\centering
\begin{tabular*}{\textwidth}{ccccccllll}
\toprule
 & & & & & & \multicolumn{4}{c}{Spearman rank correlation} \\
\cmidrule(lr){7-10}
Measure & & Mean (SD) & & IQR & &  NFC-10 & PCF-8a & SNS & MaRs-SF \\
\midrule
NFC-10 & & 25.03 (8.27) & & 20.00 -- 31.00 & & - &  &  & \\
PCF-8a   &  & 22.17 (6.29) & & 18.75 -- 26.25 & & 0.27** &  - &  &  \\
SNS   &  & 29.07 (7.48) & & 25.00 -- 35.00 & & 0.46** &  0.29** &  - &   \\
MaRs-SF & &   8.00 (2.53) & &   6.00 -- 10.00 & & -0.04 &  0.04 &  0.14* &  - \\
\bottomrule
\end{tabular*}
\captionsetup{width=1.\textwidth}
\caption{\normalfont Correlations between performance on the MaRs-IB short forms and self-report measures. IQR = interquartile range; NFC-10 = need for cognition (10-item) scale; PCF = PROMIS cognitive functioning scale (8a); SNS = subjective numeracy scale. \\ ** p < 0.001,  * p < 0.05 (not corrected for multiple comparisons)}
\end{table}
\end{document}

### 2022-04-20 manuscript/main.tex
updated manuscript

@@
 \shorttitle{IRT analysis of MaRs-IB}
-\author{Samuel Zorowitz$^1$, Nathaniel D. Daw$^{1,2}$}
+\author{Samuel Zorowitz$^1$, Gabriele Chierchia(?), Sarah-Jayne Blakemore(?), Nathaniel D. Daw$^{1,2}$}
 \affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$Department of Psychology, Princeton University, USA}
 
-\abstract{Your abstract here.}
+\abstract{Matrix reasoning tasks are among the most widely used measures of cognitive ability in the behavioral sciences, but the paucity of matrix reasoning tests in the public domain complicates their use.
+Here we present an extensive investigation and psychometric validation of the matrix reasoning item bank (MaRs-IB), an open-access set of matrix reasoning items.
+In a first study, we calibrate the psychometric functioning of the items in the MaRs-IB in a large sample adults participants (N=1501).
+Using additive multilevel item structure models, we establish that the MaRs-IB has many desirable psychometric properties: its items span a wide range of difficulty, possess medium- to-large levels of discrimination, and exhibit as intended certain associations between item attributes and functioning (e.g.
+item complexity predicts item difficulty).
+However, we also find that item clones are not psychometrically equivalent and therefore not exchangeable.
+In a second study, we demonstrate how researchers can use the item parameters we estimated to design new MaRs-IB tests using optimal item assembly.
+With these methods, we design three parallel short-form measures that exhibited good score reliability and convergent validity in a second sample of adults (N=300).
+We hope that the materials and results made available here will encourage researchers to use the MaRs-IB in their studies.}
 
@@
 In low-stakes testing settings, matrix reasoning tasks can additionally function as measures of motivation, willingness to expend mental effort, and other aspects of personality \citep{duckworth2011role, gignac2019maximum}.
-Performance on matrix reasoning tasks can also serve as an instrumental variable in studies of psychiatric populations; that is, it can be used to control for general disruptions to cognitive ability where specific domains of cognition are primarily of interest \citep{gillan2016characterizing, rouault2018psychiatric, moutoussis2021decision}.
+Performance on matrix reasoning tasks can also serve as an instrumental variable in studies of psychiatric populations; that is, it can be used to account for general disruptions to cognitive ability where specific domains of cognition are primarily of interest \citep{gillan2016characterizing, rouault2018psychiatric, moutoussis2021decision}.
 
@@
 
+\section{Acknowledgements}
+\noindent This work was funded in part by the National Center for Advancing Translational Sciences (NCATS), a component of the National Institute of Health (NIH), under award number UL1TR003017, and by an NSF Graduate Research Fellowship.
+
 \bibliography{references}

### 2022-05-09 02_Design/scripts/censor_3pl.py
checkpoint analyses

@@ -1,173 +1,0 @@
-import os, sys
-import numpy as np
-from os.path import dirname
-from cmdstanpy import CmdStanModel
-from pandas import DataFrame, concat
-ROOT_DIR = dirname(dirname(os.path.realpath(__file__)))
-
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-### Define parameters.
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-
-## I/O parameters.
-stan_model = sys.argv[1]
-
-## RNG parameters.
-seed = int(sys.argv[2])
-
-## Simulation parameters.
-n_subj = 1500               # number of total subjects
-n_item = 128                # number of total items
-n_test = 16                 # number of items per subject
-
-## Latent correlations.
-rho = np.sqrt(0.5)
-
-## Censor parameters.
-thresholds = [0.0, 0.15, 0.25, 0.35]
-
-## Sampling parameters.
-iter_warmup   = 2000
-iter_sampling = 1000
-chains = 4
-thin = 1
-parallel_chains = 4
-
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-### Simulate data.
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-np.random.seed(seed)
-
-def inv_logit(x):
-    return 1 / (1 + np.exp(-x))
-
-def zscore(x):
-    return (x - np.mean(x)) / np.std(x)
-
-## Define metadata.
-N = n_subj * n_test
-
-## Define subject-to-trial mapping.
-J = np.repeat(np.arange(n_subj), n_test)
-
-## Define item-to-trial mapping.
-K = []
-a = np.arange(0, n_item, n_item // n_test)
-b = np.arange(n_item / n_test).astype(int)
-for _ in range(n_subj): K.append(a + np.random.choice(b, n_test, replace=True))
-K = np.concatenate(K)
-
-## Define subject parameters.
-cov = [[1,rho],[rho,1]]
-theta, X1 = np.random.multivariate_normal(np.zeros(2), cov, n_subj).T
-X1 = zscore(X1).reshape(-1,1)
-_, M1 = X1.shape
-
[…]

### 2022-05-09 02_Design/scripts/simulate_irt.py

@@ -1,148 +1,0 @@
-import os, sys
-import numpy as np
-from os.path import dirname
-from cmdstanpy import CmdStanModel
-from pandas import DataFrame, concat
-from scipy.stats import norm
-ROOT_DIR = dirname(dirname(os.path.realpath(__file__)))
-
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-### Define parameters.
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-
-## I/O parameters.
-stan_model = '2pl'
-seed = int(sys.argv[1])
-
-## Simulation parameters.
-n_subj  = 800                # number of total subjects
-n_item  = 60                 # number of total items
-n_split = 4                  # number of item sets
-
-## Sampling parameters.
-iter_warmup   = 2000
-iter_sampling = 1000
-chains = 4
-thin = 1
-parallel_chains = 4
-
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-### Simulate data.
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-np.random.seed(seed)
-
-def inv_logit(x):
-    return 1 / (1 + np.exp(-x))
-    
-## Define item order.
-ix = np.arange(n_item).reshape(n_split, n_item // n_split, order='F')
-K = np.zeros((n_subj, n_item // n_split), dtype=int)
-for i in range(n_subj): 
-    ix = np.apply_along_axis(np.random.permutation, 0, ix)
-    K[i] = ix[0]
-    
-## Define metadata.
-K = K.flatten()
-J = np.repeat(np.arange(n_subj), n_item // n_split)
-N = J.size
-
-## Generate subject abilities.
-theta = np.random.normal(0, 1, n_subj)
-
-## Generate item parameters.
-cov = [[1.5,0.0],[0.0,0.25]]
-beta, alpha = np.random.multivariate_normal(np.zeros(2), cov, n_item).T
-        
-## Sort parameters.
-alpha = np.exp(alpha[np.argsort(beta)])
-beta  = beta[np.argsort(beta)]
-
-## Generate data.
-mu = inv_logit(alpha[K] * theta[J] - beta[K])
[…]

### 2022-05-09 03_Calibration/scripts/fit_irt.py

@@ -53,15 +53,9 @@ X1 = read_csv(os.path.join(ROOT_DIR, 'designs', 'X1.csv'))
 X1 = X1.apply(zscore, 0).values
 _, M1 = X1.shape
     
-## Define variables of interest.
-if stan_model in ['1pl_m1', '1plg_m1', '1pl_m2', '1plg_m2']:
-    cols = ['intercept','n_features','n_rules']
-else:
-    cols = ['intercept','n_features','n_rules','distractor','rt']
-    
 ## Define item feature matrix.
 X2 = read_csv(os.path.join(ROOT_DIR, 'designs', 'X2.csv'))
-X2 = X2[cols].apply(zscore, 0).fillna(1).values
+X2 = X2.apply(zscore, 0).fillna(1).values
 _, M2 = X2.shape
 
 ## Define item family matrix.

### 2022-05-09 03_Calibration/scripts/fit_irt_sgddm.py
upgrade to sgddm

@@ -15,6 +15,9 @@ np.random.seed(47404)
 ## I/O parameters.
 stan_model = sys.argv[1]
 
+## SGDDM paramers.
+thin = 5
+
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
 ### Load and prepare data.
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
@@ -47,6 +50,7 @@ Y = data.accuracy.values.astype(int)
 
 ## Load DataFrame.
 StanFit = read_csv(os.path.join(ROOT_DIR, 'stan_results', f'{stan_model}.tsv.gz'), sep='\t', compression='gzip')
+n_samp = len(StanFit)
 
 ## Extract parameters.
 theta = StanFit.filter(regex='theta\[').values
@@ -57,89 +61,101 @@ if not np.any(alpha): alpha = np.ones_like(beta)
 ## Define guessing rate.
 if '1plg' or '3pl' in stan_model: gamma = 0.25
 else: gamma = 0.00
-    
+   
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
 ### Posterior predictive check.
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+np.random.seed(47404)
 
 @njit
 def inv_logit(x):
     return 1. / (1 + np.exp(-x))
 
-## Define number of posterior samples.
-n_iter = len(StanFit)
+## Preallocate space.
+R_obs = np.zeros((n_samp, N))
+R_hat = np.zeros((n_samp, N))
 
-## Compute linear predictor.
-mu = np.zeros((n_iter, N))
-for n in tqdm(range(N)): mu[:,n] = alpha[:,K[n]] * theta[:,J[n]] - beta[:,K[n]]
+## Iterate over samples.
+for n in tqdm(range(N)):
     
-## Compute p(correct | theta).
-p = gamma + (1-gamma) * inv_logit(mu)
-
-## Simulate accuracy.
-y_hat = np.random.binomial(1, p)
-
-## Precompute residuals.
-R = Y - p
-R_hat = y_hat - p
-
-## Clear memory.
-del y_hat, p, theta, beta, alpha
-
+    ## Compute linear predictor.
[…]

### 2022-05-09 03_Calibration/scripts/fit_irt_x2.py

@@ -134,4 +134,4 @@ info = DataFrame(info)
 fout = os.path.join(ROOT_DIR, 'stan_results', f'{stan_model}')
 
 ## Save.
-info.to_csv(f'{fout}_ifdm.csv', index=False)
+info.to_csv(f'{fout}_x2.csv', index=False)

### 2022-05-09 manuscript/main.tex
checkpoint manuscript before edits

@@
 \usepackage{graphicx}
+\usepackage{xurl}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 \shorttitle{IRT analysis of MaRs-IB}
-\author{Samuel Zorowitz$^1$, Gabriele Chierchia(?), Sarah-Jayne Blakemore(?), Nathaniel D. Daw$^{1,2}$}
-\affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$Department of Psychology, Princeton University, USA}
+\author{Samuel Zorowitz$^1$, Gabriele Chierchia$^3$, Sarah-Jayne Blakemore$^3$, Nathaniel D. Daw$^{1,2}$}
+\affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$Department of Psychology, Princeton University, USA\\$^3$Department of Psychology, University of Cambridge, Downing Street, Cambridge, UK}
 
@@
 They are predictive of important real-world outcomes such as childhood academic performance \citep{roth2015intelligence} and performance on college entrance exams \citep{frey2004scholastic, koenig2008act}.
-In low-stakes testing settings, matrix reasoning tasks can additionally function as measures of motivation, willingness to expend mental effort, and other aspects of personality \citep{duckworth2011role, gignac2019maximum}.
-Performance on matrix reasoning tasks can also serve as an instrumental variable in studies of psychiatric populations; that is, it can be used to account for general disruptions to cognitive ability where specific domains of cognition are primarily of interest \citep{gillan2016characterizing, rouault2018psychiatric, moutoussis2021decision}.
+In low-stakes testing settings (i.e.
+where poor performance incurs marginal or no cost to the participant), matrix reasoning tasks can additionally function as measures of motivation, willingness to expend mental effort, and other aspects of personality \citep{duckworth2011role, gignac2019maximum}.
+In studies of psychiatric populations, performance on matrix reasoning tasks have also been used to account for general disruptions to cognitive ability where specific domains of cognition are primarily of interest \citep{gillan2016characterizing, rouault2018psychiatric, moutoussis2021decision}.
 
@@
 One such challenge is the problem of copyright.
-The most prominent examples of matrix reasoning tasks, such as Raven's progressive matrices \citep{raven2003raven} and the WAIS matrix subtest \citep{wechsler1999wechsler}, are not free to use and have legal restrictions against being computerized.
+The most prominent examples of matrix reasoning tasks, such as the WAIS and WASI matrix reasoning subtests \citep{wechsler1999wechsler, corporation2008wechsler}, are not free to use and have legal restrictions against being computerized.
 A second challenge is that, across all of the matrix reasoning tests in the public domain, there are relatively few unique items available.
@@
 
-Noting these challenges, \cite{chierchia2019matrix} developed and made publicly available the matrix reasoning item bank (MaRs-IB), a collection of 80 unique matrix reasoning puzzles.
+Noting these challenges, \cite{knoll2016window} developed and  \cite{chierchia2019matrix} made publicly available the matrix reasoning item bank (MaRs-IB), a collection of 80 unique matrix reasoning puzzles.
 Each puzzle in the MaRs-IB consists of a 3x3 matrix containing abstract shapes in eight out of nine cells (Figure \ref{fig:fig00}).
@@
 \cite{chierchia2019matrix} also conducted an initial study of the psychometric properties of the MaRs-IB.
-A sample of 659 participants completed a fixed-length and fixed-order test measure in which they had eight minutes to complete as many MaRs-IB items as possible.
+A sample of 659 adult, adolescent, and child participants completed a fixed-length and fixed-order test measure in which they had eight minutes to complete as many MaRs-IB items as possible.
 The authors found that performance on this test measure had good split-half and test-retest reliability.
@@
 
-Unfortunately, there are two critical limitations in this study that undermine the utility of the summary statistics.
+There are two critical limitations in this study that undermine the utility of the summary statistics.
 The first is that, given design of the test, there was relatively little data collected for the majority of the items in the MaRs-IB.
@@
 The remainder of the paper proceeds as follows.
-First we report a calibration study, in which a large sample of participants each completed a subset of items from the MaRs-IB.
+First we report a calibration study in which a large sample of participants each completed a subset of items from the MaRs-IB.
 Using their response data, we then fit a series of multilevel item structure models in order to estimate item parameters for each item (and their clones).
@@
 
-All data, code, and model outputs (including the item parameter estimtes) are publicly available at: \url{https://github.com/dawlab/mars-irt}.
+All data, code, and model outputs (including the item parameter estimates) are publicly available at: \url{https://github.com/dawlab/mars-irt}.
 
@@
 To ensure data quality, the data from multiple participants were excluded prior to analysis (see Exclusion Criteria below) leaving the data from N=1501 participants for analysis.
-In these participants, the majority identified as women (men: N=670; women: N=811; non-binary or other: N=13; rather not say: N=2) and the average age was 28.7 years old (sd = 9.9).
+In these participants, the majority identified as women (men: N=670; women: N=811; non-binary or other: N=13; rather not say: N=2) and the average age was 28.7 years old (sd = 9.9; range: 18 -- 74).
 The sample was relatively well-educated with the majority having completed a Bachelor's degree (N=507) or Master's degree or higher (N=322).
@@
 The format of the experiment and instructions were adapted from code publicly released as part of the original study (\url{https://app.gorilla.sc/openmaterials/36164}).
+The task was programmed in jsPsych \citep{de2015jspsych} and distributed using custom web-application software (available at \url{https://github.com/nivlab/nivturk}).
 
@@
 Across these participants, each item (64 total) was administered approximately 375 times (sd = 12) and each item clone (384 total) was administered approximately 62 times (sd = 7).
+
+\subsubsection{Statistical analysis}
+
+We investigated the relationship between accuracy and response time using a mixed effects (random intercepts) linear regression model.
+Participants' log-transformed response times were predicted as a function of that trial accuracy, rest score (participants' observed scores on all other items), and item difficulty (one minus the proportion correct for that item).
+The mixed-effects model was estimated using the \textit{statsmodels} python package (v0.12.2) \citep{seabold2010statsmodels}).
 
@@
 
-\item $\sum_{n=1}^N Q_{jn} \delta_{\beta n}$: the effect of the template (level 1) attributes on item difficulty, where $Q_{jn}$ is the value of attribute $n$ for item template $j$, and $\delta_{\beta n}$ is the effect of attribute $n$.
+\item $\sum_{n=1}^N Q_{jn} \delta_{\beta n}$: the effect of the template (level 1) attributes (desribed below) on item difficulty, where $Q_{jn}$ is the value of attribute $n$ for item template $j$, and $\delta_{\beta n}$ is the effect of attribute $n$.
 
@@
 In our analyses, we considered two template level (level 1) attributes: feature number, or the number of elements in each cell of an item; and rule number, or the number of changing relationships across the cells of an item (e.g.
-color change, shape change, position change).
+size change, color change, shape change, position change).
+The feature and rule number for each item was obtained via manual annotation.
+Across item templates, the modal number of features was two (range: 1 -- 4) and the modal number of rules was three (range: 1 -- 6).
+Feature and rule number were largely uncorrelated across items ($\rho$ = 0.176).
+Additional examples of items are provided in the supplementary materials.
 We predicted that both attributes would be positively associated with item difficulty; that is, all else equal, items with a great number of features or rules would be more difficult to solve.
@@
 
+It is important to note that we deliberately chose not to include shape type as a clone attribute; that is, we elected not to directly model the effects of lower-level perceptual features on item functioning.
+As previously mentioned, each item template has three alternate forms that vary in their perceptual stimuli.
+Every item clone in the MaRs-IB is comprised of stimuli drawn one of nine stimulus sets (Figure S1).
+Each stimulus set is itself made up of approximately 15 unique stimuli (some elementary shapes, like squares and circles, are shared across stimulus sets).
+These stimuli are pseudo-randomly distributed across item clones of a given stimulus set; that is, two item clones with the same stimulus set and feature number are not guaranteed to be comprised of the same stimuli.
+As such, the most rigorous way of modeling perceptual features on item functioning would have been the inclusion of approximately 45 additional attributes for both item difficulty and discrimination.
+Rather than deal with this complexity, however, we opted to leave this unmodeled, instead to be soaked up by the residual variance term.
+
 The AMIS model framework provides a natural means of accomplishing the three goals of the calibration the study.
@@
 Finally, by estimating the magnitude of residual error in item parameters from unmodeled sources of variance (e.g.
-perceptual features), we can test whether the MaRs-IB possesses certain desirable structural properties (e.g.
+lower-level perceptual features), we can test whether the MaRs-IB possesses certain desirable structural properties (e.g.
 psychometric equivalence and exchangeability among item clones; Aim 3).
@@
 Across all 384 item clones, the average proportion correct was 0.60 (sd = 0.19, IQR = 0.45 - 0.74; Figure \ref{fig:fig01}b).
-A total of 12 items (3.1\%) exhibited performance levels beneath chance, and only 22 items (5.7\%) had performance levels near ceiling ($\geq$ 90\%).
+A total of 12 item clones (3.1\%) exhibited performance levels beneath chance, and only 22 item clones (5.7\%) had performance levels near ceiling ($\geq$ 90\%).
 Across items, the proportion of correct responses was negatively correlated with rule number ($\rho$ = -0.482, p < 0.001).
@@
 
-We investigated the relationship between accuracy and response time using a mixed effects (random intercepts) linear regression model.
-Participants' log-transformed response times were predicted as a function of that trial accuracy, rest score (participants' observed scores on all other items), and item difficulty (one minus the proportion correct for that item).
-The mixed-effects model was estimated using the \textit{statsmodels} python package (v0.12.2) \citep{seabold2010statsmodels}).
+\subsubsection{Response time statistics}
 
@@
 We fit AMIS models in two waves.
-In the first wave of models, the item discrimination parameters were fixed ($\alpha = 1$).\footnote{Note that these models still included a fixed guessing parameter ($\gamma = 0.25$).
+In the first wave, the item discrimination parameters were held fixed ($\alpha = 1$).\footnote{These models still included a fixed guessing parameter ($\gamma = 0.25$).
 We also estimated models with no guessing parameter (i.e.
-one-parameter logistic or Rasch models), but model comparison indicated these models provided worse fits to the data (Table SXX).} We applied this restriction to the first set of models in order to identify the best structure for the item difficulty parameters independent of the discrimination parameters.
-In the first wave, we fit four models:
+one-parameter logistic or Rasch models), but model comparison indicated these models provided worse fits to the data (Table S1).} We applied this restriction to the first set of models in order to identify the best structure for the item difficulty parameters independent of the discrimination parameters.
+In the first wave, we fit three models:
 
@@
 
-\item Model 1: Item difficulty is a function only of item template attributes (i.e.
-feature number, rule number).
-In this model, item with identical attributes are psycho- metrically equivalent, as are all the clones belonging to an item.
+\item Model 1: Item difficulty is a function of item template attributes (i.e.
+feature number, rule number) and item clone attributes (i.e.
+distractor type, response time).
+In this model, the difficulty of an item is explained solely
+
+item with identical attributes are psycho- metrically equivalent, as are all the clones belonging to an item.
 
@@
 
-\item Model 3: Item difficulty is a function of both item template and clone attributes (i.e.
-distractor type, mean response time), as well as template-level residuals.
-Item clones may now vary in their difficulty as a function of their attributes.
-
-\item Model 4: Item difficulty is function of attributes and residuals at both levels.
+\item Model 3: Item difficulty is function of attributes and residuals at both levels.
 In this model, neither item templates nor clones with identical attributes are necessarily equally difficult, depending on the magnitude of the residual terms.
@@
 
-Feature number, rule number, and distractor type cumulatievely explained approximately 39.2\% of the variance in item discrimination across the 64 item templates.
+Feature number, rule number, and distractor type cumulatively explained approximately 39.2\% of the variance in item discrimination across the 64 item templates.
 In contrast to item difficulty, model comparison indicated there was non-negligible residual variability in item discrimination only at template level.
@@
 Our aim was to verify that observed scores on each short form conformed to our expectations based on the TCCs and TIFs.
-Successfully doing so would provide additional evidence that we were successful in calibrating the functioning of the items in the MaRs-IB.
-We also sought to establish the convergent validity by correlating observed scores on the short form measures with performance on an abbreviated version of the Raven's progressive matrices.
+Doing so would provide additional evidence of whether we were successful in calibrating the functioning of the items in the MaRs-IB.
+We also sought to establish the convergent validity by correlating observed scores on the short form measures with performance on an abbreviated version of the Raven's progressive matrices \citep{bilker2012development}.
 
@@
 The majority of participants identified as women (men: N=138; women: N=147; non-binary or other: N=13; rather not say: N=2).
-Participants were 35.0 years old (sd = 13.2) on average.
+Participants were on average 35.0 years old (sd = 13.2; range: 18 -- 76).
 The sample was relatively well-educated, with the majority having completed a Bachelor's degree (N=121) or Master's degree or higher (N=45).
@@
 Considering the attenuation of correlations between two imperfect measures \citep{spearman1961proof}, we should not expect correlations much larger than $\rho \approx 0.64$ (or $0.80 \times 0.80$).
-Therefore, this result indicates good convergent validity between the two matrix reasoning measures.
+Therefore, this result indicates satisfactory convergent validity between the two matrix reasoning measures.
 Briefly, in this second sample, neither age nor gender was credibly associated with performance on the MaRs-IB short forms (age: $\rho$ = -0.123, 95\% HDI = -0.253 -- 0.013; gender: $\rho$ = -0.025, 95\% HDI = -0.160 -- 0.108).
@@
 
-\bibliography{references}
+\bibliography{main}
 
@@
     \centering
-    \begin{tabular}{cccccccc}
+    \begin{tabular*}{\textwidth}{clcr}
     \toprule
-    Model & FE-1 & RE-1 & FE-2 & RE-2 & psis-loco & $\Delta$ psis-loco (se) \\
+    Model & Description & psis-loco & $\Delta$ psis-loco (se) \\
     \midrule
-    1 & x &   &   &   & -14516.85 & -1223.69 (44.94)\\
-    2 & x & x &   &   & -13804.74 & -511.58 (28.77) \\
-    3 & x & x & x &   & -13481.12 & -187.96 (17.25) \\
-    4 & x & x & x & x & -13293.16 & - \\
+    1 & TK  & 27741.4 & 1155.8 (60.1) \\
+    2 & TK  & 26962.1 & 376.5 (34.5) \\
+    3 & TK & 26585.5 & \multicolumn{1}{c}{-} \\
     \midrule
-    5 & x &   & x &   & -13209.38 & -2.13 (2.04) \\
-    6 & x & x & x &   & -13207.25 & - \\
-    7 & x & x & x & x & -13208.69 & 1.44 (0.85) \\
+    4 & TK & 26419.6 & 5.12 (4.07) \\
+    5 & TK & 26414.5 & \multicolumn{1}{c}{-} \\
+    6 & TK & 26417.8 & 3.29 (1.84) \\
     \bottomrule
-    \end{tabular}
-    \caption{\label{tab:2} \normalfont Goodness of fit of item response models to MaRs-IB data.
-PSIS = Pareto-smoothed importance sampling.
-Note: LOCO values are presented in original, not deviance, scale so larger numbers indicate better fit.}
+    \end{tabular*}
+    \caption{\label{tab:2}\normalfont Goodness-of-fit of item response models to MaRs-IB data.
+LOCO values are presented in deviance scale (i.e.
+smaller numbers indicate better fit).
+Abbreviations: PSIS = Pareto-smoothed importance sampling; LOCO = leave-one-cluster-out.}
     \label{table:2}
@@
     \end{tabular*}
-    \caption{\label{table:3} \normalfont Summary of performance on the MaRs-IB and Raven's progressive matrices (RPM) short form (SF) measures. IQR = interquartile range.}
+    \caption{\label{table:3}\normalfont Summary of performance on the MaRs-IB and Raven's progressive matrices (RPM) short form (SF) measures. IQR = interquartile range.}
 \end{table}
@@
 \caption{\label{fig:fig00} Example items from the MaRs-IB.
-(A) A simple item containing a single feature and one relation change (i.e.
-only the colour changes).
-(B) A harder item containing three features and six relation changes.
-In both panels, the first option is the correct solution.}
+(A) A simple item containing one feature per cell and two relation changes (i.e.
+a change of shape and its position changes across rows and columns, respectively).
+(B) A harder item containing three features and six relation changes (i.e.
+three position changes, two color changes, one shape change).
+(C/D) Variants of Items A and B, respectively, matched on complexity (i.e.
+number of features and relation changes).
+In all panels, the first option is the correct response.}
 \end{figure}

### 2022-05-09 manuscript/supplement.tex
checkpoint manuscript before edits

@@
 \shorttitle{Supplemental materials for ``IRT analysis of MaRs-IB''}
-\author{Samuel Zorowitz$^1$, Nathaniel D. Daw$^{1,2}$}
-\affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$Department of Psychology, Princeton University, USA}
+\author{Samuel Zorowitz$^1$, Gabriele Chierchia$^3$, Sarah-Jayne Blakemore$^3$, Nathaniel D. Daw$^{1,2}$}
+\affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$Department of Psychology, Princeton University, USA\\$^3$Department of Psychology, University of Cambridge, Downing Street, Cambridge, UK}
 
@@
     \centering
+    \begin{tabular*}{\textwidth}{clcr}
+    \toprule
+    Model & Description & psis-loco & $\Delta$ psis-loco (se) \\
+    \midrule
+    7 & TK  & 27741.4 & 635.3 (29.7) \\
+    8 & TK  & 26962.1 & 665.6 (32.2) \\
+    9 & TK & 26585.5 & 633.1 (33.9) \\
+    \bottomrule
+    \end{tabular*}
+    \caption{\label{tab:2}\normalfont Goodness-of-fit of item response models to MaRs-IB data.
+LOCO values are presented in deviance scale (i.e.
+smaller numbers indicate better fit).
+Abbreviations: PSIS = Pareto-smoothed importance sampling; LOCO = leave-one-cluster-out.}
+    \label{table:2}
+\end{table}
+
+\begin{table}
+\centering
 \begin{tabular*}{\textwidth}{ccccccllll}
@@
 \caption{\normalfont Correlations between performance on the MaRs-IB short forms and self-report measures.
-IQR = interquartile range; NFC-10 = need for cognition (10-item) scale; PCF = PROMIS cognitive functioning scale (8a); SNS = subjective numeracy scale.
+Abbreviations: IQR = interquartile range; NFC-10 = need for cognition (10-item) scale; PCF = PROMIS cognitive functioning scale (8a); SNS = subjective numeracy scale.
 \\ ** p < 0.001,  * p < 0.05 (not corrected for multiple comparisons)}
 \end{table}
+
+\begin{figure}
+\centering
+\includegraphics[width=1.0\textwidth]{figures/figS01.png}
+\caption{\label{fig:figS01} An example item clone from each of the nine stimulus sets.
+In each row are three clones of one item template.
+Though some elementary shapes (e.g.
+circles, squares) are shared across stimulus sets, the majority of stimuli in each set are unique.}
+\end{figure}
+
 \end{document}

### 2022-05-10 manuscript/supplement.tex
updated supplementary materials

@@
 This result suggests that the participants that did reach items later in the sequence were able to do by sacrificing accuracy for speed.
-As such, the item-level performance summary released as part of Chierchia et al.
-(2019) are likely biased indicators of item functioning.
+As such, the item-level performance summary released as part of \cite{chierchia2019matrix} are likely biased indicators of item functioning.
+
+\section{Defining a threshold for rapid guessing}
+
+In online testing environments, it is inevitable that some participants will not engage meaningfully with an experiment and instead in engage in careless or insufficient effort responding.
+On matrix reasoning tasks, one such low-effort strategy is rapid guessing wherein participants response in such a short time that there is no way they could have meaningfully considered an item \citep{wise2017rapid}.
+In sufficient quantities, the presence of rapid guesses in data can systematically bias estimates of item parameters.
+Thus, if possible, rapid guess responses should ideally be identified and removed.
+
+There are a number of approaches for identifying rapid guess responses (for a review, see \cite{wise2017rapid}).
+We opted for a threshold approach, in which responses taking less than a particular time would be denoted as rapid guesses and participants exhibiting too many rapid guessing responses would be excluded from the data.
+To define this threshold, we fit an extended version of the effort-moderated item response theory (EM-IRT) model \citep{wise2006application} to a small dataset of responses collected during piloting.
+In the EM-IRT, the probability of correct responding for participant $i$ to item $j$ is defined as the following mixture:
+
+\begin{equation*}
+    p(y_{i,j} = 1) = (1-w_{ij}) \cdot \gamma + w_{ij} \cdot \text{logit}^{-1} \left( \alpha_{j} \cdot \theta_i - \beta_{j} \right)
+\end{equation*}
+
+\noindent where $\theta_i$ is the latent ability for person $i$, and $\beta_{j}$, $\alpha_{j}$, and $\gamma_{j}$ are the difficulty, discrimination, and guessing parameters for item $j$.
+As in the main text, here we fixed the guessing parameter for every item clone to the nominal guessing rate ($\gamma_{j} = 0.25$).
+Crucially, $w_{ij}$ is a weight parameter, bounded between zero and one, that controls whether a participant is responding effortfully in accordance with their ability ($w_{ij} \rightarrow 0$) or engaging in rapid guessing responding ($w_{ij} \rightarrow 1$).
+
+Here we defined the rapid guessing weight as a function of participant's response time on that trial:
+
+\begin{equation*}
+    w_{ij} = \text{logit}^{-1} \left( \zeta_0 + z_{ij} \cdot \zeta_1 + z_{ij}^2 \cdot \zeta_2 \right)
+\end{equation*}
+
+\noindent where $z_{ij}$ is the (log-transformed) response time for participant $i$ for item $j$, and  $\zeta_n$ are regressing coefficients mapping responses times to rapid guessing weights.
+Thus, this form of the EM-IRT model learns a function in a data-driven fashion to classify responses as having originated from effortful or rapid guessing response strategies.
+
+We fit this model to data collected from a total of N=180 participants recruited from the Prolific Academic platform during piloting.
+Each participant completed one of two sets of eight items from the MaRs-IB.
+The EM-IRT model estimated within a Bayesian framework using Hamiltonian Monte Carlo as implemented in Stan (v2.22) \citep{carpenter2017stan}.
+Four separate chains with randomised start values each took 3,000 samples from the posterior.
+The first 2,000 samples from each chain were discarded.
+As such, 4,000 post-warmup samples from the joint posterior were retained.
+The $\hat{R}$ values for all parameters were equal to or less than 1.01, indicating acceptable convergence between chains, and there were no divergent transitions in any chain.
+
+The estimated weights ($w_{ij}$) across all subjects and items are plotted as a function of their corresponding response time in Figure S2.
+As can be observed, the weights quickly approach 1 for response times faster than 5 seconds.
+Interestingly, weights begin to rise again for responses after 20 seconds suggesting that participants have an internal awareness of the amount of time that has elapsed.
+
+To define a threshold for rapid guessing, we found the corresponding response time for which $w = 0.5$.
+This was at approximately 3 seconds.
+Thus, in the main experiments we defined rapid guessing as responses taking fewer than 3 seconds.
+
+\bibliography{supplement}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{figures/figS02.png}
+\caption{\label{fig:figS02} Estimates of the rapid guessing weights ($w_{ij}$) and their corresponding responses times from the effort-moderated item response theory (EM-IRT) model fit to response data from N=180 pilot participants.
+The dashed line indicates the chosen rapid guessing threshold, i.e.
+where $w = 0.5$.}
+\end{figure}
+
 \end{document}

### 2022-05-11 03_Calibration/scripts/fit_irt_sgddm.py
updated ppc

@@ -82,7 +82,7 @@ for n in tqdm(range(N)):
     mu = mu = alpha[:,K[n]] * theta[:,J[n]] - beta[:,K[n]]
     
     ## Compute p(response).
-    p = inv_logit(mu)
+    p = gamma + (1-gamma) * inv_logit(mu)
     
     ## Compute expectation.
     expected = p.mean()

### 2022-05-11 03_Calibration/scripts/fit_irt_x2.py

@@ -15,6 +15,9 @@ np.random.seed(47404)
 ## I/O parameters.
 stan_model = sys.argv[1]
 
+## Define X2 parameters.
+minlength = 17
+
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
 ### Load and prepare data.
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
@@ -47,7 +50,7 @@ Y = data.accuracy.values.astype(int)
 
 ## Load DataFrame.
 StanFit = read_csv(os.path.join(ROOT_DIR, 'stan_results', f'{stan_model}.tsv.gz'), sep='\t', compression='gzip')
-# StanFit = StanFit.loc[::10]
+n_samp = len(StanFit)
 
 ## Extract parameters.
 theta = StanFit.filter(regex='theta\[').values
@@ -58,80 +61,55 @@ if not np.any(alpha): alpha = np.ones_like(beta)
 ## Define guessing rate.
 if '1plg' or '3pl' in stan_model: gamma = 0.25
 else: gamma = 0.00
-    
+   
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
 ### Posterior predictive check.
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-print('Performing posterior predictive check.')
+np.random.seed(47404)
 
 @njit
 def inv_logit(x):
     return 1. / (1 + np.exp(-x))
 
-## Define number of posterior samples.
-n_iter = len(StanFit)
+## Preallocate space.
+Y_hat = np.zeros((n_samp, N))
 
-## Compute linear predictor.
-mu = np.zeros((n_iter, N))
-for n in tqdm(range(N)): mu[:,n] = alpha[:,K[n]] * theta[:,J[n]] - beta[:,K[n]]
+## Iterate over samples.
+for n in tqdm(range(N)):
     
-## Compute p(correct | theta).
-p = gamma + (1-gamma) * inv_logit(mu)
-
-## Simulate accuracy.
-y_hat = np.random.binomial(1, p)
-
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-### Discrepancy measure.
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-
[…]

### 2022-05-11 manuscript/main.tex
manuscript edits

@@
 Matrix reasoning tasks are among the most widely used measures of cognitive ability in the behavioral sciences.
-Much of their popularity undoubtedly reflects their versatility in use.
+Much of their popularity undoubtedly reflects their versatility of use.
 Matrix reasoning tasks are strong (albeit impure) indicators of general intelligence \citep{gignac2015raven} and working memory capacity \citep{kane2004generality, unsworth2005working}.
-They are predictive of important real-world outcomes such as childhood academic performance \citep{roth2015intelligence} and performance on college entrance exams \citep{frey2004scholastic, koenig2008act}.
+They are predictive of important real-world outcomes such as childhood academic achievement \citep{roth2015intelligence} and performance on college entrance exams \citep{frey2004scholastic, koenig2008act}.
 In low-stakes testing settings (i.e.
-where poor performance incurs marginal or no cost to the participant), matrix reasoning tasks can additionally function as measures of motivation, willingness to expend mental effort, and other aspects of personality \citep{duckworth2011role, gignac2019maximum}.
-In studies of psychiatric populations, performance on matrix reasoning tasks have also been used to account for general disruptions to cognitive ability where specific domains of cognition are primarily of interest \citep{gillan2016characterizing, rouault2018psychiatric, moutoussis2021decision}.
+where a participant incurs little or no cost for poor performance), matrix reasoning tasks can additionally function as measures of motivation, willingness to expend mental effort, and other facets of personality \citep{duckworth2011role, gignac2019maximum}.
+In studies of psychiatric populations, performance on matrix reasoning tasks have also been used to control for general disruptions to cognitive ability when specific domains of cognition are of primary interest \citep{gillan2016characterizing, rouault2018psychiatric, moutoussis2021decision}.
 
-Unfortunately there are a number of impediments to using matrix reasoning tasks as part of behavioral research.
-One such challenge is the problem of copyright.
-The most prominent examples of matrix reasoning tasks, such as the WAIS and WASI matrix reasoning subtests \citep{wechsler1999wechsler, corporation2008wechsler}, are not free to use and have legal restrictions against being computerized.
+Unfortunately there are a number of obstacles to using matrix reasoning tasks as part of behavioral research.
+One challenge is the problem of copyright.
+Many of the most prominent matrix reasoning tasks, such as the WAIS and WASI matrix reasoning subtests \citep{wechsler1999wechsler, corporation2008wechsler}, are not free to use and have legal restrictions against being computerized.
 A second challenge is that, across all of the matrix reasoning tests in the public domain, there are relatively few unique items available.
@@
 The availability of only a limited number of items raises the possibility of repeat exposure effects, which threaten the validity of these measures  \citep{ng1974applicability, bors2003effect}.
-This problem is exacerbated in the current age of online experiments, where multiple groups of researchers may be unintentionally administering the same test to the same participants recruited from the same online labor platforms.
+This problem is exacerbated in the current era of online experiments, where multiple groups of researchers may be inadvertently administering the same test to the same participants recruited from the same online labor platforms.
 
-Noting these challenges, \cite{knoll2016window} developed and  \cite{chierchia2019matrix} made publicly available the matrix reasoning item bank (MaRs-IB), a collection of 80 unique matrix reasoning puzzles.
+Noting these challenges, \cite{knoll2016window} / \cite{chierchia2019matrix} developed and made publicly available the matrix reasoning item bank (MaRs-IB), a collection of 80 unique matrix reasoning puzzles.
 Each puzzle in the MaRs-IB consists of a 3x3 matrix containing abstract shapes in eight out of nine cells (Figure \ref{fig:fig00}).
-To solve a puzzle, participants must deduce which of four response options correctly completes the matrix based on the pattern of changing relations among the shapes across cells.
-The puzzles in the MaRs-IB were designed to span a considerable range of complexity, both in the number of unique shapes and changing relations in and across cells.
-To further increase the reusability of the MaRs-IB and minimize practice effects, each puzzle (hereafter referred to as a template item) has six clones.
-An item's clones share identical structural properties (i.e.
-number of shapes, number of changing relations) but may vary in their distractors (two unique sets per item) or perceptual features (three unique shape sets per item).
-Thus, the MaRs-IB addresses several issues associated with other matrix reasoning tests.
+Across cells, some or all of these shapes change according to various arbitrary abstract rules (e.g.
+color change, position change).
+To solve a puzzle, participants deduce which of four response options correctly completes the matrix based on these rules.
+The puzzles in the MaRs-IB vary in their complexity, both in the number of elements per cell and number of rules determining the relations of elements across cells (Figure \ref{fig:fig00}a/b).
+To increase the reusability of the MaRs-IB, each puzzle (hereafter referred to as an item template) has six clones.
+A single item's clones share identical structural characteristics (i.e.
+number of elements, number of rules) but may vary in their distractors (two unique sets per template) or perceptual features (three unique shape sets per template; Figure \ref{fig:fig00}c/d).
+Thus, the MaRs-IB addresses several of the issues associated with other matrix reasoning tests.
 
-\cite{chierchia2019matrix} also conducted an initial study of the psychometric properties of the MaRs-IB.
-A sample of 659 adult, adolescent, and child participants completed a fixed-length and fixed-order test measure in which they had eight minutes to complete as many MaRs-IB items as possible.
-The authors found that performance on this test measure had good split-half and test-retest reliability.
-Moreover they found that participants' ability to solve puzzles in the MaRs-IB was moderately predictive of their performance on a digit span task and the ICAR matrix reasoning test, indicating good convergent validity.
+\cite{chierchia2019matrix} conducted an initial study of the psychometric properties of the MaRs-IB.
+A sample of 659 adult, adolescent, and child participants completed an experiment in which they had eight minutes to complete as many MaRs-IB items as possible (items were presented in the same order to all participants).
+The authors found that performance on this task had good split-half and test-retest reliability.
+Moreover, they found that participants' ability to solve MaRs-IB puzzles was moderately predictive of their performance on a working memory task and the ICAR matrix reasoning test, indicating good convergent validity.
 At the conclusion of this study, \cite{chierchia2019matrix} publicly released a summary of the difficulty (i.e.
-proportion correct responses) of each item in the MaRs-IB so that future researchers could construct new test measures of custom difficulty and duration.
+proportion correct responses) of each item template and clone in the MaRs-IB so that other researchers could construct their own test measures of custom difficulty and duration.
 
-There are two critical limitations in this study that undermine the utility of the summary statistics.
-The first is that, given design of the test, there was relatively little data collected for the majority of the items in the MaRs-IB.
+Two important limitations in the experiment conducted by \cite{chierchia2019matrix} undermine the utility of these summary statistics, and therefore impede the design of new matrix reasoning tasks using the MaRs-IB.
+The first is that, given the fixed-order task design,  there was relatively few responses collected for the majority of the items in the MaRs-IB.
 Indeed, only half (42 of 80 items) were completed by 100 or more participants.
-As such, one cannot be confident about the reported functioning of a majority of the items (an issue that is exacerbated when considering that each item has six clones).
-A second issue is participants were allowed to choose for themselves whether to prioritize accuracy (number of correct responses) or productivity (number of items reached).
-As a consequence, the response data for items that came later in the fixed-order test were increasingly likely to be produced by participants sacrificing accuracy for speed, thereby biasing the proportion correct measure for these items (see the Supplementary materials for evidence of this point).
+As such, one cannot be confident about the true difficulty of a majority of the items (an issue that is exacerbated when considering that each item has six clones).
+A second issue is participants were able to choose for themselves whether to prioritize accuracy (number of correct responses) or productivity (number of items reached).
+As a consequence, response data for items that came later in the fixed-order task are increasingly likely to have been produced by participants sacrificing accuracy for speed, thereby biasing the proportion correct measure for these items (see the supplementary materials for evidence of this effect).
 Therefore, additional investigation of the functioning and psychometric properties of the MaRs-IB is warranted.
@@
 Here we provide a more extensive study and psychometric validation of the MaRs-IB.
-The most important contribution of the current work is the characterization of (almost) every item in the MaRs-IB using item response theory (IRT) \citep{embretson2013item, de2013theory}.
+The most important contribution of the current work is the characterization of (almost) every item in the MaRs-IB using item response theory (IRT; \citealt{embretson2013item, de2013theory}).
 IRT models confer several important advantages for the purposes of psychometric validation.
-First they provide an estimate of the difficulty and discrimination of each item, where the latter is an index of how much information an item provides given a level of ability.
-Using these quantities, IRT models in turn facilitate the computation of the test information function (TIF), which measures the reliability of a test measure across a range of ability.
-Finally, IRT models make possible optimal test assembly  \citep{van1998optimal}, or the design of new tests with maximal reliability given researcher-specified constraints (e.g.
+First they provide an estimate of the \textit{difficulty} and \textit{discrimination} of each item, where the latter is an index of how much information an item provides about ability.
+Using these quantities, IRT models in turn allow experimenters to compute the test information function (TIF), which quantifies the reliability of a test (i.e.
+an assemblage of items) in measuring performance across a range of ability levels.
+Finally, IRT models make possible optimal test assembly  \citep{van1998optimal}, or the design of new tests with maximal reliability given experimenter-defined constraints (e.g.
 test duration, test difficulty).
 
-A second contribution provided here is an investigation of the relationship between the attributes of MaRs-IB items and their psychometric properties using explanatory item response models \citep{de2004explanatory, wilson2008explanatory}.
-The utility of the MaRs-IB depends in part on its items exhibiting certain expected characteristics.
-Across items we should expect complexity (e.g.
-more shapes per cell, more changing relations) to be positively associated with difficulty.
-Moreover if item clones are exchangeable (i.e.
-perfect replicas with identical psychometric properties), then we should expect no association between distractor type and difficulty.
-Thus explanatory IRT models offer an opportunity to evaluate the success of the MaRs-IB design.
-As a secondary benefit, explanatory IRT models also provide more accurate estimates of item parameters \citep{neuhaus2006separating}.
+A second contribution provided here is an investigation of how item characteristics shape the psychometric functioning of items in the MaRs-IB.
+Using explanatory item response modeling \citep{de2004explanatory, wilson2008explanatory}, we investigate how the number of elements and number of rules impact item functioning (i.e.
+difficulty, discrimination).
+We modeled these two properties because they have previously been found to be among the strongest predictors of item difficulty in matrix reasoning tasks \citep{carpenter1990one, embretson1998cognitive}.
+This analysis therefore provides another means to evaluate the success of the MaRs-IB design.
+Insofar that the number of elements and rules govern the total amount of information that must be stored in working memory in order to correctly solve an item \citep{embretson1998cognitive}, confirmation that they impact performance would also provide insight into the cognitive reasons for individual differences in performance on the MaRs-IB.
+Finally, explanatory IRT models yield more accurate estimates of item parameters \citep{neuhaus2006separating}, which in turn helps to ensure that new MaRs-IB tests designed using these item parameter estimates function as expected.
 
-A third contribution is the decomposition of variance in item functioning, into modeled and unmodeled sources, using additive multilevel item structure (AMIS) models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
+%A third contribution is the decomposition of variance in item functioning, into modeled and unmodeled sources, using additive multilevel item structure (AMIS) models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
 First we use AMIS models to quantify the residual variance in difficulty and discrimination across items after accounting for their respective attributes.
@@
 
-The purpose of the first study was to calibrate the items of the MaRs-IB using response data collected from a large number of adult participants from the general population.
+The purpose of the first study was to calibrate the items in the MaRs-IB using response data collected from a large number of adult participants from the general population.
 In particular, we sought to accomplish the following three objectives: (1) to characterize the psychometric properties of the items in the MaRs-IB using item response models; (2) to investigate the associations between item attributes and their psychometric properties, and (3) to determine the equivalence of items with similar complexities, and clones of the same item templates.
@@
 
-The data from multiple participants who completed the experiment were excluded prior to analysis for one or more of the following reasons: rapid guessing (responding in less than 3 s) on four or more trials (N=70); failing to respond on four or more trials (N=8); or minimizing the browser window to a dimension smaller than the minimal requirements (N=7).
+The data from multiple participants who completed the experiment were excluded prior to analysis for one or more of the following reasons: rapid guessing (responding in less than 3 s)\footnote{A threshold of 3 s was decided on based on pilot data.
+See the supplementary materials for details.} on four or more trials (N=70); failing to respond on four or more trials (N=8); or minimizing the browser window to a dimension smaller than the minimal requirements (N=7).
 In total, 83 of 1584 (5.2\%) participants were excluded leaving the data from N=1501 participants available for analysis.
@@
 
-\subsubsection{Goodness of fit}
+\subsubsection{Goodness-of-fit}
 
 The fit of the best-fitting model to the data was evaluated using posterior predictive model checking \citep{gelman1996posterior, levy2017bayesian}.
-A sample of predicted responses was generated for each sample of item parameters and the posterior predictive $p$ (PPP) value was computed based on two discrepancy statistics:  the Orlando and Thissen index ($S-\chi^2$) \citep{toribio2011discrepancy} and the standardized generalized dimensionality discrepancy measure (SGDDM) \citep{levy2015standardized}.
+A sample of predicted responses was generated for each sample of simulated parameters and a posterior predictive $p$ (PPP) value was computed based on two discrepancy statistics: a $\chi^2_{NC}$ discrepancy measure based on the observed score distribution \citep{sinharay2006posterior} and the standardized generalized dimensionality discrepancy measure (SGDDM; \citealt{levy2015standardized}).
 
-The $S-\chi^2$ discrepancy statistic is a measure of item fit based on the comparison of observed and expected proportions of examinees who answer an item correctly at each raw-score group.
-A global measure of fit at the test level is obtained by summing the discrepancy values over the items.
-SGDDM is a measure of model dimensionality.
-More specifically, SGDDM may be interpreted as the mean absolute conditional correlation between all non-redundant pairs of items, which allows for the identification of inter-item correlations beyond those explained by the model.
-This latter measure is important as it tests for the presence of non-negligible response dependence across items, which may otherwise lower the efficiency of future test forms due to the inclusion of nonindependent or  redundant items.
+The $\chi^2_{NC}$ discrepancy statistic is a measure of item fit based on the comparison of observed and expected proportions of participants at each score level (here we ignore scores of zero and one, which comprised 7 total participants).
+A global measure of fit at the test level is obtained by summing the discrepancy values over the groups.
+In turn, SGDDM measures the mean absolute conditional correlation between all pairs of items; that is, the SGDDM is an index of residual inter-item correlations unexplained by the model.
+The SGDDM is important in this context as it tests for local dependence across items, the presence of which might otherwise lower the efficiency of future test forms via the inclusion of nonindependent or redundant items.
+For both discrepancy statistics, the PPP value is the proportion of draws in which the posterior predictive discrepancy is equal to or higher than the realized discrepancy.
+A poor model fit to the data is indicated when the PPP-values are extreme (PPP $\leq$ 0.05 or PPP $\geq$ 0.95).
 
@@
 
-\subsubsection{Aim 2: Quantify the associations between item attributes and functioning}
+\subsubsection{Aim 2:}
 
-Next we inspected the associations between item attributes and parameters.
-As predicted, a one-unit change in feature number was associated with an increase in item difficulty ($\delta_{\beta_1}$ = 0.579, 95\% HDI = 0.389 -- 0.770), or an approximate 10.6\% reduction in accuracy for a participant of average ability.
-Similarly, a one-unit change in rule number was also associated with an increase in item difficulty ($\delta_{\beta_2}$ = 0.514, 95\% HDI = 0.378 -- 0.663), or a 9.5\% reduction in accuracy.
+%\subsubsection{Aim 2: Quantify the associations between item attributes and functioning}
 
-Looking next at the clone-level attributes, there was an unexpected association between distractor type and item difficulty ($\delta_{\beta_3}$ = 1.105, 95\% HDI = 0.940 -- 1.269).
-All else being equal, an item with MD distractors was associated with an expected 19.9\% reduction in accuracy compared to items with PD distractors.
+Next we inspected the associations between item complexity (i.e.
+number of elements and number of rules) and item functioning.
+As predicted, a one-unit change in the number of elements was associated with an increase in item difficulty ($\delta_{\beta_1}$ = 0.579, 95\% HDI = 0.389 -- 0.770), or an approximate 10.6\% reduction in accuracy for a participant of average ability.
+Similarly, a one-unit change in the number of rules was also associated with an increase in item difficulty ($\delta_{\beta_2}$ = 0.514, 95\% HDI = 0.378 -- 0.663), or a 9.5\% reduction in accuracy.
+The difference in coefficients was not credibly different than zero ($\delta_{\beta_1} - \delta_{\beta_2}$ = 0.065, 95\% HDI = -0.189 -- 0.319), indicating that both attributes exert approximately equal impact on item difficulty.
+
+Together the number of elements and number of rules explained approximately 67.6\% of the variance in difficulty across the 64 item templates.
+In other words, the majority of variability in difficulty across item templates can be attributed to their complexity.
+Individually the number of elements and number of rules explain 29.5\% and 40.6\% of the variance in item template difficulty, respectively.
+Across all 384 item clones, these two attributes explain 38.6\% of the variance in difficulty.
+
+Only the number of rules was associated with item discrimination: a one-unit change in rule number was associated with a marginal increase in discrimination ($\delta_{\alpha_2}$ = 0.020, 95\% HDI = 0.003 -- 0.037).
+There was not a credible association between the number of elements and discrimination ($\delta_{\alpha_1}$ = -0.015, 95\% HDI = -0.036 -- 0.008).
+These two attributes explained 31.9\% of the variance in discrimination across item templates.
+\subsubsection{Aim 3:}
+
+%\subsubsection{Objective 3: Decompose the variability in item functioning}
+
+Contrary to the assumption that item clones are exchangeable, we found an association between distractor type and item difficulty ($\delta_{\beta_3}$ = 1.105, 95\% HDI = 0.940 -- 1.269).
+Items with MD distractors were associated with an estimated 19.9\% reduction in accuracy compared to their equivalent items with PD distractors.
+Together, number of elements, number of rules, and distractor type explained slightly over half (52.0\%) of the variance in item difficulty.
+Regardless, a model that included a clone-level random effects term for item difficulty was preferred to models without.
+Therefore, the results indicate that item clones are not equally difficult and are thus not exchangeable.
+
+The estimated item difficulty per clone, organized by distractor type and shape set, is presented in Figure \ref{fig:fig02}.
+The relative increase in difficulty for items with MD-type distractors compared to PD-type distractors is visually apparent (Figure \ref{fig:fig02}a).
+In contrast, the pattern of item difficulties by shape set is more complicated (Figure \ref{fig:fig02}b).
+We computed post-hoc contrasts to compare item difficulty within items across shape sets.
+We found that item clones of shape set 2 were marginally more difficult on average than clones of shape set 1 ($\bar{\beta}_{ss_2} - \bar{\beta}_{ss_1}$ = 0.149, 95\% HDI = 0.054 -- 0.241); and clones of shape set 3 were marginally more difficult on average than clones of shape set 2 ($\bar{\beta}_{ss_3} - \bar{\beta}_{ss_2}$ = 0.161, 95\% HDI = 0.066 -- 0.259).
+However, the magnitude of these differences were still a fraction of the magnitude of the clone-level residual variance ($\epsilon_{\beta k}$ =  0.620, 95\% HDI = 0.538 -- 0.701).
+Thus, the sources of residual variability in item difficulty across clones is complex and likely idiosyncratic based on  perceptual features...
+
+Turning to item discrimination, we did not find a credible association between discrimination and distractor type ($\delta_{\alpha_3}$ = -0.029, 95\% HDI = -0.065 -- 0.005).
+Moreover, a model that included a clone-level random effects term for item discrimination was not preferred to a simpler model without.
+As such, there is not sufficient evidence to refute that the assumption that item clones are similarly discriminating.
+We note this result should be interpreted with caution given the relatively weak predictive improvement of the best-fitting model compared to a model with a clone-level random effects term.
+
+Finally, we inspected the relationship between mean response time and item functioning.
 As expected, there was a positive association between mean response time and item difficulty ($\delta_{\beta_4}$ = 0.541, 95\% HDI = 0.414 -- 0.671).
-
-In contrast, there was only one credible association between an item attribute and item discrimination parameters.
-A one-unit change in rule number was associated with a small increase in item discrimination $\delta_{\alpha_2}$ = 0.020, 95\% HDI = 0.003 -- 0.037).
-There were no reliable associations between item discrimination and feature number ($\delta_{\alpha_1}$ = -0.015, 95\% HDI = -0.036 -- 0.008), distractor type ($\delta_{\alpha_3}$ = -0.029, 95\% HDI = -0.065 -- 0.005), or average response time ($\delta_{\alpha_4}$ = -0.010, 95\% HDI = -0.030 -- 0.009).
-
-\subsubsection{Objective 3: Decompose the variability in item functioning}
-
-Finally, we inspected the relative contributions of modeled and unmodeled sources to variability in item functioning.
-The two modeled item-level attributes, feature number and rule number, explained approximately 65.5\% of the variance in difficulty across the 64 item templates (or 35.0\% of the variance across all 384 item clones).
-In conjunction with distractor type, the three item attributes explained nearly half (48.2\%) of the variance in item difficulty across all 384 item clones.
-Nevertheless, there was non-negligible residual variability in item difficulty at the template and clone levels as evidenced by the comparatively better fit of an item structure model that included random effects at both levels.
-That is, the results indicate that neither item templates nor clones with identical attributes can be assumed to be equal in difficulty.
-
-Feature number, rule number, and distractor type cumulatively explained approximately 39.2\% of the variance in item discrimination across the 64 item templates.
-In contrast to item difficulty, model comparison indicated there was non-negligible residual variability in item discrimination only at template level.
-Indeed, a model that included random effects at the clone level produced a comparatively worse fit to the data.
-As such, the results indicate that item templates with identical attributes cannot be assumed to be equally discriminating; however, there was not sufficient evidence to refute that item clones are equally discriminating.
-We note this result should be interpreted with caution given the relatively weak predictive improvement of the best-fitting model compared to other similar models.
+That is, response times were slower on average for more difficult items.
+There were not, however, a credible associations between item discrimination and average response time ($\delta_{\alpha_4}$ = -0.010, 95\% HDI = -0.030 -- 0.009).
 
@@
 There was also a small association between gender and ability ($\rho_2$ = 0.128, 95\% HDI = 0.078 -- 0.182), such that women performed marginally better than men.
-Average response time was positively associated with ability ($\rho_3$ = 0.427, 95\% HDI = 0.374 -- 0.478), as was $\Delta$ response time ($\rho_4$ = 0.368, 95\% HDI = 0.313 -- 0.424), suggesting a speed-accuracy trade-off in performance.
+Average response time was positively associated with ability ($\rho_3$ = 0.427, 95\% HDI = 0.374 -- 0.478), as was $\Delta$ response time ($\rho_4$ = 0.368, 95\% HDI = 0.313 -- 0.424).
+In other words, participants with higher levels of ability spent on average more time on each item and deliberated for longer on more difficult items.
+Together, these four person attributes explained 50.3\% of the variance in ability.
+These results are consistent with a speed-accuracy trade-off --- higher levels of ability in this sample reflect in part a tendency to slow down and respond carefully overall and increasingly for more challenging items.
 
-\subsubsection{Model Fit}
+\subsubsection{Goodness-of-fit}
 
-Finally, we validated the fit of the best-fitting model using several posterior predictive model checking measures.
-Across the 384 items, the average posterior predictive p-value for the $S-\chi^2$ index of item fit was $p = 0.510$.
-Only 14 of 384 items (3.64\%) indicated misfit with PPP-values less than 0.05 or greater than 0.95.
-Moreover, the PPP-value for the test of residual dependence was $p = 0.803$.
-Though this suggests some evidence for response dependence, it is within a tolerable range.
-Overall then, the best-fitting model provides an adequate fit to the data.
+Finally, we validated the fit of the best-fitting model using two posterior predictive model checking measures.
+The PPP-value for chi-square discrepancy statistic did not exceed the critical threshold ($\chi^2_{NC}$ = 15.904, PPP-value = 0.297), indicating that the model was sufficiently able to reproduce the distribution of observed scores.
+The PPP-value for the SGDDM statistic did exceed the critical threshold (PPP-value = 0.019).
+However, the mean of the realized SGDDM values (SGDDM = 0.0180) were only marginally larger than the mean of the posterior predicted SGDDM values (SGDDM = 0.0175).
+That is, the residual inter-item correlations in the data unexplained by the best-fitting model were small on average and only slightly larger than what we would expect to observe by chance.
+As such, there is little evidence for local dependence in the data.
+Overall then, we can conclude the best-fitting model provides an adequate fit to the data.
 
@@
 
-In this study, we recruited a large sample of adults to complete items from the MaRs-IB, making sure that each item (and its clones) were completed by a sizable number of participants.
+In this first study, we recruited a large sample of adults to complete items from the MaRs-IB, making sure that each item (and its clones) were completed by a sufficient number of participants.
 Using multilevel item structure models, we then estimated the psychometric properties (i.e.
 difficulty and discrimination) of each item.
-Corroborating the findings of a preliminary investigation of the MaRs-IB \citep{chierchia2019matrix}, we found that the items of the MaRs-IB vary greatly in their difficulty -- a prerequisite for measuring nonverbal reasoning across the ability spectrum.
-We also found that the items of the MaRs-IB exhibited medium-to-large levels of discrimination \citep{baker2017basics}, similar to what has been found in investigations of other measures of matrix reasoning \citep{chiesi2012using}.
+Corroborating the results of an initial investigation of the MaRs-IB \citep{chierchia2019matrix}, we found that the items in the MaRs-IB vary greatly in their difficulty -- a prerequisite for measuring nonverbal reasoning across the ability spectrum.
+We also found that the items in the MaRs-IB exhibited medium-to-large levels of discrimination, similar to what has been found in investigations of other of matrix reasoning tasks \citep{chiesi2012using, chiesi2012item, van2013shortened}.
 
-We also investigated whether the MaRs-IB exhibited several desirable properties.
-We confirmed that the difficulty of MaRs-IB items is a function of, and to a moderate extent explained by, item attributes (namely, feature number and rule number).
-Notably, we also identified that item discrimination is associated with rule number, in line with previous investigations of matrix reasoning measures \citep{embretson1999generating}.
-Crucially, we found evidence that item clones are not equally difficult.
-Unexpectedly, and contrary to the results of \cite{chierchia2019matrix}, distractor type emerged as a robust predictor of item difficulty.
-All else being equal, items were much harder when using MD-type distractors compared to PD-type distactors.
-Moroever, there was non-negligible residual variance in item difficulty across clones even after taking into consideration all modeled sources of variance.
-Therefore item clones in the MaRs-IB are not psychometrically equivalent and should not be treated as exchangeable.
+We also investigated how the structural characteristics of items impact their functioning.
+We found that the number of elements and number of rules were both positively associated with item difficulty.
+These two attributes exerted approximately equal influence on difficulty (i.e.
+a one-unit change in either was independently associated with a roughly 10\% reduction in performance).
+This finding is interesting in light of previous investigations of nonverbal reasoning tasks, which have found either the number of elements \citep{bethell1984adaptive} or number of rules \citep{mulholland1980components} to be more determinant of item difficulty.
+One possible reason for this discrepancy is that these two attributes are largely uncorrelated ($\rho$ = 0.176) across items in the MaRs-IB.
+Together the number of elements and number of rules explained 67.6\% of the variance in item difficulty across item templates (and 38.6\% of the variance across item clones), which is in line with previous investigations of matrix reasoning tasks \citep{carpenter1990one, matzen1994error}.
+These findings further validate the design of the MaRs-IB insofar that item complexity strongly determines item difficulty.
 
-In summary, the results of the calibration study confirm the MaRs-IB to be a useful and psychometrically-sound resource for measuring matrix reasoning ability.
-The items span a great range of difficulty with medium-to-large levels of discrimination, thus making the MaRs-IB suitable for measuring matrix reasoning across a wide range of ability levels.
-That item clones vary in their difficulty, and should not therefore be treated as exchangeable, does not undermine this point.
-Furthermore, the item parameter estimates derived here can be used to design new MaRs-IB measures with good reliability and potential for reuse, as we demonstrate in the next study.
+Finally, we investigated whether the item clones in the MaRs-IB are psychometrically equivalent and exchangeable.
+Crucially, we found repeated evidence that item clones are not equally difficult.
+Unexpectedly, and contrary to the results of \cite{chierchia2019matrix}, distractor type emerged as a substantial predictor of item difficulty (i.e.
+items with MD-type distractors are associated with an almost 19.9\% reduction in accuracy compared to the same items with PD-type distractors).
+This discrepancy in findings between studies may reflect our larger sample size (N=1501 versus N=659 participants) or population (adults-only versus a mixture of adults, adolescents, and children), and deserves further investigation.
+Even after controlling for the number of elements, number of rules, and distractor type, we found non-negligible residual variability in item difficulty across clones that were difficult to characterize and likely reflect both the additive and interactive effects of low-level perceptual features.
+Regardless, our results clearly indicate that item clones in the MaRs-IB are not psychometrically equivalent and should not be treated as exchangeable.
+
+In summary, the results of the calibration study confirm the MaRs-IB to be a useful and psychometrically valid resource for measuring matrix reasoning ability.
+The items vary considerably in their difficulty, and all posses medium-to-large levels of discrimination, thereby making the MaRs-IB suitable for measuring matrix reasoning across a wide range of ability levels.
+That item clones are not equivalently difficult, and should not therefore be treated as exchangeable, does not invalidate this conclusion.
+The item parameter estimates derived as part of this study are therefore suitable for use in designing new MaRs-IB measures with both good reliability and potential for reuse, as we demonstrate in the next study.
 
@@
 
-The results of the test assembly are presented in Figure \ref{fig:fig02}.
-As expected, the MIP solver selected for items that were more discriminating on average (Figure \ref{fig:fig02}a).
-Though this was not an explicit constraint, the test characteristic curves (TCCs) of the three short forms are markedly similar (Figure \ref{fig:fig02}b).
+The results of the test assembly are presented in Figure \ref{fig:fig03}.
+As expected, the MIP solver selected for items that were more discriminating on average (Figure \ref{fig:fig03}a).
+Though this was not an explicit constraint, the test characteristic curves (TCCs) of the three short forms are markedly similar (Figure \ref{fig:fig03}b).
 Based on the TCCs, the expected observed score for each measure is 7.7 out of 12 items.
-(Note that, because the lowest ability participants are expected to guess an item correctly 25\% of time, the lower asymptote of the TCCs is at a score of 3.) Finally, the TIFs across the three short forms are remarkably similar (Figure~\ref{fig:fig02}c), as is expected given the assembly constraints.
+(Note that, because the lowest ability participants are expected to guess an item correctly 25\% of time, the lower asymptote of the TCCs is at a score of 3.) Finally, the TIFs across the three short forms are remarkably similar (Figure~\ref{fig:fig03}c), as is expected given the assembly constraints.
 As such, we should expect each short form to have approximately equal precision in measuring matrix reasoning ability across the spectrum of ability.
@@
 After providing consent, participants first completed three short surveys: the 10-item need for cognition survey \citep{chiesi2018applying}, the 8-item PROMIS Cognitive Function-Short Form (8a) \citep{iverson2021normative}, and the 8-item subject numeracy scale \citep{fagerlin2007measuring}.
-These measures were included as part of exploratory analyses that are reported in the supplementary materials.
+These measures were included as part of secondary analyses  to explore the associations between personality and matrix reasoning performance (all pairwise correlations are reported in Table S2 of the supplementary materials).
 Afterwards participants completed one of the three 12-item MaRs-IB short-form measures.
 The administration procedure of the short-forms was identical to that in the first study.
-Finally, participants completed the 9-item abbreviated Raven's progressive matrices (RPM; form A) \citep{bilker2012development}.
+Finally, participants completed the 9-item abbreviated Raven's progressive matrices (RPM form A; \citealt{bilker2012development}).
 To maximize consistency in administration, the presentation format and instructions for the RPM measure was the same as for the MaRs-IB test forms.
@@
 On average, participants completed the measure in 174 s (sd = 54.0 s) and correctly solved 8.0 of 12 items (sd = 2.5).
-The distribution of observed scores was right-shifted (Figure~\ref{fig:fig03}), as was expected given the nominal guessing rate.
+The distribution of observed scores was right-shifted (Figure~\ref{fig:fig04}), as was expected given the nominal guessing rate.
 A one-way ANOVA comparing the average score across the three short-forms was not statistically significant (F(2,297) = 0.253, p = 0.777).
@@
     \centering
-    \begin{tabular*}{\textwidth}{clcr}
+    \begin{tabular*}{1.02\textwidth}{cccccccccccr}
     \toprule
-    Model & Description & psis-loco & $\Delta$ psis-loco (se) \\
+    && \multicolumn{3}{c}{Difficulty} && \multicolumn{3}{c}{Discrimination} && \multicolumn{2}{c}{LOO-CV} \\
+    \cmidrule(lr){3-5}\cmidrule(lr){7-9}\cmidrule(lr){11-12}
+    Model && {\small FE-1/2} & {\small RE-1} & {\small RE-2} && {\small FE-1/2} & {\small RE-1} & {\small RE-2} && psis-loco & $\Delta$ psis-loco (se) \\
     \midrule
-    1 & TK  & 27741.4 & 1155.8 (60.1) \\
-    2 & TK  & 26962.1 & 376.5 (34.5) \\
-    3 & TK & 26585.5 & \multicolumn{1}{c}{-} \\
+    1 && X &   &   &&   &   &   && 27741.4 & 1155.8 (60.1) \\
+    2 && X & X &   &&   &   &   && 26962.1 & 376.5 (34.5) \\
+    3 && X & X & X &&   &   &   && 26585.5 & \multicolumn{1}{c}{-} \\
     \midrule
-    4 & TK & 26419.6 & 5.12 (4.07) \\
-    5 & TK & 26414.5 & \multicolumn{1}{c}{-} \\
-    6 & TK & 26417.8 & 3.29 (1.84) \\
+    4 && X & X & X && X &   &   && 26419.6 & 5.12 (4.07) \\
+    5 && X & X & X && X & X &   && 26414.5 & \multicolumn{1}{c}{-} \\
+    6 && X & X & X && X & X & X && 26417.8 & 3.29 (1.84) \\
     \bottomrule
     \end{tabular*}
-    \caption{\label{tab:2}\normalfont Goodness-of-fit of item response models to MaRs-IB data.
-LOCO values are presented in deviance scale (i.e.
-smaller numbers indicate better fit).
+    \caption{\label{tab:2}\normalfont Goodness-of-fit of item response models to the MaRs-IB response data.
+The columns under difficulty and discrimination indicate the specification of those parameters for each model (i.e.
+the presence of level-1/2 fixed effects, level-1 random effects, and level-2 random effects).
+LOO-CV values are presented in deviance scale (i.e.
+smaller values indicate better fit).
 Abbreviations: PSIS = Pareto-smoothed importance sampling; LOCO = leave-one-cluster-out.}
@@
 \caption{\label{fig:fig00} Example items from the MaRs-IB.
-(A) A simple item containing one feature per cell and two relation changes (i.e.
-a change of shape and its position changes across rows and columns, respectively).
-(B) A harder item containing three features and six relation changes (i.e.
+(A) A simple item containing one element per cell and two rules (i.e.
+a change of shape and position).
+(B) A harder item containing three elements and six rules (i.e.
 three position changes, two color changes, one shape change).
-(C/D) Variants of Items A and B, respectively, matched on complexity (i.e.
-number of features and relation changes).
+(C/D) Alternate versions of Items A and B, respectively, matched on complexity (i.e.
+number of elements and rules).
 In all panels, the first option is the correct response.}
@@
 \includegraphics[width=1.0\textwidth]{figures/fig02.png}
-\caption{\label{fig:fig02}Results from the test assembly of three parallel MaRs-IB short form (SF) measures.
+\caption{\label{fig:fig02}}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=1.0\textwidth]{figures/fig03.png}
+\caption{\label{fig:fig03}Results from the test assembly of three parallel MaRs-IB short form (SF) measures.
 (A) The psychometric properties of the items selected for each form, compared to all remaining items.
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/fig03.png}
-\caption{\label{fig:fig03}The joint distribution of total scores on the MaRs-IB short form and abbreviated Raven's progressive matrices (RPM) measures.
+\includegraphics[width=1.0\textwidth]{figures/fig04.png}
+\caption{\label{fig:fig04}The joint distribution of total scores on the MaRs-IB short form and abbreviated Raven's progressive matrices (RPM) measures.
 The distribution of observed scores for each measure are shown in the margins.}

### 2022-05-12 manuscript/main.tex
updated model comparison

@@
 difficulty, discrimination).
-We modeled these two properties because they have previously been found to be among the strongest predictors of item difficulty in matrix reasoning tasks \citep{carpenter1990one, embretson1998cognitive}.
+We modeled these two properties because they have previously been found to be among the strongest predictors of item difficulty in matrix reasoning tasks \citep{primi2001complexity}.
 This analysis therefore provides another means to evaluate the success of the MaRs-IB design.
@@
 
-\subsubsection{Statistical analysis}
+\subsubsection{Response time analysis}
 
 We investigated the relationship between accuracy and response time using a mixed effects (random intercepts) linear regression model.
-Participants' log-transformed response times were predicted as a function of that trial accuracy, rest score (participants' observed scores on all other items), and item difficulty (one minus the proportion correct for that item).
-The mixed-effects model was estimated using the \textit{statsmodels} python package (v0.12.2) \citep{seabold2010statsmodels}).
+Trial-wise (log-transformed) response times were predicted as a function of trial accuracy, rest score (participants' observed scores on all other items), and item difficulty (one minus the proportion correct for that item).
+The mixed-effects model was estimated using the \textit{statsmodels} python package (v0.12.2;  \citealt{seabold2010statsmodels}).
 
@@
 Specifically, we used additive multilevel item structure (AMIS) models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
-In AMIS models, item parameters are defined according to a hierarchical structure in which clones descend from item templates.
-At the core of of all of the AMIS models used here is the three parameter logistic (3PL) item response model, where the probability of a correct response ($y_{ijk} = 1$) for person $i$ on item clone $k$ belonging to item template $j$ is:
+In AMIS models, item parameters are defined according to a hierarchical structure in which item clones are nested in item templates.
+At the core of all of the AMIS models used here is the three parameter logistic (3PL) item response model, where the probability of a correct response ($y_{ijk} = 1$) for person $i$ on item clone $k$ belonging to item template $j$ is:
 
@@
 \noindent where $\theta_i$ is the latent ability for person $i$, and $\beta_{jk}$, $\alpha_{jk}$, and $\gamma_{jk}$ are the difficulty, discrimination, and guessing parameters for item clone $k$ of item family $j$.
-Because the estimates of guessing parameters are often unreliable in the absence of very large amounts of data \citep{han2012fixing}, we fixed the guessing parameter for every item clone to the nominal guessing rate ($\gamma_{jk} = 0.25$).
+Because estimates of guessing parameters are often unreliable in the absence of very large amounts of response data \citep{han2012fixing}, we fixed the guessing parameter for every item to the nominal guessing rate ($\gamma_{jk} = 0.25$).
 
@@
 
-\noindent such that each item difficulty parameter was modeled as an intercept ($\mu_\beta$), reflecting the average difficulty across all items, and four additional components:
+\noindent To elaborate, item difficulty parameter was modeled as an intercept ($\mu_\beta$), reflecting the average difficulty across all items, and four additional components:
 
@@
 
-\item $\sum_{n=1}^N Q_{jn} \delta_{\beta n}$: the effect of the template (level 1) attributes (desribed below) on item difficulty, where $Q_{jn}$ is the value of attribute $n$ for item template $j$, and $\delta_{\beta n}$ is the effect of attribute $n$.
+\item $\sum_{n=1}^N Q_{jn} \delta_{\beta n}$: the effect of item template (level 1) attributes (described below) on difficulty, where $Q_{jn}$ is the value of attribute $n$ for item template $j$, and $\delta_{\beta n}$ is the effect of attribute $n$.
 
-\item $\epsilon_{\beta j}$: the template (level 1) residual, or residual variability in item difficulty of the item template unexplained by its attributes.
+\item $\epsilon_{\beta j}$: the template (level 1) residual, or the residual variability in difficulty of the item template unexplained by its attributes.
 
-\item $\sum_{m=1}^M R_{km} \delta_{\beta m}$: the effect of the clone (level 2) attributes on item difficulty, where $R_{km}$ is the value of attribute $m$ for item clone $k$, and $\delta_{\beta m}$ is the effect of attribute $m$.
+\item $\sum_{m=1}^M R_{km} \delta_{\beta m}$: the effect of item clone (level 2) attributes on difficulty, where $R_{km}$ is the value of attribute $m$ for item clone $k$, and $\delta_{\beta m}$ is the effect of attribute $m$.
 
-\item $\epsilon_{\beta k}$: the clone (level 2) residual, or residual variability in item difficulty of the item clone unexplained by its attributes.
+\item $\epsilon_{\beta k}$: the clone (level 2) residual, or residual variability in difficulty of the item clone unexplained by its attributes.
 
@@
 
-\noindent In turn, item discrimination parameters were expressed as the sum of an equivalent set of components:
+\noindent So too, item discrimination parameters were expressed as the sum of an equivalent set of components:
 
@@
 
-In our analyses, we considered two template level (level 1) attributes: feature number, or the number of elements in each cell of an item; and rule number, or the number of changing relationships across the cells of an item (e.g.
-size change, color change, shape change, position change).
-The feature and rule number for each item was obtained via manual annotation.
-Across item templates, the modal number of features was two (range: 1 -- 4) and the modal number of rules was three (range: 1 -- 6).
-Feature and rule number were largely uncorrelated across items ($\rho$ = 0.176).
-Additional examples of items are provided in the supplementary materials.
-We predicted that both attributes would be positively associated with item difficulty; that is, all else equal, items with a great number of features or rules would be more difficult to solve.
-In contrast, we had no hypotheses about the relationship between these attributes and item discrimination.
+In our analyses, we considered two item template (level 1) attributes, element number and rule number, which have previously been identified as key determinants of item difficulty in matrix reasoning tasks \citep{primi2001complexity}.
+Element number refers to the number of geometric shapes in each cell of a matrix, whereas rule number refers to the number of relationships that govern the changes of these shapes across cells.
+In the MaRs-IB, there are four principal rules: size change, color change, position change.
+Across item templates, element number ranges from one to four (median = 2) and rule number ranges from one to six (median = 3).
+Importantly, element and rule number are mostly uncorrelated across item templates ($\rho$ = 0.176).
+The feature and rule number for each item was determined via manual annotation.
 
-We also considered two clone level (level 2) attributes: distractor type, or the use of MD-type versus PD-type response alternatives; and mean response time, or the average time participants spent deliberating on that item clone.
+At the clone-level (level 2), we modeled an additional two attributes: distractor type and mean response time.
+Distractor type refers to whether a clone uses MD-type versus PD-type distractors, whereas mean response time refers to the average time participants spent deliberating on that item clone.
 We note that, in contrast to all other modeled attributes, average response time is not an intrinsic property of an item.
-We opted to include it in our models, however, because including response time as a covariate has previously been shown to improve parameter estimation \citep{bertling2018using}.
-To prevent possible collinearity between the attributes, mean response time was orthogonalized with respect to the three other item attributes; therefore, any effects of response time reflects residual structure after accounting for all other item attributes.
-Based on \cite{chierchia2019matrix}, we predicted there would be no association between distractor type and item difficulty, and a positive association between mean response time and item difficulty.
-As with the template-level attributes, we had no hypotheses about the relationship between the clone-level attributes and item discrimination.
+We elected to include it in our models, however, because including response time as a covariate has previously been shown to improve parameter estimation \citep{bertling2018using}.
+To preclude any collinearity between the attributes, mean response time was orthogonalized with respect to the three other item attributes; as such, any effects of response time reflects residual structure after accounting for all other item attributes.
 
-It is important to note that we deliberately chose not to include shape type as a clone attribute; that is, we elected not to directly model the effects of lower-level perceptual features on item functioning.
-As previously mentioned, each item template has three alternate forms that vary in their perceptual stimuli.
-Every item clone in the MaRs-IB is comprised of stimuli drawn one of nine stimulus sets (Figure S1).
-Each stimulus set is itself made up of approximately 15 unique stimuli (some elementary shapes, like squares and circles, are shared across stimulus sets).
-These stimuli are pseudo-randomly distributed across item clones of a given stimulus set; that is, two item clones with the same stimulus set and feature number are not guaranteed to be comprised of the same stimuli.
-As such, the most rigorous way of modeling perceptual features on item functioning would have been the inclusion of approximately 45 additional attributes for both item difficulty and discrimination.
-Rather than deal with this complexity, however, we opted to leave this unmodeled, instead to be soaked up by the residual variance term.
+We deliberately chose not to incorporate shape set (i.e.
+1, 2, or 3) as a clone-level attribute.
+As described above, every item template in the MaRs-IB has three unique versions that differ in the geometric shapes populating its cells.
+There are 45 unique geometric shapes in total, organized into nine stimulus sets, where each item clone draws a subset of shapes from one of the nine sets (see Figure S2 for examples).
+As such, the groups in tripartite definition of shape set do not constitute meaningful categories and we therefore do not incorporate them as fixed effects.
+Furthermore, item clones drawing from the same shape set are not always populated by the same geometric shapes, making complicated the possibility of modeling the nine stimulus sets instead.
+The most rigorous way of modeling the presence of a given geometric shape on item functioning would be to include each as a binary attribute in estimating item difficulty and discrimination.
+In order to keep the models simple, we elected to leave shape set unmodeled and accounted for by the residual variability terms instead.
 
-The AMIS model framework provides a natural means of accomplishing the three goals of the calibration the study.
-By fitting an AMIS model, we estimate the item parameters and quantify the psychometric properties for every item, and their clones, in the MaRs-IB (Aim 1).
-By modeling the effects of item attributes, we can evaluate the success of the MaRs-IB design by investigating whether item attributes influence item functioning as expected (e.g.
-a positive association between rule number and item difficulty; Aim 2).
-Finally, by estimating the magnitude of residual error in item parameters from unmodeled sources of variance (e.g.
-lower-level perceptual features), we can test whether the MaRs-IB possesses certain desirable structural properties (e.g.
-psychometric equivalence and exchangeability among item clones; Aim 3).
+We fit a series of nested AMIS models in order to identify the model that best predicted participants' response data.
+The models varied in how item parameters were specified, with each successive model allowing for greater flexibility in item parameter estimation.
+We fit the models in two waves.
+In the first, item discrimination was fixed ($\alpha = 1$)\footnote{These models still included a fixed guessing parameter ($\gamma = 0.25$).
+We also estimated models with no guessing parameter (i.e.
+one-parameter logistic or Rasch models), but model comparison indicated these models provided worse fits to the data (Table S1).} in order to identify the best structure for the item difficulty independent of item discrimination.
+In the first wave, we fit three models:
 
-We also estimated person-level abilities following an explanatory item response modeling approach \citep{wilson2008explanatory}:
+\begin{itemize}
+
+\item Model 1: Item difficulty is a function of item template attributes (level-1) and item clone attributes (level-2).
+In this model, item difficulty is predicted solely by the fixed effect contribution of item attributes (i.e.
+element number, rule number, distractor type, response time).
+
+\item Model 2: Item difficulty is a function of all item attributes (level-1/2) and level-1 residuals.
+The inclusion of the level-1 residual serves to quantify the magnitude of residual variability in item difficulty across item templates unexplained by the level-1 attributes (i.e.
+element number, rule number).
+
+\item Model 3: Item difficulty is a function of all item attributes (level-1/2) and level-1/2 residuals.
+The inclusion of the level-2 residual serves to quantify the magnitude of residual variability in item difficulty across item clones unexplained by the level-2 attributes (i.e.
+distractor type, mean response time).
+
+\end{itemize}
+
+In the second wave of model fitting, item discrimination was now a free parameter to be estimated.
+Moreover, in the second set of models, item difficulty parameters were specified according to the best-fitting model from the first wave.
+In the second wave, we fit an additional three models:
+
+\begin{itemize}
+
+    \item Model 4: Item discrimination is a function of item template attributes (level-1) and item clone attributes (level-2).
+In this model, item discrimination is predicted solely by the fixed effect contribution of item attributes (i.e.
+element number, rule number, distractor type, response time).
+
+    \item Model 5: Item discrimination is a function of all item attributes (level-1/2) and level-1 residuals.
+The inclusion of the level-1 residual serves to quantify the magnitude of residual variability in item discrimination across item templates unexplained by the level-1 attributes (i.e.
+element number, rule number).
+
+    \item Model 6: Item discrimination is a function of all item attributes (level-1/2) and level-1/2 residuals.
+The inclusion of the level-2 residual serves to quantify the magnitude of residual variability in item discrimination across item clones unexplained by the level-2 attributes (i.e.
+distractor type, mean response time).
+
+\end{itemize}
+
+The AMIS modeling framework provides a natural means of accomplishing the three goals of the calibration study.
+Through any of the models above, we obtain estimates of the the psychometric properties (i.e.
+difficulty, discrimination) of every item template and clone in the MaRs-IB (Aim 1).
+Through the level-1 fixed effects, we are able to test for associations between item complexity (i.e.
+element and rule number) and item functioning (Aim 2).
+We predicted that both attributes would be positively associated with item difficulty; that is, all else equal, items with either a great number of elements and rules would be more difficult.
+In contrast, we had no hypotheses about the impact of these attributes on item discrimination.
+
+Finally, through both the level-2 fixed and random effects, we can evaluate whether item clones in the MaRs-IB are exchangeable (Aim 3).
+For example, an association between distractor type and item functioning would entail that clones are not psychometrically equivalent.
+Moreover, if the level-2 residual variance in item difficulty or discrimination were non-negligible, that would similarly entail that item clones are not psychometrically equivalent due to one or more unmodeled features (e.g.
+low-level perceptual properties).
+
+\subsubsection{Person-level effects}
+
+In each of the six models described above, we incorporated person-level attributes as predictors of latent ability ($\theta_i$) following an explanatory item response modeling approach \citep{wilson2008explanatory}:
 
@@
 
-\noindent where $X_{ip}$ is the value of attribute $p$ for person $i$, $\rho_p$ is the partial correlation of attribute $n$, and $\epsilon_i$ is variance in ability unexplained by the person-level attributes.
+\noindent where $X_{ip}$ is the value of attribute $p$ for person $i$, $\rho_p$ is the partial correlation of person attribute $p$, and $\epsilon_i$ is variance in ability unexplained by the person-level attributes.
 
-We investigated person-level abilities as a function of four covariates: age, gender, average response time, and $\Delta$ response time.
+We estimated person-level abilities as a function of four covariates: age, gender, mean response time, and $\Delta$ response time.
 This last term reflects the degree of change in a participant's deliberation time as a function of item difficulty (defined as one minus the proportion of correct responses for that item).
+Mean response time and $\Delta$ response time were calculated for each participant simultaneously through linear regression, where a given participant's (log-transformed) response times were regressed against an intercept and item difficulty.
 All covariates were standard-scored prior to analysis with the exception of gender, which was binary coded (male = -0.5, female = 0.5).
@@
 As such, 10,000 post-warmup samples from the joint posterior were retained.
-The $\hat{R}$ values for all parameters were equal to or less than 1.01, indicating acceptable convergence between chains, and there were no divergent transitions in any chain.
+The $\hat{R}$ values for all parameters were less than 1.01, indicating acceptable convergence between chains, and there were no divergent transitions in any chain.
 
@@
 To ensure the identifiability of all models, person abilities were constrained to have a mean of zero and a variance of 1.0.
-In other words, the residual variance in ability was specified as $V(\epsilon) = 1 - \sum \rho_p^2$.
+Specifically, the residual variance of latent ability was specified as $V(\epsilon) = 1 - \sum \rho_p^2$.
 
@@
 The fit of the best-fitting model to the data was evaluated using posterior predictive model checking \citep{gelman1996posterior, levy2017bayesian}.
-A sample of predicted responses was generated for each sample of simulated parameters and a posterior predictive $p$ (PPP) value was computed based on two discrepancy statistics: a $\chi^2_{NC}$ discrepancy measure based on the observed score distribution \citep{sinharay2006posterior} and the standardized generalized dimensionality discrepancy measure (SGDDM; \citealt{levy2015standardized}).
+A sample of predicted responses was generated for each sample of simulated parameters and a posterior predictive $p$ (PPP) value was computed based on two discrepancy statistics: a chi-square $\left( \chi^2_{NC} \right)$ discrepancy measure based on the observed score distribution \citep{sinharay2006posterior} and the standardized generalized dimensionality discrepancy measure (SGDDM; \citealt{levy2015standardized}).
 
@@
 
-\subsubsection{Response time statistics}
+\subsubsection{Response time analysis}
 
-The results are summarized in Table \ref{table:1}.
-On average, participants spent 15.9 s (sd = 7.2 s) on each item.
+The results of the linear mixed-effects model of response times are summarized in Table \ref{table:1}.
+On average, participants spent 15.9 s (sd = 7.2 s) per item.
 There were significant main effects of accuracy ($\beta = 0.067$, $t = 9.84$, $p < 0.001$), rest score ($\beta = 0.125$, $t = 16.95$, $p < 0.001$), and item difficulty ($\beta = 0.137, t = 44.33, p < 0.001$).
-That is, response times were slower for correct responses, better-performing participants, and more difficult items.
+In other words, response times were slower on average for correct responses, better- performing participants, and more difficult items.
+The positive association between response time and accuracy is consistent with speed-accuracy trade-offs; that is, slower responses are associated with more careful work.
+This interpretation is corroborated by the main effect of rest score, such that more accurate participants were slower overall.
 
-There were also significant accuracy by difficulty and  accuracy by rest score interactions.
-Correct responses were even slower for more difficult items, but faster for better performing participants.
-There was also a significant difficulty by score interaction, such that better performing participants exhibited slower responses for more difficult items.
+There were also significant accuracy by difficulty ($\beta = 0.097$, $t = 14.34$, $p < 0.001$) and accuracy by rest score interactions ($\beta = -0.064$, $t = -8.76$, $p < 0.001$).
+Correct responses were even slower for more difficult items, but faster for better-performing participants.
+There was also a significant difficulty by score interaction ($\beta = 0.032$, $t = 6.55$, $p < 0.001$), such that better-performing participants exhibited even slower responses for more difficult items.
 This pattern of response time results is visible in Figure \ref{fig:fig01}c.
@@
 
-\subsubsection{Alternative item structure models}
+In sum, we found evidence of speed-accuracy trade-offs in performance in this sample.
+Accurate responses and more accurate participants were both slower overall.
+Notably, better-performing participants demonstrated adjustments to their response times as function of item difficulty; whereas the lowest scoring participants maintained the same work rate regardless of item difficulty, the highest scoring participant exhibited the largest slowing in responses as items became more challenging.
+We return to this pattern of results in the General Discussion.
 
-We fit a series of nested AMIS models in order to identify the model that best predicted participants' response data.
-The models varied in how item parameters were specified, with each successive model allowing for greater flexibility in item parameter estimation.
-For the goals of current study, it is important to identify the model that most accurately reflects variability in item functioning, as insufficiently flexible models can result in the biased estimation of item parameters \citep{lathrop2017item}.
+\subsubsection{Model comparison}
 
-We fit AMIS models in two waves.
-In the first wave, the item discrimination parameters were held fixed ($\alpha = 1$).\footnote{These models still included a fixed guessing parameter ($\gamma = 0.25$).
-We also estimated models with no guessing parameter (i.e.
-one-parameter logistic or Rasch models), but model comparison indicated these models provided worse fits to the data (Table S1).} We applied this restriction to the first set of models in order to identify the best structure for the item difficulty parameters independent of the discrimination parameters.
-In the first wave, we fit three models:
+The results of the model comparison is summarized in Table \ref{table:1}.
+Of the models fitted in the first wave, which varied only in their specification of item difficulty, Model 3 exhibited the best fit to the data.
+Indeed, Model 3 demonstrated a considerable improvement in fit over Model 1 ($\Delta$ LOCO = 1155.8, se = 60.1) and Model 2 ($\Delta$ LOCO = 376.5, se = 34.5), indicating the presence of non-negligible residual variance in item difficulty across both item templates and clones.
+As the best-fitting model in the first wave, Model 3 served as the starting point in the second wave of model fitting.
 
-\begin{itemize}
-
-\item Model 1: Item difficulty is a function of item template attributes (i.e.
-feature number, rule number) and item clone attributes (i.e.
-distractor type, response time).
-In this model, the difficulty of an item is explained solely
-
-item with identical attributes are psycho- metrically equivalent, as are all the clones belonging to an item.
-
-\item Model 2: Item difficulty is a function of item template attributes and residuals.
-In this model, item templates with identical attributes may not be equally difficult; however, all clones belonging to an item are still psychometrically equivalent.
-
-\item Model 3: Item difficulty is function of attributes and residuals at both levels.
-In this model, neither item templates nor clones with identical attributes are necessarily equally difficult, depending on the magnitude of the residual terms.
-
-\end{itemize}
-
-The goodness-of-fit of the four models is summarized in Table \ref{table:1}.
-Model 4 demonstrated a better fit of the data than Model 1 ($\Delta$ LOCO = 1223.69, se = 44.94), Model 2 ($\Delta$ LOCO 511.58, se = 28.77), and Model 3 ($\Delta$ LOCO = 187.96, se = 17.25).
-Thus, Model 4 was selected as the starting point for the second wave of model fitting.
-
-Next we estimated a second set of models in which the item discrimination parameter was a free parameter.
-For all models, the item difficulty parameters were specified according to Model 4.
-In the second wave, we fit three models:
-
-\begin{itemize}
-
-    \item Model 5: Item discrimination is a function of both template and clone attributes.
-In this model, item templates with identical attributes will be equally discriminating; however, item clones may vary in their discrimination as a function of their attributes.
-
-    \item Model 6: Item discrimination is a function of all item attributes and template-level residuals.
-In this model, item templates with identical attributes may not necessarily be equally discriminating.
-
-    \item Model 7: Item discrimination is a function of attributes and residuals at both levels.
-In this model, neither item templates nor clones with identical attributes are necessarily equally discriminating, depending on the magnitude of the residual terms.
-
-\end{itemize}
-
-The goodness-of-fit of the three models is summarized in Table \ref{table:1}.
-Notably, we did not find the most complex model (Model 7) to be the best-fitting model.
-Although Model 7 was preferred to Model 5 ($\Delta$ LOCO = 2.13, se = 2.04), it was not preferred to Model 6 ($\Delta$ LOCO = -1.44, se = 0.85).
-However, these differences in LOCO values from each model to the next were each within two standard errors of the mean, indicating only weak predictive improvement of Model 6 over the others \citep{vehtari2022cv}.
-As such we will select Model 6 as the best-fitting model overall, but proceed in discussing it with caution.
+In the second wave of model fitting, in which models varied only in their specification of item discrimination, Model 5 demonstrated the best fit to the data.
+By comparison, Model 5 demonstrated smaller improvements in fit over Model 4 ($\Delta$ LOCO = 5.12, se = 4.07) and Model 6 ($\Delta$ LOCO = 3.29, se = 1.84).
+It must be noted these differences in LOCO values from each model to the next were each within two standard errors of the mean, indicating only weak predictive improvement of Model 6 over the others \citep{vehtari2022cv}.
+As such we will select Model 5 as the best-fitting model overall, but proceed in discussing it with caution.
 We also note that Model 6 fit clearly better than Model 4 ($\Delta$ LOCO = 85.91, se = 6.33, supporting the inclusion of varying item discrimination parameters).
+We also note that all second-wave models exhibited better fits to the data than the first-wave models, supporting the inclusion of item discrimination parameters.
 
@@
 
-With 64 item templates, each with six clones, there is a vast number of possible MaRs-IB test forms we could construct.
-Returning to our motivating problem --- the availability of abbreviated matrix reasoning measures with the potential for reuse --- we decided to construct three parallel short form measures.
+Given a starting point of 64 item templates with six clones each, there is a vast number of possible MaRs-IB test forms we could construct.
+Returning to our motivating problem --- the availability of matrix reasoning measures with the potential for reuse --- we decided to construct three abbreviated parallel short form measures.
 To do so, we used mixed integer programming \citep{der2005wj} to maximize the test information function (TIF) of each short form subject to the following constraints:
@@
     \item Each short form was required to contain 12 items.
-This was chosen to minimize the administration time of a given form (2-4 minutes on average) and to achieve sufficient test reliability (Cronbach's $\alpha \geq 0.8$).
+This was chosen to minimize the administration time of a given form (2-4 minutes on average) while achieving sufficient test score reliability (Cronbach's $\alpha \geq 0.8$).
 
@@
     \item Clones selected for inclusion could either use MD-type or PD-type distractors, but not both.
-This was chosen to prevent item redundancy (i.e.
+This constraint was adopted in order to prevent item redundancy (i.e.
 including the same puzzle twice in one short form).
@@
 
-\noindent The TIF was maximized at five ability levels ($\theta = -1.0, -0.5, 0.0, 0.5, 1.0$).
-Previous simulation studies have shown that target values at three to five well-chosen $\theta$s are generally sufficient \citep{der2005wj}.
-Solutions to the mixed integer programming problem were found using the \textit{mip} python package (v1.13.0) \citep{santos2020mixed}.
+\noindent The TIF was maximized at five ability levels ($\theta = -1.0, -0.5, 0.0, 0.5, 1.0$), which previous simulation studies have shown to be generally sufficient \citep{der2005wj}.
+Solutions to the mixed integer programming problem were found using the \textit{mip} python package (v1.13.0; \citealt{santos2020mixed}).
 
@@
 Based on the TCCs, the expected observed score for each measure is 7.7 out of 12 items.
-(Note that, because the lowest ability participants are expected to guess an item correctly 25\% of time, the lower asymptote of the TCCs is at a score of 3.) Finally, the TIFs across the three short forms are remarkably similar (Figure~\ref{fig:fig03}c), as is expected given the assembly constraints.
+(Note that, because the lowest ability participants are expected to guess an item correctly 25\% of time, the lower asymptote of the TCCs is at a score of 3.) Finally, the TIFs across the three short forms are remarkably similar (Figure~\ref{fig:fig03}c), as is expected given the last assembly constraint.
 As such, we should expect each short form to have approximately equal precision in measuring matrix reasoning ability across the spectrum of ability.
@@
 Our aim was to verify that observed scores on each short form conformed to our expectations based on the TCCs and TIFs.
-Doing so would provide additional evidence of whether we were successful in calibrating the functioning of the items in the MaRs-IB.
-We also sought to establish the convergent validity by correlating observed scores on the short form measures with performance on an abbreviated version of the Raven's progressive matrices \citep{bilker2012development}.
+Doing so would provide evidence that we were successful in calibrating the functioning of the items in the MaRs-IB in the previous study.
+We also sought to demonstrate convergent validity by correlating performance on the MaRs-IB short forms with performance in the same participants on an abbreviated version of the Raven's progressive matrices \citep{bilker2012development}.
 
@@
 Finally, participants completed the 9-item abbreviated Raven's progressive matrices (RPM form A; \citealt{bilker2012development}).
-To maximize consistency in administration, the presentation format and instructions for the RPM measure was the same as for the MaRs-IB test forms.
+To maximize consistency in administration, the presentation format and instructions for the RPM task was the same as for the MaRs-IB test forms.
 
@@
 Of additional note are the associations we found between accuracy and response time.
-The best-performing participants spent more time on each item, consistent with a speed-accuracy trade-off account of performance \citep{heitz2014speed}.
-Interestingly, we also found an interaction between participant ability and item difficulty; that is, the best-performing participants increased their deliberation time as items become more difficult, whereas the worst-performing participants maintained approximately the same work rate.
-These results are consistent with a persistence account of participant ability, wherein good performance reflects in part a willingness to invest time in the solution process \citep{ranger2021effects}.
+Accurate responses and more accurate participants were slower on average, consistent with a speed-accuracy trade-off account of performance \citep{heitz2014speed}.
+Interestingly, we also found an interaction between participant ability and item difficulty: whereas the worst-performing participants maintained approximately the same work rate irrespective of item difficulty, the best-performing participants were even slower as items became more challenging.
+These results are consistent with a persistence interpretation of ability, wherein good task performance in part reflects a willingness to invest time in the solution process and poor task performance reflects a tendency to give up sooner \citep{ranger2021effects}.
 As such, the MaRs-IB may be suitable not only for measuring matrix reasoning ability but also mental effort costs \citep{kool2018mental}, opportunity costs \citep{otto2019opportunity}, or other motivational factors \citep{duckworth2011role} related to the tendency to exert effort or give up.
-The use of cognitive models may help to disentangle the relative contributions of these factors to performance on the MaRs-IB \citep{ranger2014accumulator}.
+The use of more sophisticated models may help to disentangle the relative contributions of latent ability and persistence to performance on the MaRs-IB \citep{ranger2014accumulator}.
 
@@
 \toprule
-Predictor &  Coef. (SE) & Z-score (p-value) & 95\% CI  \\
+Predictor &  Coef. (se) & Z-score (p-value) & 95\% CI  \\
 \midrule
@@
     \midrule
-    1 && X &   &   &&   &   &   && 27741.4 & 1155.8 (60.1) \\
-    2 && X & X &   &&   &   &   && 26962.1 & 376.5 (34.5) \\
-    3 && X & X & X &&   &   &   && 26585.5 & \multicolumn{1}{c}{-} \\
-    \midrule
+    1 && X &   &   &&   &   &   && 27741.4 & 1326.87 (62.88) \\
+    2 && X & X &   &&   &   &   && 26962.1 & 547.56 (37.79) \\
+    3 && X & X & X &&   &   &   && 26585.5 & 171.05 (12.65) \\
     4 && X & X & X && X &   &   && 26419.6 & 5.12 (4.07) \\

### 2022-05-12 manuscript/main.tex
updated clone residual variance args + updated supp tables

@@
 \item $\sum_{n=1}^N Q_{jn} \delta_{\beta n}$: the effect of item template (level 1) attributes (described below) on difficulty, where $Q_{jn}$ is the value of attribute $n$ for item template $j$, and $\delta_{\beta n}$ is the effect of attribute $n$.
+This component is the fixed effects contribution to item template difficulty.
 
 \item $\epsilon_{\beta j}$: the template (level 1) residual, or the residual variability in difficulty of the item template unexplained by its attributes.
+This component is the random effects contribution to item template difficulty.
 
 \item $\sum_{m=1}^M R_{km} \delta_{\beta m}$: the effect of item clone (level 2) attributes on difficulty, where $R_{km}$ is the value of attribute $m$ for item clone $k$, and $\delta_{\beta m}$ is the effect of attribute $m$.
+This component is the fixed effects contribution to item clone difficulty.
 
 \item $\epsilon_{\beta k}$: the clone (level 2) residual, or residual variability in difficulty of the item clone unexplained by its attributes.
+This component is the random effects contribution to item clone difficulty.
 
@@
 By comparison, Model 5 demonstrated smaller improvements in fit over Model 4 ($\Delta$ LOCO = 5.12, se = 4.07) and Model 6 ($\Delta$ LOCO = 3.29, se = 1.84).
+This result indicates the presence of non-negligible residual variance in item discrimination across item templates but not clones.
 It must be noted these differences in LOCO values from each model to the next were each within two standard errors of the mean, indicating only weak predictive improvement of Model 6 over the others \citep{vehtari2022cv}.
@@
 
-\subsubsection{Aim 1: Estimate item parameters for all items in the MaRs-IB}
+\subsubsection{Aim 1: Quantify the psychometric properties of items in the MaRs-IB}
 
 Across all items, the average item difficulty was $\mu_\beta = 0.177$ (95\% HDI = 0.118 -- 0.238).
-There was considerable variability in item difficulty across items (sd = 1.431, 95\% HDI = 1.370 -- 1.492), wih the smallest ($\beta$ = -3.704, 95\% HDI = -4.687 -- -2.775) and largest ($\beta$ = 3.497, 95\% HDI = 2.481 -- 4.635) item difficulty parameters spanning a wide range.
-For an average ability participant ($\theta = 0$), these parameter estimates translate to an expected average accuracy of 58.4\% across all items, and ceiling (floor) performance for the least (most) difficult item.
+There was considerable variability in item difficulty across items (sd = 1.431, 95\% HDI = 1.370 -- 1.492), with the smallest  and largest item difficulty parameters spanning a wide range ($\beta_{\min}$ = -3.704, 95\% HDI = -4.687 -- -2.775; $\beta_{\max}$ = 3.497, 95\% HDI = 2.481 -- 4.635).
+For an average ability participant ($\theta = 0$), these parameter estimates translate to an expected average accuracy of 58.4\% across all items, and ceiling (chance-level) performance for the least (most) difficult item.
 
@@
 
-\subsubsection{Aim 2:}
-
-%\subsubsection{Aim 2: Quantify the associations between item attributes and functioning}
+\subsubsection{Aim 2: Measuring the association between item complexity and functioning}
 
@@
 
-%\subsubsection{Objective 3: Decompose the variability in item functioning}
+\subsubsection{Objective 3: Investigating the exchangeability of item clones}
 
@@
 The relative increase in difficulty for items with MD-type distractors compared to PD-type distractors is visually apparent (Figure \ref{fig:fig02}a).
-In contrast, the pattern of item difficulties by shape set is more complicated (Figure \ref{fig:fig02}b).
-We computed post-hoc contrasts to compare item difficulty within items across shape sets.
-We found that item clones of shape set 2 were marginally more difficult on average than clones of shape set 1 ($\bar{\beta}_{ss_2} - \bar{\beta}_{ss_1}$ = 0.149, 95\% HDI = 0.054 -- 0.241); and clones of shape set 3 were marginally more difficult on average than clones of shape set 2 ($\bar{\beta}_{ss_3} - \bar{\beta}_{ss_2}$ = 0.161, 95\% HDI = 0.066 -- 0.259).
-However, the magnitude of these differences were still a fraction of the magnitude of the clone-level residual variance ($\epsilon_{\beta k}$ =  0.620, 95\% HDI = 0.538 -- 0.701).
-Thus, the sources of residual variability in item difficulty across clones is complex and likely idiosyncratic based on  perceptual features...
+In contrast, the pattern of item difficulties across clones by shape set is more complicated (Figure \ref{fig:fig02}b).
+After controlling for distractor type and mean response time, the magnitude of the clone-level residual variance is $\epsilon_{\beta k}$ = 0.620 (95\% HDI = 0.538 -- 0.701).
+Across the 128 unique sets of three clones, organized by item and distractor type, that level of residual variance translates into an average difference in difficulty of 0.752 (95\% HDI = 0.630 -- 0.877) between the easiest and hardest item clones of each triad, an effect that is two-thirds as large as the effect of distractor type.
+Furthermore, the shape set membership of the most difficult clone varies considerably across the 128 sets of clones (shape set 1 = 39; shape set 2 = 32; shape set 3 = 57).
+Thus, the sources of residual variability in item difficulty across clones is complex and likely reflect the idiosyncratic contributions of their low-level perceptual features.
 
@@
 
-We calculated the ordinal reliability of the MaRs-IB and RPM short-form measures using the lavaan \citep{lavaan} and semTools \citep{semtools} packages available in R.
+We calculated the ordinal reliability of the MaRs-IB and RPM short-form measures using the \textit{lavaan} \citep{lavaan} and \textit{semTools} \citep{semtools} packages available in R.
 
@@
     \end{tabular*}
-    \caption{\label{tab:2}\normalfont Goodness-of-fit of item response models to the MaRs-IB response data.
+    \caption{\label{tab:2}\normalfont Comparison of item response models fit to the MaRs-IB response data.
 The columns under difficulty and discrimination indicate the specification of those parameters for each model (i.e.

### 2022-05-12 manuscript/supplement.tex
manuscript edits
updated clone residual variance args + updated supp tables

@@
     \centering
-    \begin{tabular*}{\textwidth}{clcr}
+    \begin{tabular*}{0.45\textwidth}{ccr}
     \toprule
-    Model & Description & psis-loco & $\Delta$ psis-loco (se) \\
+    Model & psis-loco & $\Delta$ psis-loco (se) \\
     \midrule
-    7 & TK  & 27741.4 & 635.3 (29.7) \\
-    8 & TK  & 26962.1 & 665.6 (32.2) \\
-    9 & TK & 26585.5 & 633.1 (33.9) \\
+    1r & 28376.7 & 635.3 (29.7) \\
+    1 & 27741.4 & \multicolumn{1}{c}{-} \\
+    \midrule
+    2r & 27627.6 & 665.6 (32.2) \\
+    2 & 26962.1 & \multicolumn{1}{c}{-} \\
+    \midrule
+    3r & 27218.7 & 633.1 (33.9) \\
+    3 & 26585.5 & \multicolumn{1}{c}{-} \\
     \bottomrule
     \end{tabular*}
-    \caption{\label{tab:2}\normalfont Goodness-of-fit of item response models to MaRs-IB data.
-LOCO values are presented in deviance scale (i.e.
-smaller numbers indicate better fit).
+    \caption{\label{tab:2}\normalfont Comparison of the first wave of item response models to their equivalents without a guessing parameter ($\gamma = 0$).
+LOO-CV values are presented in deviance scale (i.e.
+smaller values indicate better fit).
 Abbreviations: PSIS = Pareto-smoothed importance sampling; LOCO = leave-one-cluster-out.}

### 2022-05-13 manuscript/main.tex
manuscript edits

@@
 
-A second contribution provided here is an investigation of how item characteristics shape the psychometric functioning of items in the MaRs-IB.
-Using explanatory item response modeling \citep{de2004explanatory, wilson2008explanatory}, we investigate how the number of elements and number of rules impact item functioning (i.e.
+A second contribution provided here is an analysis of how item properties shape the psychometric functioning of items in the MaRs-IB.
+Using explanatory item response modeling \citep{de2004explanatory, wilson2008explanatory} we investigate how item complexity, i.e.
+the number of elements and number of rules, explain item functioning (i.e.
 difficulty, discrimination).
-We modeled these two properties because they have previously been found to be among the strongest predictors of item difficulty in matrix reasoning tasks \citep{primi2001complexity}.
-This analysis therefore provides another means to evaluate the success of the MaRs-IB design.
-Insofar that the number of elements and rules govern the total amount of information that must be stored in working memory in order to correctly solve an item \citep{embretson1998cognitive}, confirmation that they impact performance would also provide insight into the cognitive reasons for individual differences in performance on the MaRs-IB.
-Finally, explanatory IRT models yield more accurate estimates of item parameters \citep{neuhaus2006separating}, which in turn helps to ensure that new MaRs-IB tests designed using these item parameter estimates function as expected.
+We modeled these two item attributes because they have been previously identified as among the strongest predictors of item difficulty in matrix reasoning tasks \citep{embretson1998cognitive, primi2001complexity}.
+Thus this explanatory analysis provides another means by which to validate the success of the MaRs-IB design; that is, to determine whether the MaRs-IB exhibits similar relationships between item characteristics and functioning as observed in other established matrix reasoning measures.
+As a secondary benefit, explanatory IRT models yield more accurate estimates of item parameters \citep{neuhaus2006separating}, which in turn helps to ensure that new MaRs-IB tests designed using these item parameter estimates function as expected.
 
-%A third contribution is the decomposition of variance in item functioning, into modeled and unmodeled sources, using additive multilevel item structure (AMIS) models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
-First we use AMIS models to quantify the residual variance in difficulty and discrimination across items after accounting for their respective attributes.
-At the item-level, smaller residual variances indicate that items are functioning as intended.
-We also use AMIS models to quantify the variance in difficulty and discrimination across item clones.
-If item clones are to be treated as exchangeable (i.e.
-able to be used interchangeably), then the residual variance in their item parameters must be negligible.
-AMIS models therefore provide a second means by which to evaluate the success of the MaRs-IB design.
+A contribution is an investigation of the exchangeability of item clones in the MaRs-IB, or whether clones are (approximately) psychometrically equivalent.
+Establishing whether the MaRs-IB possesses this property is critical for the design of new matrix reasoning tasks using these items.
+If item clones are exchangeable, then new parallel test forms of equivalent difficulty and discrimination can be generated simply by swapping out one set of clones for another (e.g.
+exchanging clones from shape set 1 for shape set 2).
+To test for exchangeability in the MaRs-IB, we use additive multilevel item structure (AMIS) models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
+These models enable us to quantify the variability in item functioning across clones by distractor type and shape set, which should be negligible if clones are to be treated as exchangeable.
 
@@
 The purpose of the first study was to calibrate the items in the MaRs-IB using response data collected from a large number of adult participants from the general population.
-In particular, we sought to accomplish the following three objectives: (1) to characterize the psychometric properties of the items in the MaRs-IB using item response models; (2) to investigate the associations between item attributes and their psychometric properties, and (3) to determine the equivalence of items with similar complexities, and clones of the same item templates.
+In particular, we sought to accomplish the following three aims: (1) to quantify the psychometric properties (i.e.
+difficulty, discrimination) of the items in the MaRs-IB using item response models; (2) to measure the associations between item complexity and functioning; and (3) to determiner whether item clones in the MaRs-IB are exchangeable (i.e.
+psychometrically equivalent).
 
@@
 Eight of the nine resulting cells contain abstract shapes, while one cell on the bottom right-hand side of the matrix is empty.
-Participants are instructed to ``complete the matrix'' by identifying the missing shape from among four possible alternatives.
+Participants were instructed to ``complete the matrix'' by identifying the missing shape from among four possible alternatives.
 
@@
 element and rule number) and item functioning (Aim 2).
-We predicted that both attributes would be positively associated with item difficulty; that is, all else equal, items with either a great number of elements and rules would be more difficult.
+We predicted that both attributes would be positively associated with item difficulty; that is, all else equal, items with either a greater number of elements and rules would be more difficult.
 In contrast, we had no hypotheses about the impact of these attributes on item discrimination.
@@
 number of elements and number of rules) and item functioning.
-As predicted, a one-unit change in the number of elements was associated with an increase in item difficulty ($\delta_{\beta_1}$ = 0.579, 95\% HDI = 0.389 -- 0.770), or an approximate 10.6\% reduction in accuracy for a participant of average ability.
-Similarly, a one-unit change in the number of rules was also associated with an increase in item difficulty ($\delta_{\beta_2}$ = 0.514, 95\% HDI = 0.378 -- 0.663), or a 9.5\% reduction in accuracy.
+As predicted, a one-unit change in element number was associated with an increase in item difficulty ($\delta_{\beta_1}$ = 0.579, 95\% HDI = 0.389 -- 0.770), or an approximate 10.6\% reduction in accuracy for a participant of average ability.
+Similarly, a one-unit change in rule number was also associated with an increase in item difficulty ($\delta_{\beta_2}$ = 0.514, 95\% HDI = 0.378 -- 0.663), or a 9.5\% reduction in accuracy.
 The difference in coefficients was not credibly different than zero ($\delta_{\beta_1} - \delta_{\beta_2}$ = 0.065, 95\% HDI = -0.189 -- 0.319), indicating that both attributes exert approximately equal impact on item difficulty.
 
-Together the number of elements and number of rules explained approximately 67.6\% of the variance in difficulty across the 64 item templates.
+Together element number and rule number explained approximately 67.6\% of the variance in difficulty across the 64 item templates.
 In other words, the majority of variability in difficulty across item templates can be attributed to their complexity.
@@
 These two attributes explained 31.9\% of the variance in discrimination across item templates.
-\subsubsection{Aim 3:}
 
-\subsubsection{Objective 3: Investigating the exchangeability of item clones}
+\subsubsection{Aim 3: Investigating the exchangeability of item clones}
 
@@
 Items with MD distractors were associated with an estimated 19.9\% reduction in accuracy compared to their equivalent items with PD distractors.
-Together, number of elements, number of rules, and distractor type explained slightly over half (52.0\%) of the variance in item difficulty.
+Together, element rule, rule number, and distractor type explained slightly over half (52.0\%) of the variance in item difficulty.
 Regardless, a model that included a clone-level random effects term for item difficulty was preferred to models without.
@@
 We also investigated how the structural characteristics of items impact their functioning.
-We found that the number of elements and number of rules were both positively associated with item difficulty.
+We found that element number and rule number were both positively associated with item difficulty.
 These two attributes exerted approximately equal influence on difficulty (i.e.
@@
 One possible reason for this discrepancy is that these two attributes are largely uncorrelated ($\rho$ = 0.176) across items in the MaRs-IB.
-Together the number of elements and number of rules explained 67.6\% of the variance in item difficulty across item templates (and 38.6\% of the variance across item clones), which is in line with previous investigations of matrix reasoning tasks \citep{carpenter1990one, matzen1994error}.
+Together element number and rule number explained 67.6\% of the variance in item difficulty across item templates (and 38.6\% of the variance across item clones), which is in line with previous investigations of matrix reasoning tasks \citep{carpenter1990one, matzen1994error}.
 These findings further validate the design of the MaRs-IB insofar that item complexity strongly determines item difficulty.
@@
 Performance on the MaRs-IB short-forms is summarized in Table \ref{table:3}.
-On average, participants completed the measure in 174 s (sd = 54.0 s) and correctly solved 8.0 of 12 items (sd = 2.5).
+On average, participants completed the measure in 174 s (sd = 54.0 s) out of a possible 360 s total and correctly solved 8.0 of 12 items (sd = 2.5).
 The distribution of observed scores was right-shifted (Figure~\ref{fig:fig04}), as was expected given the nominal guessing rate.
@@
 We also verified that the design of the MaRs-IB was in part successful.
-As intended, item attributes such as feature number and rule number were positively associated with, and explained much of the variability, in item difficulty.
-We also found some limitations in the design of the MaRs-IB.
-For example, distractor type was also associated with item difficulty.
-Moreover, we also found non-negligible residual variance in item difficulty across clones after controlling for item attributes.
-As a consequence, item clones in the MaRs-IB cannot therefore be treated as psychometrically equivalent and exchangeable.
+Item complexity (i.e.
+element number and rule number) was positively associated with, and explained more than half of the variability in, item difficulty.
+In contrast, we also uncovered some undesirable properties of the MaRs-IB.
+Namely, item clones exhibited large differences in difficulty by distractor type, which persisted even after controlling for this and other item attributes.
+As a consequence, item clones in the MaRs-IB cannot be assumed to be psychometrically equivalent and exchangeable.
 
 With the estimated item parameters, we then constructed three new MaRs-IB short-form measures using optimal test assembly methods.
-These abbreviated test measures were designed to measure matrix reasoning ability with maximal reliability in the general adult population under several desirable constraints (i.e.
-12 item limit, administration time of 2-4 minutes).
+These abbreviated test measures were designed to be maximally reliable under time constraints (i.e.
+administration time of 2-4 minutes).
 In a second sample of participants, we found that each short form measure had adequate reliability (Cronbach $\alpha \approx 0.8$) and good convergent validity with an abbreviated Raven's progressive matrices measure.
@@
 children, clinical samples).
-By using item response models here, however, we make possible the opportunity for future IRT ``linking'' studies.
-IRT linking describes a set of methods to establish the comparability of item and ability parameters estimated from data collected from two or more groups that differ in ability \citep{lee2018irt}.
-Future studies utilizing these methods could provide insight not only in how the functioning of the MaRs-IB might change in across populations, but also how matrix reasoning ability differs across populations.
+By using item response models in the present analyses, however, we make possible the opportunity for future IRT ``linking'' studies.
+IRT linking describes a set of methods to establish the comparability of item and ability parameters estimated from response data collected from two or more groups that differ in ability \citep{lee2018irt}.
+Future studies might exploit these methods to provide new insights, not only in how the functioning of the MaRs-IB may differ in across populations, but also in how matrix reasoning ability changes across populations.
 
-A second limitation is in our relatively limited investigation of sources of variance in item functioning.
-Here we investigated the contributions of only three item attributes.
-There were many item properties that we elected not to model (e.g.
-type of changing relation such as color change or shape change).
-Previous investigations of matrix reasoning tasks have proposed other taxonomies of item attributes that affect item functioning \citep{carpenter1990one}.
-Incorporating additional attributes into our models may have further increased the precision of our parameter estimates and might  provide additional knowledge for guiding test assembly.
-Future studies should consider more deeply investigating the possible relationships between other attributes of items in the MaRs-IB and their functioning.
+A second limitation of the present analyses is our relatively restricted investigation of the sources of variance in item functioning in the MaRs-IB.
+Previous psychometric investigations of the Raven's progressive matrices, for example, have highlighted the impact of not only the number of rules but also the types of rules on item difficulty \citep{carpenter1990one, embretson1998cognitive}.
+So too, it is possible that the four rules in the MaRs-IB are not equally difficult or discriminating (e.g.
+color changes are easier to process than position changes) and may explain part of the residual variance in item functioning across item templates.
+Other studies of matrix reasoning tasks have found large impacts of perceptual features on item difficulty \citep{primi2001complexity, primi2014developing}.
+For the reasons stated above, the current analyses did not investigate the contribution of perceptual features to item functioning in the MaRs-IB.
+Future studies should consider investigating more deeply these and other sources of variance in item functioning --- a possibility made easier by the public release of all the data collected as part of the current studies.
 
-In conclusion, we have provided an psychometric validation of the MaRs-IB, finding that it is suitable for measuring matrix reasoning ability in the general adult population.
-We hope that the materials and results presented here will encourage researchers interested in measuring matrix reasoning using the MaRs-IB in their studies.
+In conclusion, we have provided a comprehensive psychometric validation of the MaRs-IB, finding that it is suitable for measuring matrix reasoning ability in the general adult population.
+We hope that the results and materials presented here can server to guide researchers to design their own matrix reasoning tests according to their needs using the MaRs-IB.
 
@@
 \includegraphics[width=1.0\textwidth]{figures/fig02.png}
-\caption{\label{fig:fig02}}
+\caption{\label{fig:fig02} The estimated item difficulty parameters for the items in the MaRs-IB.
+Each dot represents one item clone.
+(A) Item difficulty parameters color-coded by minimal difference (MD)-type and paired difference (PD)-type distractors.
+(B) Item difficulty parameters color-coded by shape set.}
 \end{figure}

### 2022-05-13 manuscript/supplement.tex
manuscript edits

@@
 
-\section*{Speed-accuracy trade-offs in Chierchia et al. (2019)}
+\section*{Speed-accuracy trade-offs in \cite{chierchia2019matrix}}
 
-To investigate the possibility of speed-accuracy trade-offs in the original MaRs-IB data, we looked at the proportion of correct responses to the easiest items (dimension 1/2 items) as a function of the number of participants having reached that item.
-In all partitions of these data, we found robust positive correlations between proportion correct and number reached (dimension 1 items: $\rho$ = 0.514, p = 0.050; dimension 2 items: $\rho$ = 0.767, p < 0.001; combined: $\rho$ = 0.695, p < 0.001).
-This result suggests that the participants that did reach items later in the sequence were able to do by sacrificing accuracy for speed.
-As such, the item-level performance summary released as part of \cite{chierchia2019matrix} are likely biased indicators of item functioning.
+To investigate the possibility of speed-accuracy trade-offs in the MaRs-IB response data collected by \cite{chierchia2019matrix}, we looked at the proportion of correct responses to the easiest items (dimension 1 \& 2 items) as a function of the number of participants having reached that item.
+The logic is that, if items that appeared later in the fixed-order test were disproportionately reached by participants sacrificing accuracy for speed, then we should observe a positive correlation between the total number of available responses and proportion correct amongst the easiest items.
+
+We detected strong positive correlations between proportion correct and number reached (dimension 1 items: $\rho$ = 0.514, p = 0.050; dimension 2 items: $\rho$ = 0.767, p < 0.001; combined: $\rho$ = 0.695, p < 0.001).
+This result supports the hypothesis that participants that did reach items later in the test did so by prioritizing speed at the expense of accuracy.
+This suggests that the item summary statistics released as part of \cite{chierchia2019matrix} are likely biased indicators of item difficulty.
 
@@
 \midrule
-NFC-10 & & 25.03 (8.27) & & 20.00 -- 31.00 & & - &  &  & \\
-PCF-8a   &  & 22.17 (6.29) & & 18.75 -- 26.25 & & 0.27** &  - &  &  \\
-SNS   &  & 29.07 (7.48) & & 25.00 -- 35.00 & & 0.46** &  0.29** &  - &   \\
-MaRs-SF & &   8.00 (2.53) & &   6.00 -- 10.00 & & -0.04 &  0.04 &  0.14* &  - \\
+NFC-10 & & 25.03 (8.27) & & 20.00 -- 31.00 & & \multicolumn{1}{c}{-} &  &  & \\
+PCF-8a   &  & 22.17 (6.29) & & 18.75 -- 26.25 & & 0.27** &  \multicolumn{1}{c}{-} &  &  \\
+SNS   &  & 29.07 (7.48) & & 25.00 -- 35.00 & & 0.46** &  0.29** &  \multicolumn{1}{c}{-} &   \\
+MaRs-SF & &   8.00 (2.53) & &   6.00 -- 10.00 & & -0.04 &  0.04 &  0.14* &  \multicolumn{1}{c}{-} \\
 \bottomrule
@@
 \caption{\normalfont Correlations between performance on the MaRs-IB short forms and self-report measures.
-Abbreviations: IQR = interquartile range; NFC-10 = need for cognition (10-item) scale; PCF = PROMIS cognitive functioning scale (8a); SNS = subjective numeracy scale.
-\\ ** p < 0.001,  * p < 0.05 (not corrected for multiple comparisons)}
+Abbreviations: IQR = interquartile range; NFC-10 = need for cognition (10-item) scale \citep{chiesi2018applying}; PCF = PROMIS cognitive functioning scale (8a; \citealt{iverson2021normative}); SNS = subjective numeracy scale \citep{fagerlin2007measuring}.
+** p < 0.001,  * p < 0.05 (not corrected for multiple comparisons)}
 \end{table}

### 2022-05-13 manuscript/main.tex
latest draft of manuscript

@@
 In a first study, we calibrate the psychometric functioning of the items in the MaRs-IB in a large sample adults participants (N=1501).
-Using additive multilevel item structure models, we establish that the MaRs-IB has many desirable psychometric properties: its items span a wide range of difficulty, possess medium- to-large levels of discrimination, and exhibit as intended certain associations between item attributes and functioning (e.g.
-item complexity predicts item difficulty).
-However, we also find that item clones are not psychometrically equivalent and therefore not exchangeable.
+Using additive multilevel item structure models, we establish that the MaRs-IB has many desirable psychometric properties: its items span a wide range of difficulty, possess medium- to-large levels of discrimination, and exhibit clear associations between item complexity and difficulty.
+However, we also find that item clones are not always psychometrically equivalent and cannot assumed to be exchangeable.
 In a second study, we demonstrate how researchers can use the item parameters we estimated to design new MaRs-IB tests using optimal item assembly.
@@
 
-Noting these challenges, \cite{knoll2016window} / \cite{chierchia2019matrix} developed and made publicly available the matrix reasoning item bank (MaRs-IB), a collection of 80 unique matrix reasoning puzzles.
-Each puzzle in the MaRs-IB consists of a 3x3 matrix containing abstract shapes in eight out of nine cells (Figure \ref{fig:fig00}).
-Across cells, some or all of these shapes change according to various arbitrary abstract rules (e.g.
-color change, position change).
-To solve a puzzle, participants deduce which of four response options correctly completes the matrix based on these rules.
-The puzzles in the MaRs-IB vary in their complexity, both in the number of elements per cell and number of rules determining the relations of elements across cells (Figure \ref{fig:fig00}a/b).
+Noting these challenges, \cite{knoll2016window} / \cite{chierchia2019matrix} developed and made publicly available the matrix reasoning item bank (MaRs-IB), a collection of 80 matrix reasoning puzzles.
+Each puzzle in the MaRs-IB consists of a 3x3 matrix containing geometric shapes in eight out of nine cells (Figure \ref{fig:fig00}).
+Across cells, some or all of these shapes change according to a number of abstract rules.
+Based on these results, participants must deduce which of four response options correctly completes the matrix.
+The puzzles in the MaRs-IB vary in their complexity, both in the number of elements per cell and number of rules determining the relations of these elements across cells (Figure \ref{fig:fig00}a/b).
 To increase the reusability of the MaRs-IB, each puzzle (hereafter referred to as an item template) has six clones.
@@
 
-\cite{chierchia2019matrix} conducted an initial study of the psychometric properties of the MaRs-IB.
-A sample of 659 adult, adolescent, and child participants completed an experiment in which they had eight minutes to complete as many MaRs-IB items as possible (items were presented in the same order to all participants).
+\cite{chierchia2019matrix} conducted an initial study of the psychometric properties of the MaRs-IB in which a sample of 659 adult, adolescent, and child participants had eight minutes to complete as many MaRs-IB items as possible (items were presented in the same order to all participants).
 The authors found that performance on this task had good split-half and test-retest reliability.
-Moreover, they found that participants' ability to solve MaRs-IB puzzles was moderately predictive of their performance on a working memory task and the ICAR matrix reasoning test, indicating good convergent validity.
-At the conclusion of this study, \cite{chierchia2019matrix} publicly released a summary of the difficulty (i.e.
+Moreover, they found that participants' ability to solve MaRs-IB puzzles was moderately predictive of their performance on a working memory task and the ICAR matrix reasoning test, indicating satisfactory convergent validity.
+At the conclusion of this study, \cite{chierchia2019matrix} made publicly available a summary of the difficulty (i.e.
 proportion correct responses) of each item template and clone in the MaRs-IB so that other researchers could construct their own test measures of custom difficulty and duration.
@@
 Two important limitations in the experiment conducted by \cite{chierchia2019matrix} undermine the utility of these summary statistics, and therefore impede the design of new matrix reasoning tasks using the MaRs-IB.
-The first is that, given the fixed-order task design,  there was relatively few responses collected for the majority of the items in the MaRs-IB.
-Indeed, only half (42 of 80 items) were completed by 100 or more participants.
+The first is that, given the fixed-order task design,  there was relatively few responses collected for the majority of the items.
+Indeed only half (42 of 80 items) were completed by 100 or more participants.
 As such, one cannot be confident about the true difficulty of a majority of the items (an issue that is exacerbated when considering that each item has six clones).
@@
 First they provide an estimate of the \textit{difficulty} and \textit{discrimination} of each item, where the latter is an index of how much information an item provides about ability.
-Using these quantities, IRT models in turn allow experimenters to compute the test information function (TIF), which quantifies the reliability of a test (i.e.
-an assemblage of items) in measuring performance across a range of ability levels.
+Using these quantities, IRT models in turn allow experimenters to compute the test information function (TIF), which quantifies the reliability of a test in measuring performance across a range of ability levels.
 Finally, IRT models make possible optimal test assembly  \citep{van1998optimal}, or the design of new tests with maximal reliability given experimenter-defined constraints (e.g.
-test duration, test difficulty).
+test duration and/or difficulty).
 
 A second contribution provided here is an analysis of how item properties shape the psychometric functioning of items in the MaRs-IB.
-Using explanatory item response modeling \citep{de2004explanatory, wilson2008explanatory} we investigate how item complexity, i.e.
-the number of elements and number of rules, explain item functioning (i.e.
+Using explanatory item response modeling \citep{de2004explanatory, wilson2008explanatory}, we investigate how item complexity, i.e.
+the number of elements and number of rules, determine item functioning (i.e.
 difficulty, discrimination).
 We modeled these two item attributes because they have been previously identified as among the strongest predictors of item difficulty in matrix reasoning tasks \citep{embretson1998cognitive, primi2001complexity}.
-Thus this explanatory analysis provides another means by which to validate the success of the MaRs-IB design; that is, to determine whether the MaRs-IB exhibits similar relationships between item characteristics and functioning as observed in other established matrix reasoning measures.
-As a secondary benefit, explanatory IRT models yield more accurate estimates of item parameters \citep{neuhaus2006separating}, which in turn helps to ensure that new MaRs-IB tests designed using these item parameter estimates function as expected.
+Thus this explanatory analysis provides another means by which to validate the success of the MaRs-IB design; that is, to confirm whether the MaRs-IB exhibits similar relationships between item properties and functioning as observed in other established matrix reasoning measures.
+As a secondary benefit, explanatory IRT models yield more accurate estimates of item parameters \citep{neuhaus2006separating}, which in turn helps to ensure that new MaRs-IB tests, designed using the item parameter estimates from these IRT models, function as expected.
 
-A contribution is an investigation of the exchangeability of item clones in the MaRs-IB, or whether clones are (approximately) psychometrically equivalent.
+A third contribution is an investigation of the exchangeability of item clones in the MaRs-IB, or if clones are (approximately) psychometrically equivalent.
 Establishing whether the MaRs-IB possesses this property is critical for the design of new matrix reasoning tasks using these items.
-If item clones are exchangeable, then new parallel test forms of equivalent difficulty and discrimination can be generated simply by swapping out one set of clones for another (e.g.
-exchanging clones from shape set 1 for shape set 2).
+If item clones are exchangeable, then new parallel test forms of equivalent difficulty and discrimination can be generated simply by swapping out one set of item clones for another (e.g.
+exchanging clones that differ only in shape set).
 To test for exchangeability in the MaRs-IB, we use additive multilevel item structure (AMIS) models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
-These models enable us to quantify the variability in item functioning across clones by distractor type and shape set, which should be negligible if clones are to be treated as exchangeable.
+These models enable us to quantify the variability in item functioning across clones by distractor type and shape set, which must be negligible if clones are to be treated as exchangeable.
 
@@
 
-All data, code, and model outputs (including the item parameter estimates) are publicly available at: \url{https://github.com/dawlab/mars-irt}.
+All data, code, and model outputs (including the item parameter estimates) are publicly available at: \url{https://github.com/ndawlab/mars-irt}.
 
@@
 The format of the experiment and instructions were adapted from code publicly released as part of the original study (\url{https://app.gorilla.sc/openmaterials/36164}).
-The task was programmed in jsPsych \citep{de2015jspsych} and distributed using custom web-application software (available at \url{https://github.com/nivlab/nivturk}).
+The task was programmed in jsPsych \citep{de2015jspsych} and distributed to participants via the web using custom software (available at \url{https://github.com/nivlab/nivturk}).
 
-In order to ensure sufficient sampling of every item (and their clones) in the MaRs-IB, participants were administered a pseudorandomly-selected set of 16 out of 64 total items.\footnote{We did not collect new data for the 16 easiest items in the MaRs-IB (items 1, 2, 3, 4, 5, 7, 8, 9, 32, 33, 38, 41, 43, 48, 57, 68) because performance on these items was at or near ceiling.} Item sets were constructed as follows: We subdivided the 64 items into 16 sets of four based on their dimensionality (a measure of item complexity defined in \cite{chierchia2019matrix}).
+In order to ensure sufficient sampling of every item template and clone in the MaRs-IB, participants were administered a pseudorandomly-selected set of 16 out of 64 total items.\footnote{We did not collect new data for the 16 easiest items in the MaRs-IB (items 1, 2, 3, 4, 5, 7, 8, 9, 32, 33, 38, 41, 43, 48, 57, 68) because performance on these items was at or near ceiling.} Item sets were constructed as follows: We subdivided the 64 items into 16 sets of four based on their dimensionality (a measure of item complexity defined in \citealt{chierchia2019matrix}).
 Participants were randomly assigned one item from each of the 16 sets.
 As a consequence, all participants received test forms of roughly equal difficulty.
-Importantly we also counterbalanced the assignment of item clones across participants, such that we had an approximately uniform number of responses available for each item by shape set (1, 2, or 3) and distractor type (minimal difference, MD, or or paired difference, PD; see \cite{chierchia2019matrix} for details).
+Importantly we also counterbalanced the assignment of item clones across participants, such that we had an approximately uniform number of responses available for each item by shape set (1, 2, or 3) and distractor type (minimal difference, MD, or or paired difference, PD; see \citealt{chierchia2019matrix} for details).
 
@@
 
-In our analyses, we considered two item template (level 1) attributes, element number and rule number, which have previously been identified as key determinants of item difficulty in matrix reasoning tasks \citep{primi2001complexity}.
+In our analyses we considered two item template (level 1) attributes, element number and rule number, which have previously been identified as key determinants of item difficulty in matrix reasoning tasks \citep{embretson1998cognitive, primi2001complexity}.
 Element number refers to the number of geometric shapes in each cell of a matrix, whereas rule number refers to the number of relationships that govern the changes of these shapes across cells.
-In the MaRs-IB, there are four principal rules: size change, color change, position change.
+In the MaRs-IB there are four principal rules: size change, color change, position change, and shape change.
 Across item templates, element number ranges from one to four (median = 2) and rule number ranges from one to six (median = 3).
-Importantly, element and rule number are mostly uncorrelated across item templates ($\rho$ = 0.176).
-The feature and rule number for each item was determined via manual annotation.
+Element and rule number are uncorrelated across item templates ($\rho$ = 0.176), and were determined for each item via manual annotation.
 
@@
 We deliberately chose not to incorporate shape set (i.e.
-1, 2, or 3) as a clone-level attribute.
-As described above, every item template in the MaRs-IB has three unique versions that differ in the geometric shapes populating its cells.
+1, 2, or 3) as a clone-level attribute for the following reasons.
+As described above, every item template in the MaRs-IB has three unique versions that differ only in the geometric shapes populating its cells.
 There are 45 unique geometric shapes in total, organized into nine stimulus sets, where each item clone draws a subset of shapes from one of the nine sets (see Figure S2 for examples).
-As such, the groups in tripartite definition of shape set do not constitute meaningful categories and we therefore do not incorporate them as fixed effects.
+As such, shape set --- indicating whether a clone is the 1st, 2nd, or 3rd version of an item template --- is not a meaningful label and we therefore do not incorporate it into the models.
 Furthermore, item clones drawing from the same shape set are not always populated by the same geometric shapes, making complicated the possibility of modeling the nine stimulus sets instead.
-The most rigorous way of modeling the presence of a given geometric shape on item functioning would be to include each as a binary attribute in estimating item difficulty and discrimination.
+The most rigorous way of modeling the presence or absence of a given geometric shape on item functioning would be to include each as a binary attribute predicting item difficulty and discrimination.
 In order to keep the models simple, we elected to leave shape set unmodeled and accounted for by the residual variability terms instead.
@@
 
-The AMIS modeling framework provides a natural means of accomplishing the three goals of the calibration study.
-Through any of the models above, we obtain estimates of the the psychometric properties (i.e.
+The AMIS modeling framework provides a natural means of accomplishing the three objectives of the calibration study.
+Through all of the models above, we obtain estimates of the the psychometric properties (i.e.
 difficulty, discrimination) of every item template and clone in the MaRs-IB (Aim 1).
@@
 
-Finally, through both the level-2 fixed and random effects, we can evaluate whether item clones in the MaRs-IB are exchangeable (Aim 3).
-For example, an association between distractor type and item functioning would entail that clones are not psychometrically equivalent.
-Moreover, if the level-2 residual variance in item difficulty or discrimination were non-negligible, that would similarly entail that item clones are not psychometrically equivalent due to one or more unmodeled features (e.g.
+Finally, through both the level-2 fixed and random effects, we can determine whether item clones in the MaRs-IB are exchangeable (Aim 3).
+For example, an association between distractor type and either item parameter would entail that clones are not psychometrically equivalent.
+Moreover, if either level-2 residual variance term were non-negligible, that would similarly entail that not all item clones are equivalent due to one or more unmodeled features (e.g.
 low-level perceptual properties).
@@
 In other words, response times were slower on average for correct responses, better- performing participants, and more difficult items.
-The positive association between response time and accuracy is consistent with speed-accuracy trade-offs; that is, slower responses are associated with more careful work.
-This interpretation is corroborated by the main effect of rest score, such that more accurate participants were slower overall.
+The positive association between response time and accuracy is consistent with a speed-accuracy trade-off in participants' performance.
+This interpretation is corroborated by the main effect of rest score, where better-performing participants were slower overall.
 
@@
 
-In sum, we found evidence of speed-accuracy trade-offs in performance in this sample.
+In sum, we found evidence of a speed-accuracy trade-off in performance in this sample.
 Accurate responses and more accurate participants were both slower overall.
-Notably, better-performing participants demonstrated adjustments to their response times as function of item difficulty; whereas the lowest scoring participants maintained the same work rate regardless of item difficulty, the highest scoring participant exhibited the largest slowing in responses as items became more challenging.
-We return to this pattern of results in the General Discussion.
+Notably, better-performing participants demonstrated adjustments to their response times as function of item difficulty; whereas the lowest scoring participants maintained the same work rate regardless of item difficulty, the highest scoring participants showed the largest slowing in responding as items became more challenging.
+We return to these results in the General Discussion.
 
-\subsubsection{Model comparison}
+\subsubsection{Item response models}
 
@@
 By comparison, Model 5 demonstrated smaller improvements in fit over Model 4 ($\Delta$ LOCO = 5.12, se = 4.07) and Model 6 ($\Delta$ LOCO = 3.29, se = 1.84).
-This result indicates the presence of non-negligible residual variance in item discrimination across item templates but not clones.
-It must be noted these differences in LOCO values from each model to the next were each within two standard errors of the mean, indicating only weak predictive improvement of Model 6 over the others \citep{vehtari2022cv}.
+This result suggests the presence of non-negligible residual variance in item discrimination across item templates but not clones.
+It must be noted these differences in LOCO values from each model to the next were each within two standard errors of the mean, indicating only weak predictive improvement of Model 5 over the others \citep{vehtari2022cv}.
 As such we will select Model 5 as the best-fitting model overall, but proceed in discussing it with caution.
-We also note that Model 6 fit clearly better than Model 4 ($\Delta$ LOCO = 85.91, se = 6.33, supporting the inclusion of varying item discrimination parameters).
 We also note that all second-wave models exhibited better fits to the data than the first-wave models, supporting the inclusion of item discrimination parameters.
@@
 Items with MD distractors were associated with an estimated 19.9\% reduction in accuracy compared to their equivalent items with PD distractors.
-Together, element rule, rule number, and distractor type explained slightly over half (52.0\%) of the variance in item difficulty.
-Regardless, a model that included a clone-level random effects term for item difficulty was preferred to models without.
-Therefore, the results indicate that item clones are not equally difficult and are thus not exchangeable.
+Together element number, rule number, and distractor type explained slightly over half (52.0\%) of the variance in item difficulty.
+Despite this item response models that included a clone-level residual variance term for item difficulty was preferred to models without.
+Thus, the results indicate that not all item clones are not equally difficult and are thus cannot be assumed to be exchangeable.
 
-The estimated item difficulty per clone, organized by distractor type and shape set, is presented in Figure \ref{fig:fig02}.
+To better demonstrate the residual variance in item difficulty across clones, the estimated item difficulty per clone organized by distractor type and shape set is presented in Figure \ref{fig:fig02}.
 The relative increase in difficulty for items with MD-type distractors compared to PD-type distractors is visually apparent (Figure \ref{fig:fig02}a).
@@
 After controlling for distractor type and mean response time, the magnitude of the clone-level residual variance is $\epsilon_{\beta k}$ = 0.620 (95\% HDI = 0.538 -- 0.701).
-Across the 128 unique sets of three clones, organized by item and distractor type, that level of residual variance translates into an average difference in difficulty of 0.752 (95\% HDI = 0.630 -- 0.877) between the easiest and hardest item clones of each triad, an effect that is two-thirds as large as the effect of distractor type.
-Furthermore, the shape set membership of the most difficult clone varies considerably across the 128 sets of clones (shape set 1 = 39; shape set 2 = 32; shape set 3 = 57).
-Thus, the sources of residual variability in item difficulty across clones is complex and likely reflect the idiosyncratic contributions of their low-level perceptual features.
+Across the 128 unique sets of clones (i.e.
+across item and distractor type), that translates into an average difference in difficulty of 0.752 (95\% HDI = 0.630 -- 0.877) between the easiest and hardest item clones of each set.
+In other words, the magnitude of residual variance in difficulty across clones is roughly two-thirds as large as the effect of distractor type.
+Importantly, the most difficult clone in each triad varied almost uniformly by shape set (shape set 1 = 39; shape set 2 = 32; shape set 3 = 57).
+Thus, the sources of residual variability in item difficulty across clones is complex and likely reflect the idiosyncratic contributions of low-level perceptual features.
 
 Turning to item discrimination, we did not find a credible association between discrimination and distractor type ($\delta_{\alpha_3}$ = -0.029, 95\% HDI = -0.065 -- 0.005).
-Moreover, a model that included a clone-level random effects term for item discrimination was not preferred to a simpler model without.
-As such, there is not sufficient evidence to refute that the assumption that item clones are similarly discriminating.
+Moreover, a model that included a clone-level random effects term for item discrimination was not preferred to a model without.
+As such, there is not sufficient evidence to reject the assumption that item clones are equivalently discriminating.
 We note this result should be interpreted with caution given the relatively weak predictive improvement of the best-fitting model compared to a model with a clone-level random effects term.
@@
 Finally, we inspected the relationship between mean response time and item functioning.
-As expected, there was a positive association between mean response time and item difficulty ($\delta_{\beta_4}$ = 0.541, 95\% HDI = 0.414 -- 0.671).
-That is, response times were slower on average for more difficult items.
-There were not, however, a credible associations between item discrimination and average response time ($\delta_{\alpha_4}$ = -0.010, 95\% HDI = -0.030 -- 0.009).
+As expected, there was a positive association between mean response time and item difficulty ($\delta_{\beta_4}$ = 0.541, 95\% HDI = 0.414 -- 0.671) indicating response times were slower for more difficult items.
+There were not a credible association between item discrimination and average response time ($\delta_{\alpha_4}$ = -0.010, 95\% HDI = -0.030 -- 0.009).
 
@@
 Average response time was positively associated with ability ($\rho_3$ = 0.427, 95\% HDI = 0.374 -- 0.478), as was $\Delta$ response time ($\rho_4$ = 0.368, 95\% HDI = 0.313 -- 0.424).
-In other words, participants with higher levels of ability spent on average more time on each item and deliberated for longer on more difficult items.
+In other words participants with higher levels of ability spent more time on each item on average, and even longer for more difficult items.
 Together, these four person attributes explained 50.3\% of the variance in ability.
-These results are consistent with a speed-accuracy trade-off --- higher levels of ability in this sample reflect in part a tendency to slow down and respond carefully overall and increasingly for more challenging items.
+These results are consistent with a speed-accuracy trade-off.
+That is, higher levels of ability in this sample reflect in part a tendency to slow down and respond carefully overall (and especially for the most challenging items).
 
@@
 The PPP-value for the SGDDM statistic did exceed the critical threshold (PPP-value = 0.019).
-However, the mean of the realized SGDDM values (SGDDM = 0.0180) were only marginally larger than the mean of the posterior predicted SGDDM values (SGDDM = 0.0175).
-That is, the residual inter-item correlations in the data unexplained by the best-fitting model were small on average and only slightly larger than what we would expect to observe by chance.
+However, the mean of the realized SGDDM values ($\text{SGDDM}_r$ = 0.0180) were only marginally larger than the mean of the posterior predicted SGDDM values ($\text{SGDDM}_p$ = 0.0175) meaning that the residual inter-item correlations in the data were small on average and only slightly larger than what we would expect to observe by chance.
 As such, there is little evidence for local dependence in the data.
@@
 
-In this first study, we recruited a large sample of adults to complete items from the MaRs-IB, making sure that each item (and its clones) were completed by a sufficient number of participants.
+In this first study, we recruited a large sample of adults to complete items from the MaRs-IB, making sure that each item template and clones were completed by a sizable number of participants.
 Using multilevel item structure models, we then estimated the psychometric properties (i.e.
@@
 
-We also investigated how the structural characteristics of items impact their functioning.
-We found that element number and rule number were both positively associated with item difficulty.
-These two attributes exerted approximately equal influence on difficulty (i.e.
+We also investigated how item complexity impacts item functioning.
+We found that element number and rule number were both positively associated with item difficulty, the effects of which were of approximately equal magnitude (i.e.
 a one-unit change in either was independently associated with a roughly 10\% reduction in performance).
-This finding is interesting in light of previous investigations of nonverbal reasoning tasks, which have found either the number of elements \citep{bethell1984adaptive} or number of rules \citep{mulholland1980components} to be more determinant of item difficulty.
-One possible reason for this discrepancy is that these two attributes are largely uncorrelated ($\rho$ = 0.176) across items in the MaRs-IB.
+This finding is interesting in light of previous investigations of nonverbal reasoning tasks, which have found a greater influence on item difficulty from either the number of elements \citep{bethell1984adaptive} or number of rules \citep{mulholland1980components}.
+One possible reason for this discrepancy is that these two attributes are largely uncorrelated ($\rho$ = 0.176) across items in the MaRs-IB, which allows for unconfounded estimates of their effects.
 Together element number and rule number explained 67.6\% of the variance in item difficulty across item templates (and 38.6\% of the variance across item clones), which is in line with previous investigations of matrix reasoning tasks \citep{carpenter1990one, matzen1994error}.
-These findings further validate the design of the MaRs-IB insofar that item complexity strongly determines item difficulty.
+These findings further validate the design of the MaRs-IB insofar that item complexity is a primary determinant of item difficulty.
 
 Finally, we investigated whether the item clones in the MaRs-IB are psychometrically equivalent and exchangeable.
-Crucially, we found repeated evidence that item clones are not equally difficult.
+Crucially, we found evidence that item clones are not equally difficult.
 Unexpectedly, and contrary to the results of \cite{chierchia2019matrix}, distractor type emerged as a substantial predictor of item difficulty (i.e.
-items with MD-type distractors are associated with an almost 19.9\% reduction in accuracy compared to the same items with PD-type distractors).
+items with MD-type distractors are associated with a 19.9\% reduction in accuracy compared to the same items with PD-type distractors).
 This discrepancy in findings between studies may reflect our larger sample size (N=1501 versus N=659 participants) or population (adults-only versus a mixture of adults, adolescents, and children), and deserves further investigation.
-Even after controlling for the number of elements, number of rules, and distractor type, we found non-negligible residual variability in item difficulty across clones that were difficult to characterize and likely reflect both the additive and interactive effects of low-level perceptual features.
-Regardless, our results clearly indicate that item clones in the MaRs-IB are not psychometrically equivalent and should not be treated as exchangeable.
+Even after controlling for element number, rule number, and distractor type, we found non-negligible residual variability in item difficulty across clones that was difficult to characterize and likely reflects both the additive and interactive effects of low-level perceptual features.
+Regardless, our results clearly indicate that item clones in the MaRs-IB cannot be assumed to be psychometrically equivalent and should not be treated as exchangeable.
 
@@
 After providing consent, participants first completed three short surveys: the 10-item need for cognition survey \citep{chiesi2018applying}, the 8-item PROMIS Cognitive Function-Short Form (8a) \citep{iverson2021normative}, and the 8-item subject numeracy scale \citep{fagerlin2007measuring}.
-These measures were included as part of secondary analyses  to explore the associations between personality and matrix reasoning performance (all pairwise correlations are reported in Table S2 of the supplementary materials).
+These measures were included as part of exploratory analyses to measure the associations between personality and matrix reasoning ability (these correlations are reported in Table S2 of the supplementary materials).
 Afterwards participants completed one of the three 12-item MaRs-IB short-form measures.
-The administration procedure of the short-forms was identical to that in the first study.
+The administration procedure of the short-forms was identical to that in the previous study.
 Finally, participants completed the 9-item abbreviated Raven's progressive matrices (RPM form A; \citealt{bilker2012development}).
-To maximize consistency in administration, the presentation format and instructions for the RPM task was the same as for the MaRs-IB test forms.
+To maximize consistency in administration, the presentation format and instructions for the RPM task were the same as for the MaRs-IB test forms.
 
@@
 Namely, item clones exhibited large differences in difficulty by distractor type, which persisted even after controlling for this and other item attributes.
-As a consequence, item clones in the MaRs-IB cannot be assumed to be psychometrically equivalent and exchangeable.
+As a consequence, item clones in the MaRs-IB cannot be assumed to be psychometrically equivalent and should not be treated as exchangeable.
 

### 2022-05-13 01_Priors/data/study05/scripts/collate_data.py
add documentation to pilot data

@@ -1,0 +1,186 @@
+import os, json
+import numpy as np
+from os.path import dirname, join
+from pandas import DataFrame, concat, read_csv
+ROOT_DIR = dirname(dirname(os.path.realpath(__file__)))
+RAW_DIR = os.path.join(ROOT_DIR, 'raw')
+DATA_DIR = os.path.join(ROOT_DIR, 'data')
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Assemble metadata.
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## Locate files.
+files = sorted([f for f in os.listdir(RAW_DIR) if f.endswith('surveys.json')])
+
+## Main loop.
+METADATA = []
+for f in files:
+    
+    ## Define subject
+    subject = f.replace('_surveys.json','')
+    
+    ## Load JSON.
+    with open(os.path.join(RAW_DIR, f), 'r') as tmp:
+        JSON = json.load(tmp)
+        
+    ## Initialize dictionary.
+    dd = dict(subject=subject)
+    
+    ## Add metadata.
+    demo, = [dd for dd in JSON if dd['trial_type'] == 'survey-demo']
+    for k,v in demo['responses'].items(): dd[k] = v
+        
+    ## Append.
+    METADATA.append(dd)
+    
+## Concatenate and save data.
+METADATA = DataFrame(METADATA).sort_values(['subject']).rename(columns={'gender-categorical':'gender'})
+METADATA.to_csv(os.path.join(DATA_DIR, 'metadata.csv'), index=False)
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Assemble surveys.
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## Locate files.
+files = sorted([f for f in os.listdir(RAW_DIR) if f.endswith('surveys.json')])
+
+## Main loop.
+SURVEYS = []
+for f in files:
+    
+    ## Define subject
+    subject = f.replace('_surveys.json','')
+    
+    ## Load JSON.
+    with open(os.path.join(RAW_DIR, f), 'r') as tmp:
+        JSON = json.load(tmp)
+        
+    ## Initialize dictionary.
+    dd = dict(subject=subject)
+    
+    ## Gather surveys.
[…]

### 2022-05-24 manuscript/main.tex
manuscript updates

@@
 In a first study, we calibrate the psychometric functioning of the items in the MaRs-IB in a large sample adults participants (N=1501).
-Using additive multilevel item structure models, we establish that the MaRs-IB has many desirable psychometric properties: its items span a wide range of difficulty, possess medium- to-large levels of discrimination, and exhibit clear associations between item complexity and difficulty.
+Using additive multilevel item structure models, we establish that the MaRs-IB has many desirable psychometric properties: its items span a wide range of difficulty, possess medium- to-large levels of discrimination, and exhibit robust associations between item complexity and difficulty.
 However, we also find that item clones are not always psychometrically equivalent and cannot assumed to be exchangeable.
@@
 With these methods, we design three parallel short-form measures that exhibited good score reliability and convergent validity in a second sample of adults (N=300).
-We hope that the materials and results made available here will encourage researchers to use the MaRs-IB in their studies.}
+We hope that the materials, results, and tutorials made available here will encourage researchers to use the MaRs-IB in their studies.}
 
@@
 In low-stakes testing settings (i.e.
-where a participant incurs little or no cost for poor performance), matrix reasoning tasks can additionally function as measures of motivation, willingness to expend mental effort, and other facets of personality \citep{duckworth2011role, gignac2019maximum}.
+where participants incur little or no cost for poor performance), matrix reasoning tasks can additionally function as measures of motivation, willingness to expend mental effort, and other facets of personality \citep{duckworth2011role, gignac2019maximum}.
 In studies of psychiatric populations, performance on matrix reasoning tasks have also been used to control for general disruptions to cognitive ability when specific domains of cognition are of primary interest \citep{gillan2016characterizing, rouault2018psychiatric, moutoussis2021decision}.
@@
 One challenge is the problem of copyright.
-Many of the most prominent matrix reasoning tasks, such as the WAIS and WASI matrix reasoning subtests \citep{wechsler1999wechsler, corporation2008wechsler}, are not free to use and have legal restrictions against being computerized.
+Many of the most prominent matrix reasoning tasks, such as the WAIS and WASI matrix reasoning subtests \citep{wechsler1999wechsler, corporation2008wechsler}, are not free to use and have legal restrictions against digitization.
 A second challenge is that, across all of the matrix reasoning tests in the public domain, there are relatively few unique items available.
@@
 Across cells, some or all of these shapes change according to a number of abstract rules.
-Based on these results, participants must deduce which of four response options correctly completes the matrix.
+Based on these rules, participants must deduce which of four response options correctly completes the matrix.
 The puzzles in the MaRs-IB vary in their complexity, both in the number of elements per cell and number of rules determining the relations of these elements across cells (Figure \ref{fig:fig00}a/b).
 To increase the reusability of the MaRs-IB, each puzzle (hereafter referred to as an item template) has six clones.
-A single item's clones share identical structural characteristics (i.e.
+A single item's clones possess identical structural properties (i.e.
 number of elements, number of rules) but may vary in their distractors (two unique sets per template) or perceptual features (three unique shape sets per template; Figure \ref{fig:fig00}c/d).
@@
 
-\cite{chierchia2019matrix} conducted an initial study of the psychometric properties of the MaRs-IB in which a sample of 659 adult, adolescent, and child participants had eight minutes to complete as many MaRs-IB items as possible (items were presented in the same order to all participants).
+\cite{chierchia2019matrix} conducted an initial study of the psychometric properties of the MaRs-IB in which a sample of 659 adult, adolescent, and child participants had eight minutes to complete as many MaRs-IB items as possible.
+Of note, items were presented in the same order to all participants.
 The authors found that performance on this task had good split-half and test-retest reliability.
@@
 The first is that, given the fixed-order task design,  there was relatively few responses collected for the majority of the items.
-Indeed only half (42 of 80 items) were completed by 100 or more participants.
-As such, one cannot be confident about the true difficulty of a majority of the items (an issue that is exacerbated when considering that each item has six clones).
+Indeed only 42 of 80 item templates were completed by 100 or more participants.
+As such, one cannot be confident about the true difficulty of a majority of the items (an issue that is exacerbated when considering that each template has six clones).
 A second issue is participants were able to choose for themselves whether to prioritize accuracy (number of correct responses) or productivity (number of items reached).
@@
 First they provide an estimate of the \textit{difficulty} and \textit{discrimination} of each item, where the latter is an index of how much information an item provides about ability.
-Using these quantities, IRT models in turn allow experimenters to compute the test information function (TIF), which quantifies the reliability of a test in measuring performance across a range of ability levels.
-Finally, IRT models make possible optimal test assembly  \citep{van1998optimal}, or the design of new tests with maximal reliability given experimenter-defined constraints (e.g.
-test duration and/or difficulty).
+Using these quantities, IRT models in turn allow experimenters to compute the test information function (TIF), which quantifies the reliability of a test in measuring performance given a level of ability.
+Finally, IRT models make possible optimal test assembly  \citep{van1998optimal}, or the design of new tests with maximal reliability given researcher-specified constraints (e.g.
+test duration or difficulty).
 
-A second contribution provided here is an analysis of how item properties shape the psychometric functioning of items in the MaRs-IB.
-Using explanatory item response modeling \citep{de2004explanatory, wilson2008explanatory}, we investigate how item complexity, i.e.
-the number of elements and number of rules, determine item functioning (i.e.
-difficulty, discrimination).
+A second contribution provided here is an analysis of how item properties shape the psychometric functioning (i.e.
+difficulty, discrimination) of items in the MaRs-IB.
+Using explanatory item response modeling \citep{de2004explanatory, wilson2008explanatory}, we specifically investigate how item complexity, i.e.
+the number of elements and number of rules, determine item functioning.
 We modeled these two item attributes because they have been previously identified as among the strongest predictors of item difficulty in matrix reasoning tasks \citep{embretson1998cognitive, primi2001complexity}.
-Thus this explanatory analysis provides another means by which to validate the success of the MaRs-IB design; that is, to confirm whether the MaRs-IB exhibits similar relationships between item properties and functioning as observed in other established matrix reasoning measures.
-As a secondary benefit, explanatory IRT models yield more accurate estimates of item parameters \citep{neuhaus2006separating}, which in turn helps to ensure that new MaRs-IB tests, designed using the item parameter estimates from these IRT models, function as expected.
+This explanatory analysis provides another means by which to validate the design of the MaRs-IB; that is, to determine whether the MaRs-IB exhibits the same associations between item properties and functioning as found in other established matrix reasoning tests.
+As a secondary benefit, explanatory IRT models yield more precise estimates of item parameters \citep{neuhaus2006separating}, which in turn help to ensure that any new MaRs-IB tests designed based on those estimates function as expected.
 
@@
 Establishing whether the MaRs-IB possesses this property is critical for the design of new matrix reasoning tasks using these items.
-If item clones are exchangeable, then new parallel test forms of equivalent difficulty and discrimination can be generated simply by swapping out one set of item clones for another (e.g.
-exchanging clones that differ only in shape set).
-To test for exchangeability in the MaRs-IB, we use additive multilevel item structure (AMIS) models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
-These models enable us to quantify the variability in item functioning across clones by distractor type and shape set, which must be negligible if clones are to be treated as exchangeable.
+If item clones are exchangeable, then new parallel test forms of equivalent difficulty and discrimination can be generated simply by substituting a an item clone from one shape set for another.
+To test for exchangeability, we use additive multilevel item structure (AMIS) models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
+These models enable us to quantify the variability in item difficulty and discrimination across clones, by distractor type and shape set, which must be negligible if clones are to be treated as exchangeable.
 
 The remainder of the paper proceeds as follows.
-First we report a calibration study in which a large sample of participants each completed a subset of items from the MaRs-IB.
+First we report a calibration study in which a large sample of adult participants completed a subset of items from the MaRs-IB.
 Using their response data, we then fit a series of multilevel item structure models in order to estimate item parameters for each item (and their clones).
-Using these same models, we also interrogate the relationship between item attributes and functioning within and cross items in the MaRs-IB.
+Using these same models, we also interrogate the relationship between item attributes and functioning items in the MaRs-IB.
 Then we report a second validation study, in which a smaller sample of participants completed one of three new MaRs-IB short-form tests.
 These tests --- designed using optimal test assembly methods applied to the item parameters estimated during the calibration study --- were found to have good psychometric properties and convergent validity.
-This second study provides a blueprint for how to construct new MaRs-IB test measures using the item parameter estimates that have been released as part of this study.
+This second study provides a blueprint for how to construct new MaRs-IB test measures using the item parameter estimates that were derived in the calibration study.
 
-All data, code, and model outputs (including the item parameter estimates) are publicly available at: \url{https://github.com/ndawlab/mars-irt}.
+All data, code, model outputs (including the estimated item parameters), and tutorials are publicly available at: \url{https://github.com/ndawlab/mars-irt}.
 
-\clearpage
 \section{Calibration Study}
@@
 
-The purpose of the first study was to calibrate the items in the MaRs-IB using response data collected from a large number of adult participants from the general population.
+The purpose of the first study was to calibrate the items in the MaRs-IB using response data collected from a large number of adult participants recruited from the general population.
 In particular, we sought to accomplish the following three aims: (1) to quantify the psychometric properties (i.e.
-difficulty, discrimination) of the items in the MaRs-IB using item response models; (2) to measure the associations between item complexity and functioning; and (3) to determiner whether item clones in the MaRs-IB are exchangeable (i.e.
-psychometrically equivalent).
+difficulty, discrimination) of the items in the MaRs-IB using item response models; (2) to measure the associations between item complexity and functioning; and (3) to determine whether item clones in the MaRs-IB are psychometrically equivalent and exchangeable.
 
@@
 Participants were randomly assigned one item from each of the 16 sets.
-As a consequence, all participants received test forms of roughly equal difficulty.
-Importantly we also counterbalanced the assignment of item clones across participants, such that we had an approximately uniform number of responses available for each item by shape set (1, 2, or 3) and distractor type (minimal difference, MD, or or paired difference, PD; see \citealt{chierchia2019matrix} for details).
+As such, all participants received test forms of roughly equal difficulty.
+Importantly we also counterbalanced the assignment of item clones across participants, such that we had an approximately uniform number of responses available for each item by shape set (1, 2, or 3) and distractor type (minimal difference, MD, or paired difference, PD; see \citealt{chierchia2019matrix} for details).
 
@@
 
-In our analyses we considered two item template (level 1) attributes, element number and rule number, which have previously been identified as key determinants of item difficulty in matrix reasoning tasks \citep{embretson1998cognitive, primi2001complexity}.
-Element number refers to the number of geometric shapes in each cell of a matrix, whereas rule number refers to the number of relationships that govern the changes of these shapes across cells.
+In our analyses we considered two item template (level 1) attributes: element number and rule number.
+These have previously been identified as key determinants of item difficulty in matrix reasoning tasks \citep{embretson1998cognitive, primi2001complexity}.
+Element number refers to the number of geometric shapes in each cell of a matrix, whereas rule number refers to the number of relationships that govern the changes among these shapes across cells.
 In the MaRs-IB there are four principal rules: size change, color change, position change, and shape change.
@@
 Distractor type refers to whether a clone uses MD-type versus PD-type distractors, whereas mean response time refers to the average time participants spent deliberating on that item clone.
-We note that, in contrast to all other modeled attributes, average response time is not an intrinsic property of an item.
-We elected to include it in our models, however, because including response time as a covariate has previously been shown to improve parameter estimation \citep{bertling2018using}.
-To preclude any collinearity between the attributes, mean response time was orthogonalized with respect to the three other item attributes; as such, any effects of response time reflects residual structure after accounting for all other item attributes.
+We note that, in contrast to the other modeled attributes, average response time is not an intrinsic property of an item.
+We elected to include it in our models, however, because including response time as a covariate has previously been shown to improve parameter estimation \citep{bertling2018using}.\footnote{Removing mean response time as a clone-level attribute did not appreciably change the estimated item parameters; see Figure S1.} To preclude any collinearity between the attributes, mean response time was orthogonalized with respect to the three other item attributes; as such, any effects of response time reflects residual structure after accounting for all other item attributes.
 
 We deliberately chose not to incorporate shape set (i.e.
-1, 2, or 3) as a clone-level attribute for the following reasons.
-As described above, every item template in the MaRs-IB has three unique versions that differ only in the geometric shapes populating its cells.
-There are 45 unique geometric shapes in total, organized into nine stimulus sets, where each item clone draws a subset of shapes from one of the nine sets (see Figure S2 for examples).
-As such, shape set --- indicating whether a clone is the 1st, 2nd, or 3rd version of an item template --- is not a meaningful label and we therefore do not incorporate it into the models.
-Furthermore, item clones drawing from the same shape set are not always populated by the same geometric shapes, making complicated the possibility of modeling the nine stimulus sets instead.
-The most rigorous way of modeling the presence or absence of a given geometric shape on item functioning would be to include each as a binary attribute predicting item difficulty and discrimination.
-In order to keep the models simple, we elected to leave shape set unmodeled and accounted for by the residual variability terms instead.
+1, 2, or 3) as a clone-level attribute.
+The allocation of geometric shapes to item clones in the MaRs-IB is complex, such that not all clones of the same shape set share the same perceptual elements (see the supplementary materials for a full discussion).
+In other words, shape set does not constitute a meaningful nominal variable and was therefore not included in the model.
+Instead, the contribution of perceptual elements to item functioning is captured by residual variability terms of the item structure models.
 
@@
 The AMIS modeling framework provides a natural means of accomplishing the three objectives of the calibration study.
-Through all of the models above, we obtain estimates of the the psychometric properties (i.e.
+Via all of the models above, we obtain estimates of the the psychometric properties (i.e.
 difficulty, discrimination) of every item template and clone in the MaRs-IB (Aim 1).
@@
 Finally, through both the level-2 fixed and random effects, we can determine whether item clones in the MaRs-IB are exchangeable (Aim 3).
-For example, an association between distractor type and either item parameter would entail that clones are not psychometrically equivalent.
-Moreover, if either level-2 residual variance term were non-negligible, that would similarly entail that not all item clones are equivalent due to one or more unmodeled features (e.g.
-low-level perceptual properties).
+For example, a credible association between distractor type and either item parameter would entail that clones are not psychometrically equivalent.
+Moreover, if either level-2 residual variance term were non-negligible, that would similarly entail that not all item clones are equivalent due to unmodeled attributes (e.g.
+low-level perceptual features).
 
@@
 In each of the six models described above, we incorporated person-level attributes as predictors of latent ability ($\theta_i$) following an explanatory item response modeling approach \citep{wilson2008explanatory}:
-
 \begin{equation} \label{eq:4}
@@
 We estimated person-level abilities as a function of four covariates: age, gender, mean response time, and $\Delta$ response time.
-This last term reflects the degree of change in a participant's deliberation time as a function of item difficulty (defined as one minus the proportion of correct responses for that item).
-Mean response time and $\Delta$ response time were calculated for each participant simultaneously through linear regression, where a given participant's (log-transformed) response times were regressed against an intercept and item difficulty.
+This last term reflects the degree of change in a participant's response time as a function of item difficulty (defined as one minus the proportion of correct responses for that item), and was included because of its relationship to participants' performance (see Results).
+Mean response time and $\Delta$ response time were calculated for each participant simultaneously via linear regression, where a given participant's (log-transformed) response times were regressed against an intercept and item difficulty.
 All covariates were standard-scored prior to analysis with the exception of gender, which was binary coded (male = -0.5, female = 0.5).
@@
 In turn, SGDDM measures the mean absolute conditional correlation between all pairs of items; that is, the SGDDM is an index of residual inter-item correlations unexplained by the model.
-The SGDDM is important in this context as it tests for local dependence across items, the presence of which might otherwise lower the efficiency of future test forms via the inclusion of nonindependent or redundant items.
+The SGDDM is important in this context because it tests for local dependence across items, the presence of which might otherwise lower the efficiency of future test forms by including nonindependent or redundant items.
 For both discrepancy statistics, the PPP value is the proportion of draws in which the posterior predictive discrepancy is equal to or higher than the realized discrepancy.
@@
 On average, participants completed 9.6 of 16 items correctly (sd = 3.1, IQR = 8.0 -- 12.0) (Figure \ref{fig:fig01}a).
-Timing out occurred on only 1.8\% of trials.
+Timing out occurred on only 1.8\% of trials (these were subsequently coded as incorrect responses).
 A total of 43 participants (2.9\%) performed below chance (i.e.
@@
 Accurate responses and more accurate participants were both slower overall.
-Notably, better-performing participants demonstrated adjustments to their response times as function of item difficulty; whereas the lowest scoring participants maintained the same work rate regardless of item difficulty, the highest scoring participants showed the largest slowing in responding as items became more challenging.
+Notably, better-performing participants demonstrated adjustments to their response times as function of item difficulty; whereas the lowest scoring participants maintained an equivalent work rate regardless of item difficulty, the highest scoring participants showed the largest slowing in responding as items became more challenging.
 We return to these results in the General Discussion.
@@
 In the second wave of model fitting, in which models varied only in their specification of item discrimination, Model 5 demonstrated the best fit to the data.
-By comparison, Model 5 demonstrated smaller improvements in fit over Model 4 ($\Delta$ LOCO = 5.12, se = 4.07) and Model 6 ($\Delta$ LOCO = 3.29, se = 1.84).
+In contrast, Model 5 yielded smaller improvements in fit over Model 4 ($\Delta$ LOCO = 5.12, se = 4.07) and Model 6 ($\Delta$ LOCO = 3.29, se = 1.84).
 This result suggests the presence of non-negligible residual variance in item discrimination across item templates but not clones.
@@
 Together element number and rule number explained approximately 67.6\% of the variance in difficulty across the 64 item templates.
-In other words, the majority of variability in difficulty across item templates can be attributed to their complexity.
-Individually the number of elements and number of rules explain 29.5\% and 40.6\% of the variance in item template difficulty, respectively.
+That is, the majority of variability in difficulty across item templates can be attributed to their complexity.
+Individually the number of elements and number of rules respectively explain 29.5\% and 40.6\% of the variance in item template difficulty.
 Across all 384 item clones, these two attributes explain 38.6\% of the variance in difficulty.
@@
 Together element number, rule number, and distractor type explained slightly over half (52.0\%) of the variance in item difficulty.
-Despite this item response models that included a clone-level residual variance term for item difficulty was preferred to models without.
-Thus, the results indicate that not all item clones are not equally difficult and are thus cannot be assumed to be exchangeable.
+Despite this, an item response model that included a clone- level residual variance term for item difficulty was preferred to one without.
+In sum, the results clearly indicate that item clones are not equally difficult and cannot be assumed to be exchangeable.
 
-To better demonstrate the residual variance in item difficulty across clones, the estimated item difficulty per clone organized by distractor type and shape set is presented in Figure \ref{fig:fig02}.
+To make clear the residual variance in item difficulty across clones, the estimated item difficulty per clone is presented in Figure \ref{fig:fig02}.
 The relative increase in difficulty for items with MD-type distractors compared to PD-type distractors is visually apparent (Figure \ref{fig:fig02}a).
-In contrast, the pattern of item difficulties across clones by shape set is more complicated (Figure \ref{fig:fig02}b).
-After controlling for distractor type and mean response time, the magnitude of the clone-level residual variance is $\epsilon_{\beta k}$ = 0.620 (95\% HDI = 0.538 -- 0.701).
-Across the 128 unique sets of clones (i.e.
-across item and distractor type), that translates into an average difference in difficulty of 0.752 (95\% HDI = 0.630 -- 0.877) between the easiest and hardest item clones of each set.
-In other words, the magnitude of residual variance in difficulty across clones is roughly two-thirds as large as the effect of distractor type.
-Importantly, the most difficult clone in each triad varied almost uniformly by shape set (shape set 1 = 39; shape set 2 = 32; shape set 3 = 57).
-Thus, the sources of residual variability in item difficulty across clones is complex and likely reflect the idiosyncratic contributions of low-level perceptual features.
+In contrast the pattern of residual difficulty across clones, after accounting for all modeled sources of variance, is more complicated (Figure \ref{fig:fig02}b).
+The magnitude of the clone-level residual variance is $\epsilon_{\beta k}$ = 0.620 (95\% HDI = 0.538 -- 0.701), or 35.1\% of the total variance in difficulty across item clones.
+This corresponds to a mean absolute difference in difficulty of 0.705 (95\% HDI = 0.612 -- 0.786) across clones per item and distractor type, or roughly two-thirds as large as the effect of distractor type.
+Thus, the residual variability in item difficulty across clones is large, complex, and likely reflects the idiosyncratic contributions of low-level perceptual features.
 
@@
 Together, these four person attributes explained 50.3\% of the variance in ability.
-These results are consistent with a speed-accuracy trade-off.
-That is, higher levels of ability in this sample reflect in part a tendency to slow down and respond carefully overall (and especially for the most challenging items).
+These results are consistent with a speed-accuracy trade-off in performance.
+That is, higher levels of ability in this sample reflect in part a tendency to slow down and respond more carefully on average (and especially for the most challenging items).
 
@@
 
-Finally, we validated the fit of the best-fitting model using two posterior predictive model checking measures.
+Finally, we checked the fit of the best-fitting model using two posterior predictive model checking measures.
 The PPP-value for chi-square discrepancy statistic did not exceed the critical threshold ($\chi^2_{NC}$ = 15.904, PPP-value = 0.297), indicating that the model was sufficiently able to reproduce the distribution of observed scores.
 The PPP-value for the SGDDM statistic did exceed the critical threshold (PPP-value = 0.019).
-However, the mean of the realized SGDDM values ($\text{SGDDM}_r$ = 0.0180) were only marginally larger than the mean of the posterior predicted SGDDM values ($\text{SGDDM}_p$ = 0.0175) meaning that the residual inter-item correlations in the data were small on average and only slightly larger than what we would expect to observe by chance.
+However, the mean of the realized SGDDM values ($\text{SGDDM}_r$ = 0.0180) was only marginally larger than the mean of the posterior predicted SGDDM values ($\text{SGDDM}_p$ = 0.0175).
+This means that the residual inter-item correlations in the data were small on average and only slightly larger than what we would expect to observe by chance.
 As such, there is little evidence for local dependence in the data.
@@
 difficulty and discrimination) of each item.
-Corroborating the results of an initial investigation of the MaRs-IB \citep{chierchia2019matrix}, we found that the items in the MaRs-IB vary greatly in their difficulty -- a prerequisite for measuring nonverbal reasoning across the ability spectrum.
-We also found that the items in the MaRs-IB exhibited medium-to-large levels of discrimination, similar to what has been found in investigations of other of matrix reasoning tasks \citep{chiesi2012using, chiesi2012item, van2013shortened}.
+Corroborating the results of an initial investigation of the MaRs-IB \citep{chierchia2019matrix}, we found that the items in the MaRs-IB vary greatly in their difficulty --- a prerequisite for measuring nonverbal reasoning across the spectrum of ability.
+We also found that the items in the MaRs-IB exhibited medium-to-large levels of discrimination, similar to what has been found in investigations of other matrix reasoning tasks \citep{chiesi2012using, chiesi2012item, van2013shortened}.
 
-We also investigated how item complexity impacts item functioning.
+We also investigated how item complexity shapes item functioning.
 We found that element number and rule number were both positively associated with item difficulty, the effects of which were of approximately equal magnitude (i.e.
@@
 One possible reason for this discrepancy is that these two attributes are largely uncorrelated ($\rho$ = 0.176) across items in the MaRs-IB, which allows for unconfounded estimates of their effects.
-Together element number and rule number explained 67.6\% of the variance in item difficulty across item templates (and 38.6\% of the variance across item clones), which is in line with previous investigations of matrix reasoning tasks \citep{carpenter1990one, matzen1994error}.
+Together element number and rule number explained 67.6\% of the variance in difficulty across templates (and 38.6\% of the variance across item clones), which is in line with previous investigations of matrix reasoning tasks \citep{carpenter1990one, matzen1994error}.
 These findings further validate the design of the MaRs-IB insofar that item complexity is a primary determinant of item difficulty.
@@
 Crucially, we found evidence that item clones are not equally difficult.
-Unexpectedly, and contrary to the results of \cite{chierchia2019matrix}, distractor type emerged as a substantial predictor of item difficulty (i.e.
+Unexpectedly, and contrary to the results of \cite{chierchia2019matrix}, distractor type emerged as a robust predictor of item difficulty (i.e.
 items with MD-type distractors are associated with a 19.9\% reduction in accuracy compared to the same items with PD-type distractors).
-This discrepancy in findings between studies may reflect our larger sample size (N=1501 versus N=659 participants) or population (adults-only versus a mixture of adults, adolescents, and children), and deserves further investigation.
-Even after controlling for element number, rule number, and distractor type, we found non-negligible residual variability in item difficulty across clones that was difficult to characterize and likely reflects both the additive and interactive effects of low-level perceptual features.
+This discrepancy in findings between studies may reflect the much larger number of responses available per clone in our study.
+We found non-negligible residual variability in item difficulty across clones that was difficult to characterize and likely reflects the effects of low-level perceptual features.
 Regardless, our results clearly indicate that item clones in the MaRs-IB cannot be assumed to be psychometrically equivalent and should not be treated as exchangeable.
@@
 In summary, the results of the calibration study confirm the MaRs-IB to be a useful and psychometrically valid resource for measuring matrix reasoning ability.
-The items vary considerably in their difficulty, and all posses medium-to-large levels of discrimination, thereby making the MaRs-IB suitable for measuring matrix reasoning across a wide range of ability levels.
-That item clones are not equivalently difficult, and should not therefore be treated as exchangeable, does not invalidate this conclusion.
+The items vary greatly in their difficulty, and possess medium-to-large levels of discrimination, thereby making the MaRs-IB suitable for measuring matrix reasoning across a wide range of ability.
+That item clones are not equivalently difficult, and therefore should not be treated as exchangeable, does not invalidate this conclusion.
 The item parameter estimates derived as part of this study are therefore suitable for use in designing new MaRs-IB measures with both good reliability and potential for reuse, as we demonstrate in the next study.
@@
 These results are consistent with a persistence interpretation of ability, wherein good task performance in part reflects a willingness to invest time in the solution process and poor task performance reflects a tendency to give up sooner \citep{ranger2021effects}.
-As such, the MaRs-IB may be suitable not only for measuring matrix reasoning ability but also mental effort costs \citep{kool2018mental}, opportunity costs \citep{otto2019opportunity}, or other motivational factors \citep{duckworth2011role} related to the tendency to exert effort or give up.
+As such, the MaRs-IB may be suitable not only for measuring matrix reasoning ability but also mental effort costs \citep{kool2018mental}, opportunity costs \citep{payne1996time}, or other motivational factors \citep{duckworth2011role} related to the tendency to exert effort or give up.
 The use of more sophisticated models may help to disentangle the relative contributions of latent ability and persistence to performance on the MaRs-IB \citep{ranger2014accumulator}.
@@
 
-A second limitation of the present analyses is our relatively restricted investigation of the sources of variance in item functioning in the MaRs-IB.
+A second limitation of the present analyses is our relatively narrow investigation of the sources of variance in item functioning in the MaRs-IB.
 Previous psychometric investigations of the Raven's progressive matrices, for example, have highlighted the impact of not only the number of rules but also the types of rules on item difficulty \citep{carpenter1990one, embretson1998cognitive}.
@@
 color changes are easier to process than position changes) and may explain part of the residual variance in item functioning across item templates.
-Other studies of matrix reasoning tasks have found large impacts of perceptual features on item difficulty \citep{primi2001complexity, primi2014developing}.
-For the reasons stated above, the current analyses did not investigate the contribution of perceptual features to item functioning in the MaRs-IB.
-Future studies should consider investigating more deeply these and other sources of variance in item functioning --- a possibility made easier by the public release of all the data collected as part of the current studies.
+Other studies of matrix reasoning tasks have identified large contributions of perceptual features to item difficulty \citep{primi2001complexity, primi2014developing}, effects which were not explicitly considered here.
+Future studies should consider investigating these and other sources of variance in item functioning --- a possibility made easier by the public release of all the data collected as part of the current studies.
 
 In conclusion, we have provided a comprehensive psychometric validation of the MaRs-IB, finding that it is suitable for measuring matrix reasoning ability in the general adult population.
-We hope that the results and materials presented here can server to guide researchers to design their own matrix reasoning tests according to their needs using the MaRs-IB.
+We hope that the results and materials presented here will encourage researchers to design their own matrix reasoning tests, tailored to their needs, using the MaRs-IB.
+To facilitate this all tutorials for optimal test assembly using the estimated item parameters are available at: \url{https://github.com/ndawlab/mars-irt}.
 
@@
 Each dot represents one item clone.
-(A) Item difficulty parameters color-coded by minimal difference (MD)-type and paired difference (PD)-type distractors.
-(B) Item difficulty parameters color-coded by shape set.}
+(A) Item difficulty color-coded by minimal difference (MD)-type (blue) and paired difference (PD)-type (grey) distractors.
+(B) Residual item difficulty across item clones, after accounting for all other modeled sources of variance, color-coded by shape set.}
 \end{figure}

### 2022-05-24 manuscript/supplement.tex
manuscript updates

@@
 
-The estimated weights ($w_{ij}$) across all subjects and items are plotted as a function of their corresponding response time in Figure S2.
+The estimated weights ($w_{ij}$) across all subjects and items are plotted as a function of their corresponding response time in Figure S3.
 As can be observed, the weights quickly approach 1 for response times faster than 5 seconds.
@@
 Thus, in the main experiments we defined rapid guessing as responses taking fewer than 3 seconds.
+
+\section{Organization of geometric stimuli in the MaRs-IB}
+
+Every item template in the MaRs-IB has three unique versions that differ only in the geometric shapes populating its cells.
+There are 45 unique geometric shapes in total, organized into nine stimulus sets, where each item clone draws a subset of shapes from one of the nine sets (Figure S2).
+As such, shape set --- indicating whether a clone is the 1st, 2nd, or 3rd version of an item template --- is not a meaningful nominal variable.
+Furthermore, item clones drawing from the same shape set are not always populated by the same geometric shapes, making complicated the possibility of modeling the nine stimulus sets instead.
+The most rigorous way of modeling the presence or absence of a given geometric shape on item functioning would be to include each as a binary attribute predicting item difficulty and discrimination.
+In order to keep our models simple, we elected to leave shape set unmodeled and accounted for by the residual variability terms instead.
 
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/figS01.png}
-\caption{\label{fig:figS01} An example item clone from each of the nine stimulus sets.
+\includegraphics[width=\textwidth]{figures/figS01.png}
+\caption{\label{fig:figS01} Item parameter estimates for the best-fitting model (Model 5) with and without including mean response time as a clone-level attribute.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=1.0\textwidth]{figures/figS02.png}
+\caption{\label{fig:figS02} An example item clone from each of the nine stimulus sets.
 In each row are three clones of one item template.
@@
 \centering
-\includegraphics[width=0.8\textwidth]{figures/figS02.png}
-\caption{\label{fig:figS02} Estimates of the rapid guessing weights ($w_{ij}$) and their corresponding responses times from the effort-moderated item response theory (EM-IRT) model fit to response data from N=180 pilot participants.
+\includegraphics[width=0.8\textwidth]{figures/figS03.png}
+\caption{\label{fig:figS03} Estimates of the rapid guessing weights ($w_{ij}$) and their corresponding responses times from the effort-moderated item response theory (EM-IRT) model fit to response data from N=180 pilot participants.
 The dashed line indicates the chosen rapid guessing threshold, i.e.

### 2022-05-26 manuscript/main.tex
camera ready

@@
 To increase the reusability of the MaRs-IB, each puzzle (hereafter referred to as an item template) has six clones.
-A single item's clones possess identical structural properties (i.e.
+A single item's clones share the same level of complexity (i.e.
 number of elements, number of rules) but may vary in their distractors (two unique sets per template) or perceptual features (three unique shape sets per template; Figure \ref{fig:fig00}c/d).
@@
 The first is that, given the fixed-order task design,  there was relatively few responses collected for the majority of the items.
-Indeed only 42 of 80 item templates were completed by 100 or more participants.
+Indeed only 42 of 80 templates were completed by 100 or more participants.
 As such, one cannot be confident about the true difficulty of a majority of the items (an issue that is exacerbated when considering that each template has six clones).
@@
 As such, all participants received test forms of roughly equal difficulty.
-Importantly we also counterbalanced the assignment of item clones across participants, such that we had an approximately uniform number of responses available for each item by shape set (1, 2, or 3) and distractor type (minimal difference, MD, or paired difference, PD; see \citealt{chierchia2019matrix} for details).
+Importantly we also counterbalanced the assignment of item clones across participants, such that we had an approximately uniform number of responses available for each clone by shape set (1, 2, or 3) and distractor type (minimal difference, MD, or paired difference, PD; see \citealt{chierchia2019matrix} for details).
 
@@
 
-In our analyses we considered two item template (level 1) attributes: element number and rule number.
+In our analyses we considered two template (level 1) attributes: element number and rule number.
 These have previously been identified as key determinants of item difficulty in matrix reasoning tasks \citep{embretson1998cognitive, primi2001complexity}.
@@
 
-Together element number and rule number explained approximately 67.6\% of the variance in difficulty across the 64 item templates.
-That is, the majority of variability in difficulty across item templates can be attributed to their complexity.
+Together element number and rule number explained approximately 67.6\% of the variance in difficulty across the 64 templates.
+That is, the majority of variability in difficulty across templates can be attributed to their complexity.
 Individually the number of elements and number of rules respectively explain 29.5\% and 40.6\% of the variance in item template difficulty.
-Across all 384 item clones, these two attributes explain 38.6\% of the variance in difficulty.
+Across all 384 clones, these two attributes explain 38.6\% of the variance in difficulty.
 
@@
 There was not a credible association between the number of elements and discrimination ($\delta_{\alpha_1}$ = -0.015, 95\% HDI = -0.036 -- 0.008).
-These two attributes explained 31.9\% of the variance in discrimination across item templates.
+These two attributes explained 31.9\% of the variance in discrimination across templates.
 
@@
 Despite this, an item response model that included a clone- level residual variance term for item difficulty was preferred to one without.
-In sum, the results clearly indicate that item clones are not equally difficult and cannot be assumed to be exchangeable.
+In sum, the results clearly indicate that clones are not equally difficult and cannot be assumed to be exchangeable.
 
@@
 
-In conclusion, we have provided a comprehensive psychometric validation of the MaRs-IB, finding that it is suitable for measuring matrix reasoning ability in the general adult population.
+In conclusion we have provided the most comprehensive psychometric validation of the MaRs-IB, the current largest available bank of open access matrix reasoning items, finding that it is suitable for measuring matrix reasoning ability in the general adult population.
 We hope that the results and materials presented here will encourage researchers to design their own matrix reasoning tests, tailored to their needs, using the MaRs-IB.
-To facilitate this all tutorials for optimal test assembly using the estimated item parameters are available at: \url{https://github.com/ndawlab/mars-irt}.
+In support of this aspiration we have made all of the data, code, model outputs, and tutorials for optimal test assembly available at: \url{https://github.com/ndawlab/mars-irt}.
 

### 2022-05-26 manuscript/supplement.tex
camera ready

@@
 This result supports the hypothesis that participants that did reach items later in the test did so by prioritizing speed at the expense of accuracy.
-This suggests that the item summary statistics released as part of \cite{chierchia2019matrix} are likely biased indicators of item difficulty.
+This suggests that the summary statistics of later items released as part of \cite{chierchia2019matrix} are likely biased indicators of item difficulty.
 
@@
 
-We fit this model to data collected from a total of N=180 participants recruited from the Prolific Academic platform during piloting.
+We fit this model to data collected from a total of N=180 participants recruited from the Prolific Academic platform as part of a pilot experiment (independent of the experiments presented in the main text).
 Each participant completed one of two sets of eight items from the MaRs-IB.
@@
 \includegraphics[width=1.0\textwidth]{figures/figS02.png}
-\caption{\label{fig:figS02} An example item clone from each of the nine stimulus sets.
-In each row are three clones of one item template.
-Though some elementary shapes (e.g.
-circles, squares) are shared across stimulus sets, the majority of stimuli in each set are unique.}
+\caption{\label{fig:figS02} An example item clone from each of the nine stimulus sets.}
 \end{figure}

### 2022-11-17 manuscript/main.tex
manuscript updates

@@
 \usepackage[utf8x]{inputenc}
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{xurl}
-\usepackage[colorinlistoftodos]{todonotes}
+\usepackage{amsmath}            % equation formatting
+\usepackage{graphicx}           % figure formatting
+\usepackage{xurl, hyperref}     % link formatting
+\usepackage{xcolor,soul}        % highlight formatting
 
 % specify link style
-\usepackage{hyperref}
 \hypersetup{
@@
 }
+
+% define highlight style
+\colorlet{usercolorname}{yellow!00}    % use for public version
+% \colorlet{usercolorname}{yellow!30}    % use for reviewer version
+\sethlcolor{usercolorname}
 
@@
 
-% override apa6 first paragraph index
-% https://tex.stackexchange.com/questions/155028/indentation-apa6-class-for-2nd-paragraphs-after-title
-% \makeatletter
-%   \b@level@one@skip=-2.5ex plus -1ex minus -.2ex
-%   \b@level@two@skip=-2.5ex plus -1ex minus -.2ex
-% \makeatother
-
+% define title / short-title
 \title{An item response theory analysis of the Matrix Reasoning Item Bank (MaRs-IB)}
 \shorttitle{IRT analysis of MaRs-IB}
+
+% define authors
 \author{Samuel Zorowitz$^1$, Gabriele Chierchia$^3$, Sarah-Jayne Blakemore$^3$, Nathaniel D. Daw$^{1,2}$}
@@
 
-\abstract{Matrix reasoning tasks are among the most widely used measures of cognitive ability in the behavioral sciences, but the paucity of matrix reasoning tests in the public domain complicates their use.
+% define abstract
+\abstract{Matrix reasoning tasks are among the most widely used measures of cognitive ability in the behavioral sciences, but the lack of matrix reasoning tests in the public domain complicates their use.
 Here we present an extensive investigation and psychometric validation of the matrix reasoning item bank (MaRs-IB), an open-access set of matrix reasoning items.
@@
 In a second study, we demonstrate how researchers can use the item parameters we estimated to design new MaRs-IB tests using optimal item assembly.
-With these methods, we design three parallel short-form measures that exhibited good score reliability and convergent validity in a second sample of adults (N=300).
+With these methods, we design two new sets of test forms that exhibited good score reliability and convergent validity with an established measure of matrix reasoning in an independent sample of adults (N=600).
 We hope that the materials, results, and tutorials made available here will encourage researchers to use the MaRs-IB in their studies.}
 
+% start document
 \begin{document}
+
 \maketitle
@@
 Across cells, some or all of these shapes change according to a number of abstract rules.
-Based on these rules, participants must deduce which of four response options correctly completes the matrix.
+Based on these patterns, participants must deduce which of four response options correctly completes the matrix.
 The puzzles in the MaRs-IB vary in their complexity, both in the number of elements per cell and number of rules determining the relations of these elements across cells (Figure \ref{fig:fig00}a/b).
-To increase the reusability of the MaRs-IB, each puzzle (hereafter referred to as an item template) has six clones.
-A single item's clones share the same level of complexity (i.e.
-number of elements, number of rules) but may vary in their distractors (two unique sets per template) or perceptual features (three unique shape sets per template; Figure \ref{fig:fig00}c/d).
-Thus, the MaRs-IB addresses several of the issues associated with other matrix reasoning tests.
+Further increasing the reusability of the MaRs-IB, each puzzle (hereafter referred to as an item template) has six clones.
+An item's clones are equally complex (i.e.
+possess the same number of elements and rules) but may vary in their distractors (two unique sets per template) or perceptual features (three unique shape sets per template; Figure \ref{fig:fig00}c/d).
+Thus, the MaRs-IB addresses the copyright and limited reuse issues associated with other matrix reasoning tests.
 
@@
 Of note, items were presented in the same order to all participants.
-The authors found that performance on this task had good split-half and test-retest reliability.
+The authors found that measures of performance on this task had good split-half and test-retest reliability.
 Moreover, they found that participants' ability to solve MaRs-IB puzzles was moderately predictive of their performance on a working memory task and the ICAR matrix reasoning test, indicating satisfactory convergent validity.
 At the conclusion of this study, \cite{chierchia2019matrix} made publicly available a summary of the difficulty (i.e.
-proportion correct responses) of each item template and clone in the MaRs-IB so that other researchers could construct their own test measures of custom difficulty and duration.
+proportion correct responses) of every item template and clone in the MaRs-IB so that other researchers could construct their own test measures of custom difficulty and duration.
 
-Two important limitations in the experiment conducted by \cite{chierchia2019matrix} undermine the utility of these summary statistics, and therefore impede the design of new matrix reasoning tasks using the MaRs-IB.
-The first is that, given the fixed-order task design,  there was relatively few responses collected for the majority of the items.
-Indeed only 42 of 80 templates were completed by 100 or more participants.
-As such, one cannot be confident about the true difficulty of a majority of the items (an issue that is exacerbated when considering that each template has six clones).
+Two important limitations of the experiment conducted by \cite{chierchia2019matrix} undermine the utility of these summary statistics and, consequently, impede the design of new matrix reasoning tests using the MaRs-IB.
+The first is that, given the fixed-order experiment design, there were relatively few responses collected for the majority of the items.
+Indeed only 42 of 80 item templates were completed by 100 or more participants.
+As such, one cannot be confident about the true difficulty of a majority of the items (an issue that is exacerbated when considering that each item template has six clones).
 A second issue is participants were able to choose for themselves whether to prioritize accuracy (number of correct responses) or productivity (number of items reached).
@@
 IRT models confer several important advantages for the purposes of psychometric validation.
-First they provide an estimate of the \textit{difficulty} and \textit{discrimination} of each item, where the latter is an index of how much information an item provides about ability.
-Using these quantities, IRT models in turn allow experimenters to compute the test information function (TIF), which quantifies the reliability of a test in measuring performance given a level of ability.
+First they provide an estimate of the \textit{difficulty} and \textit{discrimination} of each item, where the latter is an index of the reliability of an item.
+Using these quantities, IRT models in turn allow experimenters to compute the test information function (TIF), which quantifies the reliability of a test for measuring performance given a level of ability.
 Finally, IRT models make possible optimal test assembly \citep{van1998optimal}, or the design of new tests with maximal reliability given researcher-specified constraints (e.g.
@@
 the number of elements and number of rules, determine item functioning.
-We modeled these two item attributes because they have been previously identified as among the strongest predictors of item difficulty in matrix reasoning tasks \citep{embretson1998cognitive, primi2001complexity}.
+We studied these two item attributes because they have been previously identified as among the strongest predictors of item difficulty in matrix reasoning tasks \citep{embretson1998cognitive, primi2001complexity}.
 This explanatory analysis provides another means by which to validate the design of the MaRs-IB; that is, to determine whether the MaRs-IB exhibits the same associations between item properties and functioning as found in other established matrix reasoning tests.
@@
 Establishing whether the MaRs-IB possesses this property is critical for the design of new matrix reasoning tasks using these items.
-If item clones are exchangeable, then new parallel test forms of equivalent difficulty and discrimination can be generated simply by substituting a an item clone from one shape set for another.
+If item clones are exchangeable, then new parallel test forms of equivalent difficulty and reliability can be generated simply by substituting an item clone from one shape set for another.
 To test for exchangeability, we use additive multilevel item structure (AMIS) models \citep{geerlings2011modeling, cho2014additive, lathrop2017item}.
@@
 The remainder of the paper proceeds as follows.
-First we report a calibration study in which a large sample of adult participants completed a subset of items from the MaRs-IB.
-Using their response data, we then fit a series of multilevel item structure models in order to estimate item parameters for each item (and their clones).
-Using these same models, we also interrogate the relationship between item attributes and functioning items in the MaRs-IB.
-Then we report a second validation study, in which a smaller sample of participants completed one of three new MaRs-IB short-form tests.
-These tests --- designed using optimal test assembly methods applied to the item parameters estimated during the calibration study --- were found to have good psychometric properties and convergent validity.
-This second study provides a blueprint for how to construct new MaRs-IB test measures using the item parameter estimates that were derived in the calibration study.
+First, we report a calibration study in which a large sample of adult participants completed a number of items from the MaRs-IB.
+Using their response data, we fitted a series of multilevel item structure models in order to estimate item parameters for each item (and their clones).
+Using these same models, we also interrogate the relationship between item attributes and item functioning.
+We conclude by reporting a second validation study, in which an independent sample of participants (N=600) completed one of five novel MaRs-IB test forms.
+These tests --- designed using optimal test assembly and the item parameter estimates from the first study --- were found to have good psychometric properties and convergent validity with an established measure of matrix reasoning ability.
+This second study provides a blueprint for how researchers can construct new MaRs-IB test measures based on the results of our first study.
 
@@
 
-The purpose of the first study was to calibrate the items in the MaRs-IB using response data collected from a large number of adult participants recruited from the general population.
+The purpose of the first study was to calibrate the items in the MaRs-IB using response data collected from a large number of adult participants from the general population.
 In particular, we sought to accomplish the following three aims: (1) to quantify the psychometric properties (i.e.
@@
 The design of the MaRs-IB items have been described previously \citep{chierchia2019matrix}.
-Briefly, each MaRs-IB item consists of a 3 x 3 matrix.
-Eight of the nine resulting cells contain abstract shapes, while one cell on the bottom right-hand side of the matrix is empty.
+Briefly, each MaRs-IB item consists of a 3x3 matrix.
+Eight of the nine cells contain abstract shapes, while one cell on the bottom right-hand side of the matrix is empty.
 Participants were instructed to ``complete the matrix'' by identifying the missing shape from among four possible alternatives.
@@
 
-In order to ensure sufficient sampling of every item template and clone in the MaRs-IB, participants were administered a pseudorandomly-selected set of 16 out of 64 total items.\footnote{We did not collect new data for the 16 easiest items in the MaRs-IB (items 1, 2, 3, 4, 5, 7, 8, 9, 32, 33, 38, 41, 43, 48, 57, 68) because performance on these items was at or near ceiling.} Item sets were constructed as follows: We subdivided the 64 items into 16 sets of four based on their dimensionality (a measure of item complexity defined in \citealt{chierchia2019matrix}).
+In order to ensure sufficient sampling of every item template and clone in the MaRs-IB, participants were administered a pseudorandomly-selected set of 16 out of 64 total items.\footnote{We did not collect new data for the 16 easiest items in the MaRs-IB (items 1, 2, 3, 4, 5, 7, 8, 9, 32, 33, 38, 41, 43, 48, 57, 68) because performance on these items was at or near ceiling.} Item sets were constructed as follows: We subdivided the 64 items into 16 sets of four based on their dimensionality (a measure of item complexity, defined in \citealt{chierchia2019matrix}).
 Participants were randomly assigned one item from each of the 16 sets.
 As such, all participants received test forms of roughly equal difficulty.
-Importantly we also counterbalanced the assignment of item clones across participants, such that we had an approximately uniform number of responses available for each clone by shape set (1, 2, or 3) and distractor type (minimal difference, MD, or paired difference, PD; see \citealt{chierchia2019matrix} for details).
+
+Importantly we also counterbalanced the assignment of item clones across participants, such that we had an approximately uniform number of responses available for each clone by shape set (1, 2, or 3) and distractor type (minimal difference, MD, or paired difference, PD).
+\hl{Distractor type refers to the two strategies used to generate the distractor response options in the MaRs-IB.
+The MD strategy produces distractors that are variations of the target response.
+The MD strategy has the advantage of preventing pop-out effects (where the correct response ``pops out'' among the possible responses), but has the disadvantage of theoretically allowing participants to solve an item by looking at the response options only.
+In contrast, the PD strategy produces distractors that are have at least one component in common with the target response.
+The PD strategy prevents participants from solving items by looking at the response options alone, but can potentially induce pop-out effects.
+We note that, for complex items with many elements, it is not possible to prevent both at the same time.}
 
@@
 In AMIS models, item parameters are defined according to a hierarchical structure in which item clones are nested in item templates.
-At the core of all of the AMIS models used here is the three parameter logistic (3PL) item response model, where the probability of a correct response ($y_{ijk} = 1$) for person $i$ on item clone $k$ belonging to item template $j$ is:
+The foundation of all the models used here is the three parameter logistic (3PL) item response model, where the probability of a correct response ($y_{ijk} = 1$) for person $i$ on item clone $k$ belonging to item template $j$ is:
 
@@
 Element number refers to the number of geometric shapes in each cell of a matrix, whereas rule number refers to the number of relationships that govern the changes among these shapes across cells.
-In the MaRs-IB there are four principal rules: size change, color change, position change, and shape change.
+In the MaRs-IB there are four principal rule types: size change, color change, position change, and shape change.
 Across item templates, element number ranges from one to four (median = 2) and rule number ranges from one to six (median = 3).
-Element and rule number are uncorrelated across item templates ($\rho$ = 0.176), and were determined for each item via manual annotation.
+Element and rule number are uncorrelated across item templates (Spearman's $\rho$ = 0.176), and were determined for each item via manual annotation.
 
@@
 The allocation of geometric shapes to item clones in the MaRs-IB is complex, such that not all clones of the same shape set share the same perceptual elements (see the supplementary materials for a full discussion).
-In other words, shape set does not constitute a meaningful nominal variable and was therefore not included in the model.
-Instead, the contribution of perceptual elements to item functioning is captured by residual variability terms of the item structure models.
+In other words, shape set does not constitute an interpretable categorical variable and was therefore not included in the model.
+Instead, the contribution of perceptual elements to item functioning is captured by the clone-level residual variability terms of the item structure models.
 
@@
     \item Model 1: Item difficulty is a function of item template attributes (level-1) and item clone attributes (level-2).
-In this model, item difficulty is predicted solely by the fixed effect contribution of item attributes (i.e.
+In this model, item difficulty is predicted solely by item attributes (i.e.
 element number, rule number, distractor type, response time).
@@
     \item Model 2: Item difficulty is a function of all item attributes (level-1/2) and level-1 residuals.
-The inclusion of the level-1 residual serves to quantify the magnitude of residual variability in item difficulty across item templates unexplained by the level-1 attributes (i.e.
+The inclusion of the level-1 residual serves to quantify the magnitude of residual variability in item difficulty, across item templates, unexplained by the level-1 attributes (i.e.
 element number, rule number).
@@
     \item Model 3: Item difficulty is a function of all item attributes (level-1/2) and level-1/2 residuals.
-The inclusion of the level-2 residual serves to quantify the magnitude of residual variability in item difficulty across item clones unexplained by the level-2 attributes (i.e.
+The inclusion of the level-2 residual serves to quantify the magnitude of residual variability in item difficulty, across item clones, unexplained by the level-2 attributes (i.e.
 distractor type, mean response time).
@@
 
-In the second wave of model fitting, item discrimination was now a free parameter to be estimated.
-Moreover, in the second set of models, item difficulty parameters were specified according to the best-fitting model from the first wave.
+In the second wave of model fitting, item discrimination was specified as a free parameter to be estimated.
+Moreover, in the second set of models, the structure of the item difficulty parameters were specified according to the best-fitting model from the first wave.
 In the second wave, we fit an additional three models:
@@
     \item Model 4: Item discrimination is a function of item template attributes (level-1) and item clone attributes (level-2).
-In this model, item discrimination is predicted solely by the fixed effect contribution of item attributes (i.e.
+In this model, item discrimination is predicted solely by item attributes (i.e.
 element number, rule number, distractor type, response time).
@@
     \item Model 5: Item discrimination is a function of all item attributes (level-1/2) and level-1 residuals.
-The inclusion of the level-1 residual serves to quantify the magnitude of residual variability in item discrimination across item templates unexplained by the level-1 attributes (i.e.
+The inclusion of the level-1 residual serves to quantify the magnitude of residual variability in item discrimination, across item templates, unexplained by the level-1 attributes (i.e.
 element number, rule number).
@@
     \item Model 6: Item discrimination is a function of all item attributes (level-1/2) and level-1/2 residuals.
-The inclusion of the level-2 residual serves to quantify the magnitude of residual variability in item discrimination across item clones unexplained by the level-2 attributes (i.e.
+The inclusion of the level-2 residual serves to quantify the magnitude of residual variability in item discrimination, across item clones, unexplained by the level-2 attributes (i.e.
 distractor type, mean response time).
@@
 We predicted that both attributes would be positively associated with item difficulty; that is, all else equal, items with either a greater number of elements and rules would be more difficult.
-In contrast, we had no hypotheses about the impact of these attributes on item discrimination.
+In contrast, we had no hypotheses regarding the association between these attributes on item discrimination.
 
 Finally, through both the level-2 fixed and random effects, we can determine whether item clones in the MaRs-IB are exchangeable (Aim 3).
-For example, a credible association between distractor type and either item parameter would entail that clones are not psychometrically equivalent.
-Moreover, if either level-2 residual variance term were non-negligible, that would similarly entail that not all item clones are equivalent due to unmodeled attributes (e.g.
-low-level perceptual features).
+For example, a credible association between distractor type and either item difficulty or discrimination would entail that clones are not psychometrically equivalent.
+Moreover, if we found any level-2 residual variance term were substantial, that would similarly indicate that not all item clones are equivalent due to unmodeled attributes (e.g.
+perceptual features).
 
@@
 
-All models were estimated within a Bayesian framework using Hamiltonian Monte Carlo as implemented in Stan (v2.22) \citep{carpenter2017stan}.
+All models were estimated within a Bayesian framework using Hamiltonian Monte Carlo as implemented in Stan (v2.22; \citealt{carpenter2017stan}).
 For all models, four separate chains with randomised start values each took 7500 samples from the posterior.
@@
 Specifically, the residual variance of latent ability was specified as $V(\epsilon) = 1 - \sum \rho_p^2$.
+The specification of the model priors are detailed in the supplement.
 
@@
 The goodness of fit of the models was compared using Bayesian leave-one out cross-validation \citep{vehtari2017practical}, which has been found to perform better than more traditional information criteria for comparing item response models \citep{luo2017performances}.
-We computed the conditional leave-one-cluster out (LOCO) cross-validation \citep{merkle2019bayesian}, which measures a model's ability to generalize to held out items (rather than held out responses or held out participants) --- i.e., for inferring generalization to additional items constructed using the same features rather than to additional participants sampled from the same population.
+We computed the conditional leave-one-cluster out (LOCO) cross-validation \citep{merkle2019bayesian}, which measures a model's ability to generalize to held-out items (rather than held-out responses or held-out participants) --- i.e., for inferring generalization to additional items constructed using the same features rather than to additional participants sampled from the same population.
 
@@
 
-The $\chi^2_{NC}$ discrepancy statistic is a measure of item fit based on the comparison of observed and expected proportions of participants at each score level (here we ignore scores of zero and one, which comprised 7 total participants).
+The $\chi^2_{NC}$ discrepancy statistic is a measure of item fit based on the comparison of observed and expected proportions of participants at each score level (we ignore scores $\leq 1$, achieved by only 7 participants).
 A global measure of fit at the test level is obtained by summing the discrepancy values over the groups.
 In turn, SGDDM measures the mean absolute conditional correlation between all pairs of items; that is, the SGDDM is an index of residual inter-item correlations unexplained by the model.
-The SGDDM is important in this context because it tests for local dependence across items, the presence of which might otherwise lower the efficiency of future test forms by including nonindependent or redundant items.
+The SGDDM is important in this context because it tests for local dependence across items, the presence of which might otherwise lower the efficiency of future test forms via the inclusion of nonindependent or redundant items.
 For both discrepancy statistics, the PPP value is the proportion of draws in which the posterior predictive discrepancy is equal to or higher than the realized discrepancy.
-A poor model fit to the data is indicated when the PPP-values are extreme (PPP $\leq$ 0.05 or PPP $\geq$ 0.95).
+A poor model fit to the data is indicated when the PPP-values are extreme (PPP $\leq$ 0.05).
+
+\hl{As a final validation of the best-fitting model, we also performed a parameter recovery analysis.
+In this analysis, we generated 100 artificial datasets with sampling and statistical properties matched to what we observed empirically.
+That is, we generated datasets with 1500 participants and 384 item clones (nested in 64 item templates), where the item parameters were randomly generated and matched to the observed distribution of item parameters.
+We then fit the best-fitting model to each artificial dataset, and quantified the consistency between the ground-truth and recovered model parameters.
+The complete details of this procedure can be found in the supplementary materials.}
 
@@
 
-On average, participants completed 9.6 of 16 items correctly (sd = 3.1, IQR = 8.0 -- 12.0) (Figure \ref{fig:fig01}a).
+On average, participants completed 9.6 of 16 items correctly (sd = 3.1, IQR = 8.0 -- 12.0; Figure \ref{fig:fig01}a).
 Timing out occurred on only 1.8\% of trials (these were subsequently coded as incorrect responses).
@@
 Across items, the proportion of correct responses was negatively correlated with rule number ($\rho$ = -0.482, p < 0.001).
-These results further corroborate those reported by \cite{chierchia2019matrix} in that the MaRs-IB exhibited good functioning with performance levels for only a few item clones at ceiling or floor.
+These results further corroborate those reported by \cite{chierchia2019matrix} in that most items in the MaRs-IB exhibited good functioning, with performance at ceiling or floor for only a few item clones.
 
@@
 
-There were also significant accuracy by difficulty ($\beta = 0.097$, $t = 14.34$, $p < 0.001$) and accuracy by rest score interactions ($\beta = -0.064$, $t = -8.76$, $p < 0.001$).
+There were also a significant accuracy by difficulty interaction ($\beta = 0.097$, $t = 14.34$, $p < 0.001$) and accuracy by rest score interaction ($\beta = -0.064$, $t = -8.76$, $p < 0.001$).
 Correct responses were even slower for more difficult items, but faster for better-performing participants.
@@
 
-In sum, we found evidence of a speed-accuracy trade-off in performance in this sample.
-Accurate responses and more accurate participants were both slower overall.
+In sum, we found evidence of a speed-accuracy trade-off in performance in the calibration sample.
+Both correct responses and better-performing participants were slower overall.
 Notably, better-performing participants demonstrated adjustments to their response times as function of item difficulty; whereas the lowest scoring participants maintained an equivalent work rate regardless of item difficulty, the highest scoring participants showed the largest slowing in responding as items became more challenging.
@@
 The results of the model comparison is summarized in Table \ref{table:1}.
-Of the models fitted in the first wave, which varied only in their specification of item difficulty, Model 3 exhibited the best fit to the data.
-Indeed, Model 3 demonstrated a considerable improvement in fit over Model 1 ($\Delta$ LOCO = 1155.8, se = 60.1) and Model 2 ($\Delta$ LOCO = 376.5, se = 34.5), indicating the presence of non-negligible residual variance in item difficulty across both item templates and clones.
+Of the first-wave models, which varied only in their specification of item difficulty, Model 3 exhibited the best fit to the data.
+Indeed, Model 3 demonstrated a considerable improvement in fit over Model 1 ($\Delta$ LOCO = 1155.8, se = 60.1) and Model 2 ($\Delta$ LOCO = 376.5, se = 34.5).
+This result indicates the presence of non-negligible residual variance in item difficulty across both item templates and clones.
 As the best-fitting model in the first wave, Model 3 served as the starting point in the second wave of model fitting.
 
-In the second wave of model fitting, in which models varied only in their specification of item discrimination, Model 5 demonstrated the best fit to the data.
+Of the second-wave models, which varied only in their specification of item discrimination, Model 5 demonstrated the best fit to the data.
 In contrast, Model 5 yielded smaller improvements in fit over Model 4 ($\Delta$ LOCO = 5.12, se = 4.07) and Model 6 ($\Delta$ LOCO = 3.29, se = 1.84).
@@
 As such we will select Model 5 as the best-fitting model overall, but proceed in discussing it with caution.
-We also note that all second-wave models exhibited better fits to the data than the first-wave models, supporting the inclusion of item discrimination parameters.
+We also note that all second-wave models exhibited better fits to the data than the first-wave models, supporting the estimation of item discrimination parameters.
 
@@
 There was considerable variability in item difficulty across items (sd = 1.431, 95\% HDI = 1.370 -- 1.492), with the smallest  and largest item difficulty parameters spanning a wide range ($\beta_{\min}$ = -3.704, 95\% HDI = -4.687 -- -2.775; $\beta_{\max}$ = 3.497, 95\% HDI = 2.481 -- 4.635).
-For an average ability participant ($\theta = 0$), these parameter estimates translate to an expected average accuracy of 58.4\% across all items, and ceiling (chance-level) performance for the least (most) difficult item.
+For an average ability participant ($\theta_i = 0$), these parameter estimates translate to an expected average accuracy of 58.4\% across all items, and ceiling (chance-level) performance for the least (most) difficult item.
 
@@
 These two attributes explained 31.9\% of the variance in discrimination across templates.
+Critically, these results must evaluated in light of our parameter recovery analyses (see \textit{Goodness-of-fit} below).
 
@@
 Items with MD distractors were associated with an estimated 19.9\% reduction in accuracy compared to their equivalent items with PD distractors.
-Together element number, rule number, and distractor type explained slightly over half (52.0\%) of the variance in item difficulty.
-Despite this, an item response model that included a clone- level residual variance term for item difficulty was preferred to one without.
-In sum, the results clearly indicate that clones are not equally difficult and cannot be assumed to be exchangeable.
+Together element number, rule number, and distractor type explained slightly over half (52.0\%) of the variance in difficulty across item clones.
+Regardless, the results of the model comparison demonstrated that an item response model which included a clone-level residual variance term for item difficulty was preferred to one without.
+Together, these two results clearly indicate that not all item clones are equally difficult and, therefore, item clones cannot be assumed to be exchangeable.
 
 To make clear the residual variance in item difficulty across clones, the estimated item difficulty per clone is presented in Figure \ref{fig:fig02}.
-The relative increase in difficulty for items with MD-type distractors compared to PD-type distractors is visually apparent (Figure \ref{fig:fig02}a).
+The relative increase in difficulty for items with MD-type distractors compared to PD-type distractors is easily seen (Figure \ref{fig:fig02}a).
 In contrast the pattern of residual difficulty across clones, after accounting for all modeled sources of variance, is more complicated (Figure \ref{fig:fig02}b).
@@
 This corresponds to a mean absolute difference in difficulty of 0.705 (95\% HDI = 0.612 -- 0.786) across clones per item and distractor type, or roughly two-thirds as large as the effect of distractor type.
-Thus, the residual variability in item difficulty across clones is large, complex, and likely reflects the idiosyncratic contributions of low-level perceptual features.
+Thus, the residual variability in item difficulty across clones is large, complex, and likely reflects the idiosyncratic contributions of perceptual features.
 
@@
 These results are consistent with a speed-accuracy trade-off in performance.
-That is, higher levels of ability in this sample reflect in part a tendency to slow down and respond more carefully on average (and especially for the most challenging items).
+That is, higher levels of ability in this sample reflect, in part, a tendency to slow down and respond more carefully (especially for the most challenging items).
 
@@
 
-Finally, we checked the fit of the best-fitting model using two posterior predictive model checking measures.
+We next evaluated the fit of the best-fitting model to the data using two posterior predictive model checking measures.
 The PPP-value for chi-square discrepancy statistic did not exceed the critical threshold ($\chi^2_{NC}$ = 15.904, PPP-value = 0.297), indicating that the model was sufficiently able to reproduce the distribution of observed scores.
@@
 
+\hl{Finally, we inspected the results of the parameter recovery analysis (the complete results are reported in the supplementary materials).
+Briefly, we found we were able to recover item difficulty parameters with excellent precision.
+Conversely, we observed only adequate recovery of the item discrimination parameters.
+The results may in turn explain why we detected only one credible association between item attributes and item discrimination.
+Although we were adequately-powered to detect large associations between item attributes and discrimination (i.e.
+explaining 14\% or more variance in the latter), we are poorly powered to detect smaller associations.
+This was not true for item difficulty, where we were well-powered to detect associations of the magnitude reported above.
+Importantly, we found through follow-up analyses that the less-than-perfect recovery of item discrimination parameters yielded only negligible effects on the reliability of test forms produced using optimal test assembly procedures.
+Therefore, our analyses were sufficiently powered for the primary objectives of this study.}
+
 \subsection{Discussion}
 
-In this first study, we recruited a large sample of adults to complete items from the MaRs-IB, making sure that each item template and clones were completed by a sizable number of participants.
+In this first study, we recruited a large sample of adults to complete items from the MaRs-IB, making sure that each item template and clone were completed by a sizable number of participants.
 Using multilevel item structure models, we then estimated the psychometric properties (i.e.
@@
 This finding is interesting in light of previous investigations of nonverbal reasoning tasks, which have found a greater influence on item difficulty from either the number of elements \citep{bethell1984adaptive} or number of rules \citep{mulholland1980components}.
-One possible reason for this discrepancy is that these two attributes are largely uncorrelated ($\rho$ = 0.176) across items in the MaRs-IB, which allows for unconfounded estimates of their effects.
+One possible reason for this discrepancy is that these two attributes are largely uncorrelated across items in the MaRs-IB, which allows for unconfounded estimates of their effects.
 Together element number and rule number explained 67.6\% of the variance in difficulty across templates (and 38.6\% of the variance across item clones), which is in line with previous investigations of matrix reasoning tasks \citep{carpenter1990one, matzen1994error}.
 These findings further validate the design of the MaRs-IB insofar that item complexity is a primary determinant of item difficulty.
+\hl{Finally, we found that rule number, but not element number, was associated with item discrimination.
+However, this finding must be interpreted with caution as our parameter recovery analysis revealed that our sample size was adequately powered to detect only larger effects (attributes explaining 14\% or more variance in item discrimination).
+Future studies with larger samples will be required to more thoroughly investigate the relationship between item complexity and discrimination.}
 
@@
 This discrepancy in findings between studies may reflect the much larger number of responses available per clone in our study.
-We found non-negligible residual variability in item difficulty across clones that was difficult to characterize and likely reflects the effects of low-level perceptual features.
-Regardless, our results clearly indicate that item clones in the MaRs-IB cannot be assumed to be psychometrically equivalent and should not be treated as exchangeable.
+We also found non-negligible residual variability in item difficulty across clones that was difficult to characterize and likely reflects the effects of low-level perceptual features.
+Together, our results clearly indicate that item clones in the MaRs-IB cannot be assumed to be psychometrically equivalent and should not be treated as exchangeable.
 
@@
 
-\section{Validation Study}
+% https://tex.stackexchange.com/a/240438
+\section{\texorpdfstring{\hl{Validation Study}}{}}
 
@@
 
-The purpose of the second study was to provide a demonstration of how to use the item parameter estimates derived in the previous study to design new MaRs-IB measures.
-Specifically, we use optimal test assembly methods to construct three parallel short form measures with maximal reliability under time length constraints.
+The purpose of the second study was to provide a demonstration of how to use the item parameter estimates, derived in the previous study, to design new MaRs-IB measures.
+Specifically, we use optimal test assembly \citep{van1998optimal} to design test forms that maximize test reliability given researcher-defined constraints.
 
@@
 
-Given a starting point of 64 item templates with six clones each, there is a vast number of possible MaRs-IB test forms we could construct.
-Returning to our motivating problem --- the availability of matrix reasoning measures with the potential for reuse --- we decided to construct three abbreviated parallel short form measures.
-To do so, we used mixed integer programming \citep{der2005wj} to maximize the test information function (TIF) of each short form subject to the following constraints:
+Given the starting point of 64 item templates with six clones each, there is a vast number of possible MaRs-IB test forms we could construct.
+Returning to our motivating problem --- the availability of matrix reasoning measures with the potential for reuse --- we decided to construct two new sets of test forms: a set of three short-form measures and a set of two long-form measures.
+
+To design the test measures, we used mixed integer programming \citep{der2005wj} in order to maximize the test information function (TIF) of each test subject to the following constraints:
 
@@
 
-    \item Each short form was required to contain 12 items.
-This was chosen to minimize the administration time of a given form (2-4 minutes on average) while achieving sufficient test score reliability (Cronbach's $\alpha \geq 0.8$).
+    \item Each short- and long-form test was required to contain 12 and 24 items, respectively.
+This was chosen to minimize the administration time of a given form while achieving a minimal score reliability $\geq 0.7$ and $\geq 0.8$.
 
-    \item All short forms were required to be made up of clones from the same items. This constraint was adopted in order to maximize the similarity of the short forms.
+    \item Within each test set, the difference in TIFs between tests should be minimized.
+This constraint was adopted to ensure that each test form had approximately equal psychometric properties (i.e.
+equivalent reliability across ability levels).
 
-    \item Clones selected for inclusion could either use MD-type or PD-type distractors, but not both.
+    \item For a given test, clones selected for inclusion could either use MD-type or PD-type distractors, but not both.
 This constraint was adopted in order to prevent item redundancy (i.e.
-including the same puzzle twice in one short form).
+including the same puzzle twice in one test form).
 
-    \item The difference in TIFs between short forms should be minimized.
-This constraint was adopted to ensure that each short form had approximately equal psychometric properties (i.e.
-equivalent measurement reliability across ability levels).
+    \item All three short-form tests (but not the long-form tests) were required to be made up of clones from the same items.
+This constraint was adopted in order to maximize the similarity of the short forms.
 
@@
 
-\noindent The TIF was maximized at five ability levels ($\theta = -1.0, -0.5, 0.0, 0.5, 1.0$), which previous simulation studies have shown to be generally sufficient \citep{der2005wj}.
+\noindent For both assembly procedures, the TIF was maximized at five ability levels ($\theta$ = -1.0, -0.5, 0.0, 0.5, 1.0), which previous simulation studies have shown to be generally sufficient \citep{der2005wj}.
 Solutions to the mixed integer programming problem were found using the \textit{mip} python package (v1.13.0; \citealt{santos2020mixed}).
@@
 The results of the test assembly are presented in Figure \ref{fig:fig03}.
-As expected, the MIP solver selected for items that were more discriminating on average (Figure \ref{fig:fig03}a).
-Though this was not an explicit constraint, the test characteristic curves (TCCs) of the three short forms are markedly similar (Figure \ref{fig:fig03}b).
-Based on the TCCs, the expected observed score for each measure is 7.7 out of 12 items.
-(Note that, because the lowest ability participants are expected to guess an item correctly 25\% of time, the lower asymptote of the TCCs is at a score of 3.) Finally, the TIFs across the three short forms are remarkably similar (Figure~\ref{fig:fig03}c), as is expected given the last assembly constraint.
-As such, we should expect each short form to have approximately equal precision in measuring matrix reasoning ability across the spectrum of ability.
-Indeed, the simulated marginal score reliability for each test form are similar (short form 1: $\alpha = 0.794$; short form 2: $\alpha = 0.795$; short form 3: $\alpha = 0.792$).
+As expected, the MIP solver selected for items that were more discriminating on average (Figure \ref{fig:fig03}, left column).
+Though this was not an explicit constraint, the test characteristic curves (TCCs) for each test form, within a set, were markedly similar (Figure \ref{fig:fig03}, middle column).
+Based on the TCCs, the expected average score was 7.7 out of 12 items for the short-form measures and 14.4 out of 24 items for the long-form measures.
+(Note that, because the lowest ability participants are expected to guess an item correctly 25\% of time, the lower asymptote of the TCCs is one quarter of the maximum possible score.) Finally, the TIFs for each test form within a set are hearly identical (Figure~\ref{fig:fig03}, right column), as would be expected given the last assembly constraint.
+We should therefore expect each test form within a set to have approximately equal reliability in measuring matrix reasoning ability.
+Indeed, the IRT test score reliability coefficients \citep{kim2010estimation, nicewander2018conditional} for each short form were similar (short form 1: $\rho = 0.706$; short form 2: $\rho = 0.709$; short form 3: $\rho = 0.705$), as were the coefficients for each long form (long form 1: $\rho = 0.821$; long form 2: $\rho = 0.821$).
 
-Next, we proceeded to administer each short form to a second sample of participants.
-Our aim was to verify that observed scores on each short form conformed to our expectations based on the TCCs and TIFs.
+Next, we proceeded to administer each test form to a second sample of participants.
+Our aim was to verify that observed scores on each test form conformed to our expectations based on the TCCs and TIFs.
 Doing so would provide evidence that we were successful in calibrating the functioning of the items in the MaRs-IB in the previous study.
-We also sought to demonstrate convergent validity by correlating performance on the MaRs-IB short forms with performance in the same participants on an abbreviated version of the Raven's progressive matrices \citep{bilker2012development}.
+We also sought to demonstrate convergent validity by correlating participants' total scores on the MaRs-IB test forms with their total scores on an abbreviated version of the Raven's progressive matrices \citep{bilker2012development}.
 
@@
 
-A total of N=347 participants were recruited from the Prolific Academic platform to participate in an online behavioral experiment in November, 2021.
+Two samples of participants ($N_1=347$, $N_2=360$) were recruited from the Prolific Academic platform to participate in an online behavioral experiment in November, 2021 and September, 2022.
 Participants were eligible if they currently resided in the United States and had not participated in the calibration study.
-Study duration was on average 10.8 minutes (sd = 4.6).
-Participants received monetary compensation for their time (rate USD \$10/hr) plus a bonus up to \$0.75 based on task performance.
-Participants earned on average \$2.10 USD (sd = \$0.18).
+Study duration was on average 10.8 minutes (sd = 4.6) for sample 1 and 14.3 minutes (sd = 4.9) for sample 2.
+Participants received monetary compensation for their time (rate USD \$10/hr) plus a bonus up to \$0.75 (sample 1) or \$1.50 (sample 2) based on task performance.
+Sample 1 earned on average \$2.10 USD (sd = \$0.18) and sample 2 earned on average \$3.17 USD (sd = \$0.30).
 This study was approved by the Institutional Review Board of Princeton University (\#7392), and all participants provided informed consent.
 
-To ensure data quality, the data from multiple participants were excluded prior to analysis (see Exclusion Criteria below) leaving the data from N=300 participants for analysis.
-The majority of participants identified as women (men: N=138; women: N=147; non-binary or other: N=13; rather not say: N=2).
-Participants were on average 35.0 years old (sd = 13.2; range: 18 -- 76).
-The sample was relatively well-educated, with the majority having completed a Bachelor's degree (N=121) or Master's degree or higher (N=45).
-By comparison, fewer participants endorsed having completed only some college (N=100), only a high school degree (N=33), or preferred not to say (N=1).
+To ensure data quality, the data from multiple participants were excluded prior to analysis (see Exclusion Criteria below) leaving the data from a total of 600 participants ($N_1=300$, $N_2=300$) for analysis.
+Across samples, the majority of participants identified as women (men: N=281; women: N=281; non-binary or other: N=25; rather not say: N=2).
+Participants were on average 36.6 years old (sd = 13.1; range: 18 -- 76).
+The combined sample was relatively well-educated, with the majority having completed a Bachelor's degree (N=227) or Master's degree or higher (N=76).
+By comparison, fewer participants endorsed having completed only some college (N=142), only a high school degree or lower (N=98), or preferred not to say (N=1).
 
@@
 These measures were included as part of exploratory analyses to measure the associations between personality and matrix reasoning ability (these correlations are reported in Table S2 of the supplementary materials).
-Afterwards participants completed one of the three 12-item MaRs-IB short-form measures.
-The administration procedure of the short-forms was identical to that in the previous study.
-Finally, participants completed the 9-item abbreviated Raven's progressive matrices (RPM form A; \citealt{bilker2012development}).
+Afterwards participants completed either one of the three 12-item MaRs-IB short forms (sample 1) or one of the two 24-item MaRs-IB long forms (sample 2).
+The administration procedure of the test forms was identical to that in the previous study.
+Finally, all participants completed the 9-item abbreviated Raven's progressive matrices (RPM form A; \citealt{bilker2012development}).
 To maximize consistency in administration, the presentation format and instructions for the RPM task were the same as for the MaRs-IB test forms.
@@
 
-To ensure data quality, the data from multiple participants were excluded prior to analysis for one or more of the following reasons: failing to complete all three sections of the experiment (N=17); failing one or more attention checks \citep{zorowitz2021inattentive} embedded in the self-report measures (N=21); experiencing technical difficulties during the experiment (N=11); rapid guessing on four or more items (N=3); or failing to respond on four or more items (N=1).
-In total, 47 of 347 (13.5\%) participants were excluded leaving the data from N=300 participants for analysis.
+To ensure data quality, the data from multiple participants were excluded prior to analysis for one or more of the following reasons: failing to complete all three sections of the experiment (N=27); failing one or more attention checks \citep{zorowitz2021inattentive} embedded in the self-report measures (N=49); experiencing technical difficulties during the experiment (N=24); rapid guessing on four or more items (N=24); or failing to respond on four or more items (N=2).
+In total, 107 of 707 (15.1\%) participants were excluded leaving the data from N=600 participants for analysis.
 
@@
 
-To measure the convergent validity between the MaRs-IB and RPM measures, we fit an additional explanatory item response model.
-Here participants' ability on the MaRs-IB short forms were estimated as a function of their gender, age, and observed score on the RPM (Eq \ref{eq:4}).
-The probability of a participant correctly solving a MaRs-IB was defined according to the 3PL model (Eq \ref{eq:1}), where the difficulty and discrimination parameters for an item were fixed to the posterior mean of the estimates obtained in the previous study.
-Thus, the model estimates the partial correlation between RPM performance and MaRs-IB ability controlling for the age and gender of the participants, and the differences in item functioning across test forms.
+We calculated descriptive statistics to summarize the distribution of participants' total scores on the MaRs-IB short forms, MaRs-IB long forms, and abbreviated RPM.
+The convergent validity between the MaRs-IB and RPM measures was estimated by calculating the Pearson correlation between participants' total scores on those measures (collapsing across MaRs-IB test form).
+These calculations were performed separately for the MaRs-IB short-form and long-form measures.
 
-The IRT model was estimated within a Bayesian framework using Hamiltonian Monte Carlo as implemented in Stan (v2.22) \citep{carpenter2017stan}.
-Four separate chains with randomised start values each took 7500 samples from the posterior.
+To quantify the agreement between the observed and expected MaRs-IB scores, we relied on posterior predictive model checking.
+Specifically, we calculated a $\chi^2_{NC}$ discrepancy measure \citep{sinharay2006posterior}, which compares the observed and model-predicted proportion of participants at each total score level.
+The model-predicted scores were obtained via a two-step procedure.
+First, we estimated for all participants a posterior distribution of latent ability parameters under the 3PL model, given their observed item responses and the item parameters estimated during the calibration study.
+Second, using the distribution of ability estimates, we generated a sample of predicted responses.
+We computed a posterior predictive $p$ (PPP) value based on the discrepancy statistic, where a poor fit to the data is indicated when the PPP-value is extreme (PPP $\leq$ 0.05).
+This procedure was also repeated for the abbreviated RPM measure, but the item difficulty and discrimination parameters were jointly estimated alongside the person ability parameters under the 3PL model.
+The item guessing parameters were held fixed to their nominal guessing rates.
+
+The latent ability and RPM item parameters were estimated within a Bayesian framework using Hamiltonian Monte Carlo as implemented in Stan (v2.22) \citep{carpenter2017stan}.
+Four separate chains with randomised start values each took 7,500 samples from the posterior.
 The first 5,000 samples from each chain were discarded.
@@
 The $\hat{R}$ values for all parameters were less than 1.01, indicating acceptable convergence between chains, and there were no divergent transitions in any chain.
-
-We calculated the ordinal reliability of the MaRs-IB and RPM short-form measures using the \textit{lavaan} \citep{lavaan} and \textit{semTools} \citep{semtools} packages available in R.
+The prior for the latent ability parameters was Normal(0,1).
+For the RPM response model, the prior on the item difficulty and discrimination parameters were Normal(0.0, 2.5) and Lognormal(1.0, 1.0), respectively.
 
@@
 
-Performance on the MaRs-IB short-forms is summarized in Table \ref{table:3}.
-On average, participants completed the measure in 174 s (sd = 54.0 s) out of a possible 360 s total and correctly solved 8.0 of 12 items (sd = 2.5).
-The distribution of observed scores was right-shifted (Figure~\ref{fig:fig04}), as was expected given the nominal guessing rate.
-A one-way ANOVA comparing the average score across the three short-forms was not statistically significant (F(2,297) = 0.253, p = 0.777).
-Importantly, the marginal score reliability of each measure was satisfactory (all $\alpha > 0.75$).
-In sum, performance on each MaRs-IB test measure conformed to expectations, indicating that item functioning was adequately calibrated in the previous study.
+Performance on the MaRs-IB test forms is summarized in Table \ref{table:3}.
+On average, participants completed the short-form measures in 174 s (sd = 54.0 s) out of a possible 360 s total and correctly solved 8.0 of 12 items (sd = 2.5).
+In turn, participants completed the long-form measures in 349 s (sd = 97.9 s) out of a possible 720 s total and correctly solved 15.7 of 24 items (sd = 5.1).
+A one-way ANOVA comparing the total scores across the three short forms was not statistically significant (F(2,297) = 0.253, p = 0.777); so too, the distributions of total across the two long forms were not significantly different (F(1,298) = 0.219, p = 0.640).
+Crucially, none of the posterior predictive p-values corresponding to the $\chi^2_{NC}$ discrepancy measure exceeded the critical value, indicating that the previously estimated item parameters were sufficiently able to predict the observed distribution of total scores for each test form.
+In other words, performance on each new MaRs-IB test form conformed to expectations, indicating that item functioning was well calibrated in the previous study.
 
 Performance on the abbreviated RPM measure is also summarized in Table \ref{table:3}.
-On average, participants completed the measure in 125 s (sd = 40.1 s) and responded correctly on 4.5 of 9 items (sd = 2.0).
-The proportion of correct responses on on the RPM was significantly lower than that observed for the MaRs-IB ($t = 13.028$, $p < 0.001$), suggesting that the RPM is more difficult or less biased by guessing.\footnote{In contrast to the MaRs-IB, items in the RPM have either six or eight response options.
-As such, the nominal guessing rate is expected to be lower for the RPM.} The marginal score reliability for the RPM was also adequate.
+On average, participants completed the measure in 127 s (sd = 39.6 s) and responded correctly on 4.5 of 9 items (sd = 2.0).
+The proportion of correct responses on on the RPM was significantly lower than that observed for the MaRs-IB short forms ($t = 13.028$, $p < 0.001$) and long forms ($t = 12.812$,  $p < 0.001$), indicating that the RPM is more difficult.
+Notably, the calculated IRT test reliability for the abbreviated RPM measure was smaller than that for both the MaRs-IB short- and long-form measures.
+The average correlation between scores on the RPM and MaRs-IB short-form measures was $r = 0.470$ ($p < 0.001$); between the RPM and long-form measures, the correlation was $r = 0.598$ ($p < 0.001$).
+These results indicate satisfactory convergent validity between the two families of matrix reasoning measures.
 
-The partial correlation between observed scores on the RPM and ability on the MaRs-IB was $\rho = 0.527$ (95\% HDI = 0.457 - 0.577).
-Considering the attenuation of correlations between two imperfect measures \citep{spearman1961proof}, we should not expect correlations much larger than $\rho \approx 0.64$ (or $0.80 \times 0.80$).
-Therefore, this result indicates satisfactory convergent validity between the two matrix reasoning measures.
-Briefly, in this second sample, neither age nor gender was credibly associated with performance on the MaRs-IB short forms (age: $\rho$ = -0.123, 95\% HDI = -0.253 -- 0.013; gender: $\rho$ = -0.025, 95\% HDI = -0.160 -- 0.108).
-
-To summarize, we used the estimated item parameters in conjunction with optimal assembly methods to construct three parallel MaRs-IB short forms.
-As expected, these new measures were quick to administer, exhibited score distributions in line with model predictions, and demonstrated good internal reliability and convergent validity with a second matrix reasoning measure.
-Together, these results highlight our success in accurately calibrating the functioning of MaRs-IB items, and consequently the feasibility of constructing new matrix reasoning measures.
+To summarize, we used the estimated item parameters and optimal test assembly procedures to construct two new sets of MaRs-IB test forms.
+These new measures were quick to administer, produced score distributions consistent with model predictions, possess adequate-to-good test score reliability, and demonstrated convergent validity with an established measure of matrix reasoning ability.
+Furthermore, all of these new test forms were more reliable than than an abbreviated version of the Raven's progressive matrices.
+Together, these results demonstrate our success in calibrating the functioning of MaRs-IB items and also the feasibility of constructing new reliable tests of matrix reasoning with the MaRs-IB.
 
@@
 
-Here we have provided the most comprehensive investigation of the MaRs-IB to date.
+Here we have provided a comprehensive investigation of the MaRs-IB.
 Using item structure models fit to data collected from a large sample of adult participants, we found that the MaRs-IB possesses many desirable psychometric properties.
 The items in the MaRs-IB span a considerable range of difficulty and exhibit medium-to-high levels of discrimination.
-The MaRs-IB is therefore suitable for measuring matrix reasoning across the ability spectrum in the adults.
-We also verified that the design of the MaRs-IB was in part successful.
+The MaRs-IB is therefore suitable for measuring matrix reasoning across the ability spectrum in general adult population.
+We also verified that the design of the MaRs-IB was, at least in part, successful.
 Item complexity (i.e.
-element number and rule number) was positively associated with, and explained more than half of the variability in, item difficulty.
+element number and rule number) was positively associated with, and explained the majority of variance of, item difficulty.
 In contrast, we also uncovered some undesirable properties of the MaRs-IB.
-Namely, item clones exhibited large differences in difficulty by distractor type, which persisted even after controlling for this and other item attributes.
+Namely, we observed substantial variability in item difficulty (both systematic and unsystematic) across item clones.
 As a consequence, item clones in the MaRs-IB cannot be assumed to be psychometrically equivalent and should not be treated as exchangeable.
 
-With the estimated item parameters, we then constructed three new MaRs-IB short-form measures using optimal test assembly methods.
-These abbreviated test measures were designed to be maximally reliable under time constraints (i.e.
-administration time of 2-4 minutes).
-In a second sample of participants, we found that each short form measure had adequate reliability (Cronbach $\alpha \approx 0.8$) and good convergent validity with an abbreviated Raven's progressive matrices measure.
-Moreover, the observed scores on the short-forms followed the expected distribution as predicted by their test characteristic curves.
-These results demonstrate the success of the item calibration study in producing accurate estimates of the item parameters.
+Using the item parameters estimates, we then constructed two new sets of MaRs-IB test forms using optimal test assembly methods.
+These test measures were designed to be maximally reliable under differing constraints (e.g.
+administration times of 2 -- 4 minutes vs.
+4 -- 8 minutes).
+In a second sample of participants, we found that the total scores from each test form was consistent with model predictions and predictive of performance on a second, established measure of matrix reasoning ability.
+Collectively, these results highlight the success of the item calibration study in producing accurate estimates of item functioning for the items in the MaRs-IB.
 
 Of additional note are the associations we found between accuracy and response time.
-Accurate responses and more accurate participants were slower on average, consistent with a speed-accuracy trade-off account of performance \citep{heitz2014speed}.
-Interestingly, we also found an interaction between participant ability and item difficulty: whereas the worst-performing participants maintained approximately the same work rate irrespective of item difficulty, the best-performing participants were even slower as items became more challenging.
+We found that correct responses and better-performing participants were slower on average, consistent with a speed-accuracy trade-offs in performance \citep{heitz2014speed}.
+Interestingly, we also found an interaction between participant ability and item difficulty: whereas the worst-performing participants maintained the same work rate irrespective of item difficulty, the best-performing participants spent more time deliberating as items became more challenging.
 These results are consistent with a persistence interpretation of ability, wherein good task performance in part reflects a willingness to invest time in the solution process and poor task performance reflects a tendency to give up sooner \citep{ranger2021effects}.
-As such, the MaRs-IB may be suitable not only for measuring matrix reasoning ability but also mental effort costs \citep{kool2018mental}, opportunity costs \citep{payne1996time}, or other motivational factors \citep{duckworth2011role} related to the tendency to exert effort or give up.
-The use of more sophisticated models may help to disentangle the relative contributions of latent ability and persistence to performance on the MaRs-IB \citep{ranger2014accumulator}.
+As such, the MaRs-IB may be suitable not only for measuring matrix reasoning ability but also mental effort costs \citep{kool2018mental}, opportunity costs \citep{payne1996time}, or other motivational factors \citep{duckworth2011role} related to people's tendency to exert effort or give up.
+The use of more sophisticated models (e.g.
+\citealt{ranger2014accumulator}) may help to disentangle the relative contributions of latent ability and persistence to performance on the MaRs-IB.
 
@@
 Here we analyzed response data collected from an online adult sample which was relatively young and well-educated.
-We cannot guarantee that the psychometric properties of the MaRs-IB found here will generalize to other populations or testing contexts.
-Future researchers may consider replicating the current study in other populations of interest (e.g.
+We cannot guarantee that the psychometric properties of the MaRs-IB reported here will generalize to other populations or testing contexts.
+Future researchers should consider replicating the current study in other populations of interest (e.g.
 children, clinical samples).
-By using item response models in the present analyses, however, we make possible the opportunity for future IRT ``linking'' studies.
+By using item response models in this study, however, we make possible the opportunity for future IRT ``linking'' studies.
 IRT linking describes a set of methods to establish the comparability of item and ability parameters estimated from response data collected from two or more groups that differ in ability \citep{lee2018irt}.
 Future studies might exploit these methods to provide new insights, not only in how the functioning of the MaRs-IB may differ in across populations, but also in how matrix reasoning ability changes across populations.
+\hl{Future studies involving larger samples could also provide populations norms for MaRs-IB test scores, which would enable researchers to compare the outcomes of specific individuals or groups (e.g.
+clinical groups) against the performance of the general population (of a particular country or region).}
 
@@
 So too, it is possible that the four rules in the MaRs-IB are not equally difficult or discriminating (e.g.
-color changes are easier to process than position changes) and may explain part of the residual variance in item functioning across item templates.
+color changes may be easier to process than position changes) and may explain part of the residual variance in item functioning across item templates.
 Other studies of matrix reasoning tasks have identified large contributions of perceptual features to item difficulty \citep{primi2001complexity, primi2014developing}, effects which were not explicitly considered here.
@@
 We hope that the results and materials presented here will encourage researchers to design their own matrix reasoning tests, tailored to their needs, using the MaRs-IB.
-In support of this aspiration we have made all of the data, code, model outputs, and tutorials for optimal test assembly available at: \url{https://github.com/ndawlab/mars-irt}.
+In support of this, we have made all of our data, code, and model outputs available at: \url{https://github.com/ndawlab/mars-irt}.
 
@@
 \begin{table}
+    \small
     \centering
-    \begin{tabular*}{\textwidth}{llcccc}
+    \begin{tabular*}{1.07\textwidth}{lccrccrc}
     \toprule
-    Measure & Sample & Task Time (sd) & Mean Score (sd) & IQR & Reliability ($\alpha$) \\
+    Measure & \# & Sample & \multicolumn{1}{c}{Task time (sd)} & \multicolumn{1}{c}{Mean score} (sd) & IQR & \multicolumn{1}{c}{$\chi^2_{NC}$ (ppp)} & Reliability \\
     \midrule
-    MaRs-IB SF-1 & N=103 & 174.9 s (53.3) & 7.9 (2.4) & 7 - 10 & 0.761 \\
-    MaRs-IB SF-2 & N=98  & 169.0 s (53.0) & 8.2 (2.6) & 7 - 10 & 0.810 \\
-    MaRs-IB SF-3 & N=99  & 178.5 s (55.7) & 7.9 (2.7) & 6 - 10 & 0.829 \\
-    RPM SF-A     & N=300 & 124.6 s (40.1) & 4.5 (2.0) & 3 - 9 & 0.791 \\
+    MaRs-IB SF & 1 & N=103 & 174.9 s  (53.3)  &  7.9 (2.4) &  7 - 10 & 13.6 (0.284) & 0.706 \\
+               & 2 & N=98  & 169.0 s  (53.0)  &  8.2 (2.6) &  7 - 10 & 6.6 (0.811)  & 0.709 \\
+               & 3 & N=99  & 178.5 s  (55.7)  &  7.9 (2.7) &  6 - 10 & 7.1 (0.778)  & 0.705 \\
+    \midrule
+    MaRs-IB LF & 1 & N=153  & 351.7 s (92.6)  & 15.9 (4.9) & 13 - 20 & 12.5 (0.910) & 0.821 \\
+               & 2 & N=147  & 346.8 s (103.4) & 15.6 (5.4) & 11 - 20 & 29.7 (0.120) & 0.821 \\
+    \midrule
+    RPM SF     & A & N=600  & 126.7 s (39.6)  &  4.5 (2.0) &  3 - 6  & 10.3 (0.359) & 0.613 \\
     \bottomrule
     \end{tabular*}
-    \caption{\label{table:3}\normalfont Summary of performance on the MaRs-IB and Raven's progressive matrices (RPM) short form (SF) measures. IQR = interquartile range.}
+    \caption{\label{table:3}\normalfont Summary of performance on the MaRs-IB 12-item short forms (SF), MaRs-IB 24-item long forms (LF), and Raven's progressive matrices (RPM) 9-item short form (SF) measures.
+Reliability reflects the IRT test score reliability coefficient.
+Abbreviations: IQR = interquartile range.}
 \end{table}
@@
 \includegraphics[width=1.0\textwidth]{figures/fig04.png}
-\caption{\label{fig:fig04}The joint distribution of total scores on the MaRs-IB short form and abbreviated Raven's progressive matrices (RPM) measures.
-The distribution of observed scores for each measure are shown in the margins.}
+\caption{\label{fig:fig04}The joint distribution of total scores on the abbreviated Raven's progressive matrices (RPM) and the MaRs-IB test forms.
+(A) Scores from the MaRs-IB 12-item short forms (SF).
+(B) Scores from the MaRs-IB 24-item long forms (LF).
+The distribution of observed scores for each measure are plotted in the margins.}
 \end{figure}

### 2022-11-17 manuscript/supplement.tex
manuscript updates

@@
 \usepackage[utf8x]{inputenc}
-\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage[colorinlistoftodos]{todonotes}
+\usepackage{amsmath}            % equation formatting
+\usepackage{graphicx}           % figure formatting
+\usepackage{xurl, hyperref}     % link formatting
+\usepackage{xcolor,soul}        % highlight formatting
 
-% override apa6 first paragraph index
-% https://tex.stackexchange.com/questions/155028/indentation-apa6-class-for-2nd-paragraphs-after-title
-% \makeatletter
-%   \b@level@one@skip=-2.5ex plus -1ex minus -.2ex
-%   \b@level@two@skip=-2.5ex plus -1ex minus -.2ex
-% \makeatother
+% specify link style
+\hypersetup{
+    colorlinks=true,
+    linkcolor=blue,
+    filecolor=magenta,      
+    urlcolor=cyan,
+    citecolor=blue
+}
 
+% define highlight style
+\colorlet{usercolorname}{yellow!00}    % use for public version
+% \colorlet{usercolorname}{yellow!40}    % use for reviewer version
+\sethlcolor{usercolorname}
+
+% define title / short-title
 \title{Supplemental materials for ``An item response theory analysis of the Matrix Reasoning Item Bank (MaRs-IB)''}
 \shorttitle{Supplemental materials for ``IRT analysis of MaRs-IB''}
+
+% define authors
 \author{Samuel Zorowitz$^1$, Gabriele Chierchia$^3$, Sarah-Jayne Blakemore$^3$, Nathaniel D. Daw$^{1,2}$}
@@
 
+% start document
 \begin{document}
+
 \maketitle
@@
 
+\section{\texorpdfstring{\hl{Bayesian additive multilevel item structure (AMIS) model priors}}{}}
+
+In the additive multilevel item structure (AMIS) models, there are three families of parameters: item difficulty parameters, item discrimination parameters, and person ability parameters.
+We describe the priors for each set of parameters in turn.
+The model code is publicly available at \url{https://github.com/ndawlab/mars-irt}.
+
+The item difficulty parameters can be decomposed into their fixed and random effects components.
+The fixed effects components include the grand mean, $\mu_\beta$, the effects of template-level (level-1) attributes, $\delta_{\beta_n}$, and the effects of clone-level (level-2) attributes, $\delta_{\beta_m}$.
+All three fixed effects terms were assigned standard normal priors, $\mathcal{N}(0,1)$.
+The template- and clone-level random effects components, $\epsilon_{\beta j}$ and $\epsilon_{\beta k}$, were each assigned normally distributed priors with a mean of zero and standard deviations estimated by the model, $\sigma_{\beta j}$ and $\sigma_{\beta k}$.
+These standard deviations were, in turn, assigned half-student-t distributions with a mean of zero, a standard deviation of one, and three degrees of freedom, $\text{StudentT}(3, 0, 1)$.
+
+Similarly, the item discrimination parameters can be decomposed into their fixed and random effects components.
+The fixed effects components include the grand mean, $\mu_\alpha$, the effects of template-level (level-1) attributes, $\delta_{\alpha_n}$, and the effects of clone-level (level-2) attributes, $\delta_{\alpha_m}$.
+All three fixed effects terms were assigned standard normal priors, $\mathcal{N}(0,1)$.
+The template- and clone-level random effects components, $\epsilon_{\alpha j}$ and $\epsilon_{\alpha k}$, were each assigned normally distributed priors with a mean of zero and standard deviations estimated by the model, $\sigma_{\alpha j}$ and $\sigma_{\alpha k}$.
+These standard deviations were, in turn, assigned half-student-t distributions with a mean of zero, a standard deviation of one, and three degrees of freedom, $\text{StudentT}(3, 0, 1)$.
+Importantly, item discrimination parameters were restricted to be in the range $\alpha_{jk} \in [0, 5]$.
+This was achieved by passing the clone-level discrimination parameter through the standard normal cumulative distribution function (i.e.
+scaling parameters between 0 and 1), and multiplying the result by 5.
+
+Finally, the person ability parameters can be decomposed into their fixed and random effects components.
+The fixed effects components comprise the partial correlation coefficients, $\rho_p$.
+Before transformation, the partial correlation coefficients were assigned standard normal priors, $\mathcal{N}(0,1)$.
+These values were then transformed such that the sum of their squared values could not exceed 1.
+The random effects components, $\epsilon_\theta$, were assigned a normally distributed prior with a mean of zero and a variance equal to 1 - minus the sum of squared partial correlations (i.e.
+fixing the variance of the person ability distribution to 1).
+
+\section{\texorpdfstring{\hl{Parameter recovery \& power analysis}}{}}
+
+In order to assess the robustness of our item calibration analysis, we performed a parameter recovery and power analysis.
+The objectives of this analysis were threefold.
+First, we wanted to evaluate our ability to accurately estimate person and item parameters given our sample size and experiment design.
+Second, we wanted to quantify our power to detect an association between item attributes and item functioning (i.e.
+difficulty, discrimination) for different magnitudes of association.
+Third, we wanted to determine how imprecision in our estimates of item parameters impact optimal test assembly and test score reliability.
+
+We generated 100 artificial datasets whose statistical properties were matched to what we observed empirically in the item calibration dataset.
+Specifically, each dataset was composed of simulated item responses  from 1500 ``participants'' whose latent abilities were drawn from a standard normal distribution.
+Each participant ``completed'' 16 item clones from a pool of 384 item clones (nested among 64 item templates).
+Thus, each artificial item clone was completed by an average of 62.5 participants.
+The simulated distribution of item difficulty and discrimination parameters were matched to what was observed empirically ($\mu_\beta = 0.20$, $\sigma_\beta = 1.55$; $\mu_\alpha = 1.30$, $\sigma_\alpha = 0.22$).
+The total variance explained by item attributes for item difficulty and discrimination parameters was also matched to what was observed empirically ($r^2_\beta \approx 60\%$; $r^2_\alpha \approx 40\%$).
+To estimate our statistical power to detect associations between item attributes and parameters, we specified two large effects and two small effects per item parameter.
+That is, there were two pairs of attributes that explained 20\% and 10\% of the variance in item difficulty, respectively; and there were two pairs of attributes that explained 14\% and 7\% of the variance in item discrimination, respectively.
+Based on what we observed empirically, the residual variance in item difficulty and discrimination parameters was split equally between the template- ($\epsilon_j$) and clone-levels ($\epsilon_k$).
+
+Models 5 and 6 were fitted to each simulated dataset.
+We tested Model 5 because this was the model deemed the best-fitting, parsimonious model in the item calibration study.
+We tested Model 6 because it is true data generating model for the simulations.
+Comparing these two models therefore allows us to evaluate whether selecting Model 5 produces any systematic bias in our analyses.
+All models were estimated within a Bayesian framework using Hamiltonian Monte Carlo as implemented in Stan (v2.22).
+For all models, four separate chains with randomised start values each took 4,000 samples from the posterior.
+The first 3,000 samples from each chain were discarded.
+As such, 4,000 post-warmup samples from the joint posterior were retained.
+The $\hat{R}$ values for all parameters were less than 1.02, indicating acceptable convergence between chains, and there were no divergent transitions in any chain.
+
+The results of the parameter recovery study are summarized Figure \ref{fig:figS04}.
+For both models, we observed good recovery of the latent ability parameter (Model 5: $r = 0.873$; Model 6: $r = 0.873$).
+Furthermore, we observed excellent recovery for the item difficulty parameters (Model 5: $r = 0.962$; Model 6: $r = 0.962$).
+In contrast, we observed only adequate recovery of the item discrimination parameters (Model 5: $r = 0.667$; Model 6: $r = 0.671$).
+This result is not altogether unexpected given recent research that recommended at least $N=100$ observations per item for accurate item parameter estimation \cite{konig2020optimized}.
+This result, however, must be interpreted in context.
+We demonstrate below that the less-than-ideal recovery of item discrimination parameters does not seriously impact the outputs of optimal test assembly.
+
+Next we calculated the statistical power (i.e.
+true positive rate) for detecting an association between item attributes and item parameters.
+We defined a true positive as an association between an item attribute and parameter whose 95\% highest density interval excluded zero.
+The results are summarized in the second column of Figure \ref{fig:figS04}.
+For the item difficulty parameters, we detected the large association (20\% variance explained) and small association (10\% variance explained) in all 100 simulations.
+This result indicates that we were suitably powered to detect associations between item attributes and item difficulty of the magnitude reported in the manuscript.
+In contrast, true positive rates for contrasts involving item discrimination were smaller.
+For large effects (14\% variance explained), the TPR was approximately 70\% (Model 5: TPR = 0.720; Model 6: TPR = 0.700).
+For small effects (7\% variance explained), the TPR was above 40\% (Model 5: TPR = 0.435; Model 6: TPR = 0.400).
+Thus, while our study design was moderately powered for detecting larger associations between item attributes and item discrimination, we are underpowedered for detecting smaller associations.
+As such, our mostly null findings for associations between item attributes and item discrimination should be interpreted with caution.
+(That said, if any associations do exist, they are likely small in magnitude.) Regardless, this result does not alter one of the main findings of the item calibration study; namely, item clones systematically differ in their difficulty by distractor type and are not therefore exchangeable.
+
+Finally, we investigated the effects of item parameter misestimation on test assembly.
+For each simulated dataset, we submitted the ground-truth (simulated) and model-predicted (recovered) item parameters to test assembly under the same constraints used to generate the MaRs-IB short form measures.
+We then calculated for each test form its test information function (TIF) and IRT test score reliability \cite{kim2010estimation, nicewander2018conditional}.
+Crucially, we calculated the TIF and reliability for the short forms using the true (i.e.
+not recovered) item parameter estimates; therefore, we can quantify how much the psychometric properties of test forms are impacted by suboptimal item selection due to parameter estimation noise.
+The results are summarized in the third and fourth columns of Figure \ref{fig:figS04}.
+As is expected, TIF was larger for the test forms made using the true item parameters, indicating that estimation noise can lead to suboptimal item selection.
+Despite this, estimation noise led to only marginal decrements in IRT test reliability.
+For item parameters estimated using Model 5, parameter noise resulted in an average loss of score reliability of $\Delta \rho = 0.020$ (SD = 0.014).
+For item parameters estimated using Model 6, parameter noise resulted in an average loss of score reliability of $\Delta \rho = 0.019$ (SD = 0.013).
+
+In sum, our sample size and experiment design was sufficient with regard to our study objectives.
+The results of the parameter recovery analysis demonstrate that we are able to estimate item difficulty parameters, and their association with item attributes, with excellent precision.
+In contrast, we were able to estimate item discrimination parameters, and their association with item attributes, only with moderate precision.
+Because the observed variability in item discrimination parameters were small, however, imprecision in the estimation of these parameters yields minimal impact on subsequent test assembly.
+Indeed, the loss of score reliability due to suboptimal item selection was negligible.
+
 \bibliography{supplement}
@@
 \centering
-\begin{tabular*}{\textwidth}{ccccccllll}
+\begin{tabular*}{0.9\textwidth}{cccccclll}
 \toprule
- & & & & & & \multicolumn{4}{c}{Spearman rank correlation} \\
-\cmidrule(lr){7-10}
-Measure & & Mean (SD) & & IQR & &  NFC-10 & PCF-8a & SNS & MaRs-SF \\
+ & & & & & & \multicolumn{3}{c}{Spearman rank correlation} \\
+\cmidrule(lr){7-9}
+Measure & & Mean (SD) & & IQR & &  NFC-10 & PCF-8a & SNS \\
 \midrule
-NFC-10 & & 25.03 (8.27) & & 20.00 -- 31.00 & & \multicolumn{1}{c}{-} &  &  & \\
-PCF-8a   &  & 22.17 (6.29) & & 18.75 -- 26.25 & & 0.27** &  \multicolumn{1}{c}{-} &  &  \\
-SNS   &  & 29.07 (7.48) & & 25.00 -- 35.00 & & 0.46** &  0.29** &  \multicolumn{1}{c}{-} &   \\
-MaRs-SF & &   8.00 (2.53) & &   6.00 -- 10.00 & & -0.04 &  0.04 &  0.14* &  \multicolumn{1}{c}{-} \\
+NFC-10  & & 24.44 (9.01) & & 19 -- 31 & & \multicolumn{1}{c}{-} &  &  \\
+PCF-8a  & & 22.58 (6.45) & & 18 -- 28 & & 0.28** &  \multicolumn{1}{c}{-} &   \\
+SNS     & & 29.02 (7.36) & & 25 -- 35 & & 0.42** &  0.32** &  \multicolumn{1}{c}{-}   \\
+MaRs-IB SF & &  8.00 (2.53) & &  6 -- 10 & & -0.04  &  0.04 &  0.14*  \\
+MaRs-IB LF & & 15.73 (5.13) & & 12 -- 20 & & 0.12*   &  -0.06 &  0.18*  \\
 \bottomrule
@@
 \captionsetup{width=1.\textwidth}
-\caption{\normalfont Correlations between performance on the MaRs-IB short forms and self-report measures.
+\caption{\normalfont Correlations between performance on the MaRs-IB short/long form and self-report measures.
 Abbreviations: IQR = interquartile range; NFC-10 = need for cognition (10-item) scale \citep{chiesi2018applying}; PCF = PROMIS cognitive functioning scale (8a; \citealt{iverson2021normative}); SNS = subjective numeracy scale \citep{fagerlin2007measuring}.
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=1.0\textwidth]{figures/figS04.png}
+\caption{\label{fig:figS04} Summary of results from the parameter recovery \& power analysis.
+(Column 1) Pearson correlation coefficients between the ground-truth and recovered model parameters.
+(Column 2) True positive rates for detecting an association between item attributes and item parameters.
+For the item difficulty parameters (beta), large and small effects indicate 20\% and 10\% variance explained, respectively.
+For the item discrimination parameters (alpha), large and small effects indicate 14\% and 7\% variance explained, respectively.
+(Column 3) Test information function estimates for the test forms produced using the ground-truth (simulated; blue) and recovered (orange) item parameter estimates.
+(Column 4) IRT test score reliability estimates for the test forms produced using the ground-truth (simulated; blue) and recovered (orange) item parameter estimates.}
+\end{figure}
+
 \end{document}

### 2022-11-18 manuscript/supplement.tex
camera ready docs

@@
 In contrast, we observed only adequate recovery of the item discrimination parameters (Model 5: $r = 0.667$; Model 6: $r = 0.671$).
-This result is not altogether unexpected given recent research that recommended at least $N=100$ observations per item for accurate item parameter estimation \cite{konig2020optimized}.
+This result is not altogether unexpected given recent research that recommended at least $N=100$ observations per item for accurate item parameter estimation \citep{konig2020optimized}.
 This result, however, must be interpreted in context.
@@
 For each simulated dataset, we submitted the ground-truth (simulated) and model-predicted (recovered) item parameters to test assembly under the same constraints used to generate the MaRs-IB short form measures.
-We then calculated for each test form its test information function (TIF) and IRT test score reliability \cite{kim2010estimation, nicewander2018conditional}.
+We then calculated for each test form its test information function (TIF) and IRT test score reliability \citep{kim2010estimation, nicewander2018conditional}.
 Crucially, we calculated the TIF and reliability for the short forms using the true (i.e.

### 2022-11-18 manuscript/letter.tex

% set document class
\documentclass[a4paper,notitlepage,12pt]{article}

% load packages
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[a4paper, margin=0.8in]{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage[backend=biber,style=nature]{biblatex}
\usepackage{csquotes}

% specify bibliography source
\addbibresource{letter.bib}
\renewcommand*{\bibfont}{\small}

% specify link style
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    citecolor=blue
}

% specify some custom commands
\renewcommand{\baselinestretch}{1.2}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand*{\Affilfont}{\normalsize}

% set up the author block
\setlength{\affilsep}{2em}   % set the space between author and affiliation
\author[1]{Samuel Zorowitz}
\author[3]{Gabriele Chierchia}
\author[3]{Sarah-Jayne Blakemore}
\author[1,2]{Nathaniel D. Daw}
\affil[1]{Princeton Neuroscience Institute, Princeton University, USA}
\affil[2]{Department of Psychology, Princeton University, USA}
\affil[3]{Department of Psychology, University of Cambridge, Downing Street, Cambridge, UK}

% specify the title
\title{Response to reviewers for ``An item response theory analysis of the Matrix Reasoning Item Bank (MaRs-IB)''}

% turn date off
\date{}

% Define paragraph formatting
\setlength{\parindent}{0em}
\setlength{\parskip}{1em}

\begin{document}

\maketitle

\section*{Formatting note}

Formatting note: in the document below, reviewers’ comments are indicated in bold typeface. Quoted text from the revised manuscript is presented in quotation blocks; within these excerpts, text that is new is indicated in italicized typeface. All changes to text in the revised manuscript are detailed below. Furthermore, changes to text in the manuscript have been highlighted (note: highlighted section headers indicate all text in that section is new or has been revised).

\section*{Reviewer \#1}

\textbf{This paper proposes the validation of a matrix item bank to assess cognitive ability and provides three parallel forms with some evidence of reliability and convergent validity. For this purpose, it starts from a public matrix item bank of items with 4 response options, described in Chierchia (2019). The main contribution is that here a larger sample is used (1501 people vs. 659 in the original study), as well as a better application design (since in Chierchia's study, the test was applied under time-limited conditions which led to a high presence of missing values in many of the items). The tested bank is composed of 64 tested item templates and 384 derived clones.}

\textbf{In general, I believe that this study can be useful for researchers by making available a calibrated tool for the evaluation of abstract reasoning. In addition, the results found concerning the prediction of difficulty in this type of item are in line with previous studies. On the other hand, the paper is well written and the analyses, quite sophisticated, have been carried out correctly. It is also of particular value that the authors provide the code and scripts. The modeling of multilevel item structure, the application of item response theory, and the use of response times to improve the estimation of item parameters also add value.}

\textbf{However, I have some concerns regarding the tests generated, which I describe below.}

% Reviewer #1, major point #1

\textbf{First, several of the item response options have been developed using the ``minimum difference'' strategy. The use of this strategy poses a great potential threat to the validity of the test, since the correct option can be identified by looking at the answer options only (the correct answer is the one that is minimally different from the rest of the options). Even though this problem is not detected in this empirical sample, it is potentially an important problem. On the other hand, since the article evaluates the performance of different strategies in the construction of response options, it should devote more space to explaining these strategies (rather than referring to another paper). It is also striking that the final tests are composed almost exclusively of PD items, which would indicate that the MD items are far from optimal.}

We agree with the reviewer that the validity of items with MD-type distractors would be compromised if participants solved them just by looking at the response options, and that this is an important issue to consider when evaluating these items. However, our results suggest that this is not a major issue in practice given that, in this dataset, those items are on average more difficult than items with PD-type distractors (as the reviewer themselves notes). Moreover, we suspect there is a simpler reason why our new test forms are primarily composed of items with PD-type distractors: when selecting items to maximize test information, all else equal, more difficult items contribute less information than easier items. This is because participants' responses to difficult items are more likely to be contaminated by guessing, and are thus less informative about participants' latent ability \cite{lord1968analysis}. As a consequence, an optimal test assembly procedure maximizing test information (i.e. test reliability) is more likely to select the easier, PD-type item clones.

Regardless, we agree with the reviewer that it would be useful to include additional text describing these issues in the manuscript rather than referring readers to other articles. As such, we have expanded the methods to include the following text (page 9, first full paragraph):

\begin{displayquote}
Importantly we also counterbalanced the assignment of item clones across participants, such that we had an approximately uniform number of responses available for each clone by shape set (1, 2, or 3) and distractor type (minimal difference, MD, or paired difference, PD). \textit{Distractor type refers to the two strategies used to generate the distractor response options in the MaRs-IB. The MD strategy produces distractors that are variations of the target response. The MD strategy has the advantage of preventing pop-out effects (where the correct response ``pops out'' among the possible responses), but has the disadvantage of theoretically allowing participants to solve an item by looking at the response options only. In contrast, the PD strategy produces distractors that are have at least one component in common with the target response. The PD strategy prevents participants from solving items by looking at the response options alone, but can potentially induce pop-out effects. We note that, for complex items with many elements, it is not possible to prevent both at the same time.}
\end{displayquote}

% Reviewer #1, major point #2

\textbf{Second, the small sample sizes (considering the complexity of the models applied) are striking. Although the total sample is 1501 persons, the 64 item templates have been applied to subsamples of about 375 persons (a somewhat small size) and the item clones to samples of about 62 subjects (see page 9 of 54). Although auxiliary information is used in the calibration of item parameters (response times, item characteristics, etc.), this sample does not seem to be sufficient, and may certainly lead to low power for some contrasts.}

We thank the reviewer for this important point and agree that we should have included further justification of our sample size and clear discussion about the implications of the sample size for the study's various goals and conclusions. We have therefore conducted a parameter recovery and power analysis, which we now report in the Supplementary Materials (see ``Parameter recovery \& power analysis'' section beginning on page S6). The complete details of the methods and results of the analysis can be found there. In general, we find that the sample was adequately powered for the analyses underlying our positive conclusions (about associations between item characteristics and item difficulty) and, importantly, for guiding the assembly of reliable test forms; but less well-powered for testing associations between item characteristics and item discrimination (where we indeed had not reported positive results).

In slightly more detail, based on our study design, the new analyses found that our statistical models were capable of accurately estimating item difficulty parameters with high precision, but were able to accurately estimate item discrimination parameters with only adequate precision. Accordingly, we found that we were well-powered to detect moderate-to-large associations between item attributes and item difficulty, but only adequately powered to detect large associations (i.e. explaining 14\% or more variance) between item attributes and item discrimination. The latter results have minimal impact on the main conclusions of our study. We were more than sufficiently powered to detect associations between item attributes and item difficulty; therefore, we can be confident in our conclusion that items clones differ systematically in difficulty by distractor type and are not therefore exchangeable. Furthermore, as we now demonstrate in the Supplementary Materials, the somewhat noisier estimation of item discrimination parameters leads only to a negligible loss in the reliability of test forms generated from our optimal test assembly procedure. Therefore, we believe our sample size and experiment design was sufficient for the primary objectives of the item calibration study.

We now point out the implications of these power considerations at appropriate points in the manuscript, and in particular stress the caveat that asking more detailed questions about item discrimination would require a larger followup sample. Concretely, in the \textit{Goodness-of-fit} section of the methods for the Calibration Study, we have included the following paragraph (page 16):

\begin{displayquote}
\textit{As a final validation of the best-fitting model, we also performed a parameter recovery analysis. In this analysis, we generated 100 artificial datasets with sampling and statistical properties matched to what we observed empirically. That is, we generated datasets with 1500 participants and 384 item clones (nested in 64 item templates), where the item parameters were randomly generated and matched to the observed distribution of item parameters. We then fit the best-fitting model to each artificial dataset, and quantified the consistency between the ground-truth and recovered model parameters. The complete details of this procedure can be found in the supplementary materials.} 
\end{displayquote}

In the \textit{Goodness-of-fit} section of the results for the Calibration Study, we have included the following paragraph (page 22):

\begin{displayquote}
\textit{Finally, we inspected the results of the parameter recovery analysis (the complete results are reported in the supplementary materials). Briefly, we found we were able to recover item difficulty parameters with excellent precision. Conversely, we observed only adequate recovery of the item discrimination parameters. The results may in turn explain why we detected only one credible association between item attributes and item discrimination. Although we were adequately-powered to detect large associations between item attributes and discrimination (i.e. explaining 14\% or more variance in the latter), we are poorly powered to detect smaller associations. This was not true for item difficulty, where we were well-powered to detect associations of the magnitude reported above. Importantly, we found through follow-up analyses that the less-than-perfect recovery of item discrimination parameters yielded only negligible effects on the reliability of test forms produced using optimal test assembly procedures. Therefore, our analyses were sufficiently powered for the primary objectives of this study.}
\end{displayquote}

Finally, the discussion of the Calibration Study, we have included the following sentences (page 23):

\begin{displayquote}
We also investigated how item complexity shapes item functioning. We found that element number and rule number were both positively associated with item difficulty, the effects of which were of approximately equal magnitude (i.e. a one-unit change in either was independently associated with a roughly 10\% reduction in performance). This finding is interesting in light of previous investigations of nonverbal reasoning tasks, which have found a greater influence on item difficulty from either the number of elements (Bethell-Fox, Lohman, \& Snow, 1984) or number of rules (Mulholland, Pellegrino, \& Glaser, 1980). One possible reason for this discrepancy is that these two attributes are largely uncorrelated across items in the MaRs-IB, which allows for unconfounded estimates of their effects. Together element number and rule number explained 67.6\% of the variance in difficulty across templates (and 38.6\% of the variance across item clones), which is in line with previous investigations of matrix reasoning tasks  (P. A. Carpenter, Just, \& Shell, 1990; Matzen, Van der Molen, \& Dudink, 1994). These findings further validate the design of the MaRs-IB insofar that item complexity is a primary determinant of item difficulty. \textit{Finally, we found that rule number, but not element number, was associated with item discrimination. However, this finding must be interpreted with caution as our parameter recovery analysis revealed our sample size was adequately powered to detect only larger effects (attributes explaining 14\% or more variance in item discrimination). Future studies with larger samples will be required to more thoroughly investigate the relationship between item complexity and discrimination.}
\end{displayquote}

% Reviewer #1, major point #3

\textbf{Third, some norms should be provided to be useful for researchers. For this, the test should be applied to more representative samples. I see the author's analysis as a first step.}

We wholeheartedly agree with the reviewer that populations norms for MaRs-IB test scores would be helpful in order for researchers to compare the outcomes of specific individuals or groups against the performance of the general population. However, we believe that providing population norms is outside the scope of the aims of the current article. Here, we are aiming to address a growing need for individual differences researchers in this current era of online convenience samples: a paucity of unique and reusable matrix-reasoning items for measuring cognitive ability. To address this need, we have provided an in-depth demonstration that the MaRs-IB is both psychometrically sound and ideal for designing parallel test forms for online convenience samples.

Regardless, we agree that representative normative samples would also be useful for many other reasons. Therefore we mention in the discussion that this is a limitation we should address in future work (page 31):

\begin{displayquote}
The current investigation of the MaRs-IB is not without its own limitations. One notable limitation is our sample. Here we analyzed response data collected from an online adult sample which was relatively young and well-educated. We cannot guarantee that the psychometric properties of the MaRs-IB reported here will generalize to other populations or testing contexts. Future researchers should consider replicating the current study in other populations of interest (e.g. children, clinical samples). By using item response models in this study, however, we make possible the opportunity for future IRT ``linking'' studies. IRT linking describes a set of methods to establish the comparability of item and ability parameters estimated from response data collected from two or more groups that differ in ability (Lee \& Lee, 2018). Future studies might exploit these methods to provide new insights, not only in how the functioning of the MaRs-IB may differ in across populations, but also in how matrix reasoning ability changes across populations. \textit{Future studies involving larger samples could also provide populations norms for MaRs-IB test scores, which would enable researchers to compare the outcomes of specific individuals or groups (e.g. clinical groups) against the performance of the general population (of a particular country or region)}.
\end{displayquote}

% Reviewer #1, major point #4

\textbf{Fourth, in the validation study, three forms of 12 items are proposed. This seems a very small number of items if adequate psychometric properties are to be obtained. The authors report that they have used the ordinal alpha coefficient, which may be an overestimate of reliability as it does not estimate the reliability of the observed score (instead, provides the reliability of the sum of the continuous variables underlying the items; see Chalmers, 2008; Flora, 2020). Authors should report the Green and Yang coefficient (also provided by the semTools package). For example, the values of these coefficients (omega2 or omega3) are below 0.75 in several of the tests (in the first test, omega3 = 0.65). It would also be important to report the traditional measures of model fit for these three tests (RMSEA, CFI, etc.). The CFI for the third test does not seem adequate (i.e., it is lesser than 0.9). It seems a limitation that the proposed tests do not have fully satisfactory psychometric properties.}

We thank the reviewer for this crucial feedback. After reading Flora (2020) \cite{flora2020your}, we agree that the ordinal alpha coefficient is not a measure of score reliability and is therefore not useful to report. We are not enthusiastic however, about instead calculating score reliability using the Green and Yang coefficient. The GY coefficient is calculated based on pairwise item tetrachoric correlations using a probit response function. This neglects that our data are contaminated by guessing (i.e. item response functions are characterized by a nonzero lower asymptote). In our view, it is more appropriate to calculate and report the IRT test reliability coefficient \cite{kim2010estimation, nicewander2018conditional}, which is based on the empirical item response functions. Based on the IRT test reliability coefficient, we find that the three short-form measures possess only adequate score reliabilities $(\rho > 0.7$). Thus, the reviewer was right to suspect that we overestimated the reliability of the MaRs-IB short-form measures. The main text has been updated to reflect this.

Given that each of the parallel short-form measures possess less-than-desirable score reliability lower, we have designed two 24-item long-form measures. Based on the IRT test reliability coefficient, these long-form measures exhibit acceptable score reliability $(\rho > 0.8$). We then subjected these two measures to the same empirical test as the short-form measures. That is, we recruited an additional N=300 participants to complete the two long-form measures and the abbreviated Raven's progressive matrices measure. The methods and results of this additional study have been incorporated into the Validation Studies section (starting on page 24).

The reviewer also suggests reporting model fit indices for all the MaRs-IB test form measures. We agree with the reviewer that such information is important to include in the manuscript. In doing so, however, we look for an alternative to the suggestion of reporting traditional (SEM-based) measures of model fit (e.g. RMSEA, CFI, TLI). This is because a number of studies have demonstrated the limitations of these indices for categorical data \cite{reise2014evaluating}, especially for evaluating models involving dichotomous data \cite{clark2018model}, small numbers of participants, and few degrees of freedom \cite{kenny2015performance}. Accordingly, we instead rely on an established posterior predictive model check: the $\chi^2_{NC}$ discrepancy measure, which compares the observed and model-predicted proportion of participants at each total score level  \cite{sinharay2006posterior}. Using this discrepancy measure, we find that the distribution of observed test scores for each test form is not credibly different from model predictions using the item parameter estimates from the calibration study (all posterior predictive p-values $> 0.05$).

In sum, based on the reviewer's excellent feedback, we have improved the psychometric evaluation of our test form measures. We have replaced estimates of reliability using the ordinal alpha coefficient with the more appropriate IRT test reliabiltiy coefficient. In response to the lower-than-anticipated reliability of the 12-item short form measure, we have designed and in an additional study validated two 24-item long-form measures with good reliability $(\rho > 0.8$). Thus, we provide researchers with two possible choices of MaRs-IB test forms that balance reliability and administration time trade-offs. Finally, we demonstrate that the observed score distributions of each test form conforms to expectations based on the results of the item calibration study.

\textbf{Minor issues:}

% Reviewer #1, minor point #1

\textbf{1) The prior distributions used should be reported.}

We thank the author for this point. We have now described the priors for the additive multilevel item structure (AMIS) models in the supplement (beginning on page S4). Moreover, we have added a sentence to the methods to alert readers to this section (page 15).

% Reviewer #1, minor point #2

\textbf{2) In the test assembly, performed by linear programming, the constraint is that the alpha coefficient should be greater than or equal to 0.8, but the reported values are slightly less than 0.8.}

We thank the reviewer for this point. We have updated the text to reflect the revised reliability estimates (page 25):

\begin{displayquote}
\textit{Each short form was required to contain 12 items. This was chosen to minimize the administration time of a given form (2-4 minutes on average) while achieving a score reliability $\geq 0.7$.}
\end{displayquote}

\section*{Reviewer \#2}

\textbf{I think the paper is of high quality. However, I found it difficult to understand for readers. I suggest Authors to simplify some parts. Thanks.}

We thank the reviewer for the positive review. We have attempted to clean up and simplify the manuscript in parts. For example, we have simplified and shortened our description of the additive multilevel item structure models. We have also rewritten and streamlined our reporting of the second (i.e. validation) study.

\printbibliography

\end{document}

### 2022-11-18 manuscript/main.tex

@@
-\documentclass[a4paper,man,natbib]{apa6}
+\documentclass[a4paper,man,natbib,noextraspace]{apa6}
 
@@
 However, we also find that item clones are not always psychometrically equivalent and cannot assumed to be exchangeable.
-In a second study, we demonstrate how researchers can use the item parameters we estimated to design new MaRs-IB tests using optimal item assembly.
-With these methods, we design two new sets of test forms that exhibited good score reliability and convergent validity with an established measure of matrix reasoning in an independent sample of adults (N=600).
-We hope that the materials, results, and tutorials made available here will encourage researchers to use the MaRs-IB in their studies.}
+In a second study, we demonstrate how experimenters can use the estimated item parameters to design new matrix reasoning tests using optimal item assembly.
+Specifically, we design and validate two new sets of test forms in an independent sample of adults (N=600).
+We find these new tests possess good reliability and convergent validity with an established measure of matrix reasoning.
+We hope that the materials and results made available here will encourage experimenters to use the MaRs-IB in their research.}
 
@@
 All data, code, model outputs (including the estimated item parameters), and tutorials are publicly available at: \url{https://github.com/ndawlab/mars-irt}.
+The MaRs-IB stimuli are publicly available at: \url{https://osf.io/g96f4/}.
 
@@
 Specifically, the residual variance of latent ability was specified as $V(\epsilon) = 1 - \sum \rho_p^2$.
-The specification of the model priors are detailed in the supplement.
+\hl{The specification of the model priors are detailed in the supplement.}
 
